---
title: "Alpha diversity across treatment periods"
author: "Sneha Sundar"
date: "`r format(Sys.time(), "%a %b %d %X %Y")`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Load required packages
```{r, message=FALSE, error=FALSE}
library(tidyverse)
library(phyloseq)
library(speedyseq)
library(ampvis2)
library(microbiome)
library(here)
library(ggpubr)

options(getClass.msg=FALSE) # https://github.com/epurdom/clusterExperiment/issues/66
#this fixes an error message that pops up because the class 'Annotated' is defined in two different packages
```

### Load required functions
```{r}

'%!in%' <- function(x,y)!('%in%'(x,y))

source("https://raw.githubusercontent.com/fconstancias/DivComAnalyses/master/R/phyloseq_taxa_tests.R")
source("https://raw.githubusercontent.com/fconstancias/DivComAnalyses/master/R/phyloseq_normalisation.R")
source("https://raw.githubusercontent.com/fconstancias/DivComAnalyses/master/R/phyloseq_alpha.R")
source("https://raw.githubusercontent.com/fconstancias/DivComAnalyses/master/R/phyloseq_beta.R")
source("https://raw.githubusercontent.com/fconstancias/DivComAnalyses/master/R/phyloseq_heatmap.R")
```


### Load phyloseq object

```{r}
ps = "data/processed/physeq_update_11_1_21.RDS"

ps %>% 
  here::here() %>%
  readRDS() %>%
  phyloseq_get_strains_fast() %>%
  phyloseq_remove_chloro_mitho() -> physeq

```


### We want only continuous samples that are not IR2
```{r}
physeq %>% 
  subset_samples(Experiment == "Continuous") %>% 
  subset_samples(Reactor != "IR2") -> ps_polyFermS
```


### Rarefying the data
```{r message = FALSE}
ps_polyFermS %>%  rarefy_even_depth(sample.size = 4576,rngseed = 123) -> ps_rare
```

### Filtering

Taking only the unenriched samples, getting rid of dubious sample.

```{r}
ps_rare %>%
  subset_samples(Enrichment == "NotEnriched") %>%
  subset_samples(Sample_description %!in% c("TR5-15", "TR4-1")) -> ps_rare
```

### Adding treatment period to the metadata

```{r}
sam <- ps_rare %>% 
  sample_data() %>% as.matrix() %>% as.data.frame() %>% 
  mutate(Day_of_Treatment = as.numeric(Day_of_Treatment))


sam <- sam %>%
  mutate(
    Period = case_when(
      Day_of_Treatment == -2 | Day_of_Treatment == -3 | Day_of_Treatment == -6 | Day_of_Treatment == -7 ~ "pret",
      Day_of_Treatment < 10 & Day_of_Treatment >= 0 ~ "t1",
      Day_of_Treatment < 20 & Day_of_Treatment >= 10 ~ "t2",
      Day_of_Treatment < 30 & Day_of_Treatment >= 20 ~ "t3",
      Day_of_Treatment < 40 & Day_of_Treatment >= 30 ~ "t4",
      Day_of_Treatment < 50 & Day_of_Treatment >= 40 ~ "t5",
    )
  )

#remember you have na in the rows 


```

### Adding modified metadata to physeq object and removing the samples. 

```{r}
sample_data(ps_rare) <- sam

ps_rare <- ps_rare %>% subset_samples(!is.na(Period))
```

### Compute alpha diversity metrics.

```{r}
ps_rare %>%
  phyloseq_alphas(phylo = FALSE) -> alpha_df

alpha_df
```

## CTX specific graphs

```{r}
ctx_alpha_df <- alpha_df %>% filter(Reactor %in% c("CR","TR1","TR2","TR3"))

```





```{r}
ctx_shannon <- ggboxplot(ctx_alpha_df, x = "Period", y = "diversity_shannon", color = "Reactor_Treatment"
          , add = "jitter")

pdf(here("output","metabarcoding","exploratory","ctx_shannon.pdf"), width=10,height=6)
  ctx_shannon
dev.off()

```

```{r}
ctx_obs <- ggboxplot(ctx_alpha_df, x = "Period", y = "Observed", color = "Reactor_Treatment"
          , add = "jitter")

pdf(here("output","metabarcoding","exploratory","ctx_obs.pdf"), width=10,height=6)
  ctx_obs
dev.off()
```

```{r}
ctx_eveness <- ggboxplot(ctx_alpha_df, x = "Period", y = "evenness_pielou", color = "Reactor_Treatment"
          , add = "jitter")

pdf(here("output","metabarcoding","exploratory","ctx_eveness.pdf"), width=10,height=6)
  ctx_eveness
dev.off()
```


## VAN specific graphs

```{r}
van_alpha_df <- alpha_df %>% filter(Reactor %in% c("CR","TR4","TR5","TR6"))

```





```{r}
van_shannon<-ggboxplot(van_alpha_df, x = "Period", y = "diversity_shannon", color = "Reactor_Treatment"
          , add = "jitter")


pdf(here("output","metabarcoding","exploratory","van_shannon.pdf"), width=10,height=6)
  van_shannon
dev.off()
```

```{r}
van_obs <- ggboxplot(van_alpha_df, x = "Period", y = "Observed", color = "Reactor_Treatment"
          , add = "jitter")


pdf(here("output","metabarcoding","exploratory","van_obs.pdf"), width=10,height=6)
  van_obs
dev.off()
```

```{r}
van_eveness <- ggboxplot(van_alpha_df, x = "Period", y = "evenness_pielou", color = "Reactor_Treatment"
          , add = "jitter")

pdf(here("output","metabarcoding","exploratory","van_eveness.pdf"), width=10,height=6)
  van_eveness
dev.off()

```

