---
title: " beta-diversity "
author: "Florentin"
date: '`r format(Sys.time(), "%d %B, %Y")`'

output: 
  html_document: 
    toc: yes
    keep_md: yes
---
```{r setup, include=FALSE}
rm(list = ls()) # remove all the object before starting
knitr::opts_chunk$set(echo = TRUE)
```

#### Load required packages

```{r,packages,message=FALSE}
library(tidyverse)
library(phyloseq)
library(speedyseq)
library(ggrepel)
library(ampvis2)
library(plotly)
library(microbiome)
options(getClass.msg=FALSE) # https://github.com/epurdom/clusterExperiment/issues/66
#this fixes an error message that pops up because the class 'Annotated' is defined in two different packages
```

#### Load functions from Github

```{r}
'%!in%' <- function(x,y)!('%in%'(x,y))

source("https://raw.githubusercontent.com/fconstancias/DivComAnalyses/master/R/phyloseq_taxa_tests.R")
source("https://raw.githubusercontent.com/fconstancias/DivComAnalyses/master/R/phyloseq_normalisation.R")
source("https://raw.githubusercontent.com/fconstancias/DivComAnalyses/master/R/phyloseq_alpha.R")
source("https://raw.githubusercontent.com/fconstancias/DivComAnalyses/master/R/phyloseq_beta.R")
source("https://raw.githubusercontent.com/fconstancias/DivComAnalyses/master/R/phyloseq_heatmap.R")
```


#### Load physeq object


```{r}
ps = "data/processed/physeq_update_11_1_21.RDS"

ps %>% 
  here::here() %>%
  readRDS() %>%
  phyloseq_get_strains_fast() %>%
  phyloseq_remove_chloro_mitho() -> physeq



physeq %>% 
  subset_samples(Experiment == "Continuous") %>% 
  subset_samples(Paul %!in% c("Paul")) %>%
  subset_samples(Reactor != "IR2") -> ps_PolyFermS

```  

We will be analysing only the PolyFermS samples here so take a subset of the physeq object.  
```{r}
physeq %>% 
  subset_samples(Experiment == "Continuous") %>% 
  subset_samples(Paul %!in% c("Paul")) %>%
  subset_samples(Reactor != "IR2") -> ps_polyFermS

sample_data(ps_polyFermS)$Reactor <- fct_relevel(sample_data(ps_polyFermS)$Reactor, "IR1", "CR", "TR1", "TR2","TR3", "TR4", "TR5", "TR6") 

sample_data(ps_polyFermS)$Treatment <- fct_relevel(sample_data(ps_polyFermS)$Treatment, "UNTREATED",  "CTX+HV292.1", "CTX","HV292.1","VAN+CCUG59168", "VAN",  "CCUG59168") 

sample_data(ps_polyFermS)$Reactor_Treatment <- fct_relevel(sample_data(ps_polyFermS)$Reactor_Treatment, "IR1_UNTREATED","CR_UNTREATED", "CR_CTX", "CR_VAN", "TR1_CTX+HV292.1","TR2_CTX", "TR3_HV292.1", "TR5_VAN+CCUG59168", "TR4_VAN", "TR6_CCUG59168") 

ps_polyFermS %>% 
  rarefy_even_depth(sample.size = 4576,
                    rngseed = 123) -> ps_polyFermS_rare
```

Compute beta-div metrics:

```{r}
ps_polyFermS_rare %>%
  phyloseq_compute_bdiv(norm = "pc",
                        phylo = TRUE,
                        seed = 123) -> bdiv_list
```


```{r,fig.align="center", warning = FALSE, message = FALSE, results = FALSE}
ps_polyFermS_rare  %>%
  subset_samples(Enrichment == "NotEnriched") %>%
  phyloseq_plot_bdiv(bdiv_list,
                     m = "CoDa",
                     seed = 123) -> coda
  
coda$PCA$layers[[1]] = NULL

coda$PCA + geom_point(size=2,
                   aes(color = Reactor_Treatment, 
                       fill = Reactor_Treatment,
                       shape = NULL,
                       alpha = Day_of_Treatment)) + 
  theme_light() +
  geom_path(arrow = arrow(type = "open", angle = 30, length = unit(0.15, "inches")),
              size = 0.05, linetype = "dashed", inherit.aes = TRUE, aes(group=Reactor_Treatment, color = Reactor_Treatment), show.legend = FALSE) +
  scale_alpha_continuous(range=c( 0.9, 0.3)) + scale_color_viridis_d(na.value = "black") + 
  scale_fill_viridis_d(na.value = "black") + 
  # scale_shape_manual(values = c(8, 21, 22, 23, 24, 16, 15, 18, 17)) + 
  theme_classic() +
    labs(col=NULL, fill = NULL, shape = NULL) + guides(shape=FALSE) -> p1

p1
```

```{r,fig.align="center"}
p1 %>%
  ggplotly() -> p1ply

p1ply
```

```{r,fig.align="center", warning = FALSE, message = FALSE, results = FALSE}

htmlwidgets::saveWidget(as_widget(p1ply), 
  paste0(here::here(),
                    "/data/processed/",
       "beta_",
       format(Sys.time(), "%Y%b%d"),".html"))
```

Visualize distance 'trajectories' :

We will use the library use dist to get it in the dataframe we want for plotting [useDist](https://cran.r-project.org/web/packages/usedist/readme/README.html)
```{r}
ps = coda$physeq_clr

ps %>% 
  otu_table() %>%
  vegan::vegdist(method = "eucl") -> mtest

mtest %>%
  usedist::dist_groups(ps %>% get_variable("Reactor_Treatment")) -> dist_df

ps %>% 
    sample_data() %>%
    data.frame() %>%
    rownames_to_column('tmpid') %>%
    dplyr::select(tmpid, Day_from_Inoculum) -> meta_df
```

```{r}
left_join(
  dist_df,
  meta_df %>%
    dplyr::rename("varGroup1" = Day_from_Inoculum),
  by = c("Item1" = "tmpid")) %>%
  left_join(
    meta_df %>%
    dplyr::rename("varGroup2" = Day_from_Inoculum),
  by = c("Item2" = "tmpid")) -> test
```

Please carefully check if the code makes sense 

```{r}
test %>% 
  head()
```

```{r}
test %>%
  dplyr::filter(grepl("Within", Label)) %>%
  dplyr::group_by(Label, Item1, Item2, varGroup1, varGroup2) %>%
  dplyr::arrange(varGroup1, varGroup2) %>%
  dplyr::mutate(firsto = dplyr::first(Distance)) %>%
  dplyr::ungroup() %>%
  dplyr::group_by(Label, Item1, varGroup1) %>%
  dplyr::summarise(next_dist = first(firsto)) %>%
  dplyr::arrange(varGroup1, Label) -> test_prev

test_prev %>%
  DT::datatable()
```

Please carefully check if the code makes sense

```{r}
test_prev %>%
  ggplot(aes(x = as.factor(varGroup1) , y = next_dist)) +
  geom_point(size=1, alpha=0.6, aes(colour = Label, group=Label),
            position=position_jitterdodge(dodge.width=0.9)) + 
  geom_path(arrow = arrow(type = "open", angle = 30, length = unit(0.1, "inches")),
              size = 0.4, linetype = "dashed", inherit.aes = TRUE, aes(group=Label, color = Label),
            position=position_jitterdodge(dodge.width=0.9)) +
  # geom_boxplot(aes(group = varGroup2 %>% as.factor()),
               # fill = "transparent",
               # outlier.colour = NA,alpha=0.4) +
  facet_grid(Label ~ ., scales = "fixed") +
      # ggrepel::geom_text_repel(cex=2,
      #                      aes(label= Group1),
      #                      segment.color = 'black',
      #                      segment.size = 0.5,
      #                      # nudge_x =  -4,
      #                      # nudge_y = 0,
      #                      df_all_t0 %>% filter(Day2 == 6 & metric=="wunifrac") %>% unique()) +
  theme_bw() + xlab("Day") + ylab("Distance to prev") +
    geom_vline(xintercept = c(23,39), 
             color="black", alpha=0.4) -> tmp

print(tmp)
```


```{r}
tmp + facet_null()
```


```{r}
ref = 24

test %>%
  dplyr::filter(grepl("Within", Label)) %>%
  dplyr::group_by(Label) %>%
  dplyr::arrange(varGroup1, varGroup2) %>%
  # dplyr::mutate(firsto = dplyr::first(Distance)) %>%
  dplyr::filter(varGroup1 == !!ref &
                varGroup2 >  !!ref) %>%
    dplyr::arrange(varGroup2) -> test_ref
```
Please carefully check if the code/rationale makes sense 


I run some checks for the distance to ref:

```{r}

test_ref %>%
  DT::datatable()
```

```{r}

sample_to_check <- c("CR-1-S116", "CR-3-S98", "IR1-24-S295", "IR1-81-S298")

subset_samples(ps, 
               sample_names(ps) %in% sample_to_check) %>%
  sample_data() %>%
  data.frame() %>%
  dplyr::select(Sample_description,Day_of_Connection, Day_of_Treatment, Day_from_Inoculum) %>%
  DT::datatable()

```

```{r}

mtest %>%
  as.matrix() -> mtest_mat

mtest_mat[sample_to_check,sample_to_check]

```

Please carefully check

```{r}
test_ref %>%
  ggplot(aes(x = as.factor(varGroup2) , y = Distance)) +
  geom_point(size=1, alpha=0.6, aes(colour = Group1, group=Group1),
            position=position_jitterdodge(dodge.width=0.9)) + 
  geom_path(arrow = arrow(type = "open", angle = 30, length = unit(0.1, "inches")),
              size = 0.4, linetype = "dashed", inherit.aes = TRUE, aes(group=Group1, color = Group1),
            position=position_jitterdodge(dodge.width=0.9)) +
  # geom_boxplot(aes(group = varGroup2 %>% as.factor()),
               # fill = "transparent",
               # outlier.colour = NA,alpha=0.4) +
  facet_grid(Label ~ ., scales = "fixed") +
      # ggrepel::geom_text_repel(cex=2,
      #                      aes(label= Group1),
      #                      segment.color = 'black',
      #                      segment.size = 0.5,
      #                      # nudge_x =  -4,
      #                      # nudge_y = 0,
      #                      df_all_t0 %>% filter(Day2 == 6 & metric=="wunifrac") %>% unique()) +
  theme_bw() + xlab("Day") + ylab(paste0("Distance to ref :", ref)) +
    geom_vline(xintercept = c(23,39),
             color="black", alpha=0.4) -> tmp


print(tmp + theme(legend.position = "none") )
```

Not sure why there is two values for day 26.
Not sure why the vertical line is 61 here since I used the same command as the previous plot...


```{r}

tmp + facet_null()

```

Next step: ref (should be easy with the   dplyr::filter(grepl("Within", Label)))

```{r}
paste0(here::here(),
       "/data/processed/",
       "beta",
       "_",
       format(Sys.time(), "%Y%b%d")
       ,".RData") %>% save.image()
# load("/Users/fconstan/Projects/EZe/ASV/260420.RData")
```

```{r}
sessionInfo()
```

