---
title: "Resistome_visualization"
author: "Florentin Constancias"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: 
  html_document: 
    toc: yes
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Load required packages

```{r,packages,message=FALSE}
library(tidyverse)
library(phyloseq)
library(speedyseq)
library(ggrepel)
library(here)
library(microViz)
library(RColorBrewer)
library(vegan)
library(randomcoloR)
options(getClass.msg=FALSE) # https://github.com/epurdom/clusterExperiment/issues/66
#this fixes an error message that pops up because the class 'Annotated' is defined in two different packages
```

#### Source required functions

```{r message=FALSE}
rm(list = ls())

'%!in%' <- function(x,y)!('%in%'(x,y))

source("https://raw.githubusercontent.com/fconstancias/DivComAnalyses/master/R/phyloseq_taxa_tests.R")
source("https://raw.githubusercontent.com/fconstancias/DivComAnalyses/master/R/phyloseq_normalisation.R")
source("https://raw.githubusercontent.com/fconstancias/DivComAnalyses/master/R/phyloseq_alpha.R")
source("https://raw.githubusercontent.com/fconstancias/DivComAnalyses/master/R/phyloseq_beta.R")
source("https://raw.githubusercontent.com/fconstancias/DivComAnalyses/master/R/phyloseq_heatmap.R")
```

# Import data and define variables:

```{r}
physeq = readRDS("~/Projects/ETH/Alessia/mobilome/ps_combined.rds")
out_pptx = "~/Projects/ETH/Alessia/mobilome/graph.pptx"


physeq %>%
  tax_table() %>%
  data.frame() %>%
  mutate(Resistance_Mechanism_multi = ifelse(grepl(";", Resistance_Mechanism), "multi-mech", Resistance_Mechanism)) %>% 
  as.matrix() -> tax_table(physeq)


physeq %>% 
  tax_table() %>% 
  data.frame() -> tax_mapping

# physeq %>% 
#   tax_table() %>% 
#   data.frame() %>% 
#   mutate_if(is.character, as.factor) %>% 
#   as.matrix() -> tax_table(physeq)

# 
# physeq %>%
#   tax_table() %>%
#   data.frame() %>%
#   mutate_all(funs(str_replace(., "[^[:alnum:]]", "_"))) %>%
#   mutate_all(funs(str_replace(., "[[:punct:]]", "_"))) %>%
#   as.matrix() -> tax_table(physeq)

physeq %>% 
  otu_table() %>% 
  data.frame() %>% 
  replace(is.na(.), 0)  %>% 
  # mutate_all(funs(str_replace(., "[^[:alnum:]]", ""))) %>%
  mutate_all(as.numeric) %>% 
  as.matrix() %>% 
  otu_table(taxa_are_rows = TRUE) -> otu_table(physeq) 
# df[,'values'] <- as.numeric(df[,'values'])
# otu_table(physeq) <- tmp


```


```{r}

all.equal(taxa_names(physeq %>%  tax_table()), taxa_names(physeq %>%  otu_table()))
all.equal(taxa_names(physeq), taxa_names(physeq %>%  otu_table()))
```


```{r}
physeq_tmp <- physeq
tax_table(physeq_tmp) <- tax_table(physeq_tmp)[,c("Resistance_Mechanism_multi", "Best_Hit_ARO")]
# tax_table(physeq_tmp) <- tax_table(physeq_tmp)[,c("Best_Hit_ARO", "Drug_Class")]

physeq_tmp %>% 
  tax_glom(., taxrank = "Best_Hit_ARO") %>% 
  tax_glom(., taxrank = "Resistance_Mechanism_multi") -> physeq_res_multi

taxa_names(physeq_res_multi) <- tax_table(physeq_res_multi)[,"Resistance_Mechanism_multi"]

physeq_res_multi
```

```{r}
physeq_tmp <- physeq
tax_table(physeq_tmp) <- tax_table(physeq_tmp)[,c("Drug_Class_multi", "Best_Hit_ARO")]
# tax_table(physeq_tmp) <- tax_table(physeq_tmp)[,c("Best_Hit_ARO", "Drug_Class")]

physeq_tmp %>% 
  tax_glom(., taxrank = "Best_Hit_ARO") %>% 
  tax_glom(., taxrank = "Drug_Class_multi") -> physeq_drug_multi

taxa_names(physeq_drug_multi) <- tax_table(physeq_drug_multi)[,"Drug_Class_multi"]

physeq_drug_multi
```


```{r}
physeq_tmp <- physeq
tax_table(physeq_tmp) <- tax_table(physeq_tmp)[,c("Drug_Class", "Best_Hit_ARO")]
# tax_table(physeq_tmp) <- tax_table(physeq_tmp)[,c("Best_Hit_ARO", "Drug_Class")]

physeq_tmp %>% 
  tax_glom(., taxrank = "Best_Hit_ARO") %>% 
  tax_glom(., taxrank = "Drug_Class") -> physeq_drug

taxa_names(physeq_drug) <- tax_table(physeq_drug)[,"Drug_Class"]

physeq_drug
```

```{r}
physeq_tmp <- physeq

tax_table(physeq_tmp) <- tax_table(physeq_tmp)[,c("Best_Hit_ARO", "Drug_Class")]

physeq_tmp %>% 
  tax_glom(., taxrank = "Best_Hit_ARO") -> physeq_AMRgn

taxa_names(physeq_AMRgn) <- tax_table(physeq_AMRgn)[,"Best_Hit_ARO"]

physeq_AMRgn
```


<!-- Since we somehow can't use tax_glom, we have to do it manually: -->

<!-- ```{r} -->
<!-- as(tax_table(physeq), "matrix") %>%  -->
<!--   as.data.frame() %>% -->
<!--   rownames_to_column('ASV') -> tax_table_ps_combined -->

<!-- as(otu_table(physeq), "matrix") %>%  -->
<!--   as.data.frame() %>% -->
<!--   rownames_to_column('ASV') -> otu_table_ps_combined -->

<!-- physeq %>%  -->
<!--   sample_data() %>%  -->
<!--   data.frame() %>%  -->
<!--   rownames_to_column('sample_id') ->  meta_ps_combined -->

<!-- otu_table_ps_combined %>% -->
<!--   # left_join(refseq_df, by = 'ASV') %>% -->
<!--   left_join(tax_table_ps_combined, by = c("ASV" = "ASV")) %>% -->
<!--   dplyr::select(ASV, everything()) -> merged_ps_combined -->
<!-- ``` -->


<!-- ```{r} -->
<!-- merged_ps_combined %>%  -->
<!--   pivot_longer(cols = sample_names(physeq), -->
<!--                names_to = "sample") %>%  -->
<!--   group_by(sample, Best_Hit_ARO) %>%  -->
<!--   summarise(Best_Hit_ARO_value = sum(value)) %>%  -->
<!--   replace(is.na(.), 0) %>%  -->
<!--   dplyr::ungroup() %>%  -->
<!--   pivot_wider(names_from = "sample", -->
<!--               values_from =  Best_Hit_ARO_value ) -> gene_df -->

<!-- gene_df -->
<!-- ``` -->


Define colors for Drug classes:

```{r}
set.seed(123)
treat_col <- ggpubr::get_palette(palette = "lancet",
                           k = get_variable(physeq_drug_multi, "Treatment") %>%  
                              levels() %>% 
                             length())

names(treat_col) <- get_variable(physeq_drug_multi, "Treatment") %>%  
                              levels()
```


Define colors for Treatment - no matter concentration / fermentation / reactors:

```{r}
myCol <- viridis::viridis(n = get_variable(physeq_drug_multi, "Treatment") %>%  
                              levels() %>% 
                             length() + 20)
myCol

scales::show_col(myCol,
                 cex_label = 0.5)
```

```{r}
myCol <- ggpubr::get_palette(palette = "npg",
                             k = get_variable(physeq_drug_multi, "Treatment") %>%  
                              levels() %>% 
                             length() + 3)
myCol

scales::show_col(myCol,
                 cex_label = 0.5)
```


```{r}
myCol <- ggpubr::get_palette(palette = "jama",
                             k = get_variable(physeq_drug_multi, "Treatment") %>%  
                              levels() %>% 
                             length() + 6)
myCol

scales::show_col(myCol,
                 cex_label = 0.5)
```

```{r}
myCol <- ggpubr::get_palette(palette = "lancet",
                             k = get_variable(physeq_drug_multi, "Treatment") %>%  
                              levels() %>% 
                             length() + 5)
myCol

scales::show_col(myCol,
                 cex_label = 0.5)
```

```{r}
treat_col = get_variable(physeq_drug_multi, "Treatment") %>%  
                              levels()


# treat_col <- c("#1B1919", "#7C8181", "#00468B", "#2589AE","#029AAF", "#463161", "#7C66A2", "#ad99cf")

treat_col <- c("#1B1919", "#7C8181", "#00468B", "#3085AD", "#42B540", "#9E172E", "#B43C15", "#AD7987")


names(treat_col) <- get_variable(physeq_drug_multi, "Treatment") %>%  
                              levels()

scales::show_col(treat_col,
                 cex_label = 0.5)
```

```{r}

set.seed(123)
col <- distinctColorPalette(ntaxa(physeq_drug_multi))

names(col) <- taxa_names(physeq_drug_multi)
```


```{r, fig.height=50, fig.width=20}
physeq_AMRgn %>%
  microbiome::transform(transform = "log10p") %>% 
  speedyseq::psmelt() %>% 
  left_join(tax_mapping,
            by = c("Best_Hit_ARO" = "Best_Hit_ARO"),
            suffix = c("_x", "")) %>% 
  mutate(Abundance = na_if(Abundance, 0)) %>%
  mutate(logTPM = log10(Abundance + 1)) %>%
  separate_rows(., Drug_Class,sep = '; ',convert = TRUE) %>% 
  ggplot(mapping = aes(x = as.factor(Day_of_Treatment),
                       y = Best_Hit_ARO ,
                       fill = logTPM,
  )) +
  facet_grid(Drug_Class ~ Model + Treatment + Fermentation + Antibiotic_mg.mL , scales = "free", space = "free") + #,  labeller = labeller(Drug_Class_multi = label_wrap_gen(width = 15)))  +
  # facet_grid(. ~ Reactor_Treatment, scales = "free", space = "free", drop = TRUE,labeller = labeller(Drug.Class = label_wrap_gen(width = 15))) +
  #we adjust the vertical labels such that it is printed correctly. 
  geom_tile() +
  theme_light()  +
  # scale_fill_viridis_c(name = "TPM",
  #                      na.value = "transparent", trans = scales::pseudo_log_trans(sigma = 0.1),
  #                      # trans =  scales::pseudo_log_trans(sigma = 1),
  #                      breaks = c(1, 10, 50, 100, 400), labels = c(1, 10, 50, 100, 400),
  #                      limits = c(0, 500)) + #need to make sure that we include max value of tpm as upper limit
  scale_fill_viridis_c(na.value = "transparent") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5),
        strip.text.x = element_text(size=8),
        #change the size of the label and the angle so the text is printed horizontally
        strip.text.y = element_text(size = 7,angle=0)) -> perma_bbuble

perma_bbuble +
  theme(legend.position = "bottom") -> perma_bbuble 

perma_bbuble +
  scale_y_discrete(label = function(x) stringr::str_trunc(x, 34, side = "center")) ->perma_bbuble

perma_bbuble

perma_bbuble %>%
  export::graph2ppt(append = TRUE, width = 20, height = 50,
                    file = out_pptx)
```


```{r, fig.height=28, fig.width=20}
physeq_AMRgn %>%
  microbiome::transform(transform = "log10p") %>% 
  speedyseq::psmelt() %>% 
  left_join(tax_mapping,
            by = c("Best_Hit_ARO" = "Best_Hit_ARO"),
            suffix = c("_x", "")) %>% 
  mutate(Best_Hit_ARO = fct_reorder(Best_Hit_ARO, Abundance, .desc=FALSE)) %>% 
  mutate(Abundance = na_if(Abundance, 0)) %>%
  mutate(logTPM = log10(Abundance + 1))  %>% 
  # df$Item = factor(df$Item, levels = a_lot$Item[order(a_lot$percent)])

# tmp$variable <- factor(melted$variable, levels=rev(levels(melted$variable)))
  # mutate(Best_Hit_ARO =  fct_reorder(Best_Hit_ARO, logTPM)) %>% 
  # mutate(Best_Hit_ARO =  fct_relevel(Best_Hit_ARO, logTPM)) %>% 
  # separate_rows(., Drug_Class,sep = '; ',convert = TRUE) %>% 
  ggplot(mapping = aes(x = as.factor(Day_of_Treatment),
                       y = Best_Hit_ARO,
                       fill = logTPM,
  )) +
  facet_grid(.  ~ Model + Treatment + Fermentation + Antibiotic_mg.mL , scales = "free", space = "free") + #,  labeller = labeller(Drug_Class_multi = label_wrap_gen(width = 15)))  +
  # facet_grid(. ~ Reactor_Treatment, scales = "free", space = "free", drop = TRUE,labeller = labeller(Drug.Class = label_wrap_gen(width = 15))) +
  #we adjust the vertical labels such that it is printed correctly. 
  geom_tile() +
  theme_light()  +
  # scale_fill_viridis_c(name = "TPM",
  #                      na.value = "transparent", trans = scales::pseudo_log_trans(sigma = 0.1),
  #                      # trans =  scales::pseudo_log_trans(sigma = 1),
  #                      breaks = c(1, 10, 50, 100, 400), labels = c(1, 10, 50, 100, 400),
  #                      limits = c(0, 500)) + #need to make sure that we include max value of tpm as upper limit
  scale_fill_viridis_c(na.value = "transparent") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5),
        strip.text.x = element_text(size=8),
        #change the size of the label and the angle so the text is printed horizontally
        strip.text.y = element_text(size = 7,angle=0)) -> perma_bbuble

perma_bbuble +
  theme(legend.position = "bottom") +
  scale_y_discrete(label = function(x) stringr::str_trunc(x, 34, side = "center"))  -> perma_bbuble

perma_bbuble

perma_bbuble %>%
  export::graph2ppt(append = TRUE, width = 20, height = 28,
                    file = out_pptx)
```


```{r, fig.height=28, fig.width=20}
physeq_AMRgn %>%
  microbiome::transform(transform = "log10p") %>% 
  speedyseq::psmelt() %>% 
  left_join(tax_mapping,
            by = c("Best_Hit_ARO" = "Best_Hit_ARO"),
            suffix = c("_x", "")) %>% 
  mutate(Best_Hit_ARO = fct_reorder(Best_Hit_ARO, Abundance, .desc=FALSE)) %>%
  mutate(Abundance = na_if(Abundance, 0)) %>%
  mutate(logTPM = log10(Abundance + 1))  %>% 
  # df$Item = factor(df$Item, levels = a_lot$Item[order(a_lot$percent)])

# tmp$variable <- factor(melted$variable, levels=rev(levels(melted$variable)))
  # mutate(Best_Hit_ARO =  fct_reorder(Best_Hit_ARO, logTPM)) %>% 
  # mutate(Best_Hit_ARO =  fct_relevel(Best_Hit_ARO, logTPM)) %>% 
  # separate_rows(., Drug_Class,sep = '; ',convert = TRUE) %>% 
  ggplot(mapping = aes(x = as.factor(Day_of_Treatment),
                       y = Best_Hit_ARO,
                       fill = logTPM,
  )) +
  facet_grid(Drug_Class_multi  ~  Model + Treatment + Fermentation + Antibiotic_mg.mL , scales = "free", space = "free") + #,  labeller = labeller(Drug_Class_multi = label_wrap_gen(width = 15)))  +
  # facet_grid(. ~ Reactor_Treatment, scales = "free", space = "free", drop = TRUE,labeller = labeller(Drug.Class = label_wrap_gen(width = 15))) +
  #we adjust the vertical labels such that it is printed correctly. 
  geom_tile() +
  theme_light()  +
  # scale_fill_viridis_c(name = "TPM",
  #                      na.value = "transparent", trans = scales::pseudo_log_trans(sigma = 0.1),
  #                      # trans =  scales::pseudo_log_trans(sigma = 1),
  #                      breaks = c(1, 10, 50, 100, 400), labels = c(1, 10, 50, 100, 400),
  #                      limits = c(0, 500)) + #need to make sure that we include max value of tpm as upper limit
  scale_fill_viridis_c(na.value = "transparent") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5),
        strip.text.x = element_text(size=8),
        #change the size of the label and the angle so the text is printed horizontally
        strip.text.y = element_text(size = 7,angle=0)) -> perma_bbuble

perma_bbuble +
  theme(legend.position = "bottom") +
  scale_y_discrete(label = function(x) stringr::str_trunc(x, 34, side = "center"))  -> perma_bbuble

perma_bbuble

perma_bbuble %>%
  export::graph2ppt(append = TRUE, width = 20, height = 28,
                    file = out_pptx)


# perma_bbuble + facet_null() +  facet_grid(.  ~  Model + Treatment + Fermentation + Antibiotic_mg.mL , scales = "free", space = "free") +
#  ggside::geom_ysidetile(aes(x = 0, yfill = col[perma_bbuble$data$Drug_Class_multi]), show.legend = TRUE) -> test_annot
# 
# 
# test_annot
```

pheatmap test:

all
```{r, fig.height=30, fig.width=30}
physeq_AMRgn %>%
  speedyseq::tax_glom("Best_Hit_ARO") %>% 
  microbiome::transform(transform = "log10p") %>% 
  speedyseq::psmelt() %>% 
  # left_join(tax_mapping,
  #           by = c("Best_Hit_ARO" = "Best_Hit_ARO"),
  #           suffix = c("_x", "")) %>% 
  # mutate(Sample = fct_reorder(Sample, Model,Treatment,Fermentation,Antibiotic_mg.mL, Day_of_Treatment )) %>%
  mutate(Abundance = na_if(Abundance, 0)) %>%
  mutate(logTPM = log10(Abundance + 1)) -> pheatmap_df


pheatmap_df %>% 
  # replace(is.na(.), 0)  %>% 
  # select(Sample, Best_Hit_ARO, logTPM) %>% 
  group_by(Sample, Best_Hit_ARO) %>%
  pivot_wider(names_from =  Sample, Best_Hit_ARO,
              values_from = logTPM) %>% 
  column_to_rownames("Best_Hit_ARO") -> pheatmap_samples
  
pheatmap_samples %>% 
  replace(is.na(.), 0)  %>% 
  t() %>% 
  vegdist(na.rm = TRUE) %>% 
  hclust() -> clust_t

pheatmap_samples %>% 
    replace(is.na(.), 0)  %>%
    vegdist(na.rm = TRUE) %>% 
  # cor() %>% 
  hclust() -> clust

 pheatmap_samples %>% 
  pheatmap::pheatmap(cluster_rows = clust ,
                     cluster_cols = clust_t,
                     na_col = "transparent",
                     annotation_row = tax_mapping %>% rownames_to_column('tmp_col') %>% 
                       filter(Best_Hit_ARO %in% taxa_names(physeq_AMRgn)) %>% 
                       distinct(Best_Hit_ARO, .keep_all = TRUE) %>% 
                       column_to_rownames("Best_Hit_ARO") %>% 
                       select(Drug_Class_multi, Resistance_Mechanism_multi),
                     annotation_colors = list(Drug_Class_multi = col,
                                              Treatment = treat_col),
                     annotation_col = physeq_AMRgn %>% 
                       sample_data() %>% 
                       data.frame() %>% 
                       select(Model,Treatment,Fermentation,Antibiotic_mg.mL, Day_of_Treatment)) -> p_pheat
 
 
 p_pheat %>% 
   print()
 
 
 p_pheat %>%
  export::graph2ppt(append = TRUE, width = 30, height = 30,
                    file = out_pptx)
```

chick
```{r, fig.height=30, fig.width=30}
physeq_AMRgn %>%
  speedyseq::tax_glom("Best_Hit_ARO") %>% 
  subset_samples(Model == "Chicken") %>% 
  filter_taxa(function(x) sum(x > 0) > 0, TRUE) -> ps_tmp

ps_tmp %>% 
  microbiome::transform(transform = "log10p") %>% 
  speedyseq::psmelt() %>% 
  # left_join(tax_mapping,
  #           by = c("Best_Hit_ARO" = "Best_Hit_ARO"),
  #           suffix = c("_x", "")) %>% 
  # mutate(Sample = fct_reorder(Sample, Model,Treatment,Fermentation,Antibiotic_mg.mL, Day_of_Treatment )) %>%
  mutate(Abundance = na_if(Abundance, 0)) %>%
  mutate(logTPM = log10(Abundance + 1)) -> pheatmap_df


pheatmap_df %>% 
  # replace(is.na(.), 0)  %>% 
  # select(Sample, Best_Hit_ARO, logTPM) %>% 
  group_by(Sample, Best_Hit_ARO) %>%
  pivot_wider(names_from =  Sample, Best_Hit_ARO,
              values_from = logTPM) %>% 
  column_to_rownames("Best_Hit_ARO") -> pheatmap_samples
  
pheatmap_samples %>% 
  replace(is.na(.), 0)  %>% 
  t() %>% 
  vegdist(na.rm = TRUE) %>% 
  hclust() -> clust_t

# pheatmap_samples %>% 
#     replace(is.na(.), 0)  %>%
#     vegdist(na.rm = TRUE) %>% 
#   # cor() %>% 
#   hclust() -> clust

 pheatmap_samples %>% 
  pheatmap::pheatmap(cluster_rows = FALSE ,
                     cluster_cols = clust_t,
                     na_col = "transparent",
                     annotation_row = tax_mapping %>% rownames_to_column('tmp_col') %>% 
                       filter(Best_Hit_ARO %in% taxa_names(physeq_AMRgn)) %>% 
                       distinct(Best_Hit_ARO, .keep_all = TRUE) %>% 
                       column_to_rownames("Best_Hit_ARO") %>% 
                       select(Drug_Class_multi, Resistance_Mechanism_multi),
                     annotation_colors = list(Drug_Class_multi = col)) -> p_pheat #,
                     # annotation_col = ps_tmp %>% 
                     #   sample_data() %>% 
                     #   data.frame() %>% 
                     #   select(Model,Treatment,Fermentation,Antibiotic_mg.mL, Day_of_Treatment)) -> p_pheat
 
 
 p_pheat %>% 
   print()
 
  p_pheat %>%
  export::graph2ppt(append = TRUE, width = 20, height = 20,
                    file = out_pptx)
```

Human
```{r, fig.height=30, fig.width=30}
physeq_AMRgn %>%
  speedyseq::tax_glom("Best_Hit_ARO") %>% 
  subset_samples(Model == "Human") %>% 
  filter_taxa(function(x) sum(x > 0) > 0, TRUE) -> ps_tmp

ps_tmp %>% 
  microbiome::transform(transform = "log10p") %>% 
  speedyseq::psmelt() %>% 
  # left_join(tax_mapping,
  #           by = c("Best_Hit_ARO" = "Best_Hit_ARO"),
  #           suffix = c("_x", "")) %>% 
  # mutate(Sample = fct_reorder(Sample, Model,Treatment,Fermentation,Antibiotic_mg.mL, Day_of_Treatment )) %>%
  mutate(Abundance = na_if(Abundance, 0)) %>%
  mutate(logTPM = log10(Abundance + 1)) -> pheatmap_df


pheatmap_df %>% 
  # replace(is.na(.), 0)  %>% 
  # select(Sample, Best_Hit_ARO, logTPM) %>% 
  group_by(Sample, Best_Hit_ARO) %>%
  pivot_wider(names_from =  Sample, Best_Hit_ARO,
              values_from = logTPM) %>% 
  column_to_rownames("Best_Hit_ARO") -> pheatmap_samples
  
pheatmap_samples %>% 
  replace(is.na(.), 0)  %>% 
  t() %>% 
  vegdist(na.rm = TRUE) %>% 
  hclust() -> clust_t

pheatmap_samples %>%
    replace(is.na(.), 0)  %>%
    vegdist(na.rm = TRUE) %>%
  # cor() %>%
  hclust() -> clust

 pheatmap_samples %>% 
  pheatmap::pheatmap(cluster_rows = clust ,
                     cluster_cols = clust_t,
                     na_col = "transparent",
                     annotation_row = tax_mapping %>% rownames_to_column('tmp_col') %>% 
                       filter(Best_Hit_ARO %in% taxa_names(physeq_AMRgn)) %>% 
                       distinct(Best_Hit_ARO, .keep_all = TRUE) %>% 
                       column_to_rownames("Best_Hit_ARO") %>% 
                       select(Drug_Class_multi, Resistance_Mechanism_multi),
                     annotation_colors = list(Drug_Class_multi = col,
                                              Treatment = treat_col),
                     annotation_col = ps_tmp %>%
                       sample_data() %>%
                       data.frame() %>%
                       select(Model,Treatment,Fermentation,Antibiotic_mg.mL, Day_of_Treatment)) -> p_pheat
 
 
 p_pheat %>% 
   print()
 
   p_pheat %>%
  export::graph2ppt(append = TRUE, width = 20, height = 20,
                    file = out_pptx)
```


```{r, fig.height=16, fig.width=18}
# see Sneha's & Hannah's code for "plot.df.sep <- separate_rows(plot.df, Drug.Class,sep = '; ',convert = TRUE)"

physeq_AMRgn %>%
  microbiome::transform(transform = "log10p") %>% 
  speedyseq::psmelt() %>% 
    left_join(tax_mapping,
            by = c("Best_Hit_ARO" = "Best_Hit_ARO"),
            suffix = c("_x", "")) %>%
    mutate(Abundance = na_if(Abundance, 0)) %>%
  mutate(logTPM = log10(Abundance + 1))  %>% 
  separate_rows(., Resistance_Mechanism,sep = '; ',convert = TRUE) %>% 
  ggplot(aes(x = as.factor(Day_of_Treatment), y = Abundance, fill = Resistance_Mechanism)) +   
  geom_bar( stat = "identity", colour="black", size=0.025) +
  facet_grid(Drug_Class_multi  ~ Model +  Treatment + Fermentation + Antibiotic_mg.mL , scales = "free", space = "free_x") +
  ggtitle("ARG Type Abundance per Sample") +
  xlab("Day (Treatment)")  +
  ylab("log10 TPM") + 
  theme_light() -> p_mec

p_mec %>% 
  ggpubr::set_palette("jco") -> p_mec

p_mec

p_mec %>%
  export::graph2ppt(append = TRUE, width = 20, height = 18,
                    file = out_pptx)
```

```{r, fig.height=6, fig.width=14}
# see Sneha's & Hannah's code for "plot.df.sep <- separate_rows(plot.df, Drug.Class,sep = '; ',convert = TRUE)"

physeq_AMRgn %>%
  microbiome::transform(transform = "log10p") %>% 
  speedyseq::psmelt() %>% 
    left_join(tax_mapping,
            by = c("Best_Hit_ARO" = "Best_Hit_ARO"),
            suffix = c("_x", "")) %>%
    mutate(Abundance = na_if(Abundance, 0)) %>%
  mutate(logTPM = log10(Abundance + 1))  %>% 
  separate_rows(., Resistance_Mechanism,sep = '; ',convert = TRUE) %>% 
  ggplot(aes(x = as.factor(Day_of_Treatment), y = Abundance, fill = Resistance_Mechanism_multi)) +   
  geom_bar( stat = "identity", colour="black", size=0.025) +
  facet_grid(.  ~ Model +  Treatment + Fermentation + Antibiotic_mg.mL , scales = "free", space = "free_x") +
  ggtitle("ARG Type Abundance per Sample") +
  xlab("Day (Treatment)")  +
  ylab("log10 TPM") + 
  theme_light() -> p_mec

p_mec %>% 
  ggpubr::set_palette("jco") -> p_mec

p_mec

p_mec %>%
  export::graph2ppt(append = TRUE, width = 18, height = 6,
                    file = out_pptx)
```


```{r, fig.height=6, fig.width=14}
# see Sneha's & Hannah's code for "plot.df.sep <- separate_rows(plot.df, Drug.Class,sep = '; ',convert = TRUE)"

physeq_AMRgn %>%
  microbiome::transform(transform = "log10p") %>% 
  speedyseq::psmelt() %>% 
    left_join(tax_mapping,
            by = c("Best_Hit_ARO" = "Best_Hit_ARO"),
            suffix = c("_x", "")) %>%
    mutate(Abundance = na_if(Abundance, 0)) %>%
  mutate(logTPM = log10(Abundance + 1))  %>% 
  # separate_rows(., Resistance_Mechanism,sep = '; ',convert = TRUE) %>% 
  ggplot(aes(x = as.factor(Day_of_Treatment), y = Abundance, fill = Resistance_Mechanism_multi)) +   
  geom_bar( stat = "identity", colour="black", size=0.025) +
  facet_grid(.  ~ Model +  Treatment + Fermentation + Antibiotic_mg.mL , scales = "free", space = "free_x") +
  ggtitle("ARG Type Abundance per Sample") +
  xlab("Day (Treatment)")  +
  ylab("log10 TPM") + 
  theme_light() -> p_mec

p_mec %>% 
  ggpubr::set_palette("jco") -> p_mec

p_mec

p_mec %>%
  export::graph2ppt(append = TRUE, width = 18, height = 6,
                    file = out_pptx)
```


```{r, fig.height=6, fig.width=14}
# see Sneha's & Hannah's code for "plot.df.sep <- separate_rows(plot.df, Drug.Class,sep = '; ',convert = TRUE)"

physeq_AMRgn %>%
  microbiome::transform(transform = "log10p") %>% 
  speedyseq::psmelt() %>% 
    left_join(tax_mapping,
            by = c("Best_Hit_ARO" = "Best_Hit_ARO"),
            suffix = c("_x", "")) %>%
    mutate(Abundance = na_if(Abundance, 0)) %>%
  mutate(logTPM = log10(Abundance + 1))  %>% 
  separate_rows(., Resistance_Mechanism,sep = '; ',convert = TRUE) %>%
  ggplot(aes(x = as.factor(Day_of_Treatment), y = Abundance, fill = Resistance_Mechanism)) +   
  geom_bar( stat = "identity", position = "fill", colour="black", size=0.025) +
  facet_grid(.  ~ Model +  Treatment + Fermentation + Antibiotic_mg.mL , scales = "free", space = "free_x") +
  ggtitle("ARG Type Abundance per Sample") +
  xlab("Day (Treatment)")  +
  ylab("Proportion") + 
  theme_light() -> p_mec

p_mec %>% 
  ggpubr::set_palette("jco") -> p_mec

p_mec

p_mec %>%
  export::graph2ppt(append = TRUE, width = 18, height = 6,
                    file = out_pptx)
```

```{r, fig.height=6, fig.width=14}
# https://newbedev.com/how-to-control-ordering-of-stacked-bar-chart-using-identity-on-ggplot2
physeq_AMRgn %>%
  microbiome::transform(transform = "log10p") %>% 
  speedyseq::psmelt() %>% 
    left_join(tax_mapping,
            by = c("Best_Hit_ARO" = "Best_Hit_ARO"),
            suffix = c("_x", "")) %>%
    mutate(Abundance = na_if(Abundance, 0)) %>%
  mutate(logTPM = log10(Abundance + 1))  %>% 
  # separate_rows(., Resistance_Mechanism,sep = '; ',convert = TRUE) %>% 
  ggplot(aes(x = as.factor(Day_of_Treatment), y = Abundance, fill = Resistance_Mechanism_multi)) +   
  geom_bar( stat = "identity", position = "fill", colour="black", size=0.00125) +
  facet_grid(.  ~ Model +  Treatment + Fermentation + Antibiotic_mg.mL , scales = "free", space = "free_x") +
  ggtitle("ARG Type Abundance per Sample") +
  xlab("Day (Treatment)")  +
  ylab("Proportion") + 
  theme_light() -> p_mec

p_mec %>% 
  ggpubr::set_palette("jco") -> p_mec

p_mec

p_mec %>%
  export::graph2ppt(append = TRUE, width = 18, height = 6,
                    file = out_pptx)
```


```{r, fig.height=6, fig.width=14}
physeq_drug_multi %>%
  microbiome::transform(transform = "log10p") %>% 
  speedyseq::psmelt() %>% 
  ggplot(aes(x = as.factor(Day_of_Treatment), y = Abundance, fill = Drug_Class_multi)) +   
  geom_bar( stat = "identity", colour="black", size=0.025) +
  scale_fill_manual(values=col) +
  #geom_bar(stat="identity",colour=NA,size=0) +
  facet_grid( ~ Model +  Treatment + Fermentation + Antibiotic_mg.mL , scales = "free", space = "free_x") +
  ggtitle("ARG Type Abundance per Sample") +
  xlab("Day (Treatment)")  +
  ylab("log10 TPM") + 
  theme_light() -> p_drug

p_drug

p_drug %>%
  export::graph2ppt(append = TRUE, width = 18, height = 6,
                    file = out_pptx)
```

```{r, fig.height=6, fig.width=14}

physeq_drug_multi %>%
  microbiome::transform(transform = "log10p") %>% 
  speedyseq::psmelt() %>% 
  ggplot(aes(x = as.factor(Day_of_Treatment), y = Abundance, fill = Drug_Class_multi)) +   
  geom_bar( stat = "identity", position="fill", colour="black", size=0.025) +
  scale_fill_manual(values=col) +
  #geom_bar(stat="identity",colour=NA,size=0) +
  facet_grid(. ~ Model + Treatment + Fermentation + Antibiotic_mg.mL , scales = "free", space = "free_x", drop = TRUE) +
  ggtitle("ARG Type Abundance per Sample") +
  xlab("Day (Treatment)")  +
  ylab("Proportion") +
  theme_light() -> p_drug_prop

p_drug_prop

p_drug_prop %>%
  export::graph2ppt(append = TRUE, width = 18, height = 6,
                    file = out_pptx)

```


```{r, include= FALSE, eval= FALSE}
# tmp <- physeq_drug
# tax_table(tmp) <- tax_table(tmp)[,c("Best_Hit_ARO")]
# 
# 
# tmp %>%
#   tax_fix() %>%
#   phyloseq_validate() %>% 
#   microbiome::transform(transform = "log10p") %>% 
#   comp_barplot(
#     tax_level = "Best_Hit_ARO", n_taxa = 20, other_name = "Other",
#     # taxon_renamer = function(x) stringr::str_remove(x, " [ae]t rel."),
#     palette = distinct_palette(n = 20, add = "grey90"),
#     merge_other = FALSE, bar_outline_colour = "darkgrey"
#   ) +
#   # coord_flip() +
#   facet_wrap(  ~  Treatment + Fermentation + Antibiotic_mg.mL) +
#   labs(x = NULL, y = NULL) +
#   theme(axis.text.y = element_blank(), axis.ticks.y = element_blank())
```

```{r, include= TRUE, eval= TRUE}
physeq_AMRgn_round <- physeq_AMRgn
round(otu_table(physeq_AMRgn_round)) -> otu_table(physeq_AMRgn_round)


physeq_AMRgn_round %>% 
  phyloseq_alphas(phylo = FALSE) -> alpha_df

```


```{r, fig.width= 18, fig.height= 6}
# plot panels for each treatment
alpha_df %>% 
    pivot_longer(cols = all_of(c("Observed", "diversity_shannon")), values_to = "value", names_to = 'alphadiversiy', values_drop_na  = TRUE) %>%
    mutate(alphadiversiy = fct_relevel(alphadiversiy, c("Observed", "diversity_shannon"))) %>% 
  # filter(!is.na(Fermentation)) %>% 
  # filter(Treatment != "DONOR") %>%
  ggplot(aes(x = Day_of_Treatment, y = value, color = Treatment)) +
  geom_point() + 
  geom_line(linetype = "dashed",  size = 0.25) +
  labs(title = "AMR Gene Richness",
       y = "Richness", x = "Days of Treatment") +
  scale_color_manual(values = treat_col,
                     na.value = "black") +
  facet_grid(alphadiversiy  ~  Model + Treatment + Fermentation + Antibiotic_mg.mL , scales = "free", space = "fixed", drop = TRUE) +
  theme_light() -> p_alpha

p_alpha

p_alpha %>%
  export::graph2ppt(append = TRUE, width = 20, height = 6,
                    file = out_pptx)
```

Beta Diversity:

```{r, fig.width= 8, fig.height= 6}

# sample_data(physeq)$Reactor_Treatment <- fct_relevel(sample_data(physeq)$Reactor_Treatment, "DONOR", "CR_UNTREATED", "TR1_CTX+HV292.1",
#                                                      "TR2_CTX","TR3_HV292.1", "TR4_VAN", "TR5_VAN+CCUG59168", "TR6_CCUG59168") 
# 
# sample_data(physeq)$Treatment <- fct_relevel(sample_data(physeq)$Treatment, "DONOR", "UNTREATED",  "CTX+HV292.1", "CTX","HV292.1","VAN+CCUG59168", "VAN",  "CCUG59168") 


# physeq %>% 
#   rarefy_even_depth(sample.size = 43,
#                     rngseed = 123) -> phyloseq_rare

physeq_AMRgn %>%
  phyloseq_compute_bdiv(norm = "pc",
                        phylo = FALSE,
                        seed = 123) -> bdiv_list

physeq_AMRgn  %>%
  # subset_samples(Treatment != "DONOR") %>%
  phyloseq_plot_bdiv(dlist = bdiv_list,
                     seed = 123,
                     axis1 = 1,
                     axis2 = 2)  -> pcoa

# phyloseq_plot_bdiv(bdiv_list,
#                    # m = "CoDa",
#                    seed = 123) -> coda
# 
pcoa$wjaccard$layers = NULL

pcoa$wjaccard + geom_point(size=3,
                           aes(color = Treatment, 
                               fill = NULL,
                               shape = Fermentation %>%  as.factor(),
                               alpha = Day_of_Treatment)) + 
  theme_light() +
  geom_path(arrow = arrow(type = "open", angle = 30, length = unit(0.15, "inches")),
            size = 0.05, linetype = "dashed", inherit.aes = TRUE, aes(group=Reactor_Treatment), show.legend = FALSE) +
  scale_alpha_continuous(range=c( 0.9, 0.3)) +
  scale_fill_manual(values = treat_col,
                     na.value = "black") +
  scale_color_manual(values = treat_col,
                     na.value = "black") +
  scale_shape_manual(values = c(15, 19), na.value =  17) + 
  theme_classic() -> p1 # +
  # labs(col=NULL, fill = NULL, shape = NULL) + guides(shape=TRUE) -> p1

p1 + facet_grid(Model ~.)

p1 %>%
  export::graph2ppt(append = TRUE, width = 8, height = 6,
                    file = out_pptx)
```

```{r, fig.width= 8, fig.height= 6}

physeq_AMRgn %>% 
  tax_table() %>% 
  data.frame() %>% 
  rownames_to_column("id") %>% 
  mutate(gene = id) %>% 
  column_to_rownames("id") %>% 
  as.matrix %>% 
  tax_table() -> tax_table(physeq_AMRgn)



physeq_AMRgn  %>%
  # subset_samples(Treatment != "DONOR") %>% 
  phyloseq_add_taxa_vector(dist = bdiv_list$wjaccard,
                           phyloseq = .,
                           figure_ord = p1,
                           tax_rank_plot = "Best_Hit_ARO", taxrank_glom = "Best_Hit_ARO",
                           top_r = 10, fact = 0.6) -> pco_env

# pco_env$plot %>% 
#   export::graph2ppt(append = TRUE, width = 8, height = 6,
#                     file = out_pptx)

pco_env$signenvfit %>% 
  DT::datatable()
```

```{r, fig.width= 8, fig.height= 6}
physeq_drug_multi  %>%
  # subset_samples(Treatment != "DONOR") %>% 
  phyloseq_add_taxa_vector(dist = bdiv_list$wjaccard,
                           phyloseq = .,
                           figure_ord = p1,
                           tax_rank_plot = "Drug_Class_multi", taxrank_glom = "Drug_Class_multi",
                           top_r = 10, fact = 0.6) -> pco_env


pco_env$plot + facet_grid(Model ~ .) -> pco_env$plot

pco_env$plot

pco_env$plot  %>%
  export::graph2ppt(append = TRUE, width = 8, height = 6,
                    file = out_pptx)

pco_env$signenvfit %>% 
  DT::datatable()
```

Make the heatmap pretty 
```{r, fig.width= 20, fig.height= 4}
physeq %>% 
  # speedyseq::tax_glom("Best_Hit_ARO") %>% 
  speedyseq::psmelt() %>% 
  rename(TPM = Abundance) %>% 
  mutate(TPM = na_if(TPM, 0)) %>%
  # mutate(logTPM = log10(TPM + 1)) %>% 
  # filter(Drug_Class %in% c("cephalosporin", "glycopeptide")) %>% 
  # mutate(Drug_Class = fct_relevel(Drug_Class, c("cephalosporin", "glycopeptide"))) %>% 
  filter(Best_Hit_ARO %in% c("vanZA", "vanYA", "vanXA", "vanA", "vanHA", "vanRA", "CTX-M-1")) %>% 
  # mutate(Classification = fct_relevel(Classification, c("MAG", "Chromosome", "Plasmid", "Uncertain - plasmid or chromosomal"))) %>% 
  ggplot(mapping = aes(x = Day_of_Treatment ,
                       y = Best_Hit_ARO ,
                       fill = TPM,
  )) +
  facet_grid(Drug_Class ~  Model + Treatment + Fermentation + Antibiotic_mg.mL, scales = "free",  space = "fixed", drop = TRUE) +
  # facet_grid(. ~ Treatment,  scales = "free", space = "free", drop = TRUE) +
  # facet_grid(. ~ Reactor_Treatment, scales = "free", space = "free", drop = TRUE,labeller = labeller(Drug_Class = label_wrap_gen(width = 15))) +
  #we adjust the vertical labels such that it is printed correctly. 
  geom_tile() +
  theme_bw()  +
  # scale_fill_viridis_c(name = "TPM",
  #                      na.value = "transparent", trans = scales::pseudo_log_trans(sigma = 0.1),
  #                      # trans =  scales::pseudo_log_trans(sigma = 1),
  #                      breaks = c(1, 10, 50, 100, 400), labels = c(1, 10, 50, 100, 400),
  #                      limits = c(0, 500)) + #need to make sure that we include max value of tpm as upper limit
  scale_fill_viridis_c(na.value = "transparent") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5),
        strip.text.x = element_text(size=8),
        #change the size of the label and the angle so the text is printed horizontally
        strip.text.y = element_text(size = 7,angle=0)) -> perma_bbuble

perma_bbuble +
  theme(legend.position = "bottom") -> perma_bbuble

perma_bbuble


perma_bbuble %>%
  export::graph2ppt(append = TRUE, width = 16, height = 4,
                    file = out_pptx)
```

What's next? 
- MAG, plasmid, Chromosome ... proportions.
- SQM contig taxonomy coloring/facetting.


```{r}

source("https://raw.githubusercontent.com/fconstancias/DivComAnalyses/master/R/phyloseq_normalisation.R")


ps_tmp_tax <- physeq


ps_tmp_tax %>% 
  tax_table() %>% 
  data.frame() %>% 
  # select(Tax) %>% 
  separate(Tax, into = c("Kingdom", "Phylum", "Class", "Order", "Family", "Genus", "Species"), sep = ";") %>% 
  mutate(across(c("Kingdom", "Phylum", "Class", "Order", "Family", "Genus", "Species"), gsub, pattern = "[a-z]_", replacement = "")) %>% 
  as.matrix() -> tax_table(ps_tmp_tax)

ps_tmp_tax_2 <- ps_tmp_tax

ps_tmp_tax %>% 
   tax_table() %>% 
  data.frame() %>% 
  select(c("Kingdom", "Phylum", "Class", "Order", "Family", "Genus", "Species"))  %>% 
    as.matrix() -> tax_table(ps_tmp_tax_2)

ps_tmp_tax_2 %>%   
  phyloseq_get_strains_fast(species = FALSE) %>% 
  tax_table() %>% 
  data.frame()  %>% 
    mutate(Strain =  gsub(" human_megahit.*","",Strain)) %>% 
    mutate(Strain =  gsub(" chick_megahit.*","",Strain)) -> full_tax

ps_tmp_tax %>% 
  tax_table() %>% 
  data.frame() %>% 
  rownames_to_column("ASV") %>% 
  # select(-Tax) %>% 
  left_join(full_tax %>% 
            select(Strain) %>% rownames_to_column("ASV"),
             by = c("ASV" = "ASV")) %>% 
  select(-Predicted_DNA) %>% 
  column_to_rownames("ASV") %>% 
  as.matrix() -> tax_table(ps_tmp_tax)

  
  
  
tax_table(ps_tmp_tax) %>% 
    data.frame() %>% 
  select(c("Best_Hit_ARO", "Strain")) %>% 
  head()

ps_tmp_tax %>% 
  saveRDS("~/Projects/ETH/Alessia/mobilome/ps_combined_tax_clean.rds")


as(tax_table(ps_tmp_tax), "matrix") %>% 
  as.data.frame() %>%
  rownames_to_column('ASV') -> tax_table_ps_combined

as(otu_table(ps_tmp_tax), "matrix") %>% 
  as.data.frame() %>%
  rownames_to_column('ASV') -> otu_table_ps_combined

ps_tmp_tax %>% 
  sample_data() %>% 
  data.frame() %>% 
  rownames_to_column('sample_id') ->  meta_ps_combined

otu_table_ps_combined %>%
  # left_join(refseq_df, by = 'ASV') %>%
  left_join(tax_table_ps_combined, by = c("ASV" = "ASV")) %>%
  dplyr::select(ASV, everything()) -> merged_ps_combined


merged_ps_combined %>% 
  write_tsv("~/Projects/ETH/Alessia/mobilome/ps_combined_tax_clean_rds.tsv")


```

```{r}
tax_table(ps_tmp_tax) %>% 
    data.frame() %>% 
  select(c("Best_Hit_ARO", "Strain")) %>% 
  group_by(Strain) %>% 
  add_count() %>% 
  arrange(-n)
```

Define colors for Taxonomic information:

```{r}
ps_tmp_tax %>%  speedyseq::psmelt() %>%  pull(Strain) %>%  unique() -> strains
set.seed(456)
col_strains <- distinctColorPalette(length(strains),
                               runTsne = FALSE)

names(col_strains) <- strains
```

```{r, fig.height=16, fig.width=18}
# see Sneha's & Hannah's code for "plot.df.sep <- separate_rows(plot.df, Drug.Class,sep = '; ',convert = TRUE)"

ps_tmp_tax %>%
  microbiome::transform(transform = "log10p") %>% 
  speedyseq::psmelt() %>% 
    # left_join(tax_mapping,
    #         by = c("Best_Hit_ARO" = "Best_Hit_ARO"),
    #         suffix = c("_x", "")) %>%
    mutate(Abundance = na_if(Abundance, 0)) %>%
  mutate(logTPM = log10(Abundance + 1))  %>% 
  # separate_rows(., Resistance_Mechanism,sep = '; ',convert = TRUE) %>% 
  ggplot(aes(x = as.factor(Day_of_Treatment), y = Abundance, fill = Strain)) +  
  geom_bar( stat = "identity", colour="black", size=0.025) +
  facet_grid(Drug_Class_multi  ~ Model +  Treatment + Fermentation + Antibiotic_mg.mL , scales = "free", space = "free_x") +
  ggtitle("ARG Type Abundance per Sample") +
  xlab("Day (Treatment)")  +
  ylab("log10 TPM") + 
  theme_light() +
  scale_fill_manual(name = "", values = col_strains) +
  theme(legend.position = "bottom") -> p_mec

# p_mec %>% 
#   ggpubr::set_palette("jco") -> p_mec

p_mec

p_mec %>%
  export::graph2ppt(append = TRUE, width = 20, height = 18,
                    file = out_pptx)
```

```{r, fig.height=16, fig.width=18}

ps_tmp_tax %>%
  microbiome::transform(transform = "log10p") %>% 
  speedyseq::psmelt() %>% 
    # left_join(tax_mapping,
    #         by = c("Best_Hit_ARO" = "Best_Hit_ARO"),
    #         suffix = c("_x", "")) %>%
    mutate(Abundance = na_if(Abundance, 0)) %>%
  mutate(logTPM = log10(Abundance + 1))  %>% 
  # separate_rows(., Resistance_Mechanism,sep = '; ',convert = TRUE) %>% 
  ggplot(aes(x = as.factor(Day_of_Treatment), y = Abundance, fill = Drug_Class_multi)) +  
  geom_bar( stat = "identity", colour="black", size=0.025) +
  facet_grid( Strain   ~ Model +  Treatment + Fermentation + Antibiotic_mg.mL , scales = "free", space = "free_x") +
  ggtitle("ARG Type Abundance per Sample") +
  xlab("Day (Treatment)")  +
  ylab("log10 TPM") + 
  theme_light() +
  scale_fill_manual(name = "", values = col) +
  theme(legend.position = "bottom") -> p_mec

# p_mec %>% 
#   ggpubr::set_palette("jco") -> p_mec

p_mec

p_mec %>%
  export::graph2ppt(append = TRUE, width = 20, height = 18,
                    file = out_pptx)
```

```{r, fig.height=16, fig.width=18}
# see Sneha's & Hannah's code for "plot.df.sep <- separate_rows(plot.df, Drug.Class,sep = '; ',convert = TRUE)"

ps_tmp_tax %>%
  microbiome::transform(transform = "log10p") %>% 
  speedyseq::psmelt() %>% 
    # left_join(tax_mapping,
    #         by = c("Best_Hit_ARO" = "Best_Hit_ARO"),
    #         suffix = c("_x", "")) %>%
    mutate(Abundance = na_if(Abundance, 0)) %>%
  mutate(logTPM = log10(Abundance + 1))  %>% 
  # separate_rows(., Resistance_Mechanism,sep = '; ',convert = TRUE) %>% 
  ggplot(aes(x = as.factor(Day_of_Treatment), y = Abundance, fill = Strain)) +  
  geom_bar( stat = "identity", colour="black", position = "fill", size=0.05) +
  facet_grid(Drug_Class_multi  ~ Model +  Treatment + Fermentation + Antibiotic_mg.mL , scales = "free", space = "free_x") +
  ggtitle("ARG Type Abundance per Sample") +
  xlab("Day (Treatment)")  +
  ylab("Proportion") + 
  theme_light() +
  scale_fill_manual(name = "", values = col_strains) +
  theme(legend.position = "bottom") -> p_mec

# p_mec %>% 
#   ggpubr::set_palette("jco") -> p_mec

p_mec

p_mec %>%
  export::graph2ppt(append = TRUE, width = 20, height = 18,
                    file = out_pptx)
```

```{r, fig.height=6, fig.width=14}
# see Sneha's & Hannah's code for "plot.df.sep <- separate_rows(plot.df, Drug.Class,sep = '; ',convert = TRUE)"

ps_tmp_tax %>%
  microbiome::transform(transform = "log10p") %>% 
  speedyseq::psmelt() %>% 
    # left_join(tax_mapping,
    #         by = c("Best_Hit_ARO" = "Best_Hit_ARO"),
    #         suffix = c("_x", "")) %>%
    mutate(Abundance = na_if(Abundance, 0)) %>%
  mutate(logTPM = log10(Abundance + 1))  %>% 
  # separate_rows(., Resistance_Mechanism,sep = '; ',convert = TRUE) %>% 
  ggplot(aes(x = as.factor(Day_of_Treatment), y = Abundance, fill = Strain)) +  
  geom_bar( stat = "identity", colour="black", size=0.025) +
  facet_grid(.  ~ Model +  Treatment + Fermentation + Antibiotic_mg.mL , scales = "free", space = "free_x") +
  ggtitle("ARG Type Abundance per Sample") +
  xlab("Day (Treatment)")  +
  ylab("log10 TPM") + 
  theme_light() +
  scale_fill_manual(name = "", values = col_strains) +
  theme(legend.position = "bottom") -> p_mec

# p_mec %>% 
#   ggpubr::set_palette("jco") -> p_mec

p_mec

p_mec %>%
  export::graph2ppt(append = TRUE, width = 18, height = 6,
                    file = out_pptx)
```

```{r, fig.height=6, fig.width=14}
# see Sneha's & Hannah's code for "plot.df.sep <- separate_rows(plot.df, Drug.Class,sep = '; ',convert = TRUE)"

ps_tmp_tax %>%
  microbiome::transform(transform = "log10p") %>% 
  speedyseq::psmelt() %>% 
    # left_join(tax_mapping,
    #         by = c("Best_Hit_ARO" = "Best_Hit_ARO"),
    #         suffix = c("_x", "")) %>%
    mutate(Abundance = na_if(Abundance, 0)) %>%
  mutate(logTPM = log10(Abundance + 1))  %>% 
  # separate_rows(., Resistance_Mechanism,sep = '; ',convert = TRUE) %>% 
  # ggplot(aes(x = as.factor(Day_of_Treatment), y = Abundance, fill = fct_reorder(Strain, logTPM))) +   
  ggplot(aes(x = as.factor(Day_of_Treatment), y = Abundance, fill = reorder(Strain, logTPM, FUN=median, na.rm = TRUE))) +
  geom_bar( stat = "identity", position="fill", colour="black", size=0.025) +
  facet_grid(.  ~ Model +  Treatment + Fermentation + Antibiotic_mg.mL , scales = "free", space = "free_x") +
  ggtitle("ARG Type Abundance per Sample") +
  xlab("Day (Treatment)")  +
  ylab("Proportion") + 
  theme_light() +
  scale_fill_manual(name = "", values = col_strains) +
  theme(legend.position = "bottom") -> p_mec

# p_mec %>% 
#   ggpubr::set_palette("jco") -> p_mec

p_mec

p_mec %>%
  export::graph2ppt(append = TRUE, width = 18, height = 6,
                    file = out_pptx)
```

```{r}
sessionInfo()
```

