---
title: "Comparing conditions"
output: html_document
---

```{r setup, include=FALSE}
rm(list = ls()) # remove all the object before starting
knitr::opts_chunk$set(echo = TRUE)
```


```{r message=FALSE}
library(tidyverse)
library(phyloseq)
library(speedyseq)
library(ggrepel)
library(ampvis2)
library(plotly)
options(getClass.msg=FALSE)

source("https://raw.githubusercontent.com/fconstancias/DivComAnalyses/master/R/phyloseq_normalisation.R") 
 
source("https://raw.githubusercontent.com/fconstancias/DivComAnalyses/master/R/phyloseq_heatmap.R") 
```


Loading phyloseq object
```{r}
#directory where the phyloseq object is 

ps = "/Users/snehasundar/Desktop/amchick/metabarcoding analysis_9_10_2020/physeq_update.RDS"


#get strain level information
ps %>% 
  readRDS() %>% 
  phyloseq_get_strains() -> physeq

```

### CTX Treated

Transform sample counts (important to do it before filtering out ASVs otherwise the transformation will not be using all the asvs and hence will be wrong), then subseting only the samples relevant to us (Control and ctx treated batch samples) and then filtering out the ASVs not present in the samples we subsetted. 
```{r}
physeq  %>% 
  transform_sample_counts(function(x) x/sum(x) * 100) %>% 
  subset_samples(Experiment =="Batch" & Treatment %in%  c('UNTREATED','CTX','CTX+HV292.1'))  %>% 
  filter_taxa(function(x) sum(x > 0) > 0, TRUE) -> physeq_batch_ecoli
```

Get taxa common to the treatment conditions (CTX and CTX + strain)
```{r}
#taxa present in CTX treated samples
taxa_CTX <- physeq_batch_ecoli %>% subset_samples(Treatment=='CTX') %>% filter_taxa(function(x) sum(x > 0) > 0, TRUE) %>% taxa_names()

#taxa present in CTX + strain treated samples
taxa_CTX_strain <- physeq_batch_ecoli %>% subset_samples(Treatment=='CTX+HV292.1') %>% filter_taxa(function(x) sum(x > 0) > 0, TRUE) %>% taxa_names()

#common taxa
taxa_common_ecoli<-intersect(taxa_CTX, taxa_CTX_strain)


#otu table should only have the ASVs common to the two conditions
otu_table(physeq_batch_ecoli)<-otu_table(physeq_batch_ecoli)[taxa_common_ecoli,] 



```


Plotting the heatmap . 
```{r,fig.dim=c(10,8)}
#reset the factors in the order we want 
sample_data(physeq_batch_ecoli)$Enrichment <- fct_relevel(sample_data(physeq_batch_ecoli)$Enrichment,'NotEnriched','Enriched')

sample_data(physeq_batch_ecoli)$Treatment <- fct_relevel(sample_data(physeq_batch_ecoli)$Treatment,'UNTREATED','CTX','CTX+HV292.1')

sample_data(physeq_batch_ecoli)$Batch_hour <- fct_relevel(sample_data(physeq_batch_ecoli)$Batch_hour,'0','7','24')

physeq_batch_ecoli %>% phyloseq_ampvis_heatmap(transform="Null", group_by = "Treatment",facet_by =c("Batch_hour","Enrichment"), tax_aggregate = "Species",tax_add=NULL,ntax=length(taxa_common_ecoli) ) -> p.ecoli

p.ecoli

```
# VAN Treated 

Transform sample counts (important to do it before filtering out ASVs otherwise the transformation will not be using all the asvs and hence will be wrong), then subseting only the samples relevant to us (Control and van treated batch samples) and then filtering out the ASVs not present in the samples we subsetted.

```{r}
physeq %>% 
  transform_sample_counts(function(x) x/sum(x) * 100) %>% 
  subset_samples(Experiment=="Batch" & Treatment %in% c('UNTREATED','VAN','VAN+CCUG59168')) %>% filter_taxa(function(x) sum(x > 0) > 0, TRUE) -> physeq_batch_enterococcus
```

Get taxa common to the treatment conditions (VAN and VAN + strain)

```{r}
#taxa present in VAN treated samples
taxa_van <- physeq_batch_enterococcus %>% subset_samples(Treatment=='VAN') %>%  filter_taxa(function(x) sum(x > 0) > 0, TRUE) %>% 
  taxa_names()

#taxa present in VAN+strain treated samples
taxa_van_strain <- physeq_batch_enterococcus %>% subset_samples(Treatment=='VAN+CCUG59168')%>%
filter_taxa(function(x) sum(x > 0) > 0, TRUE) %>% 
taxa_names()

#common taxa 
taxa_common_enterococcus<-intersect(taxa_van, taxa_van_strain)

#otu table should only have the ASVs common to the two conditions
otu_table(physeq_batch_enterococcus)<-otu_table(physeq_batch_enterococcus)[taxa_common_enterococcus,] 

```


Plotting the heatmap . 
```{r,fig.dim=c(10,8)}
#reset the factors in the order we want 
sample_data(physeq_batch_enterococcus)$Enrichment <- fct_relevel(sample_data(physeq_batch_enterococcus)$Enrichment,'NotEnriched','Enriched')

sample_data(physeq_batch_enterococcus)$Treatment <- fct_relevel(sample_data(physeq_batch_enterococcus)$Treatment,'UNTREATED','VAN','VAN+CCUG59168')

sample_data(physeq_batch_enterococcus)$Batch_hour <- fct_relevel(sample_data(physeq_batch_enterococcus)$Batch_hour,'0','7','24')

physeq_batch_enterococcus %>% phyloseq_ampvis_heatmap(transform="Null", group_by = "Treatment",facet_by =c("Batch_hour","Enrichment"), tax_aggregate = "Species",tax_add=NULL,ntax=length(taxa_common_enterococcus) ) -> p.enterococcus

p.enterococcus

```
