---
title: "ASV table of relative abundances"
author: "Sneha Sundar"
date: "`r format(Sys.time(), "%a %b %d %X %Y")`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Load required packages
```{r, message=FALSE, error=FALSE}
library(tidyverse)
library(phyloseq)
library(speedyseq)
library(here)

options(getClass.msg=FALSE) # https://github.com/epurdom/clusterExperiment/issues/66
#this fixes an error message that pops up because the class 'Annotated' is defined in two different packages

```

Load required functions
```{r}

'%!in%' <- function(x,y)!('%in%'(x,y))

source("https://raw.githubusercontent.com/fconstancias/DivComAnalyses/master/R/phyloseq_taxa_tests.R")
source("https://raw.githubusercontent.com/fconstancias/DivComAnalyses/master/R/phyloseq_normalisation.R")
source("https://raw.githubusercontent.com/fconstancias/DivComAnalyses/master/R/phyloseq_alpha.R")
source("https://raw.githubusercontent.com/fconstancias/DivComAnalyses/master/R/phyloseq_beta.R")
source("https://raw.githubusercontent.com/fconstancias/DivComAnalyses/master/R/phyloseq_heatmap.R")
```
Load phyloseq object

```{r}
ps = "data/processed/physeq_update_11_1_21.RDS"

ps %>% 
  here::here() %>%
  readRDS() %>%
  phyloseq_get_strains_fast() %>%
  phyloseq_remove_chloro_mitho() -> physeq


```

We want only continuous sample and getting rid of dubious samples. 
```{r}
physeq %>% 
  subset_samples(Experiment == "Continuous") %>% 
  subset_samples(Enrichment == "NotEnriched") %>% 
  subset_samples(Sample_description %!in% c("TR5-15", "TR4-1")) %>% 
  subset_samples(Reactor != "IR2") -> ps_polyFermS
```

Rarefying the data
```{r message = FALSE}
ps_polyFermS %>%  rarefy_even_depth(sample.size = 4576,rngseed = 123) -> ps_rare

ps_rare
```

Convert the rarefied phyloseq object to relative abundances (percentages) and convert to a dataframe. 
```{r}
asv_table <-ps_rare  %>% 
  transform_sample_counts(function(x) round((x/sum(x))*100,2)) %>% 
  otu_table() %>% 
  as.matrix() %>% 
  as.data.frame() %>% 
  rownames_to_column('ASV')
```

Get the taxonomic information as a dataframe. 
```{r}
tax_table_ps_rare <- ps_rare %>% tax_table() %>% as.matrix() %>% as.data.frame() %>% 
  rownames_to_column("ASV")
```

Combine the the asv table and the tax table to get a table that shows the relative abundance of each asv in all the samples along with the taxonomic annotation of each ASV. 
```{r}
rel_abund_asv_table <- tax_table_ps_rare %>% 
  inner_join(asv_table)

```
Make into a CSV file. 

```{r}
outpath <- here("data","processed","rel_abund_asv_table_chicken.csv")
write.table(rel_abund_asv_table,file = outpath,sep=",",row.names = FALSE)      
```

