---
title: "Enriched heatmap and beta diversity plots"
author: "Sneha Sundar"
date: "16/11/2020"
output: html_document
---


```{r setup, include=TRUE}
rm(list = ls()) # remove all the object before starting
knitr::opts_chunk$set(echo = TRUE)
```

### Loading packages and functions from github:


```{r packages,warning=FALSE,message=FALSE,error=FALSE}
library(tidyverse)
library(phyloseq)
library(speedyseq)
library(ggrepel)
library(ampvis2)
library(plotly)
library(zCompositions)
options(getClass.msg=FALSE) # https://github.com/epurdom/clusterExperiment/issues/66

#this fixes an error message that pops up because the class 'Annotated' is defined in two different packages

```


The package is in progress but we do not actually need one, we can directly source the functions from github - **internet connection needed**:
```{r sourcing, results=FALSE, warning=FALSE, include=TRUE,message=FALSE}

source("https://raw.githubusercontent.com/fconstancias/DivComAnalyses/master/R/phyloseq_normalisation.R") 

source("https://raw.githubusercontent.com/fconstancias/DivComAnalyses/master/R/phyloseq_beta.R") 
source("https://raw.githubusercontent.com/fconstancias/DivComAnalyses/master/R/phyloseq_heatmap.R") 


```


### Loading phyloseq object

```{r}
#directory where the phyloseq object is 


ps = "/Users/snehasundar/Desktop/amchick/metabarcoding analysis_9_10_2020/physeq_update.RDS"


#get strain level information
ps %>% 
  readRDS() -> physeq

physeq %>% phyloseq_get_strains_fast() -> physeq

```


### Preparing the data for better plotting

Make factor variables out of the columns in the order we want (if you order it ggplot will appropriately plot the facets in this order).
```{r}
sample_data(physeq)$Experiment <- fct_relevel(sample_data(physeq)$Experiment, "NTC","Mock", "HV292.1", "CCUG59168", "Cecum","Continuous") 

sample_data(physeq)$Reactor <- fct_relevel(sample_data(physeq)$Reactor,"negative","positive","STRAIN", "DONOR", "IR1", "IR2", "CR", "TR1", "TR2", "TR3", "TR4", "TR5", "TR6","Batch3","Batch4")



sample_data(physeq)$Reactor_Treatment <- fct_relevel(sample_data(physeq)$`Reactor_Treatment`,"negative","positive","STRAIN","DONOR", "IR1_UNTREATED","IR2_UNTREATED","CR_UNTREATED","CR_CTX","CR_VAN","TR1_CTX+HV292.1", "TR2_CTX", "TR3_HV292.1", "TR4_VAN","TR5_VAN+CCUG59168","TR6_CCUG59168","Batch3_UNTREATED","Batch3_CTX+HV292.1","Batch3_CTX","Batch3_HV292.1","Batch4_UNTREATED","Batch4_VAN","Batch4_VAN+CCUG59168","Batch4_CCUG59168")

sample_data(physeq)$Enrichment <- fct_relevel(sample_data(physeq)$Enrichment, "NotEnriched","Enriched") 
```

We want to plot only the continuous data so filtering it out. We also do not want reactor IR2
```{r}
physeq %>% subset_samples(Experiment=="Continuous"& Reactor!='IR2') -> physeq.cont
```

## Beta diversity plots

Ordination method : RDA

```{r, message = FALSE}
physeq.cont %>%
  rarefy_even_depth(sample.size = 4576,rngseed = 123) %>%
   phyloseq_plot_bdiv(dlist = NULL, 
                     m = "CoDa", 
                     seed = 123,
                     axis1 = 1,
                     axis2 = 2) -> p
p$PCA$layers[[1]] = NULL
```


```{r, message = FALSE}
p$PCA + geom_point(size=2,
                   aes(color = Reactor_Treatment, 
                       shape = Enrichment,
                       alpha = Day_of_Treatment)) + 
  theme_light() -> plot_RDA

plot_RDA<-plot_RDA + scale_alpha_continuous(range=c(0.3, 0.9)) + scale_color_viridis_d(na.value = "black") + theme_classic()

plot_RDA
```
```{r,warning=FALSE}
p$PCA + geom_point(size=2,
                   aes(color = Reactor_Treatment, 
                       shape = Enrichment,
                       alpha = Day_of_Treatment)) + 
  theme_light() +
  geom_path(arrow = arrow(type = "open", angle = 30, length = unit(0.1, "inches")),
              size = 0.1, linetype = "solid", inherit.aes = TRUE, aes(group=Reactor_Treatment, color = Reactor_Treatment, fill = Reactor_Treatment)) +
  scale_alpha_continuous(range=c( 0.9, 0.3)) + scale_color_viridis_d(na.value = "black") +  theme_classic() -> pl
  
pl
```

## Family level abundances

### Heatmaps of relative abundance of Families in unenriched samples 


```{r message = TRUE, warning= TRUE}

physeq.cont %>% subset_samples(Enrichment == "NotEnriched") %>%
  subset_samples(sample_sums(physeq) > 4576) %>%
subset_samples(Reactor_Treatment %in% c("CR_UNTREATED","TR1_CTX+HV292.1","TR2_CTX","TR3_HV292.1")) %>% phyloseq_ampvis_heatmap(transform='percent',group_by = 'Day_of_Treatment',facet_by = 'Reactor_Treatment',tax_aggregate = "Family",tax_add = NULL,ntax = 350) -> p1.ecoli


p1.ecoli + facet_grid( ~ Reactor_Treatment , scales = "free", space = "free") + 
  scale_fill_viridis_c(breaks = c(0,  0.01, 1, 10, 50, 75, 100), 
                       labels = c(0,  0.01, 1, 10, 50, 75, 100), 
                       trans = scales::log1p_trans(),
                       na.value = 'transparent') -> p1.ecoli




physeq.cont %>% subset_samples(Enrichment == "NotEnriched") %>% 
  subset_samples(sample_sums(physeq) > 4576) %>% 
  subset_samples(Reactor_Treatment %in% c("CR_UNTREATED","TR4_VAN","TR5_VAN+CCUG59168","TR6_CCUG59168")) %>% phyloseq_ampvis_heatmap(transform='percent',group_by = 'Day_of_Treatment',facet_by = 'Reactor_Treatment',tax_aggregate = "Family",tax_add = NULL,ntax = 350) -> p1.ccug


p1.ccug + facet_grid( ~ Reactor_Treatment , scales = "free", space = "free") + 
  scale_fill_viridis_c(breaks = c(0,  0.01, 1, 10, 50, 75, 100), 
                       labels = c(0,  0.01, 1, 10, 50, 75, 100), 
                       trans = scales::log1p_trans(),
                       na.value = 'transparent') -> p1.ccug
```

### Heatmaps of relative abundance of Families in unenriched samples (showing only reactors relevant to E.coli strain)

```{r fig.height=8,fig.width=12, fig.align = "center", message = TRUE, warning= TRUE}
p1.ecoli
```

### Heatmaps of relative abundance of Families in unenriched samples (showing only reactors relevant to Enterococcus strain)



```{r fig.height=8, fig.width=12 ,fig.align = "center", message = TRUE, warning= TRUE}
p1.ccug

```



### Heatmaps of relative abundance of Families in enriched samples 


```{r message = TRUE, warning= TRUE}


physeq.cont %>% subset_samples(Enrichment == "Enriched") %>%
    #subset_samples(sample_sums(physeq) > 4576) %>% 
subset_samples(Reactor_Treatment %in% c("CR_CTX","TR1_CTX+HV292.1","TR2_CTX","TR3_HV292.1")) %>% phyloseq_ampvis_heatmap(transform='percent',group_by = 'Day_of_Treatment',facet_by = 'Reactor_Treatment',tax_aggregate = "Family",tax_add = NULL,ntax = 350) -> p2.ecoli


p2.ecoli + facet_grid( ~ Reactor_Treatment , scales = "free", space = "free") + 
  scale_fill_viridis_c(breaks = c(0,  0.01, 1, 10, 50, 75, 100), 
                       labels = c(0,  0.01, 1, 10, 50, 75, 100), 
                       trans = scales::log1p_trans(),
                       na.value = 'transparent') -> p2.ecoli




physeq.cont %>% subset_samples(Enrichment == "Enriched") %>% 
  #subset_samples(sample_sums(physeq) > 4576) %>% 
  subset_samples(Reactor_Treatment %in% c("CR_VAN","TR4_VAN","TR5_VAN+CCUG59168","TR6_CCUG59168")) %>% phyloseq_ampvis_heatmap(transform='percent',group_by = 'Day_of_Treatment',facet_by = 'Reactor_Treatment',tax_aggregate = "Family",tax_add = NULL,ntax = 350) -> p2.ccug


p2.ccug + facet_grid( ~ Reactor_Treatment , scales = "free", space = "free") + 
  scale_fill_viridis_c(breaks = c(0,  0.01, 1, 10, 50, 75, 100), 
                       labels = c(0,  0.01, 1, 10, 50, 75, 100), 
                       trans = scales::log1p_trans(),
                       na.value = 'transparent') -> p2.ccug
```


### Heatmaps of relative abundance of famililes in enriched samples (showing only reactors relevant to E.coli strain)

```{r fig.height=4, fig.width=7 ,fig.align = "center", message = TRUE, warning= TRUE}
p2.ecoli
```


### Heatmaps of relative abundance of Families in enriched samples (showing only reactors relevant to Enterococcus strain)

```{r fig.height=4, fig.width=7 ,fig.align = "center", message = TRUE, warning= TRUE}
p2.ccug

```
## Genus level information


```{r message = TRUE, warning= TRUE}

physeq.cont %>% subset_samples(Enrichment == "NotEnriched") %>%
  subset_samples(sample_sums(physeq) > 4576) %>%
subset_samples(Reactor_Treatment %in% c("CR_UNTREATED","TR1_CTX+HV292.1","TR2_CTX","TR3_HV292.1")) %>% phyloseq_ampvis_heatmap(transform='percent',group_by = 'Day_of_Treatment',facet_by = 'Reactor_Treatment',tax_aggregate = "Genus",tax_add = NULL,ntax = 350) -> p1.ecoli


p1.ecoli + facet_grid( ~ Reactor_Treatment , scales = "free", space = "free") + 
  scale_fill_viridis_c(breaks = c(0,  0.01, 1, 10, 50, 75, 100), 
                       labels = c(0,  0.01, 1, 10, 50, 75, 100), 
                       trans = scales::log1p_trans(),
                       na.value = 'transparent') -> p1.ecoli




physeq.cont %>% subset_samples(Enrichment == "NotEnriched") %>% 
  subset_samples(sample_sums(physeq) > 4576) %>% 
  subset_samples(Reactor_Treatment %in% c("CR_UNTREATED","TR4_VAN","TR5_VAN+CCUG59168","TR6_CCUG59168")) %>% phyloseq_ampvis_heatmap(transform='percent',group_by = 'Day_of_Treatment',facet_by = 'Reactor_Treatment',tax_aggregate = "Genus",tax_add = NULL,ntax = 350) -> p1.ccug


p1.ccug + facet_grid( ~ Reactor_Treatment , scales = "free", space = "free") + 
  scale_fill_viridis_c(breaks = c(0,  0.01, 1, 10, 50, 75, 100), 
                       labels = c(0,  0.01, 1, 10, 50, 75, 100), 
                       trans = scales::log1p_trans(),
                       na.value = 'transparent') -> p1.ccug
```
### Heatmaps of relative abundance of Genus in unenriched samples (showing only reactors relevant to E.coli strain)


```{r fig.height=15, fig.width=20 ,fig.align = "center", message = TRUE, warning= TRUE}
p1.ecoli
```

### Heatmaps of relative abundance of Genus in unenriched samples (showing only reactors relevant to Enterococcus strain)


```{r fig.height=15, fig.width=20 ,fig.align = "center", message = TRUE, warning= TRUE}
p1.ccug

```
### Heatmaps of relative abundance of Families in enriched samples 


```{r message = TRUE, warning= TRUE}


physeq.cont %>% subset_samples(Enrichment == "Enriched") %>%
    #subset_samples(sample_sums(physeq) > 4576) %>% 
subset_samples(Reactor_Treatment %in% c("CR_CTX","TR1_CTX+HV292.1","TR2_CTX","TR3_HV292.1")) %>% phyloseq_ampvis_heatmap(transform='percent',group_by = 'Day_of_Treatment',facet_by = 'Reactor_Treatment',tax_aggregate = "Genus",tax_add = NULL,ntax = 350) -> p2.ecoli


p2.ecoli + facet_grid( ~ Reactor_Treatment , scales = "free", space = "free") + 
  scale_fill_viridis_c(breaks = c(0,  0.01, 1, 10, 50, 75, 100), 
                       labels = c(0,  0.01, 1, 10, 50, 75, 100), 
                       trans = scales::log1p_trans(),
                       na.value = 'transparent') -> p2.ecoli




physeq.cont %>% subset_samples(Enrichment == "Enriched") %>% 
  #subset_samples(sample_sums(physeq) > 4576) %>% 
  subset_samples(Reactor_Treatment %in% c("CR_VAN","TR4_VAN","TR5_VAN+CCUG59168","TR6_CCUG59168")) %>% phyloseq_ampvis_heatmap(transform='percent',group_by = 'Day_of_Treatment',facet_by = 'Reactor_Treatment',tax_aggregate = "Genus",tax_add = NULL,ntax = 350) -> p2.ccug


p2.ccug + facet_grid( ~ Reactor_Treatment , scales = "free", space = "free") + 
  scale_fill_viridis_c(breaks = c(0,  0.01, 1, 10, 50, 75, 100), 
                       labels = c(0,  0.01, 1, 10, 50, 75, 100), 
                       trans = scales::log1p_trans(),
                       na.value = 'transparent') -> p2.ccug
```


### Heatmaps of relative abundance of Genus in enriched samples (showing only reactors relevant to E.coli strain)

```{r fig.height=7, fig.width=6 ,fig.align = "center", message = TRUE, warning= TRUE}
p2.ecoli
```



### Heatmaps of relative abundance of Genus in enriched samples (showing only reactors relevant to Enterococcus strain)

```{r fig.height=7, fig.width=6 ,fig.align = "center", message = TRUE, warning= TRUE}
p2.ccug

```