---
title: "Increasing_ASVs"
author: "Sneha Sundar"
date: "`r format(Sys.time(), "%a %b %d %X %Y")`"
output: html_document
---


Load required packages
```{r, message=FALSE, error=FALSE}
library(tidyverse)
library(phyloseq)
library(speedyseq)
library(ggrepel)
library(ampvis2)
library(plotly)
library(microbiome)
library(here)
library(UpSetR)
options(getClass.msg=FALSE) # https://github.com/epurdom/clusterExperiment/issues/66
#this fixes an error message that pops up because the class 'Annotated' is defined in two different packages
library(scales)
```

Load required functions
```{r}

'%!in%' <- function(x,y)!('%in%'(x,y))

source("https://raw.githubusercontent.com/fconstancias/DivComAnalyses/master/R/phyloseq_taxa_tests.R")
source("https://raw.githubusercontent.com/fconstancias/DivComAnalyses/master/R/phyloseq_normalisation.R")
source("https://raw.githubusercontent.com/fconstancias/DivComAnalyses/master/R/phyloseq_alpha.R")
source("https://raw.githubusercontent.com/fconstancias/DivComAnalyses/master/R/phyloseq_beta.R")
source("https://raw.githubusercontent.com/fconstancias/DivComAnalyses/master/R/phyloseq_heatmap.R")
```

Load phyloseq object

#NOTE @Alessia  you might need to change this if you do not have the a local version of the github amchick repo. If it does not work, let me/Florentin know. 
```{r}
ps = "data/processed/physeq_update_11_1_21.RDS"

ps %>% 
  here::here() %>%
  readRDS() %>%
  phyloseq_get_strains_fast() %>%
  phyloseq_remove_chloro_mitho() -> physeq


```


We want only continuous samples. 
```{r}
physeq %>% 
  subset_samples(Experiment == "Continuous") %>% 
  subset_samples(Paul %!in% c("Paul")) %>%
  subset_samples(Reactor != "IR2") -> ps_polyFermS
```

Rarefying the data
```{r message = FALSE}
ps_polyFermS %>%  rarefy_even_depth(sample.size = 4576,rngseed = 123) -> ps_rare
```

Taking only the unenriched samples, getting rid of dubious samples and taking only the the days we are interested in (8 days before treatment and 15 days after treatment)
```{r}
ps_rare %>%
  subset_samples(Enrichment == "NotEnriched") %>%
  subset_samples(Sample_description %!in% c("TR5-15", "TR4-1")) %>% # heatmap reveals they might be coming from VAN treated or enrichments
  subset_samples(Day_from_Inoculum >= 30 & Day_from_Inoculum <= 54) -> ps_polyFermS_rare


```

Normalize the otu counts with qPCR data
```{r}
ps_polyFermS_rare %>% phyloseq_density_normalize(value_idx  = "GeneCopyNumberperML") -> norm_ps_polyFermS_rare

taxa<-norm_ps_polyFermS_rare %>% taxa_names()


```

Scaling function log10(1+x). This is so that we can use a log scale in the graphs but still be able to handle 0 counts. 
```{r}
log10p_trans<-function(){trans_new(name="log10p",
                                   transform=function(x){return(log10(x+1))},inverse=function(xt){return((10^xt) - 1)})}
  
#sample_data(physeq)[['Reactor_Treatment']]
```



We need to filter out ASVs that are not stably present in the reactor.This is how we are going to do so:

Say there are 30 samples in TR5. If at least 10 samples have a non-zero value for ASVxxx the particular ASV is considered. Otherwise that ASV is discarded and will not be considered in the analysis to identify trends for TR5. 

```{r}
#function that checks whether a particular ASV should be kept

#taxa_name <- name of ASV to check
#sub <- phyloseq object
#fraction <- min threshold (default 0.33) this is the fraction of samples that need to have a non-zero value for that ASV in order to keep it. 
keep_ASV<- function(taxa_name,sub,fraction=0.33){
  return(sum(otu_table(sub)[taxa_name]>0) >= floor(nsamples(sub)*fraction))
   }


remove_ASVs_from_reactor <- function(physeq, fraction, col_name, value_to_subset){
  #take only the samples we want to look at .
  is_value_to_subset = sample_data(physeq)[[col_name]] == value_to_subset
  subset_physeq<- prune_samples(is_value_to_subset,physeq)
  #remove taxa that are not present in all samples
  subset_physeq <- prune_taxa(taxa_sums(subset_physeq)>0,subset_physeq) 
  
  #apply function `keep_ASV` to all the taxa in the phyloseq object. We will get a logical vector indicating whether or not we should consider the ASV in our analysis. 
  keep<-sapply(taxa_names(subset_physeq),keep_ASV,subset_physeq,fraction)
  
  #remove the ASVS that need to be discarded from the phyloseq object
  return(prune_taxa(keep,subset_physeq))
}
```

An example of how `remove_ASVs_from_reactor` works 
```{r}
sub<-remove_ASVs_from_reactor(norm_ps_polyFermS_rare,0.33,'Reactor_Treatment','TR1_CTX+HV292.1')

```

Function to get the trend (Increasing, decreasing or constant) of a particular ASV
```{r}
get_trend_ASV <- function(taxaname, #name of ASV
                          physeq, 
                          time_col,
                          facet_col ){
  

  out<-plot_taxa_abundances_over_time(physeq = physeq, taxa_level = "Species", 
                                                  taxa_to_plot= taxaname,
                                                  time_column = time_col,
                                                  other_columns_to_plot=NULL ,
                                                  axis_transform=FALSE,
                                                  transformation = 'log1p',
                                                  plot_total = FALSE,
                                                  data_facet1=facet_col,
                                                  data_facet2=NULL,
                                                  n_facet_row=4,
                                                  n_facet_col=4)




#perform the linear regression and get the slope
return(coef(lm(out$plot.df[['Abundances']]~out$plot.df[[time_col]]))[2])
  
}

```

For all the ASVs present in each reactor (after we filtered out the unstable ASVs), identify the trend as increasing, decreasing or constant . 
```{r}
#get all the reactor names 
reactors<-norm_ps_polyFermS_rare %>% get_variable('Reactor_Treatment') %>% as.vector() %>% unique()

#list object to store the trends of ASVs for each reactor
reactor_trends = vector("list", length(reactors))
#all the ASVs present in each reactor 
ASV_sets <- vector("list", length(reactors))

names(reactor_trends) <- reactors
names(ASV_sets) <- reactors

for(reactor in reactors){
  
  sub<-remove_ASVs_from_reactor(norm_ps_polyFermS_rare,0.33,'Reactor_Treatment',reactor)
  
  lm_coef<-sapply(taxa_names(sub),get_trend_ASV,sub,'Day_from_Inoculum','Reactor_Treatment')
  names(lm_coef) <- tax_table(sub)[taxa_names(sub),'Strain']
  
  trend = list('Increasing' = names(which(lm_coef >0)),'Decreasing'=names(which(lm_coef <0)),'Constant'=names(which(lm_coef==0)))
  ASV_sets[[reactor]] = names(lm_coef)
  reactor_trends[[reactor]] = trend
}


```

`reactor_trends` holds the information of the trend of all the non-zero, stable ASVs present in each reactor.

You can access each reactor's information like this. 
```{r}
reactor_trends$CR_UNTREATED
```

# Visualizing the venn diagram using an upset plot

## For Van treated reactors
```{r}

(upset_van <-upset(fromList(ASV_sets[c(1,6,7,8)]), order.by = "freq",nsets = 4,nintersects=NA,keep.order=TRUE))
```

## For CTX treated reactors
```{r}
(upset_ctx<-upset(fromList(ASV_sets[c(1,3,4,5)]), order.by = "freq",nsets = 4,nintersects=NA,keep.order=TRUE))


```



Assign taxa as rownames of dataframe used by the upset plot so we can start making the venn diagram . 
```{r}

asvs_ctx<-purrr::reduce(ASV_sets[c(1,3,4,5)],union)
asvs_van<-purrr::reduce(ASV_sets[c(1,6,7,8)],union)
rownames(upset_ctx$New_data) <- asvs_ctx
rownames(upset_van$New_data) <- asvs_van

```

```{r}
ctx_df<-upset_ctx$New_data
van_df <- upset_van$New_data
```


# Function to query set combinations and identify the trends of the asv in that set

```{r}
identify_asv_trends_in_set <-function(set_combo, #what reactors should we consider
                                      upset_df, #dataframe to use in upset format
                                      vars, #variable names in upset_df
                                      trends_object)  
{
  
  #names of reactors we are considering  
  reactors <- vars[set_combo %>% as.logical()]
  #names of ASV present in set combination
  asvs_tmp<-upset_df %>% filter(.data[[vars[[1]]]] == set_combo[[1]],.data[[vars[[2]]]] == set_combo[[2]], .data[[vars[[3]]]] == set_combo[[3]], .data[[vars[[4]]]] == set_combo[[4]]) %>% rownames()
  
  #function to identify the trend of a particular ASV in a reactor
  what_trend <- function(asv,reactor_name,reactor_trends){
  if(asv %in% reactor_trends[[reactor_name]][['Increasing']]){return('Increasing')}
  if(asv %in% reactor_trends[[reactor_name]][['Decreasing']]){return('Decreasing')}
  if(asv %in% reactor_trends[[reactor_name]][['Constant']]){return('Constant')}
  }
  
  #obtain result as a dataframe
  result <- data.frame(row.names = asvs_tmp)
  for(reactor_name in reactors){
  trends<-sapply(asvs_tmp,what_trend,reactor_name,reactor_trends)
  
  result <- cbind(result,trends)
  }
  colnames(result) <- reactors
 return(result)
  
}
```

#Identify trends of ASVs in CTX treated reactors

This is the order to keep in mind
```{r}

colnames(ctx_df)
```
# If you have run all the code before this, you just need to change the set_combo argument to whatever reactor you want to look at (1 for considering the reactor, 0 for not considering it)

Example: If I want to consider the ASVs in CR, TR3 then `set_combo = c(1,0,0,1)

#some combinations might throw an error. This means there is no ASV belonging to that set. 
```{r}
(t_ctx<-identify_asv_trends_in_set(set_combo=c(1,0,0,1),upset_df = ctx_df,vars = colnames(ctx_df),trends_object = reactor_trends))


```
#Identify trends of ASV in VAN treated reactors

Order to keep in mind. 
```{r}
colnames(van_df)
```

```{r}
(t_van<-identify_asv_trends_in_set(set_combo=c(1,1,1,1),upset_df = van_df,vars = colnames(van_df),trends_object = reactor_trends))
```











