---
title: " presence-absence-enrichments "
author: "Florentin"
date: '`r format(Sys.time(), "%d %B, %Y")`'

output: 
  html_document: 
    toc: yes
    keep_md: yes
---
```{r setup, include=TRUE}
rm(list = ls()) # remove all the object before starting
knitr::opts_chunk$set(echo = TRUE)
```



```{r packages,warning=FALSE,message=FALSE,error=FALSE}
library(tidyverse)
library(phyloseq)
library(speedyseq)
library(ggrepel)
library(ampvis2)
library(plotly)
library(zCompositions)
options(getClass.msg=FALSE) # https://github.com/epurdom/clusterExperiment/issues/66

#this fixes an error message that pops up because the class 'Annotated' is defined in two different packages
```

The package is in progress but we do not actually need one, we can directly source the functions from github - **internet connection needed**:
```{r sourcing, results=FALSE, warning=FALSE, include=TRUE}
source("https://raw.githubusercontent.com/fconstancias/DivComAnalyses/master/R/phyloseq_taxa_tests.R")
source("https://raw.githubusercontent.com/fconstancias/DivComAnalyses/master/R/phyloseq_normalisation.R")
source("https://raw.githubusercontent.com/fconstancias/DivComAnalyses/master/R/phyloseq_alpha.R")
source("https://raw.githubusercontent.com/fconstancias/DivComAnalyses/master/R/phyloseq_beta.R")
source("https://raw.githubusercontent.com/fconstancias/DivComAnalyses/master/R/phyloseq_heatmap.R")
```


# Load data nicely (using here::here()):


```{r}
ps = "data/processed/physeq_update_23_11.RDS"

ps %>% 
  here::here() %>%
  readRDS() %>%
  phyloseq_get_strains_fast() %>%
  phyloseq_remove_chloro_mitho() -> physeq
```  


Extract enriched data and add Block variable:

```{r, message = FALSE}
physeq %>%
  subset_samples(Experiment == "Continuous" &
                   Enrichment == "Enriched" ) %>%
  rarefy_even_depth(sample.size = 2574,rngseed = 123) -> physeq_enriched

sample_data(physeq_enriched)$Block = ifelse(sample_data(physeq_enriched)$Treatment %in% c("CTX", "CTX+HV292.1", "HV292.1"), "Exp CTX", "Exp Van")
```


# Heatmaps to visualize ASV level community: Mean values first


```{r fig.height=16, fig.width=10 ,fig.align = "center", message = TRUE, warning= TRUE}
physeq_enriched %>%
  phyloseq_ampvis_heatmap(transform = "percent",
                          group_by = "Reactor_Treatment",
                          facet_by = c("Reactor_Treatment","Enrichment", "Block", "Reactor", "Treatment", "Experiment" ),
                          tax_aggregate = "Species",
                          tax_add = NULL,
                          ntax  = 600) -> p

p + facet_grid( ~ Block + Reactor_Treatment , scales = "free", space = "free") + 
  scale_fill_viridis_c(breaks = c(0,  0.01, 1, 10, 50, 75, 100), 
                       labels = c(0,  0.01, 1, 10, 50, 75, 100), 
                       trans = scales::pseudo_log_trans(sigma = 0.001),
                       na.value = 'transparent') -> p2
p2
```
all data.

```{r fig.height=20, fig.width=30 ,fig.align = "center", message = TRUE, warning= TRUE}
physeq_enriched %>%
  phyloseq_ampvis_heatmap(transform = "percent",
                          group_by = "Day_of_Treatment",
                          facet_by = c("Reactor_Treatment","Enrichment", "Block", "Reactor", "Treatment", "Experiment" ),
                          tax_aggregate = "Species",
                          tax_add = NULL,
                          ntax  = 600) -> p

p + facet_grid( ~ Block + Reactor_Treatment, scales = "free", space = "free") + 
  scale_fill_viridis_c(breaks = c(0,  0.01, 1, 10, 50, 75, 100), 
                       labels = c(0,  0.01, 1, 10, 50, 75, 100), 
                       trans = scales::pseudo_log_trans(sigma = 0.001),
                       na.value = 'transparent') -> p2
p2
```


# Let's consider presence absence: what is not present in control but not in other ones.

## Define a function:

```{r, message = FALSE}
venn_enrichment <- function(physeq_enrichment,
                              physeq_reactor,
                              venn_column,
                              cut_a = 0,
                              cut_f = 100,
                              groups = c("TR6", "TR5", "TR5_TR6"),
                              group = "Day_of_Treatment",
                              facet_by = c("Reactor_Treatment","Enrichment", "Reactor", "Treatment", "Experiment"),
                              ntax = 600)
{
  
  require(UpSetR); require(MicrobiotaProcess)
  physeq_enrichment -> tmp2
  tax_table(physeq_enrichment) <- tax_table(physeq_enrichment)[,c("Kingdom", "Phylum", "Class", "Order", "Family", "Genus", "Strain")]
  colnames(tax_table(physeq_enrichment)) = c("Kingdom", "Phylum", "Class", "Order", "Family", "Genus", "Species")
  
  physeq_enrichment %>%
    phyloseq_to_ampvis2() %>%
    amp_venn(group_by = venn_column, cut_a = cut_a, cut_f = cut_f, normalise = FALSE, detailed_output = TRUE) -> venn_p
  
  physeq_enrichment %>%
    MicrobiotaProcess::get_upset(factorNames = venn_column) %>%
    UpSetR::upset(order.by = "freq", empty.intersections = "on") -> up #https://www.rdocumentation.org/packages/ggupset/versions/0.3.0
  
  
  venn_p$Otutable %>%
    dplyr::select(-Kingdom:-Genus) %>%
    dplyr::filter(Shared %in% !!groups) -> df
  
  df %>%
    dplyr::pull(OTU) -> ASV
  
  
  prune_taxa(ASV,
             tmp2) %>%
    phyloseq_ampvis_heatmap(transform = "none",
                            group_by = group,
                            facet_by = facet_by,
                            tax_aggregate = "Species",
                            tax_add = NULL,
                            ntax  = ntax) +
    scale_fill_viridis_c(breaks = c(0,  0.01, 1, 10, 50, 75, 100),
                         labels = c(0,  0.01, 1, 10, 50, 75, 100),
                         trans = scales::pseudo_log_trans(sigma = 0.001),
                         na.value = 'transparent') -> p
  
  #
  physeq_reactor %>%
    transform_sample_counts(function(x) x/sum(x) * 100) %>%
    plot_taxa_abundances_over_time(taxa_level = "Strain",
                                   taxa_to_plot = df %>% pull("Species"),
                                   time_column = group,
                                   axis_transform = FALSE,
                                   transformation = FALSE,
                                   data_facet1 = "Reactor") -> p2
  
  return(list("venn_plot" = venn_p$plot + ggtitle(paste0("cut_a: ", cut_a, " cut_f: ", cut_f)),
              "upset_plot" = up,
              "filter_df" = df,
              "plot_enrichment" = p,
              "plot_reactor" = p2$plot + ylab("Proportion (%)") + theme_light()
  ))
}
```


global function
```{r, message = FALSE}
venn_enrichment_all <- function(physeq_enrichment,
                              physeq_reactor,
                              venn_column,
                              cut_a = 0,
                              cut_f = 100,
                              group = "Day_of_Treatment",
                              facet_by = c("Reactor_Treatment","Enrichment", "Reactor", "Treatment", "Experiment"),
                              ntax = 600)
{
  
  require(UpSetR); require(MicrobiotaProcess)
  
  venn_enrichment(physeq_enrichment = physeq_enriched %>%
                    subset_samples(Block == "Exp CTX") %>%
                    subset_samples(Reactor %in% c("CR", "TR1", "TR3")) %>%
                    transform_sample_counts(function(x) x/sum(x) * 100) ,
                  physeq_reactor = physeq %>%
                    subset_samples(Experiment == "Continuous" &
                                     Enrichment == "NotEnriched") %>%
                    subset_samples(Reactor %in% c("CR", "TR1", "TR3")) %>%
                    rarefy_even_depth(sample.size = 2574,rngseed = 123),
                  venn_column = "Reactor",
                  cut_a = cut_a,
                  cut_f = cut_f,
                  groups = c("TR1", "TR3", "TR1_TR3")) -> CTX
  
  
  venn_enrichment(physeq_enrichment = physeq_enriched %>%
                    subset_samples(Block == "Exp Van") %>%
                    subset_samples(Reactor %in% c("CR", "TR5", "TR6")) %>%
                    transform_sample_counts(function(x) x/sum(x) * 100) ,
                  physeq_reactor = physeq %>%
                    subset_samples(Experiment == "Continuous" &
                                     Enrichment == "NotEnriched") %>%
                    subset_samples(Reactor %in% c("CR", "TR5", "TR6")) %>%
                    rarefy_even_depth(sample.size = 2574,rngseed = 123),
                  venn_column = "Reactor",
                  cut_a = cut_a,
                  cut_f = cut_f,
                  groups = c("TR6", "TR5", "TR5_TR6")) -> VAN
  
  return(list("CTX" = CTX,
              "VAN" = VAN
  ))
}
```

```{r, message = FALSE}
venn_enrichment_all() -> a
```

```{r, message = FALSE}
venn_enrichment_all(cut_a = 1,
                    cut_f = 90) -> b
```

```{r, message = FALSE}
venn_enrichment_all(cut_a = 0.1,
                    cut_f = 80) -> c
```
```{r, message = FALSE}
venn_enrichment_all(cut_a = 0,
                    cut_f = 80) -> d
```


Summary:

VAN:

```{r fig.height=6, fig.width=6 ,fig.align = "center", message = TRUE, warning= TRUE}
ggpubr::ggarrange(a$VAN$venn_plot,
                  b$VAN$venn_plot,
                  c$VAN$venn_plot,
                  d$VAN$venn_plot)
```
```{r fig.height=10, fig.width=12 ,fig.align = "center", message = TRUE, warning= TRUE}
ggpubr::ggarrange(a$VAN$plot_enrichment,
                  b$VAN$plot_enrichment,
                  c$VAN$plot_enrichment,
                  d$VAN$plot_enrichment,
                  align = "hv",
                  common.legend = TRUE)
```
```{r fig.height=5, fig.width=7 ,fig.align = "center", message = TRUE, warning= TRUE}
ggpubr::ggarrange(a$VAN$plot_reactor + scale_y_sqrt(),
                  b$VAN$plot_reactor + scale_y_sqrt(),
                  c$VAN$plot_reactor + scale_y_sqrt(),
                  d$VAN$plot_reactor + scale_y_sqrt(),
                  legend = "bottom",
                  common.legend = TRUE)
```



CTX:

```{r fig.height=6, fig.width=6 ,fig.align = "center", message = TRUE, warning= TRUE}
ggpubr::ggarrange(a$CTX$venn_plot,
                  b$CTX$venn_plot,
                  c$CTX$venn_plot,
                  d$CTX$venn_plot)
```
```{r fig.height=10, fig.width=12 ,fig.align = "center", message = TRUE, warning= TRUE}
ggpubr::ggarrange(a$CTX$plot_enrichment,
                  b$CTX$plot_enrichment,
                  c$CTX$plot_enrichment,
                  d$CTX$plot_enrichment,
                  align = "hv",
                  common.legend = TRUE)
```
```{r fig.height=5, fig.width=7 ,fig.align = "center", message = TRUE, warning= TRUE}
ggpubr::ggarrange(a$CTX$plot_reactor + scale_y_sqrt(),
                  b$CTX$plot_reactor + scale_y_sqrt(),
                  c$CTX$plot_reactor + scale_y_sqrt(),
                  d$CTX$plot_reactor + scale_y_sqrt(),
                  legend = "bottom",
                  common.legend = TRUE)
```


```{r fig.height=6, fig.width=14 ,fig.align = "center", message = TRUE, warning= TRUE}
c$CTX$plot_reactor   + scale_y_log10() + facet_grid(Reactor ~ Data) -> p 

p  + theme(legend.position="none")
```

```{r, message = FALSE}
ggpubr::get_legend(p) %>% ggpubr::as_ggplot()
```




```{r fig.height=6, fig.width=14 ,fig.align = "center", message = TRUE, warning= TRUE}
c$VAN$plot_reactor + scale_y_log10() + facet_grid(Reactor ~ Data) -> p 

p  + theme(legend.position="none")
```

```{r, message = FALSE}
ggpubr::get_legend(p) %>% ggpubr::as_ggplot()
```




# Next: 

- ~~summarise with diff cutoffs~~
- ~~upsetR plots : https://rdrr.io/github/YuLab-SMU/MicrobiotaProcess/man/get_upset.html~~
- not presence absence but fold change (with Sneha using Deseq2)