---
title: "receiver_taxa_with_qpcr"
author: "Sneha Sundar"
date: "09/11/2020"
output: html_document
---

```{r setup, include=FALSE}
rm(list = ls())
knitr::opts_chunk$set(echo = TRUE)
```

### Loading required packages

```{r,packages,message=FALSE}
library(tidyverse)
library(phyloseq)
library(plotly)
library(patchwork)
source("/Users/snehasundar/Desktop/amchick/metabarcoding analysis_9_10_2020/receiver_taxa_analysis/plot_asvs_over_time.R")
source("https://raw.githubusercontent.com/fconstancias/DivComAnalyses/master/R/phyloseq_normalisation.R")
source("/Users/snehasundar/Desktop/amchick/metabarcoding analysis_9_10_2020/receiver_taxa_analysis/test2.R")
```


### Loading phyloseq object

```{r}
#directory where the phyloseq object is 

ps = "/Users/snehasundar/Desktop/amchick/metabarcoding analysis_9_10_2020/physeq_update.RDS"

ps %>% 
  readRDS() -> physeq

```

### Preparing the data for better plotting

Make factor variables out of the columns in the order we want (if you order it ggplot will appropriately plot the facets in this order).
```{r}
sample_data(physeq)$Experiment <- fct_relevel(sample_data(physeq)$Experiment, "NTC","Mock", "HV292.1", "CCUG59168", "Cecum","Continuous") 

sample_data(physeq)$Reactor <- fct_relevel(sample_data(physeq)$Reactor,"negative","positive","STRAIN", "DONOR", "IR1", "IR2", "CR", "TR1", "TR2", "TR3", "TR4", "TR5", "TR6","Batch3","Batch4")



sample_data(physeq)$Reactor_Treatment <- fct_relevel(sample_data(physeq)$`Reactor_Treatment`,"negative","positive","STRAIN","DONOR", "IR1_UNTREATED","IR2_UNTREATED","CR_UNTREATED","CR_CTX","CR_VAN","TR1_CTX+HV292.1", "TR2_CTX", "TR3_HV292.1", "TR4_VAN","TR5_VAN+CCUG59168","TR6_CCUG59168","Batch3_UNTREATED","Batch3_CTX+HV292.1","Batch3_CTX","Batch3_HV292.1","Batch4_UNTREATED","Batch4_VAN","Batch4_VAN+CCUG59168","Batch4_CCUG59168")

sample_data(physeq)$Enrichment <- fct_relevel(sample_data(physeq)$Enrichment, "NotEnriched","Enriched") 
```


### Filter out the samples with low reads


```{r}

physeq%>% 
subset_samples(sample_sums(physeq) > 4576) -> physeq

```

### What are the strain-specific taxa?

Create a subset of physeq that contains only the two strains - CCUG-59168 (a strain of *Enterococcus faecalis*) and HV292.1 (a strain of *Escherichia coli*)
```{r}
physeq %>% 
  subset_samples(Experiment %in% c("CCUG59168","HV292.1")) -> strain_physeq
```


Obtain the ASVs specific to the two strains 
```{r}
#store otu table as dataframe in `strain_ASV`
otu_table(strain_physeq) %>% as.matrix() %>% as.data.frame() %>% rownames_to_column("ASV") -> strain_ASV

CCUG_ASVs <- strain_ASV$ASV[strain_ASV$`CCUG-59168-S352`>1000]
HV292_ASVs <- strain_ASV$ASV[strain_ASV$`HV292-1-S350`>1000]

CCUG_ASVs

HV292_ASVs
```


We want to plot only the continuous data so filtering it out. We also do not want reactor IR2
```{r}
physeq %>% subset_samples(Experiment=="Continuous"& Reactor!='IR2') -> physeq.cont


```



### Relative abundances plots


We plot for each Reactor:Treatment combination the percentage of ASVs that are specific to the *E.coli* strain HV292.1 against the `Day_from_Inoculum`. Color of points and lines indicates the particular ASV or the sum of both. The shape of the point tells us whether the percentage was 0 or not. The first vertical gray bar indicates the start of connection to second-stage reactor. The second vertical gray bar indicates the start of treatment. The data was transformed with $log_{10}(1+x)$, where $x$ indicates percentage of strain-specific ASVs. This particular transformation was chosen in order to handle the percantages that were 0 in the data.


```{r,fig.width=10, fig.height=9,eval=FALSE}

result_relative_HV292 <- plot_asvs_over_time(physeq=physeq.cont,asvs_to_plot=HV292_ASVs,plot_sum=TRUE,plot_relative_abundances=TRUE,plot_absolute_abundances=FALSE,log_transform=TRUE,data_facet1='Reactor_Treatment',data_facet2='Enrichment',time_column='Day_from_Inoculum',n_facet_col=2)

#annotating the plots with time points 
result_relative_HV292$plot <- result_relative_HV292$plot + geom_vline(xintercept = c(23,39),color="black",alpha=0.4)+ggtitle("Pecentage of HV292 specific ASV") + theme_light()

result_relative_HV292$plot
```



We plot for each Reactor:Treatment combination the percentage of ASVs that are specific to the *Enterococcus faecium* strain, CCUG-59168, against the `Day_from_Inoculum`. Color of points and lines indicates the particular ASV or the sum of both. The shape of the point tells us whether the percentage was 0 or not. The first vertical gray bar indicates the start of connection to second-stage reactor. The second vertical gray bar indicates the start of treatment. The second vertical gray bar indicates the start of treatment. The data was transformed with $log_{10}(1+x)$, where $x$ indicates percentage of strain-specific ASVs. This particular transformation was chosen in order to handle the percantages that were 0 in the data. 


```{r,fig.width=10, fig.height=9,eval=FALSE}

result_relative_CCUG <- plot_asvs_over_time(physeq=physeq.cont,asvs_to_plot=CCUG_ASVs,plot_sum=TRUE,plot_relative_abundances=TRUE,plot_absolute_abundances=FALSE,log_transform=TRUE,data_facet1='Reactor_Treatment',data_facet2='Enrichment',time_column='Day_from_Inoculum',n_facet_col=2)

#annotating the plots with time points 

result_relative_CCUG$plot <- result_relative_CCUG$plot + geom_vline(xintercept = c(23,39),color="black",alpha=0.4)+ggtitle("Pecentage of CCUG specific ASV") + theme_light()


result_relative_CCUG$plot
```


### Plotting absolute abundances of these strain-specific ASV over time (we use qPCR data of total bacterial count to estimate number of copies of strain-specific ASVs)


```{r,fig.width=10, fig.height=9,eval=FALSE}

result_abs_HV292<-plot_asvs_over_time(physeq=physeq.cont,asvs_to_plot=HV292_ASVs,plot_sum=TRUE,plot_relative_abundances = FALSE,plot_absolute_abundances=TRUE,qcpr_data_column_in_metadata = "GeneCopyNumberperML",log_transform=TRUE,data_facet1='Reactor_Treatment',data_facet2='Enrichment',time_column='Day_from_Inoculum',n_facet_col=2)

result_abs_HV292$plot <- result_abs_HV292$plot + geom_vline(xintercept = c(23,39),color="black",alpha=0.4)+ggtitle("Absolute abundances of HV292 specific ASV") + theme_light()

result_abs_HV292$plot

```

```{r,fig.width=10, fig.height=9,eval=FALSE}

result_abs_CCUG <- plot_asvs_over_time(physeq=physeq.cont,asvs_to_plot=CCUG_ASVs,plot_sum=TRUE,plot_relative_abundances = FALSE,plot_absolute_abundances=TRUE,qcpr_data_column_in_metadata = "GeneCopyNumberperML",log_transform=TRUE,data_facet1='Reactor_Treatment',data_facet2='Enrichment',time_column='Day_from_Inoculum',n_facet_col=2)

result_abs_CCUG$plot <- result_abs_CCUG$plot + geom_vline(xintercept = c(23,39),color="black",alpha=0.4)+ggtitle("Absolute abundances of CCUG specific ASV") + theme_light()

result_abs_CCUG$plot.df
```


### The following two plots DO NOT have the copy numbers of our strains as determined from qPCR !!

CCUG specific asv
```{r,fig.width=20, fig.height=12,eval=FALSE}

result_abs_CCUG$plot + result_relative_CCUG$plot 
```
HV292  specific asv

```{r,fig.width=20, fig.height=12,eval=FALSE}
result_abs_HV292$plot + result_relative_HV292$plot

```


```{r,eval=FALSE}
physeq.cont.df<-physeq.cont %>% sample_data() %>% as.matrix() %>% as.data.frame() %>% mutate(Day_from_Inoculum = as.numeric(Day_from_Inoculum)) %>% mutate(CCUG59168_Copy_Number_permL= as.numeric(CCUG59168_Copy_Number_permL)) %>% mutate(
HV292.1_Copy_Number_permL=as.numeric(
HV292.1_Copy_Number_permL)) %>% arrange(Day_from_Inoculum)


abs_CCUG_qPCR<-result_abs_CCUG$plot +
geom_point(data=physeq.cont.df ,mapping=aes(x=Day_from_Inoculum,y=log10(1+CCUG59168_Copy_Number_permL),shape=CCUG59168_Copy_Number_permL==0),color="#CC79A7") 


abs_HV292_qPCR <- result_abs_HV292$plot + geom_point(data=physeq.cont.df ,mapping=aes(x=Day_from_Inoculum,y=log10(1+HV292.1_Copy_Number_permL),shape=HV292.1_Copy_Number_permL==0),color="#CC79A7") 


```


### HV292  specific asv with strain specific qPCR data

```{r fig.width=10, fig.height=9,eval=FALSE}
abs_HV292_qPCR
```




```{r,fig.width=20, fig.height=10,,eval=FALSE}
abs_HV292_qPCR + result_relative_HV292$plot

```


### CCUG specific asv with strain specific qPCR data

```{r fig.width=10, fig.height=9,eval=FALSE}
abs_CCUG_qPCR
```


```{r,fig.width=20, fig.height=12,eval=FALSE}

abs_CCUG_qPCR + result_relative_CCUG$plot 
```




```{r}
result_family_CCUG<-plot_taxa_absolute_abundances_over_time(physeq=physeq.cont,taxa_level="Family",taxa_to_plot =c("Enterococcaceae"),datacol_for_density_normalization="GeneCopyNumberperML",    other_columns_to_plot=c("CCUG59168_Copy_Number_permL","VAN_Copy_Number_permL"),
                                        log_transform=TRUE,
                                        data_facet1='Reactor_Treatment',
                                        data_facet2='Enrichment',
                                        time_column='Day_from_Inoculum',n_facet_col=2)
```


```{r fig.width=11, fig.height=9}
result_family_CCUG$plot + ggtitle(label="Absolute abundances of ASVs specific to Family Enterococcaceae ") + geom_vline(xintercept = c(23,39),color="black",alpha=0.4) + theme_light()
```
```{r}
result_family_HV292<-plot_taxa_absolute_abundances_over_time(physeq=physeq.cont,taxa_level="Family",taxa_to_plot =c("Enterobacteriaceae"),datacol_for_density_normalization="GeneCopyNumberperML",    other_columns_to_plot=c("HV292.1_Copy_Number_permL","CTX_Copy_Number_permL"),
                                        log_transform=TRUE,
                                        data_facet1='Reactor_Treatment',
                                        data_facet2='Enrichment',
                                        time_column='Day_from_Inoculum',n_facet_col=2)
```


```{r fig.width=11, fig.height=9}
result_family_HV292$plot + ggtitle("Absolute abundances of ASVs specific to Family Enterobacteriaceae")+ geom_vline(xintercept = c(23,39),color="black",alpha=0.4) + theme_light()
```


```{r}
result_genus_CCUG<-plot_taxa_absolute_abundances_over_time(physeq=physeq.cont,taxa_level="Genus",taxa_to_plot =c("Enterococcus"),datacol_for_density_normalization="GeneCopyNumberperML",    other_columns_to_plot=c("CCUG59168_Copy_Number_permL","VAN_Copy_Number_permL"),
                                        log_transform=TRUE,
                                        data_facet1='Reactor_Treatment',
                                        data_facet2='Enrichment',
                                        time_column='Day_from_Inoculum',n_facet_col=2)
```


```{r fig.width=11, fig.height=9}
result_genus_CCUG$plot + ggtitle(label="Absolute abundances of ASVs specific to Genus Enterococcus") + geom_vline(xintercept = c(23,39),color="black",alpha=0.4) + theme_light()
```




```{r}
result_genus_HV292<-plot_taxa_absolute_abundances_over_time(physeq=physeq.cont,taxa_level="Genus",taxa_to_plot =c("Escherichia-Shigella"),datacol_for_density_normalization="GeneCopyNumberperML",    other_columns_to_plot=c("HV292.1_Copy_Number_permL","CTX_Copy_Number_permL"),
                                        log_transform=TRUE,
                                        data_facet1='Reactor_Treatment',
                                        data_facet2='Enrichment',
                                        time_column='Day_from_Inoculum',n_facet_col=2)
```


```{r fig.width=11, fig.height=9}
result_genus_HV292$plot + ggtitle("Absolute abundances of ASVs specific to Genus Escherichia-Shigella") + geom_vline(xintercept = c(23,39),color="black",alpha=0.4) + theme_light()
```


