---
title: "Receiver taxa analysis"
author: "Sneha"
date: "23/10/2020"
output: html_document
---

```{r setup, include=FALSE}
rm(list = ls())
knitr::opts_chunk$set(echo = TRUE)
```

### Loading required packages

```{r,packages,message=FALSE}
library(tidyverse)
library(phyloseq)
library(plotly)
```


### Loading phyloseq object

```{r}
#directory where the phyloseq object is 

ps = "/Users/snehasundar/Desktop/amchick/metabarcoding analysis_9_10_2020/physeq_update.RDS"

ps %>% 
  readRDS() -> physeq
```

### Filter out samples having low reads

```{r}
physeq %>% 
subset_samples(sample_sums(physeq) > 4576) -> physeq



```


### What are the strain-specific taxa?

Create a subset of physeq that contains only the two strains - CCUG-59168 (a strain of *Enterococcus faecalis*) and HV292.1 (a strain of *Escherichia coli*)
```{r}
(physeq %>% 
  subset_samples(Experiment %in% c("CCUG59168","HV292.1")) -> strain_physeq)
```


Obtain the ASVs specific to the two strains 
```{r}
#store otu table as dataframe in `strain_ASV`
otu_table(strain_physeq) %>% as.matrix() %>% as.data.frame() %>% rownames_to_column("ASV") -> strain_ASV

strain_ASV %>%
  DT::datatable()


CCUG_ASVs <- strain_ASV$ASV[strain_ASV$`CCUG-59168-S352`>1000]
HV292_ASVs <- strain_ASV$ASV[strain_ASV$`HV292-1-S350`>1000]

CCUG_ASVs

HV292_ASVs
```




### Normalization to percentages

Normalizing the counts for each ASV to percentages
```{r}

physeq %>%
  transform_sample_counts(function(x) x/sum(x) * 100) ->normalized_physeq

```

### Preparing the dataset for plotting

Create a dataframe with the % strain specific ASVs for each sample that we want to plot
```{r}
#make otu table filterable by ASV
normalized_physeq %>%  otu_table() %>% as.matrix() %>% as.data.frame() %>% rownames_to_column("ASV") -> otu.table


#add columns containing the percent of strain specific ASVs to sample data and make sure all the columns that should be numeric are numeric
physeq %>% 
  sample_data %>% 
  as.matrix() %>% 
  as.data.frame() %>% 
  rownames_to_column("SampleID") %>%  

  mutate(ASV20=as.numeric(otu.table[20,-1]),ASV21=as.numeric(otu.table[21,-1]),ASV84=as.numeric(otu.table[84,-1]),ASV86=as.numeric(otu.table[86,-1])) %>%  

  mutate(Day_of_Connection=as.numeric(Day_of_Connection),Day_of_Treatment=as.numeric(Day_of_Treatment),Day_from_Inoculum=as.numeric(Day_from_Inoculum))  %>% 
  
  mutate(ASV_HV292=ASV20+ASV21,ASV_CCUG=ASV84+ASV86) -> sample.data
  
  

#filter out batch and IR2
sample.data %>% filter(Experiment != "Batch", Reactor != "IR2") -> sample.data.cont #this is 

#what we will use for plotting

head(sample.data.cont)

```

Resetting factor levels in the order we want

```{r}

sample.data.cont$Experiment <- fct_relevel(sample.data.cont$Experiment,"Mock", "HV292.1", "CCUG59168", "Cecum","Continuous")

#creating a new column with treatment and reactor combined for better readability in the graph
sample.data.cont %>% unite("Reactor_Treatment",Reactor:Treatment,remove = FALSE) ->sample.data.cont


sample.data.cont$`Reactor_Treatment`[sample.data.cont$`Reactor_Treatment`=="DONOR_DONOR"] <- "DONOR"

sample.data.cont$`Reactor_Treatment`[sample.data.cont$`Reactor_Treatment`=="STRAIN_STRAIN"] <- "STRAIN"

sample.data.cont$`Reactor_Treatment`[sample.data.cont$`Reactor_Treatment`=="positive_positive"] <- "positive"

sample.data.cont$Reactor_Treatment <- fct_relevel(sample.data.cont$`Reactor_Treatment`,"positive","STRAIN","DONOR", "IR1_UNTREATED","CR_UNTREATED","CR_CTX","CR_VAN","TR1_CTX+HV292.1", "TR2_CTX", "TR3_HV292.1", "TR4_VAN","TR5_VAN+CCUG59168","TR6_CCUG59168")

```

### Plotting HV292 specific ASVs:

Creating a the dataframe for plotting with all the columns we need (the differnt aesthetics)

```{r}
sample.data.cont %>% 
  filter(Reactor %in% c("IR1","CR","TR1", "TR2", "TR3", "TR4","TR5","TR6")) %>% arrange(Day_from_Inoculum) %>% 
  select(Sample_description,Reactor,Treatment,Reactor_Treatment,Experiment,Enrichment,Day_from_Inoculum,ASV20,ASV21,ASV_HV292) -> plot.df1

  
plot.df1 %>% pivot_longer(cols = starts_with("ASV"),names_to = "ASV",values_to = "Percent_ASV" ) -> plot.df1

head(plot.df1)
```

We plot for each Reactor:Treatment combination the percentage of ASVs that are specific to the *E.coli* strain HV292.1 against the `Day_from_Inoculum`. Color of points and lines indicates the particular ASV or the sum of both. The shape of the point tells us whether the percentage was 0 or not. The first vertical gray bar indicates the start of connection to second-stage reactor. The second vertical gray bar indicates the start of treatment. The data was transformed with $log_{10}(1+x)$, where $x$ indicates percentage of strain-specific ASVs. This particular transformation was chosen in order to handle the percantages that were 0 in the data. 
```{r,fig.width=10, fig.height=12}
ggplot(data=plot.df1,mapping=aes(color=ASV,shape=Percent_ASV==0))+
          geom_point(mapping=aes(x=Day_from_Inoculum,y=log10(1+Percent_ASV)),alpha=0.8)+
           geom_path(mapping=aes(x=Day_from_Inoculum,y=log10(1+Percent_ASV),group=ASV),alpha=0.8) +
         facet_wrap(Reactor_Treatment~Enrichment,ncol=2) -> plot_HV292

plot_HV292 + geom_vline(xintercept = c(23,39),color="black",alpha=0.4)+ggtitle("Pecentage of HV292 specific ASV") + theme_light()
```

### Plotting CCUG specific ASVs:

Creating a the dataframe for plotting with all the columns we need (the differnt aesthetics)

```{r}
sample.data.cont %>% 
  filter(Reactor %in% c("IR1","CR","TR1", "TR2", "TR3", "TR4","TR5","TR6")) %>% arrange(Day_from_Inoculum) %>% 
  select(Sample_description,Reactor,Treatment,Reactor_Treatment,Experiment,Enrichment,Day_from_Inoculum,ASV84,ASV86,ASV_CCUG) -> plot.df2

  
plot.df2 %>% pivot_longer(cols = starts_with("ASV"),names_to = "ASV",values_to = "Percent_ASV" ) -> plot.df2

head(plot.df2)
```
We plot for each Reactor:Treatment combination the percentage of ASVs that are specific to the *Enterococcus faecium* strain, CCUG-59168, against the `Day_from_Inoculum`. Color of points and lines indicates the particular ASV or the sum of both. The shape of the point tells us whether the percentage was 0 or not. The first vertical gray bar indicates the start of connection to second-stage reactor. The second vertical gray bar indicates the start of treatment. The second vertical gray bar indicates the start of treatment. The data was transformed with $log_{10}(1+x)$, where $x$ indicates percentage of strain-specific ASVs. This particular transformation was chosen in order to handle the percantages that were 0 in the data. 

```{r,fig.width=10, fig.height=12}
ggplot(data=plot.df2,mapping=aes(color=ASV,shape=Percent_ASV==0))+
          geom_point(mapping=aes(x=Day_from_Inoculum,y=log10(1+Percent_ASV)),alpha=0.6)+
           geom_path(mapping=aes(x=Day_from_Inoculum,y=log10(1+Percent_ASV),group=ASV),alpha=0.6) +
         facet_wrap(Reactor_Treatment~Enrichment,ncol=2) -> plot_CCUG

plot_CCUG + geom_vline(xintercept = c(23,39),color="black",alpha=0.4) + ggtitle("Pecentage of CCUG specific ASV") + theme_light()
```


```{r,eval=FALSE,include=FALSE}
ggplot(data=plot.df,mapping=aes(Group=Reactor),position_dodge(0.2))+
  geom_point( mapping=aes(x=Day_from_Inoculum,y=Percent_strain_specific_ASV,fill=Enrichment,color=ASV,shape=is.0),alpha=0.6)  + geom_path(data=plot.df %>%  filter(Enrichment=="NotEnriched"),mapping=aes(x=Day_from_Inoculum,y=Percent_strain_specific_ASV,group=ASV,color=ASV),size = 0.3)+
  geom_path(data=plot.df %>%  filter(Enrichment=="Enriched"),mapping=aes(x=Day_from_Inoculum,y=Percent_strain_specific_ASV,group=ASV,color=ASV),size = 0.3)+
scale_fill_manual(values=c('black',NA),guide = guide_legend(override.aes = list(shape = 21)))  +
scale_shape_manual(values=c(21,22,23)) + facet_wrap(~Reactor,ncol=2)+
 scale_y_log10()


```

```{r,eval=FALSE,include=FALSE}
plot.df %>% filter(Enrichment=="Enriched",Reactor=="CR") %>% 
  ggplot(mapping=aes(x=Day_from_Inoculum,y=Percent_strain_specific_ASV,group=ASV,color=ASV)) +
  geom_point(mapping = aes(shape=is.0,fill=Treatment)) + scale_fill_manual(values=c('black',NA),guide = guide_legend(override.aes = list(shape = 21))) + scale_shape_manual(values=c(21,22))
```

```{r}
for(i in HV292_ASVs){
  pri<-otu.table[otu.table$ASV==i,-1]
  pri
}
  

```

