---
title: "Resistome data exploration"
author: "Hannah Li Hägi & Florentin Constancias "
date: " `r format(Sys.time(), '%B %d, %Y')` "
output: 
  html_document: 
    toc: yes
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# knitr::opts_chunk$set(dev = "ragg_png") # https://ragg.r-lib.org/index.html
knitr::opts_chunk$set(fig.width = 6) #https://r4ds.had.co.nz/graphics-for-communication.html#figure-sizing
knitr::opts_chunk$set(fig.asp = 0.618)
knitr::opts_chunk$set(fig.show = "hold")
knitr::opts_chunk$set(fig.show = "70%")
knitr::opts_chunk$set(fig.align = "center")
```

```{r}

rm(list= ls())
# gc()
require("tidyverse")
require('speedyseq')


'%!in%' <- function(x,y)!('%in%'(x,y))
source("https://raw.githubusercontent.com/fconstancias/DivComAnalyses/master/R/phyloseq_varia.R")  # source the functions from github.

```


```{r}
export_ppt_width = NULL
export_ppt_height = NULL
out_pptx = here::here("output/plots_2.pptx")
```

# Load data:

## 16S:

```{r}
"data/processed/16S/ps_silva_dada2_human_chicken_all_fct_polyferms.RDS" %>% 
  here::here() %>% 
  readRDS() %>% 
  filter_taxa(function(x) sum(x > 0) > 0, TRUE) -> ps_16S
```

## Export metadata:

To be merged with phyloseq data. TODO: Would be better to do that before in the gene phyloseq data generation.

```{r}
ps_16S %>% 
  sample_data() %>% 
  data.frame() %>% 
  filter(!is.na(metagenomic_sample_name)) %>% 
  rownames_to_column('tmp') %>% 
  select(-tmp) -> metadata
```

## phyloseq object of SQM + Pathofact + bins + ... with Num_Gi_pc quantif:

```{r}
# "data/processed/exp1_full_gene_catalog_phyloseq.RDS" %>% 
#   here::here() %>% 
#   readRDS() %>% 
#   .[["Num_Gi_pc"]] -> ps_coass
# 
# ps_coass
```

## extract on only AMR gene catalogue - qauntification on all gene catalogue:

```{r}
# ps_coass %>% 
#   subset_taxa(!is.na(PathoFact_AMR_ARG)) %>% 
#   subset_taxa(PathoFact_AMR_AMR_category != "unclassified") -> ps_AMR
```

```{r}
"data/processed/exp1_ps_AMR.RDS" %>% 
  here::here() %>% 
  readRDS() -> ps_AMR
```

```{r}
# ps_AMR %>% 
#   tax_table() %>% 
#   data.frame() %>% 
#   glimpse()
```

I hvae to do that otherwise it is not stored as proper chr when filtering... no clue what is happening here.


```{r}
ps_AMR %>%
  tax_table() %>%
  data.frame() %>%
  mutate(PathoFact_AMR_Resistance_mechanism_multi = ifelse(grepl(";", PathoFact_AMR_Resistance_mechanism), "multi-mech", PathoFact_AMR_Resistance_mechanism)) %>%
  mutate(PathoFact_AMR_AMR_sub_class_multi = ifelse(grepl(";", PathoFact_AMR_AMR_sub_class), "multidrug", PathoFact_AMR_AMR_sub_class)) %>%
  mutate(ANVIO_CONCOCT_HQ_clean = ifelse(PathoFact_AMR_MGE_prediction == "plasmid", NA, ANVIO_CONCOCT_HQ)) %>%
  mutate(GTDB_tax_CONCOCT_clean = ifelse(PathoFact_AMR_MGE_prediction == "plasmid", NA, GTDB_tax_CONCOCT)) %>%
  mutate(PathoFact_AMR_MGE_prediction_clean = ifelse(!is.na(ANVIO_CONCOCT_HQ), "MAG", PathoFact_AMR_MGE_prediction)) %>%
  mutate(PathoFact_AMR_MGE_prediction_clean_2 = case_when(grepl("ambiguous", PathoFact_AMR_MGE_prediction_clean) ~ "ambiguous", TRUE ~ PathoFact_AMR_MGE_prediction_clean)) %>%
  mutate(PathoFact_AMR_MGE_prediction_clean_2 = fct_relevel(PathoFact_AMR_MGE_prediction_clean_2, c("MAG", "chromosome", "plasmid", "phage", "ambiguous", "unclassified"))) %>%
  # mutate(Treatment = factor(Treatment, levels = c)) %>%
  # mutate(Day_of_Treatment = fct_relevel(Day_of_Treatment, c(NA, "-3", "-1", "1" , "3", "4", "5","7", "30", "47", "48")))  %>% 
  
  as.matrix() -> tax_table(ps_AMR)
```



```{r}
# ps_AMR %>% 
#   tax_table() %>% 
#   colnames()
```

# Define astetics:

## Treatment colors:

<https://colorbrewer2.org/#type=diverging&scheme=RdYlBu&n=6>
```{r, out.width = "50%",fig.width = 4.3}  
treat_col <- c("#3CB371", "#7C8181",'#d73027','#fc8d59','#fee090','#4575b4','#91bfdb', '#e0f3f8')

names(treat_col) <- get_variable(ps_AMR, "Treatment") %>%  
  levels()

treat_col %>% 
  scales::show_col(cex_label = 0.5)
```

## Drug classes colors:

```{r, out.width = "50%",fig.width = 4.3}  
ps_AMR %>% 
  generate_color_palette(var = "PathoFact_AMR_AMR_sub_class_multi", seed = 280386) -> drug_classe_multi
```


```{r, out.width = "50%",fig.width = 4.3}  
# ps_AMR %>% 
#   generate_color_palette(var = "Model_type", pal = "jco") -> model_type

```

```{r, out.width = "50%",fig.width = 4.3}  
ps_AMR %>% 
  generate_color_palette(var = "PathoFact_AMR_Resistance_mechanism_multi",  pal = "npg") -> resistance_type
```

```{r, out.width = "50%",fig.width = 4.3}  
ps_AMR %>% 
  generate_color_palette(var = "PathoFact_AMR_MGE_prediction_clean_2", seed = 72) -> MGE
```

## Others:

<https://r-graphics.org/recipe-scatter-shapes>
```{r}
fermentaion_shape <- c(22, 21)
names(fermentaion_shape) <- sample_data(ps_AMR)$Fermentation %>%  levels()

model_shape <- c(22, 21)
names(model_shape) <- sample_data(ps_AMR)$Model %>%  levels()

antibio_shape <- c(22, 21)
names(antibio_shape) <- sample_data(ps_AMR)$Antibiotic %>%  levels()


conc_stroke <- c(1, 2, 3, 4, 5)
names(conc_stroke) <- sample_data(ps_AMR)$Antibiotic_mg.mL %>%  levels()
# scale_discrete_manual(
#   aesthetics = "stroke",
#   values = c(`A` = 2, `B` = 1, `C` = 2)
# )

theme_set(theme_classic() + theme(legend.position = 'bottom'))
```

save all ascetics parameters as Rdata which can be loaded afterwards:

```{r}
save(fermentaion_shape, 
     dataframe2, dataframe3, file = "location.filename.Rdata")

Then you can

load("location.filename.RData") 

```

<https://r4ds.had.co.nz/graphics-for-communication.html#figure-sizing>

<https://www.tidyverse.org/blog/2020/08/taking-control-of-plot-scaling/>


## Others:
```{r}
ps_AMR %>% 
  tax_table() %>% 
  data.frame() %>% 
  rownames_to_column("ORF_ID") -> tax_mapping
```


```{r}
ps_AMR %>% 
  sample_data() %>% 
  data.frame() %>% 
  rownames_to_column("sample_id") -> metadata_mapping
```


```{r}
tax_mapping %>% 
  select(c("PathoFact_AMR_AMR_sub_class", "PathoFact_AMR_AMR_sub_class_multi",
           "PathoFact_AMR_ARG")) %>% 
  distinct(PathoFact_AMR_ARG, .keep_all = TRUE) %>% 
  add_count(PathoFact_AMR_ARG) %>% 
  arrange(-n) %>% 
  column_to_rownames('PathoFact_AMR_ARG')

```

```{r}
# ps_AMR %>% 
#   physeq_sel_tax_table(c("PathoFact_AMR_AMR_sub_class", "PathoFact_AMR_AMR_sub_class_multi", "rgi_CARD_Best_Hit_ARO",
#                          "PathoFact_AMR_ARG")) %>% 
ps_AMR %>% 
  physeq_sel_tax_table(c("PathoFact_AMR_ARG")) %>% 
  speedyseq::tax_glom(., taxrank = "PathoFact_AMR_ARG") %>%
  # physeq_simplify_tax("PathoFact_AMR_ARG")
  psmelt() %>% 
  group_by(Sample,PathoFact_AMR_ARG) %>% 
  add_count() %>% 
  ungroup() %>% 
  select(-OTU, -Sample:-Treatment_Dose) %>% 
  arrange(-n)
```

```{r}
ps_AMR %>%
  physeq_sel_tax_table(c("PathoFact_AMR_AMR_sub_class", "PathoFact_AMR_AMR_sub_class_multi",
                         "PathoFact_AMR_ARG")) %>%
  speedyseq::tax_glom(., taxrank = "PathoFact_AMR_ARG") %>% 
  # tax_table() %>% 
  # data.frame()
  psmelt() %>% 
  group_by(Sample,PathoFact_AMR_ARG) %>% 
  add_count() %>% 
  ungroup() %>% 
  select(-OTU, -Sample:-Treatment_Dose) %>% 
  arrange(-n)
```



```{r}
ps_AMR %>% 
  physeq_sel_tax_table(c("Contig_ID","ANVIO_CONCOCT_HQ_clean", "PathoFact_AMR_AMR_sub_class", "PathoFact_AMR_AMR_sub_class_multi", 
                         "GTDB_tax_CONCOCT_clean", 
                         "PathoFact_AMR_MGE_prediction_clean", "rgi_CARD_Best_Hit_ARO",
                         "PathoFact_AMR_ARG"))  %>% 
  # speedyseq::tax_glom(., taxrank = "PathoFact_AMR_MGE_prediction_clean")  %>% tax_table()
  # speedyseq::tax_glom(., taxrank = "PathoFact_AMR_Resistance_mechanism_multi") %>%   #%>% tax_table()
  speedyseq::tax_glom(., taxrank = "PathoFact_AMR_ARG") -> phyloseq_AMR_MAG_MGE

phyloseq_AMR_MAG_MGE
```

```{r}
ps_AMR %>% 
  tax_table() %>% 
  data.frame()
```

```{r}
phyloseq_AMR_MAG_MGE %>% 
  tax_table() %>% 
  data.frame()
```


Looks good, let's go.


## Reproducing https://www.nature.com/articles/s41467-022-29919-9.pdf


<file:///Users/fconstan/Documents/GitHub/NRP72-FBT/Figures/Figures_beta.html>

### Fig. 2a

```{r}
phyloseq_AMR_MAG_MGE %>% 
  # physeq_sel_tax_table(c("PathoFact_AMR_AMR_sub_class", "PathoFact_AMR_AMR_sub_class_multi", 
  #                        "PathoFact_AMR_MGE_prediction_clean", "rgi_CARD_Best_Hit_ARO",
  #                        "PathoFact_AMR_ARG"))  %>% 
  physeq_sel_tax_table(c("PathoFact_AMR_AMR_sub_class", "PathoFact_AMR_AMR_sub_class_multi", "PathoFact_AMR_ARG"))  %>% 
  # physeq_sel_tax_table(c("PathoFact_AMR_ARG"))  %>% 
  speedyseq::tax_glom(., taxrank = "PathoFact_AMR_ARG")  %>% 
  filter_taxa(function(x) sum(x > 0) > 0, TRUE) %>% 
  speedyseq::psmelt() %>% 
  select(-Lactose_mM:-Total_SCFA_mM) %>% 
  group_by(Sample) %>% 
  summarise(sum_abundance = sum(Abundance), .groups = "drop") %>% 
  mutate(sum_abundance = na_if(sum_abundance, 0)) %>%
  left_join(metadata_mapping,
            by = c("Sample" = "sample_id")) %>% 
  # filter(Model == "Chicken") %>% 
  mutate(Day_of_Treatment_num = as.double(Day_of_Treatment)) %>% 
  # mutate(Day_of_Treatment = fct_relevel(Day_of_Treatment, c(NA, "-3", "-1", "1" , "3", "4", "5","7", "30", "47", "48")))  %>% 
  mutate(Antibiotic_mg.mL = factor(Antibiotic_mg.mL, levels = c(NA, 20, 200,90, 600))) %>%
  ggplot(aes_string(x = "Day_of_Treatment", y = "sum_abundance", color = "Treatment", fill  = "Treatment", shape = "Antibiotic_mg.mL")) + #shape = Fermentation
  geom_point(size = 2.5) + 
  geom_line(linetype = 2,  size = 0.5,  aes_string(group = "interaction(Model, Reactor_Treatment_Dose)")) +
  labs(y = "sum Rnum_Gi", x = "Day_of_Treatment") +
  scale_color_manual(name = "", values = treat_col,
                     na.value = "black") +
  scale_fill_manual(name = "", values = treat_col,
                    na.value = "black") +
  scale_shape_manual(name = "" ,values = c(15,16,18,19), na.value =  17) +
  # scale_alpha_continuous(name = "", range=c(0.6, 1), na.value =  1) +
  geom_vline(xintercept = 0, size = 0.5, linetype = 1, color = "grey60") +
  theme_light() + 
  scale_y_continuous(labels = function(x) format(x, scientific = TRUE)) -> p1

p1 + facet_grid(as.formula(paste0("Antibiotic_mg.mL ~  Model  ")), scales = "free", space = "fixed", drop = TRUE) -> ptmp

ptmp

ptmp %>%
  export::graph2ppt(append = TRUE, width = 8, height = 4,
                    file = out_pptx)

p1 + facet_grid(as.formula(paste0(". ~  Model  ")), scales = "free", space = "fixed", drop = TRUE) -> ptmp

ptmp

ptmp %>%
  export::graph2ppt(append = TRUE, width = 8, height = 4,
                    file = out_pptx)
```

```{r}
# plot_univ_NRP72 <- function(df, x = "Day_of_Treatment", y = "value",  point_size = 2.5, color = "Treatment", fill  = "Treatment", shape = "Antibiotic", alpha = "Antibiotic_mg.mL", facet_formula = paste0("alphadiversiy ~  Model "), ylab = "Resistome alpha-diversity", xlab = "Days (Treatment)", measures = c("Observed"), path_group = "interaction(Model, Fermentation, Reactor_Treatment)"){
#   
#   df %>% 
#     pivot_longer(cols = all_of(measures), values_to = "value", names_to = 'alphadiversiy', values_drop_na  = TRUE) %>%
#     mutate(alphadiversiy = fct_relevel(alphadiversiy, measures)) -> df_ready
#   
#   df_ready %>% 
#     ggplot(aes_string(x = x, y = y, color = color, fill = fill, shape = shape, alpha = alpha)) + #shape = Fermentation
#     geom_point(size = point_size) + 
#     geom_line(linetype = 2,  size = 0.5,  aes_string(group = path_group)) +
#     labs(y = ylab, x = xlab) +
#     scale_color_manual(name = "", values = treat_col,
#                        na.value = "black") +
#     scale_fill_manual(name = "", values = treat_col,
#                       na.value = "black") +
#     facet_grid(as.formula(facet_formula), scales = "free", space = "fixed", drop = TRUE) +
#     scale_shape_manual(name = "" ,values = antibio_shape, na.value =  17) +
#     scale_alpha_discrete(name = "", range=c(0.6, 1), na.value =  0.6) + 
#     geom_vline(xintercept = 0, size = 0.5, linetype = 1, color = "grey60") -> p_alpha_both
#   
#   return(p_alpha_both)
# }
```

### Heatmap:

file:///Users/fconstan/Documents/GitHub/NRP72-FBT/tmp/human_resistome_visualization.html

```{r}
# phyloseq_AMR_MAG_MGE %>% 
#   # physeq_sel_tax_table(c("PathoFact_AMR_AMR_sub_class", "PathoFact_AMR_AMR_sub_class_multi", 
#   #                        "PathoFact_AMR_MGE_prediction_clean", "rgi_CARD_Best_Hit_ARO",
#   #                        "PathoFact_AMR_ARG"))  %>% 
#   physeq_sel_tax_table(c("PathoFact_AMR_ARG"))  %>% 
#   # physeq_sel_tax_table(c("PathoFact_AMR_ARG"))  %>% 
#   speedyseq::tax_glom(., taxrank = "PathoFact_AMR_ARG") %>% 
#   speedyseq::psmelt() %>% 
#   # left_join(tax_mapping,
#   #           by = c("Best_Hit_ARO" = "Best_Hit_ARO"),
#   #           suffix = c("_x", "")) %>% 
#   # mutate(Sample = fct_reorder(Sample, Model,Treatment,Fermentation,Antibiotic_mg.mL, Day_of_Treatment )) %>%
#   mutate(Abundance = na_if(Abundance, 0)) %>%
#   mutate(logTPM = log10(Abundance + 1)) -> pheatmap_df
# 
# 
# pheatmap_df %>% 
#   # replace(is.na(.), 0)  %>% 
#   # select(Sample, Best_Hit_ARO, logTPM) %>% 
#   group_by(Sample, PathoFact_AMR_ARG) %>%
#   pivot_wider(names_from =  Sample, PathoFact_AMR_ARG,
#               values_from = logTPM) %>% 
#   column_to_rownames("PathoFact_AMR_ARG") -> pheatmap_samples
# 
# pheatmap_samples %>% 
#   replace(is.na(.), 0)  %>% 
#   t() %>% 
#   vegan::vegdist(na.rm = TRUE) %>% 
#   hclust() -> clust_t
# 
# # pheatmap_samples %>% 
# #     replace(is.na(.), 0)  %>%
# #     vegdist(na.rm = TRUE) %>% 
# #   # cor() %>% 
# #   hclust() -> clust
# 
# pheatmap_samples %>% 
#   pheatmap::pheatmap(cluster_rows = FALSE ,
#                      cluster_cols = clust_t,
#                      na_col = "transparent",
#                      annotation_row = tax_mapping %>% rownames_to_column('tmp_col') %>% 
#                        filter(PathoFact_AMR_ARG %in% rownames(pheatmap_samples)) %>% 
#                        distinct(PathoFact_AMR_ARG, .keep_all = TRUE) %>% 
#                        column_to_rownames("PathoFact_AMR_ARG") %>% 
#                        select(PathoFact_AMR_Resistance_mechanism_multi, PathoFact_AMR_AMR_sub_class_multi),
#                      # annotation_colors = list("PathoFact_AMR_AMR_sub_class_multi" = col,
#                      #                          "Treatment" = treat_col),
#                      annotation_col = pheatmap_df %>% 
#                        select(Model,Treatment,Fermentation,Antibiotic_mg.mL, Day_of_Treatment)) -> p_pheat
# 
# 
# p_pheat %>% 
# print()
```

Histogram:

AMR drug classes

```{r}
# https://newbedev.com/how-to-control-ordering-of-stacked-bar-chart-using-identity-on-ggplot2

phyloseq_AMR_MAG_MGE %>% 
  # physeq_sel_tax_table(c("PathoFact_AMR_AMR_sub_class", "PathoFact_AMR_AMR_sub_class_multi", 
  #                        "PathoFact_AMR_MGE_prediction_clean", "rgi_CARD_Best_Hit_ARO",
  #                        "PathoFact_AMR_ARG"))  %>% 
  physeq_sel_tax_table(c("PathoFact_AMR_AMR_sub_class", "PathoFact_AMR_AMR_sub_class_multi","PathoFact_AMR_ARG"))  %>% 
  # physeq_sel_tax_table(c("PathoFact_AMR_ARG"))  %>% 
  speedyseq::tax_glom(., taxrank = "PathoFact_AMR_ARG")  %>% 
  filter_taxa(function(x) sum(x > 0) > 0, TRUE) %>% 
  speedyseq::psmelt() %>% 
  select(-Lactose_mM:-Total_SCFA_mM) %>% 
  # left_join(tax_mapping,
  #         by = c("PathoFact_AMR_ARG" = "PathoFact_AMR_ARG"),
  #         suffix = c("_x", "")) %>%
  mutate(Abundance = na_if(Abundance, 0)) %>%
  mutate(logTPM = log10(Abundance + 1))  %>% 
  # physeq_AMRgn %>%
  #   microbiome::transform(transform = "log10p") %>% 
  #   speedyseq::psmelt() %>% 
  #   left_join(tax_mapping,
  #             by = c("Best_Hit_ARO" = "Best_Hit_ARO"),
  #             suffix = c("_x", "")) %>%
  #   mutate(Abundance = na_if(Abundance, 0)) %>%
  #   mutate(logTPM = log10(Abundance + 1))  %>% 
  # separate_rows(., Resistance_Mechanism,sep = '; ',convert = TRUE) %>% 
  ggplot(aes(x = as.factor(Day_of_Treatment), y = Abundance, fill = reorder(PathoFact_AMR_AMR_sub_class_multi, Abundance, FUN=median, na.rm = TRUE))) +   
  geom_bar( stat = "identity", position = "fill", colour="black", size=0.0125) +
  facet_grid(Model ~ Treatment + Antibiotic_mg.mL , scales = "free", space = "free_x") +
  ggtitle("ARG Type Abundance per Sample") +
  xlab("Day (Treatment)")  +
  ylab("Proportion") + 
  theme_light() +
  ggpubr::rotate_x_text(60) -> p_mec

p_mec %>%
  ggpubr::set_palette(drug_classe_multi)-> p_mec

p_mec %>%
  export::graph2ppt(append = TRUE, width = 12, height = 8,
                    file = out_pptx)
```

```{r}
# https://newbedev.com/how-to-control-ordering-of-stacked-bar-chart-using-identity-on-ggplot2

phyloseq_AMR_MAG_MGE %>% 
  # physeq_sel_tax_table(c("PathoFact_AMR_AMR_sub_class", "PathoFact_AMR_AMR_sub_class_multi", 
  #                        "PathoFact_AMR_MGE_prediction_clean", "rgi_CARD_Best_Hit_ARO",
  #                        "PathoFact_AMR_ARG"))  %>% 
  physeq_sel_tax_table(c("PathoFact_AMR_AMR_sub_class_multi","PathoFact_AMR_ARG"))  %>% 
  # physeq_sel_tax_table(c("PathoFact_AMR_ARG"))  %>% 
  speedyseq::tax_glom(., taxrank = "PathoFact_AMR_ARG")  %>% 
  filter_taxa(function(x) sum(x > 0) > 0, TRUE) %>% 
  speedyseq::psmelt() %>% 
  select(-Lactose_mM:-Total_SCFA_mM) %>% 
  # left_join(tax_mapping,
  #         by = c("PathoFact_AMR_ARG" = "PathoFact_AMR_ARG"),
  #         suffix = c("_x", "")) %>%
  mutate(Abundance = na_if(Abundance, 0)) %>%
  mutate(logTPM = log10(Abundance + 1))  %>% 
  # physeq_AMRgn %>%
  #   microbiome::transform(transform = "log10p") %>% 
  #   speedyseq::psmelt() %>% 
  #   left_join(tax_mapping,
  #             by = c("Best_Hit_ARO" = "Best_Hit_ARO"),
  #             suffix = c("_x", "")) %>%
  #   mutate(Abundance = na_if(Abundance, 0)) %>%
  #   mutate(logTPM = log10(Abundance + 1))  %>% 
  # separate_rows(., Resistance_Mechanism,sep = '; ',convert = TRUE) %>% 
  ggplot(aes(x = as.factor(Day_of_Treatment), y = Abundance, fill = PathoFact_AMR_AMR_sub_class_multi)) +   
  geom_bar( stat = "identity", colour="black", size=0.0125) +
  facet_grid(Model ~ Treatment + Antibiotic_mg.mL , scales = "free", space = "free_x") +
  ggtitle("ARG Type Abundance per Sample") +
  xlab("Day (Treatment)")  +
  ylab("Rnum_Gi") + 
  theme_light() +
  ggpubr::rotate_x_text(60) -> p_mec

p_mec %>%
  ggpubr::set_palette(drug_classe_multi)-> p_mec

p_mec

p_mec %>%
  export::graph2ppt(append = TRUE, width = 12, height = 8,
                    file = out_pptx)
```

Genetic material

```{r}
# https://newbedev.com/how-to-control-ordering-of-stacked-bar-chart-using-identity-on-ggplot2

phyloseq_AMR_MAG_MGE %>% 
  physeq_sel_tax_table(c("PathoFact_AMR_MGE_prediction_clean",
                         # "PathoFact_AMR_AMR_sub_class_multi",
                         "PathoFact_AMR_ARG"))  %>%
  # physeq_sel_tax_table(c("PathoFact_AMR_ARG"))  %>% 
  speedyseq::tax_glom(., taxrank = "PathoFact_AMR_ARG")  %>% 
  filter_taxa(function(x) sum(x > 0) > 0, TRUE) %>% 
  speedyseq::psmelt() %>% 
  mutate(PathoFact_AMR_MGE_prediction_clean_2 = case_when(grepl("ambiguous", PathoFact_AMR_MGE_prediction_clean) ~ "ambiguous", TRUE ~ PathoFact_AMR_MGE_prediction_clean)) %>% 
  mutate(PathoFact_AMR_MGE_prediction_clean_2 = fct_relevel(PathoFact_AMR_MGE_prediction_clean_2, c("MAG", "chromosome", "plasmid", "phage", "ambiguous", "unclassified"))) %>% 
  select(-Lactose_mM:-Total_SCFA_mM) %>% 
  # left_join(tax_mapping,
  #         by = c("PathoFact_AMR_ARG" = "PathoFact_AMR_ARG"),
  #         suffix = c("_x", "")) %>%
  mutate(Abundance = na_if(Abundance, 0)) %>%
  mutate(logTPM = log10(Abundance + 1))  %>% 
  # physeq_AMRgn %>%
  #   microbiome::transform(transform = "log10p") %>% 
  #   speedyseq::psmelt() %>% 
  #   left_join(tax_mapping,
  #             by = c("Best_Hit_ARO" = "Best_Hit_ARO"),
  #             suffix = c("_x", "")) %>%
  #   mutate(Abundance = na_if(Abundance, 0)) %>%
  #   mutate(logTPM = log10(Abundance + 1))  %>% 
  # separate_rows(., Resistance_Mechanism,sep = '; ',convert = TRUE) %>% 
  ggplot(aes(x = as.factor(Day_of_Treatment), y = Abundance, fill = PathoFact_AMR_MGE_prediction_clean_2)) +   
  geom_bar( stat = "identity", position = "fill", colour="black", size = 0.0125) +
  facet_grid(Model ~ Treatment + Antibiotic_mg.mL , scales = "free", space = "free_x") +
  ggtitle("ARG Type Abundance per Sample") +
  xlab("Day (Treatment)")  +
  ylab("Proportion") + 
  theme_light() +
  ggpubr::rotate_x_text(60) -> p_mec

p_mec %>%
  ggpubr::set_palette(MGE)-> p_mec

p_mec

p_mec %>%
  export::graph2ppt(append = TRUE, width = 12, height = 8,
                    file = out_pptx)
```

```{r}
phyloseq_AMR_MAG_MGE %>% 
  physeq_sel_tax_table(c("PathoFact_AMR_MGE_prediction_clean",
                         # "PathoFact_AMR_AMR_sub_class_multi",
                         "PathoFact_AMR_ARG"))  %>%
  # physeq_sel_tax_table(c("PathoFact_AMR_ARG"))  %>% 
  speedyseq::tax_glom(., taxrank = "PathoFact_AMR_ARG")  %>% 
  filter_taxa(function(x) sum(x > 0) > 0, TRUE) %>% 
  speedyseq::psmelt() %>% 
  mutate(PathoFact_AMR_MGE_prediction_clean_2 = case_when(grepl("ambiguous", PathoFact_AMR_MGE_prediction_clean) ~ "ambiguous", TRUE ~ PathoFact_AMR_MGE_prediction_clean)) %>% 
  mutate(PathoFact_AMR_MGE_prediction_clean_2 = fct_relevel(PathoFact_AMR_MGE_prediction_clean_2, c("MAG", "chromosome", "plasmid", "phage", "ambiguous", "unclassified"))) %>% 
  select(-Lactose_mM:-Total_SCFA_mM) %>% 
  # left_join(tax_mapping,
  #         by = c("PathoFact_AMR_ARG" = "PathoFact_AMR_ARG"),
  #         suffix = c("_x", "")) %>%
  mutate(Abundance = na_if(Abundance, 0)) %>%
  mutate(logTPM = log10(Abundance + 1))  %>% 
  # physeq_AMRgn %>%
  #   microbiome::transform(transform = "log10p") %>% 
  #   speedyseq::psmelt() %>% 
  #   left_join(tax_mapping,
  #             by = c("Best_Hit_ARO" = "Best_Hit_ARO"),
  #             suffix = c("_x", "")) %>%
  #   mutate(Abundance = na_if(Abundance, 0)) %>%
  #   mutate(logTPM = log10(Abundance + 1))  %>% 
  # separate_rows(., Resistance_Mechanism,sep = '; ',convert = TRUE) %>% 
  ggplot(aes(x = as.factor(Day_of_Treatment), y = Abundance, fill = PathoFact_AMR_MGE_prediction_clean_2)) +   
  geom_bar( stat = "identity", colour="black", size = 0.0125) +
  facet_grid(Model ~ Treatment + Antibiotic_mg.mL , scales = "free", space = "free_x") +
  ggtitle("ARG Type Abundance per Sample") +
  xlab("Day (Treatment)")  +
  # ylab("Proportion") + 
  theme_light() +
  ggpubr::rotate_x_text(60) -> p_mec

p_mec %>%
  ggpubr::set_palette(MGE)-> p_mec

p_mec

p_mec %>%
  export::graph2ppt(append = TRUE, width = 12, height = 8,
                    file = out_pptx)
```

```{r}
phyloseq_AMR_MAG_MGE %>% 
  physeq_sel_tax_table(c("PathoFact_AMR_ARG",
                         # "PathoFact_AMR_AMR_sub_class_multi",
                         "GTDB_tax_CONCOCT_clean"))  %>%
  subset_taxa(!is.na("GTDB_tax_CONCOCT_clean")) %>% 
  # physeq_sel_tax_table(c("PathoFact_AMR_ARG"))  %>% 
  speedyseq::tax_glom(., taxrank = "GTDB_tax_CONCOCT_clean")  %>% 
  filter_taxa(function(x) sum(x > 0) > 0, TRUE) %>% 
  microbiome::transform("compositional") %>% 
  speedyseq::psmelt() %>% 
  separate(GTDB_tax_CONCOCT_clean, into = c("domain", "phylum", "class", "order", "family", "genus", "species"), sep  = ";" , remove = FALSE) -> df
```

```{r}
df %>% 
  # filter(Model == "Human") %>% 
  filter(genus %in% c("g__Enterococcus", "g__Escherichia")) %>% 
  mutate(Abundance = na_if(Abundance, 0)) %>%
  mutate(logTPM = log10(Abundance + 1))  %>% 
  # separate_rows(., Resistance_Mechanism,sep = '; ',convert = TRUE) %>% 
  # ggplot(aes(x = as.factor(Day_of_Treatment), y = Abundance, fill = fct_reorder(Strain, logTPM))) +   
  ggplot(aes(x = as.factor(Day_of_Treatment), y = Abundance, fill = GTDB_tax_CONCOCT_clean)) +#reorder(genus, Abundance, FUN=median, na.rm = TRUE))) +
  geom_bar( stat = "identity", colour="black", size=0.025) +
  facet_grid(Model  ~ Treatment + Fermentation + Antibiotic_mg.mL , scales = "free", space = "free_x", drop = TRUE) +
  ggtitle("ARG Type Abundance per Sample") +
  xlab("Day (Treatment)")  +
  # ylab("Proportion") + 
  theme_light() +
  # scale_fill_manual(name = "", values = col_strains) +
  theme(legend.position = "bottom") +
  ggpubr::rotate_x_text(60) -> p_mec

# p_mec %>% 
#   ggpubr::set_palette("jco") -> p_mec

p_mec

p_mec %>%
  export::graph2ppt(append = TRUE, width = 12, height = 8,
                    file = out_pptx)
```



### Fig. 2b

```{r}
phyloseq_AMR_MAG_MGE %>% 
  physeq_sel_tax_table(c("PathoFact_AMR_AMR_sub_class_multi",
                         "PathoFact_AMR_ARG"))  %>%
  # physeq_sel_tax_table(c("PathoFact_AMR_ARG"))  %>% 
  speedyseq::tax_glom(., taxrank = "PathoFact_AMR_ARG")  %>% 
  filter_taxa(function(x) sum(x > 0) > 0, TRUE) %>% 
  speedyseq::psmelt() -> df  #%>% 
# filter(grepl('unclassified', PathoFact_AMR_AMR_sub_class_multi))  -> df

results <- vector("list", length(df$PathoFact_AMR_AMR_sub_class_multi %>%  unique()))
names(results) = c(df$PathoFact_AMR_AMR_sub_class_multi %>%  unique())

for (dd in df$PathoFact_AMR_AMR_sub_class_multi %>%  unique())
{
  print(dd)
  df %>% 
    filter(PathoFact_AMR_AMR_sub_class_multi == !!dd) %>% 
    select(-Lactose_mM:-Total_SCFA_mM) %>% 
    group_by(Sample, PathoFact_AMR_AMR_sub_class_multi) %>% 
    summarise(sum_abundance = sum(Abundance), .groups = "drop") %>% 
    mutate(sum_abundance = na_if(sum_abundance, 0)) %>%
    left_join(metadata_mapping,
              by = c("Sample" = "sample_id")) %>% 
    # filter(Model == "Chicken") %>% 
    # mutate(Day_of_Treatment_num = as.double(Day_of_Treatment)) %>% 
    # mutate(Day_of_Treatment = fct_relevel(Day_of_Treatment, c(NA, "-3", "-1", "1" , "3", "4", "5","7", "30", "47", "48")))  %>% 
    mutate(Antibiotic_mg.mL = factor(Antibiotic_mg.mL, levels = c(NA, 20, 200,90, 600))) %>%
    ggplot(aes_string(x = "Day_of_Treatment", y = "sum_abundance", color = "Treatment", fill  = "Treatment", shape = "Antibiotic_mg.mL")) + #shape = Fermentation
    geom_point(size = 2.5) + 
    geom_line(linetype = 2,  size = 0.5,  aes_string(group = "interaction(Model, Reactor_Treatment_Dose)")) +
    labs(y = paste0("sum Rnum_Gi ") , x = "Day_of_Treatment") +
    scale_color_manual(name = "", values = treat_col,
                       na.value = "black") +
    scale_fill_manual(name = "", values = treat_col,
                      na.value = "black") +
    scale_shape_manual(name = "" ,values = c(15,16,18,19), na.value =  17) +
    # scale_alpha_continuous(name = "", range=c(0.6, 1), na.value =  1) +
    geom_vline(xintercept = 0, size = 0.5, linetype = 1, color = "grey60") +
    ggtitle(dd) +
    theme_light() + 
    scale_y_continuous(labels = function(x) format(x, scientific = TRUE)) -> p1
  
  # p1 + facet_grid(as.formula(paste0("Antibiotic_mg.mL + PathoFact_AMR_AMR_sub_class_multi ~  Model  ")), scales = "fixed", space = "fixed", drop = TRUE)
  # 
  # p1 + facet_wrap(as.formula(paste0("PathoFact_AMR_AMR_sub_class_multi ~  Model  ")),scales = "free"
  #)
  results[[dd]] <- p1 + facet_grid(as.formula(paste0("Antibiotic_mg.mL ~  Model  ")),  scales = "free", space = "fixed", drop = TRUE)
  
  results[[dd]]  %>% 
    export::graph2ppt(append = TRUE, width = 8, height = 4,
                      file = out_pptx)
  
  print(p1 + facet_grid(as.formula(paste0(" Antibiotic_mg.mL ~  Model  ")), scales = "free", space = "fixed", drop = TRUE))
  print(p1 + facet_grid(as.formula(paste0("PathoFact_AMR_AMR_sub_class_multi ~  Model  ")), scales = "free", space = "fixed", drop = TRUE))
  
}
```



Vancomycin is a glycopeptide antibiotic used in the prophylaxis and treatment of infections caused by Gram-positive bacteria. Vancomycin inhibits the synthesis of peptidoglycan, the major component of the cell wall of gram-positive bacteria. Its mechanism of action is unusual in that it acts by binding precursors of peptidoglycan, rather than by interacting with an enzyme. -> glycopeptide antibiotic

```{r}
results$`glycopeptide antibiotic`

results$`glycopeptide antibiotic` %>% 
  export::graph2ppt(append = TRUE, width = 8, height = 4,
                    file = out_pptx)
```

Cefotaxime is a semisynthetic cephalosporin taken parenterally. It is resistant to most beta-lactamases and active against Gram-negative rods and cocci due to its aminothiazoyl and methoximino functional groups. -> cephalosporin

```{r}
results$cephalosporin 

results$cephalosporin %>% 
  export::graph2ppt(append = TRUE, width = 8, height = 4,
                    file = out_pptx)
```

### Fig. 2c

```{r}
phyloseq_AMR_MAG_MGE %>% 
  physeq_sel_tax_table(c("GTDB_tax_CONCOCT_clean",
                         "PathoFact_AMR_AMR_sub_class_multi",
                         "PathoFact_AMR_ARG"))  %>%
  # physeq_sel_tax_table(c("PathoFact_AMR_ARG"))  %>% 
  speedyseq::tax_glom(., taxrank = "PathoFact_AMR_ARG")  %>% 
  filter_taxa(function(x) sum(x > 0) > 0, TRUE) %>% 
  speedyseq::psmelt() %>% 
  # filter(!is.na(CONCOCT_bin )) %>% 
  separate(GTDB_tax_CONCOCT_clean, into = c("domain", "phylum", "class", "order", "family", "genus", "species"), sep  = ";" ) -> df

```


### Fig. 3

```{r}
phyloseq_AMR_MAG_MGE %>% 
  physeq_sel_tax_table(c("PathoFact_AMR_MGE_prediction_clean",
                         "PathoFact_AMR_AMR_sub_class_multi",
                         "PathoFact_AMR_ARG"))  %>%
  # physeq_sel_tax_table(c("PathoFact_AMR_ARG"))  %>% 
  speedyseq::tax_glom(., taxrank = "PathoFact_AMR_ARG")  %>% 
  filter_taxa(function(x) sum(x > 0) > 0, TRUE) %>% 
  speedyseq::psmelt() %>% 
  mutate(PathoFact_AMR_MGE_prediction_clean_2 = case_when(grepl("ambiguous", PathoFact_AMR_MGE_prediction_clean) ~ "ambiguous", TRUE ~ PathoFact_AMR_MGE_prediction_clean)) %>% 
  mutate(PathoFact_AMR_MGE_prediction_clean_2 = fct_relevel(PathoFact_AMR_MGE_prediction_clean_2, c("MAG", "chromosome", "plasmid", "phage", "ambiguous", "unclassified"))) -> df  #%>% 
# filter(grepl('unclassified', PathoFact_AMR_AMR_sub_class_multi))  -> df

results <- vector("list", length(df$PathoFact_AMR_MGE_prediction_clean_2 %>%  unique()))
names(results) = c(df$PathoFact_AMR_MGE_prediction_clean_2 %>%  unique())

for (dd in df$PathoFact_AMR_MGE_prediction_clean_2 %>%  unique())
{
  print(dd)
  df %>% 
    filter(PathoFact_AMR_MGE_prediction_clean_2 == !!dd) %>% 
    select(-Lactose_mM:-Total_SCFA_mM) %>% 
    group_by(Sample, PathoFact_AMR_MGE_prediction_clean_2) %>% 
    summarise(sum_abundance = sum(Abundance), .groups = "drop") %>% 
    mutate(sum_abundance = na_if(sum_abundance, 0)) %>%
    left_join(metadata_mapping,
              by = c("Sample" = "sample_id")) %>% 
    # filter(Model == "Chicken") %>% 
    mutate(Day_of_Treatment_num = as.double(Day_of_Treatment)) %>% 
    # mutate(Day_of_Treatment = fct_relevel(Day_of_Treatment, c(NA, "-3", "-1", "1" , "3", "4", "5","7", "30", "47", "48")))  %>% 
    mutate(Antibiotic_mg.mL = factor(Antibiotic_mg.mL, levels = c(NA, 20, 200,90, 600))) %>%
    ggplot(aes_string(x = "Day_of_Treatment", y = "sum_abundance", color = "Treatment", fill  = "Treatment", shape = "Antibiotic_mg.mL")) + #shape = Fermentation
    geom_point(size = 2.5) + 
    geom_line(linetype = 2,  size = 0.5,  aes_string(group = "interaction(Model, Reactor_Treatment_Dose)")) +
    labs(y = paste0("sum Rnum_Gi ") , x = "Day_of_Treatment") +
    scale_color_manual(name = "", values = treat_col,
                       na.value = "black") +
    scale_fill_manual(name = "", values = treat_col,
                      na.value = "black") +
    scale_shape_manual(name = "" ,values = c(15,16,18,19), na.value =  17) +
    # scale_alpha_continuous(name = "", range=c(0.6, 1), na.value =  1) +
    geom_vline(xintercept = 0, size = 0.5, linetype = 1, color = "grey60") +
    ggtitle( dd ) +
    theme_light() -> p1
  
  results[[dd]] <- p1 + facet_grid(as.formula(paste0("Antibiotic_mg.mL ~  Model  ")), scales = "free", space = "fixed", drop = TRUE)
  
  
  results[[dd]] %>% 
    export::graph2ppt(append = TRUE, width = 8, height = 4,
                      file = out_pptx)
  
  print(p1 + facet_grid(as.formula(paste0("Antibiotic_mg.mL  ~  Model  ")), scales = "free", space = "fixed", drop = TRUE))
}
# print(p1 + facet_grid(as.formula(paste0("PathoFact_AMR_AMR_sub_class_multi ~  Model  ")), scales = "free", space = "free", drop = TRUE))


```

```{r}
results
```

per MGE per class...


```{r}
phyloseq_AMR_MAG_MGE %>% 
  physeq_sel_tax_table(c("PathoFact_AMR_MGE_prediction_clean",
                         "PathoFact_AMR_AMR_sub_class_multi",
                         "PathoFact_AMR_ARG"))  %>%
  # physeq_sel_tax_table(c("PathoFact_AMR_ARG"))  %>% 
  speedyseq::tax_glom(., taxrank = "PathoFact_AMR_ARG")  %>% 
  filter_taxa(function(x) sum(x > 0) > 0, TRUE) %>% 
  speedyseq::psmelt() %>% 
  mutate(PathoFact_AMR_MGE_prediction_clean_2 = case_when(grepl("ambiguous", PathoFact_AMR_MGE_prediction_clean) ~ "ambiguous", TRUE ~ PathoFact_AMR_MGE_prediction_clean)) %>% 
  mutate(PathoFact_AMR_MGE_prediction_clean_2 = fct_relevel(PathoFact_AMR_MGE_prediction_clean_2, c("MAG", "chromosome", "plasmid", "phage", "ambiguous", "unclassified"))) -> df  #%>%
```


```{r}
for(MGE  in c("MAG", "chromosome", "plasmid", "phage", "ambiguous", "unclassified"))
{
  
  print(MGE)
  df %>% 
    filter(PathoFact_AMR_MGE_prediction_clean_2 == !!MGE) -> df_tmp
  
  results <- vector("list", length(df_tmp$PathoFact_AMR_AMR_sub_class_multi %>%  unique()))
  
  names(results) = c(df_tmp$PathoFact_AMR_AMR_sub_class_multi %>% unique())
  
  
  
  
  for (dd in df_tmp$PathoFact_AMR_AMR_sub_class_multi %>%  unique())
  {
    print(dd)
    df_tmp %>% 
      filter(PathoFact_AMR_AMR_sub_class_multi == !!dd) %>% 
      select(-Lactose_mM:-Total_SCFA_mM) %>% 
      group_by(Sample, PathoFact_AMR_AMR_sub_class_multi) %>% 
      summarise(sum_abundance = sum(Abundance), .groups = "drop") %>% 
      mutate(sum_abundance = na_if(sum_abundance, 0)) %>%
      left_join(metadata_mapping,
                by = c("Sample" = "sample_id")) %>% 
      # filter(Model == "Chicken") %>% 
      mutate(Day_of_Treatment_num = as.double(Day_of_Treatment)) %>% 
      # mutate(Day_of_Treatment = fct_relevel(Day_of_Treatment, c(NA, "-3", "-1", "1" , "3", "4", "5","7", "30", "47", "48")))  %>% 
      mutate(Antibiotic_mg.mL = factor(Antibiotic_mg.mL, levels = c(NA, 20, 200,90, 600))) %>%
      ggplot(aes_string(x = "Day_of_Treatment", y = "sum_abundance", color = "Treatment", fill  = "Treatment", shape = "Antibiotic_mg.mL")) + #shape = Fermentation
      geom_point(size = 2.5) + 
      geom_line(linetype = 2,  size = 0.5,  aes_string(group = "interaction(Model, Reactor_Treatment_Dose)")) +
      labs(y = paste0("sum Rnum_Gi ") , x = "Day_of_Treatment") +
      scale_color_manual(name = "", values = treat_col,
                         na.value = "black") +
      scale_fill_manual(name = "", values = treat_col,
                        na.value = "black") +
      scale_shape_manual(name = "" ,values = c(15,16,18,19), na.value =  17) +
      # scale_alpha_continuous(name = "", range=c(0.6, 1), na.value =  1) +
      geom_vline(xintercept = 0, size = 0.5, linetype = 1, color = "grey60") +
      ggtitle(paste0( dd, "- ", MGE )) +
      theme_light() -> p1
    
    # p1 + facet_grid(as.formula(paste0("Antibiotic_mg.mL + PathoFact_AMR_AMR_sub_class_multi ~  Model  ")), scales = "fixed", space = "fixed", drop = TRUE)
    # 
    # p1 + facet_wrap(as.formula(paste0("PathoFact_AMR_AMR_sub_class_multi ~  Model  ")),scales = "free"
    #)
    
    results[[dd]] <- p1 + facet_grid(as.formula(paste0("Antibiotic_mg.mL ~  Model  ")),  scales = "free", space = "fixed", drop = TRUE)
    
    results[[dd]] %>% 
      export::graph2ppt(append = TRUE, width = 8, height = 4,
                        file = out_pptx)
    
    # print(p1 + facet_grid(as.formula(paste0(" Antibiotic_mg.mL ~  Model  ")), scales = "fixed", space = "fixed", drop = TRUE))
    # 
    # print(p1 + facet_grid(as.formula(paste0("PathoFact_AMR_AMR_sub_class_multi ~  Model  ")), scales = "free", space = "free", drop = TRUE)
  }
  assign(MGE, results)
}
```


```{r}
MAG
```

```{r}
chromosome
```

```{r}
plasmid
```

```{r}
phage
```



```{r}
# phyloseq_AMR_MAG_MGE %>% 
#   physeq_sel_tax_table(c("PathoFact_AMR_MGE_prediction_clean",
#                          "PathoFact_AMR_AMR_sub_class_multi",
#                          "PathoFact_AMR_ARG"))  %>%
#   # physeq_sel_tax_table(c("PathoFact_AMR_ARG"))  %>% 
#   speedyseq::tax_glom(., taxrank = "PathoFact_AMR_ARG")  %>% 
#   filter_taxa(function(x) sum(x > 0) > 0, TRUE) %>% 
#   speedyseq::psmelt() %>% 
#   mutate(PathoFact_AMR_MGE_prediction_clean_2 = case_when(grepl("ambiguous", PathoFact_AMR_MGE_prediction_clean) ~ "ambiguous", TRUE ~ PathoFact_AMR_MGE_prediction_clean)) %>% 
#   mutate(PathoFact_AMR_MGE_prediction_clean_2 = fct_relevel(PathoFact_AMR_MGE_prediction_clean_2, c("MAG", "chromosome", "plasmid", "phage", "ambiguous", "unclassified"))) -> df  #%>% 
# # filter(grepl('unclassified', PathoFact_AMR_AMR_sub_class_multi))  -> df
# 
# results <- vector("list", length(df$PathoFact_AMR_AMR_sub_class_multi %>%  unique()))
# names(results) = c(df$PathoFact_AMR_AMR_sub_class_multi %>%  unique())
# 
# for (dd in df$PathoFact_AMR_MGE_prediction_clean_2 %>%  unique())
# {
#   print(dd)
#   df %>% 
#     filter(PathoFact_AMR_AMR_sub_class_multi == !!dd) %>% 
#     
#     results <- vector("list", length(df$PathoFact_AMR_MGE_prediction_clean_2 %>%  unique()))
#   names(results) = c(df$PathoFact_AMR_MGE_prediction_clean_2 %>%  unique())
#   
#   for (dd in df$PathoFact_AMR_MGE_prediction_clean_2 %>%  unique())
#   {
#     print(dd)
#     df %>% 
#       filter(PathoFact_AMR_MGE_prediction_clean_2 == !!dd) %>% 
#       select(-Lactose_mM:-Total_SCFA_mM) %>% 
#       group_by(Sample, PathoFact_AMR_MGE_prediction_clean_2) %>% 
#       summarise(sum_abundance = sum(Abundance), .groups = "drop") %>% 
#       mutate(sum_abundance = na_if(sum_abundance, 0)) %>%
#       left_join(metadata_mapping,
#                 by = c("Sample" = "sample_id")) %>% 
#       # filter(Model == "Chicken") %>% 
#       mutate(Day_of_Treatment_num = as.double(Day_of_Treatment)) %>% 
#       mutate(Day_of_Treatment = fct_relevel(Day_of_Treatment, c(NA, "-3", "-1", "1" , "3", "4", "5","7", "30", "47", "48")))  %>% 
#       mutate(Antibiotic_mg.mL = factor(Antibiotic_mg.mL, levels = c(NA, 20, 200,90, 600))) %>%
#       ggplot(aes_string(x = "Day_of_Treatment", y = "sum_abundance", color = "Treatment", fill  = "Treatment", shape = "Antibiotic_mg.mL")) + #shape = Fermentation
#       geom_point(size = 2.5) + 
#       geom_line(linetype = 2,  size = 0.5,  aes_string(group = "interaction(Model, Reactor_Treatment_Dose)")) +
#       labs(y = paste0("sum Rnum_Gi ") , x = "Day_of_Treatment") +
#       scale_color_manual(name = "", values = treat_col,
#                          na.value = "black") +
#       scale_fill_manual(name = "", values = treat_col,
#                         na.value = "black") +
#       scale_shape_manual(name = "" ,values = c(15,16,18,19), na.value =  17) +
#       # scale_alpha_continuous(name = "", range=c(0.6, 1), na.value =  1) +
#       geom_vline(xintercept = 0, size = 0.5, linetype = 1, color = "grey60") +
#       ggtitle( dd ) +
#       theme_light() -> p1
#     
#     results[[dd]] <- p1 + facet_grid(as.formula(paste0("Antibiotic_mg.mL ~  Model  ")), scales = "fixed", space = "fixed", drop = TRUE)
#     
#     print(p1 + facet_grid(as.formula(paste0("Antibiotic_mg.mL  ~  Model  ")), scales = "free", space = "fixed", drop = TRUE))
#   }
#   # print(p1 + facet_grid(as.formula(paste0("PathoFact_AMR_AMR_sub_class_multi ~  Model  ")), scales = "free", space = "free", drop = TRUE))
#   
# }
```



```{r}
devtools::session_info()
```

