---
title: "Merge human gene catalogue + quantification "
author: "Hannah Li Hägi & Florentin Constancias "
date: " `r format(Sys.time(), '%B %d, %Y')` "
output: 
  html_document: 
    toc: yes
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
rm(list= ls())
library('dplyr')
library("tidyverse")
library('here')
# library("SQMtools")
library("phyloseq")
library("readxl")


'%!in%' <- function(x,y)!('%in%'(x,y))

#source("https://raw.githubusercontent.com/fconstancias/DivComAnalyses/master/R/phyloseq_taxa_tests.R")
source("https://raw.githubusercontent.com/fconstancias/DivComAnalyses/master/R/phyloseq_normalisation.R")
source("https://raw.githubusercontent.com/fconstancias/DivComAnalyses/master/R/phyloseq_alpha.R")
source("https://raw.githubusercontent.com/fconstancias/DivComAnalyses/master/R/phyloseq_beta.R")
source("https://raw.githubusercontent.com/fconstancias/DivComAnalyses/master/R/phyloseq_heatmap.R")
```

# Define inputs:

## tables/ data:

```{r}
SQMR <- "data/raw/SQM/human1_SQM.rds"

metadata <- "data/raw/25.11.2021_metadata_updated.tsv"


RGI_CARD <- "data/raw/CARD/rgi_contig_human1.txt"
Viralverify <- "data/raw/mobilome/viralverify/01.human1_result_table.csv"
isescan <- "data/raw/mobilome/isescan/01.SqueezeHuman1.fasta.tsv"
PathoFact_AMR <- "data/processed/PathoFact/humann1/AMR/AMR_MGE_prediction_sample_l1000_report.tsv"
PathoFact_Tox <- "data/processed/PathoFact/humann1/Tox/Toxin_prediction_sample_l1000_report.tsv"
anvio_nobin_summary <- "data/processed/anvio/human1/summary_no_bins/bin_by_bin/all/all-gene_calls.txt"

DAS_collection <- "data/processed/anvio/human1/DAS.txt"
Manual_bins <- "data/processed/anvio/human1/MAGS/concoct_10_collection_refined_2.txt"

ANVIO_SUMMARY_DAS  <- "data/processed/anvio/human1/MAGS/SUMMARY/DAS/bins_summary.txt"
ANVIO_SUMMARY_CONCOCT <- "data/processed/anvio/human1/MAGS/SUMMARY/concoct_10_refined_2/bins_summary.txt"

GTDB_tax <- "data/processed/gtdb/GTDB_all.xlsx"
SQM_DAS <- "data/raw/SQM/19.human1.bintable"

# contigs.lgt.tsv <- "~/Documents/MSc_Thesis-NRP72/data/HGT/human.contigs.lgt.tsv"
# contigs.no_lgt.tsv <- "~/Documents/MSc_Thesis-NRP72/data/HGT/human.contigs.no_lgt.tsv"
# contigs.unclassified.tsv <- "~/Documents/MSc_Thesis-NRP72/data/HGT/human.contigs.unclassified.tsv"

```

## parameters:

```{r}
Model_type <- c("protein homolog model", "protein variant model", "protein overexpression model", "rRNA gene variant model")
Cut_Off <- c("Strict", "Perfect")
besid = 95
Percent_length = 0.80
```

# Gene/annotation - contig classification:

## Load SQM:
```{r}
SQMR %>% 
  here::here() %>% 
  readRDS() -> human
```

## Load metadata:

```{r}
metadata %>% 
  here::here() %>% 
  read_tsv() %>% 
  data.frame() %>%
  filter(!is.na(metagenomic_sample_name), Model == "Human") -> human_meta
```

Modify metadata: **this should be done once on the metadata**

```{r}
# create Period column
human_meta %>%
  mutate(Reactor_Treatment = base::replace(Reactor_Treatment, Reactor_Treatment == "TR_2CTX", "TR2_CTX")) %>%
  mutate(Reactor_Treatment_Dose = if_else(!is.na(`Antibiotic_mg.mL`), paste0(Reactor_Treatment, `Antibiotic_mg.mL`), Reactor_Treatment)) %>% 
  mutate(Treatment_Dose = if_else(!is.na(`Antibiotic_mg.mL`), paste0(Treatment, `Antibiotic_mg.mL`), Treatment)) %>%
  mutate(
    Period = case_when(
      Day_of_Treatment <= 0  ~ "pret",
      Day_of_Treatment > 0 ~ as.character(Day_of_Treatment)
    )) %>%
  mutate(Period = base::replace(Period, Treatment == "DONOR", "pret")) %>%
  mutate(Day_of_Treatment = addNA(Day_of_Treatment)) %>%
  column_to_rownames("metagenomic_sample_name") -> human_meta

#human_meta$Treatment_Dose <- factor(human_meta$Treatment_Dose, ordered = TRUE, 
#                                 levels = c("DONOR", "UNTREATED", "CTX+HV292.120", "CTX+HV292.1200",
#                                            "CTX20", "CTX200", "HV292.1", "VAN+CCUG5916890", "VAN+CCUG59168600",
#                                            "VAN90", "VAN600", "CCUG59168"))

```



## PathoFact:

### PathoFact_AMR:

```{r}
PathoFact_AMR %>% 
  here::here() %>% 
  read_tsv() %>% 
  select(-Contig, -Contig_ID, -ORF_ID) %>% 
  na_if("-") %>% 
  na_if("n/a") -> PathoFact_AMR

colnames(PathoFact_AMR) <- paste0("PathoFact_AMR_", colnames(PathoFact_AMR))


PathoFact_AMR %>% 
  rename(ORF_ID = PathoFact_AMR_ORF) -> PathoFact_AMR
```

### PathoFact_Tox:

```{r}
PathoFact_Tox %>% 
  here::here() %>% 
  read_tsv() %>% 
  select(-ORF_ID, -Number_of_hits, -Signal_peptide, -Toxin_prediction) %>% 
  na_if("-") %>% 
  na_if("n/a") -> PathoFact_Tox

colnames(PathoFact_Tox) <- paste0("PathoFact_Tox_", colnames(PathoFact_Tox))

PathoFact_Tox %>% 
  rename(PathoFact_Tox_Toxin_classification = PathoFact_Tox_Toxin_confidence_level,
         ORF_ID = PathoFact_Tox_ORF) -> PathoFact_Tox
```


## rgi CARD:

```{r}
# Resistome data:
RGI_CARD %>% 
  here::here() %>% 
  read_tsv() %>% 
  # read.delim() %>%
  select(-ORF_ID) %>%
  dplyr::filter(Model_type %in% Model_type ,
                Cut_Off %in% Cut_Off,
                Nudged %!in% c("TRUE"),
                Best_Identities > besid, # plot distribution of bestID vs percentage length ref and color bitscore and symbol class AMR
                `Percentage Length of Reference Sequence` > Percent_length) %>%
  separate(Contig, c("Megahit", "Contig", "Split"), sep = "_", fill = "right") %>%
  mutate(ORF_ID = paste(Megahit, Contig, paste(Start, Stop, sep = "-"), sep = "_")) %>%
  #mutate(Contig = paste(Megahit, Contig, sep = "_")) %>%
  select(-c(Megahit, Split, CARD_Protein_Sequence, Predicted_DNA, Predicted_Protein, Contig)) %>% 
  na_if("NA") %>% 
  na_if("n/a") %>% 
  na_if("<NA>") %>% 
  na_if("") %>% 
  select(-Start, -Stop, -Orientation) %>% 
  select(ORF_ID, everything()) -> card

colnames(card) <- paste0("rgi_CARD_", colnames(card))

card %>% 
  rename(ORF_ID = rgi_CARD_ORF_ID) -> card

card %>% 
  head() %>% 
  DT::datatable()
```

## Waffle HGT: 


```{r}
# TODO move to more appropriate section

# # HGT events
# contigs.lgt.tsv %>% 
#   read_tsv() %>% 
#   select(CONTIG_NAME, CALL) -> hgt
# 
# 
# contigs.no_lgt.tsv %>% 
#   read_tsv() %>% 
#   select(CONTIG_NAME, CALL) -> no_hgt
# 
# contigs.unclassified.tsv %>% 
#   read_tsv() %>% 
#   select(CONTIG_NAME, CALL) -> unclassified_hgt
# 
# hgt_events <- rbind(hgt, no_hgt, unclassified_hgt)
```

## Viralverify:

```{r}
# Viralverify 
Viralverify  %>%
  read_csv() %>% 
  rename(Contig_ID = `Contig name`,
         viralverify_Prediction = Prediction) %>% 
  select(Contig_ID, viralverify_Prediction) -> viralverify
```


# isescan:

```{r}
isescan %>%
  read_tsv() %>% 
  group_by(seqID) %>%
  add_count(cluster, name = "count_IS_contig") %>% 
  top_n(1, wt = `E-value4copy`) %>%
  top_n(1, wt = isBegin) %>%
  mutate(IS = "yes") %>%
  select(seqID:cluster, type, count_IS_contig) -> top_isescan

colnames(top_isescan) <- paste0("isescan_", colnames(top_isescan))

top_isescan %>% 
  rename(Contig_ID = isescan_seqID) -> top_isescan


top_isescan %>% 
  head() %>% 
  DT::datatable()
```


## Combine all:

```{r}
human$orfs$table %>%
  rownames_to_column("ORF_ID") %>% 
  data.frame() %>% 
  na_if("") %>% 
  # left_join(
  # human$orfs$tax %>% 
  #   as.data.frame() %>% 
  #   rownames_to_column("ORF_ID") %>%
  #   separate(ORF_ID, c("Megahit", "Contig", "Range"), sep = "_", fill = "right", remove = FALSE) %>%
  #   # mutate(Contig = as.numeric(Contig)) %>%
  #   mutate(Contig = paste(Megahit, Contig, sep = "_")) %>% 
  #   left_join(human$orfs$seqs %>%
  #     as.data.frame() %>%
  #   rownames_to_column("ORF_ID")) %>% 
  #   select(-Megahit) %>% 
#   rename(seq = ".",
#          Contig_ID = Contig,
#          superkingdom_ORF = superkingdom,
#          phylum_ORF = phylum,
#          class_ORF= class,
#          order_ORF = order,
#          family_orf = family,
#          genus_orf = genus,
#          species_orf = species)
# ) %>% 
select(-Method, -Molecule) %>% 
  rename(Contig_ID = `Contig.ID`,
         Tax_orf = Tax) %>% 
  select(-TPM.D.1:-Hits) -> orf

# orf %>% 
#   write_tsv("~/Desktop/unitedhuman1.tsv")
colnames(orf) <- paste0("SQM_", colnames(orf))

orf %>% 
  rename(Contig_ID = SQM_Contig_ID,
         ORF_ID = SQM_ORF_ID) -> orf

orf %>% 
  head() %>% 
  DT::datatable()
```

```{r}
human$contigs$table %>%
  rownames_to_column("Contig_ID") %>% 
  data.frame() %>% 
  rename(Tax_contig = `Tax`,
         Disparity_contig = Disparity,
         GC_perc_contig = GC.perc,
         Length_contig = Length,
         Num_genes_contig = Num.genes) %>% 
  select(-Coverage.D.1:-Raw.read.count.HV.24, -Bin.ID) -> contigs


colnames(contigs) <- paste0("SQM_", colnames(contigs))

contigs %>% 
  rename(Contig_ID = SQM_Contig_ID) -> contigs

contigs %>% 
  head() %>% 
  DT::datatable()
```


```{r}
orf %>% 
  left_join(card,
            by = c("ORF_ID" = "ORF_ID")) %>% 
  left_join(PathoFact_AMR) %>% 
  left_join(PathoFact_Tox) %>% 
  left_join(viralverify,
            by = c("Contig_ID" = "Contig_ID")) %>% 
  left_join(top_isescan) %>% 
  left_join(contigs,
            by = c("Contig_ID" = "Contig_ID")) -> orf_contig

orf_contig %>% 
  head() %>% 
  DT::datatable()
```
## Test anvio no bin:

```{r}
anvio_nobin_summary %>% 
  here::here() %>%  
  read_tsv() -> anvi_no_bin_annot
```

Link anvio gene id with SQM gene id:
```{r}
anvi_no_bin_annot %>% 
  mutate(orf_id = paste0(contig, "_", start+1, "-", stop)) %>% 
  select(-gene_callers_id) -> anvi_no_bin_annot

colnames(anvi_no_bin_annot) <- paste0("anvio_", colnames(anvi_no_bin_annot))

anvi_no_bin_annot %>% 
  select(anvio_orf_id, everything()) %>% 
  select(-anvio_contig, -anvio_start, -anvio_stop, -anvio_direction ,-anvio_dna_sequence) -> anvi_no_bin_annot
```

```{r}
# anvi_no_bin_annot$anvio_orf_id %>% 
#   head()
```

```{r}
# orf_contig %>%
#   filter(Contig_ID == "megahit_37") %>%
#   distinct(ORF_ID, .keep_all = TRUE)
```

```{r}
# orf_contig %>%
#   filter(ORF_ID == "megahit_37_124-318")
```
```{r}
# anvi_no_bin_annot %>% 
#     filter(anvio_contig == "megahit_37") %>%
#   distinct(anvio_orf_id, .keep_all = TRUE)
```

```{r}
# anvi_no_bin_annot %>% 
#     filter(anvio_orf_id == "megahit_37_124-318")
```

```{r}
orf_contig %>% 
  left_join(anvi_no_bin_annot,
            by = c("ORF_ID" = "anvio_orf_id")) -> orf_contig

```

```{r}
# orf_contig %>%
#   filter(anvio_contig == "megahit_37")
```

## MAG informations:

### Collections:

#### SQM DAS collection:

```{r}
DAS_collection %>% 
  here::here() %>% 
  read_tsv(col_names = c("split", "DAS_bin")) %>% 
  separate(split, into = c("cont", "cont_num", "split", "split_id"), sep = "_", remove = FALSE) %>% 
  mutate(Contig_ID = paste0(cont,"_", cont_num)) %>% 
  distinct(Contig_ID, .keep_all = TRUE) %>% 
  select(Contig_ID, DAS_bin) -> DAS

DAS %>% 
  head() %>% 
  DT::datatable()
```

#### Manual MAGs CONCOCT:

```{r}
Manual_bins %>% 
  read_tsv(col_names = c("split", "CONCOCT_bin")) %>% 
  separate(split, into = c("cont", "cont_num", "split", "split_id"), sep = "_", remove = FALSE) %>% 
  mutate(Contig_ID = paste0(cont,"_", cont_num)) %>% 
  distinct(Contig_ID, .keep_all = TRUE) %>% 
  select(Contig_ID, CONCOCT_bin) -> CONCOCT

CONCOCT  %>% 
  head() %>% 
  DT::datatable()
```

Generate fake bin for all contigs -> export kegg annotations from anvio?

```{r}
# Manual_bins %>%
#   read_tsv(col_names = c("split", "bin")) %>%
#   mutate(bin = "all") %>% 
#   data.frame() ->  Manual_bins_tmp
# 
# # names(Manual_bins_tmp) <- NULL
# 
# Manual_bins_tmp %>% 
#   write_tsv(here::here("data/processed/anvio/human1/collection_all_splits.tsv"), col_names = FALSE)
```

Let's add the contig <-> bin to the merged data:

```{r}
orf_contig %>% 
  left_join(CONCOCT) %>% 
  left_join(DAS) -> orf_contig_2
```

### MAGs summary:

#### MAGs GDTB taxonomy:

```{r}
GTDB_tax %>%
  read_excel() %>%
  filter(dataset == "human1") %>%
  mutate(user_genome = str_remove_all(user_genome, c("human1_|-contigs|refined_2_"))) %>%
  filter(str_detect(user_genome, c("DAS|concoct"))) %>%
  select(user_genome, classification) %>%
  # separate(classification, 
  #          into = c("domain", "phylum", "class", "order", "family", "genus", "species"), 
  #          sep = ";", 
  #          fill = "right") %>% 
  rename(GTDB_tax = classification) %>% 
  mutate(user_genome = str_remove_all(user_genome, "concoct_10_")) %>% 
  mutate(user_genome = str_remove_all(user_genome, "DAS_")) -> GTDB_tax

# TODO clean taxonomy and add highest resolution.

GTDB_tax %>% 
  head() %>% 
  DT::datatable()
```

```{r}
orf_contig_2 %>% 
  left_join(GTDB_tax,
            by = c("CONCOCT_bin" = "user_genome")) %>%  #,
  # suffix = c("", "CONCOCT")) %>% 
  rename(GTDB_tax_CONCOCT = GTDB_tax) %>% 
  left_join(GTDB_tax,
            by = c("DAS_bin" = "user_genome")) %>% #,
  # suffix = c("", "_DAS")) %>% 
  rename(GTDB_tax_DAS = GTDB_tax) -> orf_contig_3
```


```{r}
## CONCOCT tax info

# GTDB_tax %>%
#   filter(str_detect(user_genome, "concoct")) %>%
#   mutate(user_genome = str_remove_all(user_genome, "concoct_10_")) -> CONCOCT_tax
# 
# colnames(CONCOCT_tax) <- paste0("CONCOCT_", colnames(CONCOCT_tax))

```

#### SQM DAS info:

```{r}
SQM_DAS %>% 
  read_tsv(skip = 1) -> SQM_DAS

colnames(SQM_DAS) <- paste0("SQM_DAS_", colnames(SQM_DAS))

SQM_DAS %>%
  mutate(SQM_DAS_HQ = ifelse(SQM_DAS_Completeness >= 80 & SQM_DAS_Contamination <= 10, "HQ", NA)) %>% 
  select(-SQM_DAS_Method, -`SQM_DAS_Coverage D-1`:-`SQM_DAS_TPM HV-24`) -> SQM_DAS_HQ

# filter(SQM_DAS_Completeness >= 80,
#        SQM_DAS_Contamination <= 10) %>% 
# mutate(SQM_DAS_bins_HQ = `SQM_DAS_Bin ID`) -> SQM_DAS_HQ

SQM_DAS_HQ %>% 
  head() %>% 
  DT::datatable()
```

```{r}
orf_contig_3 %>% 
  left_join(SQM_DAS_HQ,
            by = c("DAS_bin" = "SQM_DAS_Bin ID")) -> orf_contig_3
```

#### ANVIO DAS info:

```{r}
ANVIO_SUMMARY_DAS %>% 
  here::here() %>% 
  read_tsv() %>% 
  select(-t_domain:-t_species) -> ANVIO_SUMMARY_DAS

colnames(ANVIO_SUMMARY_DAS) <- paste0("ANVIO_SUMMARY_DAS_", colnames(ANVIO_SUMMARY_DAS))

ANVIO_SUMMARY_DAS %>%
  # filter(ANVIO_SUMMARY_DAS_percent_completion >= 80,
  #        ANVIO_SUMMARY_DAS_percent_redundancy <= 10) %>% 
  mutate(ANVIO_DAS_HQ = ifelse(ANVIO_SUMMARY_DAS_percent_completion >= 80 & ANVIO_SUMMARY_DAS_percent_redundancy <= 10, "HQ", NA)) -> ANVIO_SUMMARY_DAS

ANVIO_SUMMARY_DAS %>% 
  head() %>% 
  DT::datatable()
```

```{r}
orf_contig_3 %>% 
  left_join(ANVIO_SUMMARY_DAS,
            by = c("DAS_bin" = "ANVIO_SUMMARY_DAS_bins")) -> orf_contig_3
```

#### ANVIO CONCOCT info:

```{r}
ANVIO_SUMMARY_CONCOCT %>% 
  here::here() %>% 
  read_tsv() %>% 
  select(-t_domain:-t_species) -> ANVIO_SUMMARY_CONCOCT

colnames(ANVIO_SUMMARY_CONCOCT) <- paste0("ANVIO_SUMMARY_CONCOCT_", colnames(ANVIO_SUMMARY_CONCOCT))

ANVIO_SUMMARY_CONCOCT %>% 
  mutate(ANVIO_CONCOCT_HQ = ifelse(ANVIO_SUMMARY_CONCOCT_percent_completion >= 80 & ANVIO_SUMMARY_CONCOCT_percent_redundancy <= 10, "HQ", NA)) -> ANVIO_SUMMARY_CONCOCT

ANVIO_SUMMARY_CONCOCT %>% 
  head() %>% 
  DT::datatable()
```

```{r}
orf_contig_3 %>% 
  left_join(ANVIO_SUMMARY_CONCOCT,
            by = c("CONCOCT_bin" = "ANVIO_SUMMARY_CONCOCT_bins")) -> orf_contig_3
```


Rename ORF and contig ID:
```{r}
orf_contig_3 %>% 
  mutate(ORF_ID = paste0("h_", ORF_ID),
         Contig_ID = paste0("h_", Contig_ID)) -> full_gene_catalogue
```

##  Extract quantitative information from SQM:

### ORFs:


```{r}
human$orfs$table %>%
  rownames_to_column("ORF_ID") %>% 
  data.frame() %>% 
  select(ORF_ID, TPM.D.1:Raw.base.count.HV.24) %>% 
  # select(starts_with("TPM"), "ORF_ID") %>% 
  column_to_rownames("ORF_ID") -> quant_orfs

# colnames(quant) <- colnames(quant) %>% 
#   str_replace_all("//.", "//_")

rownames(quant_orfs) <- paste0("h_",
                               rownames(quant_orfs))

quant_orfs %>% 
  rownames_to_column("ORF_ID") -> quant_orfs

quant_orfs %>% 
  head() %>% 
  DT::datatable()

# colnames(human_otu_tpm) <- colnames(human_otu_tpm) %>% str_remove_all("TPM ") %>% str_replace_all("-", "_")
# rownames(human_otu_tpm) <- paste0("h_", rownames(human_otu_tpm))
# 
# # Coverage
# human$orfs$table %>% as.data.frame() %>% select(starts_with("Coverage")) -> human_otu_cov
# 
# colnames(human_otu_cov) <- colnames(human_otu_cov) %>% str_remove_all("Coverage ") %>% str_replace_all("-", "_")
# rownames(human_otu_cov) <- paste0("h_", rownames(human_otu_cov))
# 
# # Raw reads
# human$orfs$table %>% as.data.frame() %>% select(starts_with("Raw read count")) -> human_otu_rr
# 
# colnames(human_otu_rr) <- colnames(human_otu_rr) %>% str_remove_all("Raw read count ") %>% str_replace_all("-", "_")
# rownames(human_otu_rr) <- paste0("h_", rownames(human_otu_rr))
```

### contigs:

```{r}
human$contigs$table %>%
  rownames_to_column("Contig_ID") %>% 
  data.frame() %>% 
  select(Contig_ID, Coverage.D.1:Raw.read.count.HV.24) %>% 
  # select(starts_with("TPM"), "ORF_ID") %>% 
  column_to_rownames("Contig_ID") -> quant_contigs

# colnames(quant_contigs) <- colnames(quant_contigs) %>% 
#   str_replace_all("//.", "//_")

rownames(quant_contigs) <- paste0("h_",
                                  rownames(quant_contigs))

quant_contigs %>% 
  rownames_to_column("Contig_ID") -> quant_contigs

# colnames(human_otu_tpm) <- colnames(human_otu_tpm) %>% str_remove_all("TPM ") %>% str_replace_all("-", "_")
# rownames(human_otu_tpm) <- paste0("h_", rownames(human_otu_tpm))
# 
# # Coverage
# human$orfs$table %>% as.data.frame() %>% select(starts_with("Coverage")) -> human_otu_cov
# 
# colnames(human_otu_cov) <- colnames(human_otu_cov) %>% str_remove_all("Coverage ") %>% str_replace_all("-", "_")
# rownames(human_otu_cov) <- paste0("h_", rownames(human_otu_cov))
# 
# # Raw reads
# human$orfs$table %>% as.data.frame() %>% select(starts_with("Raw read count")) -> human_otu_rr
# 
# colnames(human_otu_rr) <- colnames(human_otu_rr) %>% str_remove_all("Raw read count ") %>% str_replace_all("-", "_")
# rownames(human_otu_rr) <- paste0("h_", rownames(human_otu_rr))


quant_contigs %>% 
  head() %>% 
  DT::datatable()

```

Combine quantification data with annotations:

```{r}
colnames(quant_orfs) <- paste0("ORF_" ,colnames(quant_orfs))

colnames(quant_contigs) <- paste0("contig_" ,colnames(quant_contigs))
```


```{r}
full_gene_catalogue %>% 
  left_join(quant_orfs,
            by = c("ORF_ID" = "ORF_ORF_ID")) -> full_gene_catalogue_orf_quant

full_gene_catalogue_orf_quant %>% 
  left_join(quant_contigs,
            by = c("Contig_ID" = "contig_Contig_ID")) -> full_quant_full_gene_catalogue
```

```{r}
full_gene_catalogue_orf_quant %>% 
  select(ORF_ID, SQM_Length.NT, starts_with("ORF_Raw.read.count.")) %>% 
  pivot_longer(cols = starts_with("ORF_Raw.read.count."),
               values_to = "read_counts", 
               names_to = "sample") %>% 
  mutate(Num_Gi = read_counts / SQM_Length.NT) %>% 
  select(-read_counts) %>% 
  pivot_wider(names_from =  sample,
              values_from = Num_Gi) %>% 
  replace(is.na(.), 0) %>% 
  rename_at(
    # select all variables with "factor.sample" in the name
    vars(contains("ORF_Raw.read.count."))
    # use stringr::str_replace to remove factor.sample.
    #   you could do the same with base::gsub()
    , funs(str_replace(., "ORF_Raw.read.count.", "ORF_Num_Gi."))) %>% 
  select(-SQM_Length.NT) -> full_gene_catalogue_orf_Num_Gi

full_gene_catalogue_orf_Num_Gi %>%  
  head() %>% 
  DT::datatable()
```

```{r}
full_gene_catalogue_orf_Num_Gi %>% 
  pivot_longer(cols = starts_with("ORF_Num_Gi"),
               values_to = "Num_Gi", 
               names_to = "sample") %>% 
  group_by(sample) %>% 
  mutate(Num_Gi_pc = Num_Gi / sum(Num_Gi)) %>% 
  ungroup() %>% 
  select(-Num_Gi) %>% 
  pivot_wider(names_from =  sample,
              values_from = Num_Gi_pc) %>% 
  replace(is.na(.), 0) %>% 
  rename_at(
    # select all variables with "factor.sample" in the name
    vars(contains("ORF_Num_Gi."))
    # use stringr::str_replace to remove factor.sample.
    #   you could do the same with base::gsub()
    , funs(str_replace(., "ORF_Num_Gi.", "ORF_Num_Gi_pc."))) -> full_gene_catalogue_orf_Num_Gi_pc

full_gene_catalogue_orf_Num_Gi_pc %>%  
  head() %>% 
  DT::datatable()
```

```{r}
full_gene_catalogue_orf_quant %>% 
  left_join(full_gene_catalogue_orf_Num_Gi_pc,
            by = c("ORF_ID" = "ORF_ID")) %>% 
  left_join(full_gene_catalogue_orf_Num_Gi,
            by = c("ORF_ID" = "ORF_ID")) -> full_gene_catalogue_orf_quant_all_metrics

```


Generate phyloseq objects:


sample_data:

```{r}
human_meta
```


tax_table:

```{r}
full_gene_catalogue_orf_quant_all_metrics %>% 
  column_to_rownames("ORF_ID") %>% 
  select(Contig_ID:ANVIO_CONCOCT_HQ) %>% 
  as.matrix() %>% 
  tax_table() -> tax
```


ORFs:

TPM:


```{r}
full_gene_catalogue_orf_quant_all_metrics %>% 
  column_to_rownames("ORF_ID") %>% 
  select(ORF_TPM.D.1:ORF_TPM.HV.24) %>% 
  rename_at(
    # select all variables with "factor.sample" in the name
    vars(contains("ORF_TPM."))
    # use stringr::str_replace to remove factor.sample.
    #   you could do the same with base::gsub()
    , funs(str_replace(., "ORF_TPM.", ""))) %>% 
  # stringr::str_replace_all("\\s","_")
  rename_at(
    # select all variables with "factor.sample" in the name
    vars(contains("."))
    , funs(str_replace_all(., "\\.", "_")))  %>% 
  as.matrix() %>% 
  otu_table(taxa_are_rows = TRUE) -> tpm
```


Num_Gi_pc

```{r}
full_gene_catalogue_orf_quant_all_metrics %>% 
  column_to_rownames("ORF_ID") %>% 
  select(starts_with("ORF_Num_Gi_pc.")) %>% 
  rename_at(
    # select all variables with "factor.sample" in the name
    vars(contains("ORF_Num_Gi_pc."))
    # use stringr::str_replace to remove factor.sample.
    #   you could do the same with base::gsub()
    , funs(str_replace(., "ORF_Num_Gi_pc.", ""))) %>% 
  # stringr::str_replace_all("\\s","_")
  rename_at(
    # select all variables with "factor.sample" in the name
    vars(contains("."))
    , funs(str_replace_all(., "\\.", "_")))  %>% 
  as.matrix() %>% 
  otu_table(taxa_are_rows = TRUE) -> Num_Gi_pc
```

Num_Gi

```{r}
full_gene_catalogue_orf_quant_all_metrics %>% 
  column_to_rownames("ORF_ID") %>% 
  select(starts_with("ORF_Num_Gi.")) %>% 
  rename_at(
    # select all variables with "factor.sample" in the name
    vars(contains("ORF_Num_Gi."))
    # use stringr::str_replace to remove factor.sample.
    #   you could do the same with base::gsub()
    , funs(str_replace(., "ORF_Num_Gi.", ""))) %>% 
  # stringr::str_replace_all("\\s","_")
  rename_at(
    # select all variables with "factor.sample" in the name
    vars(contains("."))
    , funs(str_replace_all(., "\\.", "_")))  %>% 
  as.matrix() %>% 
  otu_table(taxa_are_rows = TRUE) -> Num_Gi
```

Coverage

```{r}
full_gene_catalogue_orf_quant_all_metrics %>% 
  column_to_rownames("ORF_ID") %>% 
  select(starts_with("ORF_Coverage.")) %>% 
  rename_at(
    # select all variables with "factor.sample" in the name
    vars(contains("ORF_Coverage."))
    # use stringr::str_replace to remove factor.sample.
    #   you could do the same with base::gsub()
    , funs(str_replace(., "ORF_Coverage.", ""))) %>% 
  # stringr::str_replace_all("\\s","_")
  rename_at(
    # select all variables with "factor.sample" in the name
    vars(contains("."))
    , funs(str_replace_all(., "\\.", "_")))  %>% 
  as.matrix() %>% 
  otu_table(taxa_are_rows = TRUE) -> cov
```

ORF_Raw.read

```{r}
full_gene_catalogue_orf_quant_all_metrics %>% 
  column_to_rownames("ORF_ID") %>% 
  select(starts_with("ORF_Raw.read.count.")) %>% 
  rename_at(
    # select all variables with "factor.sample" in the name
    vars(contains("ORF_Raw.read.count."))
    # use stringr::str_replace to remove factor.sample.
    #   you could do the same with base::gsub()
    , funs(str_replace(., "ORF_Raw.read.count.", ""))) %>% 
  # stringr::str_replace_all("\\s","_")
  rename_at(
    # select all variables with "factor.sample" in the name
    vars(contains("."))
    , funs(str_replace_all(., "\\.", "_")))  %>% 
  as.matrix() %>% 
  otu_table(taxa_are_rows = TRUE) -> read_count
```



Create phyloseq objects and tsv files:

```{r}
dim(tpm)

physeq_tpm <- phyloseq(tpm,
                             tax,
                             human_meta %>% sample_data())


physeq_tpm
```

```{r}
dim(cov)

physeq_cov <- phyloseq(cov,
                             tax,
                             human_meta %>% sample_data())

physeq_cov
```


```{r}
dim(Num_Gi_pc)

physeq_Num_Gi_pc <- phyloseq(Num_Gi_pc,
                                   tax,
                                   human_meta %>% sample_data())

physeq_Num_Gi_pc
```

```{r}
dim(Num_Gi)

physeq_Num_Gi<- phyloseq(Num_Gi,
                               tax,
                               human_meta %>% sample_data())

physeq_Num_Gi
```


```{r}
dim(read_count)

physeq_read_count <- phyloseq(read_count,
                                    tax,
                                    human_meta %>% sample_data())

physeq_read_count
```


```{r}
phyloseq_list_out = list("read_count" = physeq_read_count,
                "Num_Gi" = physeq_Num_Gi,
                "Num_Gi_pc" = physeq_Num_Gi_pc,
                "cov" = physeq_cov,
                "tpm" = physeq_tpm)

```


Export outputs:

```{r}

phyloseq_list_out %>% 
  saveRDS(here::here("data/processed/human_full_gene_catalog_phyloseq.RDS"))

full_gene_catalogue_orf_quant_all_metrics %>% 
  saveRDS(here::here("data/processed/human_full_gene_catalog_full_metrics_table.RDS"))

full_gene_catalogue_orf_quant_all_metrics %>% 
  write_tsv(here::here("data/processed/human_full_gene_catalog_full_metrics_table.tsv.gz"))
```


Export only AMR and Tox:

```{r}
full_gene_catalogue_orf_quant_all_metrics %>% 
  filter(!is.na(PathoFact_AMR_ARG) | !is.na(rgi_CARD_ARO) | !is.na(PathoFact_Tox_Toxin_classification)) %>% 
  write_tsv(here::here("data/processed/human_full_gene_catalog_full_metrics_table_AMR_tox.tsv"))
```



We might need to clarify some cases before saving: e.g.: plasmid but in MAG...


```{r}
# 
# 
# left_join(card, by = "ORF_ID") %>%
#   left_join(hgt_events, by = c("Contig" = "CONTIG_NAME")) %>%
#   left_join(viralverify, by = c("Contig" = "Contig name")) %>%
#   left_join(top_isescan, by = c("Contig" = "seqID")) %>%
#   left_join(DAS, by = c("Contig" = "cont_id")) %>%
#   left_join(DAS_tax, by = c("DAS_bin" ="DAS_user_genome")) %>%
#   left_join(DAS_HQ, by = c("DAS_bin" = "DAS_bins_HQ"), keep = TRUE) %>%
#   left_join(CONCOCT, by = c("Contig" = "cont_id")) %>%
#   left_join(CONCOCT_tax, by = c("CONCOCT_bin" ="CONCOCT_user_genome")) %>%
#   left_join(CONCOCT_HQ, by = c("CONCOCT_bin" = "CONCOCT_bins_HQ"), keep = TRUE) %>%
#   select(-c(Megahit, Range)) %>%
#   column_to_rownames("ORF_ID") %>%
#   mutate(Classification_mobilome_DAS = case_when(
#     Prediction == "Plasmid" ~ "Plasmid",
#     !is.na(DAS_bin) ~ "MAG",
#     TRUE ~ Prediction)) %>%
#   mutate(Classification_mobilome_DAS_HQ = case_when(
#     Prediction == "Plasmid" ~ "Plasmid",
#     !is.na(DAS_bins_HQ) ~ "MAG",
#     TRUE ~ Prediction)) %>%
#   mutate(Classification_mobilome_CONCOCT = case_when(
#     Prediction == "Plasmid" ~ "Plasmid",
#     !is.na(CONCOCT_bin) ~ "MAG",
#     TRUE ~ Prediction)) %>%
#   mutate(Classification_mobilome_CONCOCT_HQ = case_when(
#     Prediction == "Plasmid" ~ "Plasmid",
#     !is.na(CONCOCT_bins_HQ) ~ "MAG",
#     TRUE ~ Prediction)) %>%
#   as.matrix(rownames = TRUE) -> human_tax
# 
# 
# rownames(human_tax) <- paste0("h_",rownames(human_tax))
# 
# # Check for hgt events in resistome
# human_tax %>% as.data.frame() %>% filter(!is.na(ARO), CALL == "lgt")


```




```{r}
devtools::session_info()
```

