---
title: "Merge SQM ORF abundances with RGI annotations"
author: "Sneha Sundar"
date: "04/08/2021"
output: html_document
---

Load required packages
```{r}
library('dplyr')
library("tidyverse")
library('here')

```
File paths to the SQM R object, formatted RGI annotation file and output merged file. 

```{r}
in_sqm_r <- here('sqm_results','chicken_SQM.rds')
in_rgi <- here('rgi_results','rgi_ORF_chicken_formatted.txt')
out_file <- here('data','sqm_rgi_chicken.txt')
```


Load the SqueezeMeta R object. It is quite big so need to find the best way to do this without having to store this object in my system . 
```{r}
chicken = readRDS(in_sqm_r)
```

Exploring the R Object: orfs$table seems to have all the info needed - TPM, read abundances, base abundances, coverage
```{r}
chicken$orfs$table[1:5,]
```

Load the RGI results that is in a .txt file. However this text file needs to be formatted a bit before we can load as a dataframe correctly. I wrote a small bash script to do that and ran it. We will now load this formatted .txt file. This holds the ORF_ID and corresponding antibiotic resistance gene annotation in a tab-delimited format. 
```{r}
rgi_result <- read.table(in_rgi,sep='\t', header=TRUE,quote = "")

rgi_result %>% head()
```
Some cleaning up of the ORF abundance table. 
```{r}
orfs_table<-chicken$orfs$table %>% rownames_to_column('ORF_ID')

```

Joining the SQM orf abundance table with the rgi annotations. This will give us a huge dataframe that contains all the ORFs identified by SQM having corresponding RGI annotation. 
```{r}
orf_rgi_abundances <- left_join(rgi_result,orfs_table,by='ORF_ID')
```

Save as a tab-delimited file . 
```{r}
write.table(orf_rgi_abundances,out_file,sep="\t",row.names=FALSE)
```

