---
title: "Resistome data - chicken 1 & 2"
author: "Florentin Constancias "
date: " `r format(Sys.time(), '%B %d, %Y')` "
output: 
  html_document: 
    toc: yes
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# knitr::opts_chunk$set(dev = "ragg_png") # https://ragg.r-lib.org/index.html
knitr::opts_chunk$set(fig.width = 6) #https://r4ds.had.co.nz/graphics-for-communication.html#figure-sizing
knitr::opts_chunk$set(fig.asp = 0.618)
knitr::opts_chunk$set(fig.show = "hold")
knitr::opts_chunk$set(fig.show = "70%")
knitr::opts_chunk$set(fig.align = "center")
```

```{r}
rm(list= ls())

gc()

require("tidyverse")
require('speedyseq')
require("microViz")

'%!in%' <- function(x,y)!('%in%'(x,y))

source("https://raw.githubusercontent.com/fconstancias/DivComAnalyses/master/R/phyloseq_taxa_tests.R")
source("https://raw.githubusercontent.com/fconstancias/DivComAnalyses/master/R/phyloseq_normalisation.R")
source("https://raw.githubusercontent.com/fconstancias/DivComAnalyses/master/R/phyloseq_alpha.R")
source("https://raw.githubusercontent.com/fconstancias/DivComAnalyses/master/R/phyloseq_beta.R")
source("https://raw.githubusercontent.com/fconstancias/DivComAnalyses/master/R/phyloseq_heatmap.R")
source("https://raw.githubusercontent.com/fconstancias/DivComAnalyses/master/R/phyloseq_varia.R")
source("https://raw.githubusercontent.com/fconstancias/metabaRpipe-source/master/Rscripts/functions.R")
```


```{r}
export_ppt_width = NULL
export_ppt_height = NULL

out_xlsx = "~/Desktop/plots_manuscript_chickens.xlsx"
out_pptx = "~/Desktop/plots_manuscript_chickens.pptx"

```

# Load data:

## Aestethics:

```{r}
load( here::here("Figures/Rastetics.Rdata"))
```


## 16S:

```{r}
# "data/processed/16S/ps_silva_dada2_human_chicken_all_fct_polyferms.RDS" %>% 
#   here::here() %>% 
#   readRDS() %>% 
#   filter_taxa(function(x) sum(x > 0) > 0, TRUE) -> ps_16S
```

## resistome:

```{r}
# "data/processed/chicken1-2_anvio-bin-full_gene_catalog_phyloseq.RDS" %>% 
#   here::here() %>% 
#   readRDS() -> ps
# 
# ps

```


```{r}
# ps$Num_Gi_pc  %>% 
#   microViz::ps_mutate(Day_of_Treatment = as.double(Day_of_Treatment)) %>% 
#   subset_samples(Day_of_Treatment < 34 | is.na(Day_of_Treatment)) -> ps1
# 
# subset_samples(ps1, sample_names(ps1) %!in% c("D3", "D19")) %>% 
#   filter_taxa(function(x) sum(x > 0) > 0, TRUE) -> ps2
# 
# ps2
```


```{r}
# ps_AMR %>%
#   physeq_sel_tax_table(c("Contig.ID","ARG","AMR_sub_class", "AMR_sub_class_multi",
#                          "Resistance_mechanism","Resistance_mechanism_multi",
#                          "MGE_prediction", "MGE_prediction_clean","bin_name", "bin_name_clean",
#                          "plasmid_score", "topology_plas","plasmids_mobilog_pc","virus_score","topology_phage", "taxonomy", "Bacteriophages_mobilog_pc",
#                          "IntegrativeEle_mobilog_pc")) -> ps_AMRclean
# 
# ps_AMRclean
```



save phyloseq objects:

```{r}
"data/processed/chicken1-2_anvio-bin-clean_AMR_phyloseq.RDS" %>% 
  here::here() %>% 
  readRDS() -> ps_AMR

"data/processed/chicken1-2_anvio-bin-clean_phyloseq.RDS" %>% 
  here::here() %>% 
  readRDS() -> ps2_clean 
```



identify the plasmid contigs so we can remove them from the ANVIO collection:


```{r}
ps_AMR %>%
  subset_taxa(ARG %in% c("VANA", "CTX-M-1")) -> ps_inoc_AMR

ps_inoc_AMR %>%
  tax_table(.) %>%
  data.frame() %>% pull(Contig.ID) -> contig_inoc_plas_id

```

```{r}
ps2_clean %>% 
  subset_taxa(Contig.ID %in% contig_inoc_plas_id) -> ps_inoc_plas

ps_inoc_plas
```


```{r}
ps_inoc_AMR %>% 
  # ps_mutate(Day_of_Treatment = as.double(Day_of_Treatment)) %>% 
  ps_arrange(-Day_of_Treatment) %>% 
  physeq_sel_tax_table("ARG") %>% 
  # microbiome::transform("log") %>% 
  microViz::comp_barplot(tax_level = "ARG",
                         tax_transform_for_plot = "identity",
                         label = "Day_of_Treatment", # name an alternative variable to label axes
                         n_taxa = 20, # give more taxa unique colours
                         merge_other = FALSE, # split the "other" category to display alpha diversity
                         bar_width = 0.9, # reduce the bar width to 70% of one row
                         bar_outline_colour = "grey5",
                         sample_order = "default") +#,# is the default (use NA to remove outlines)
  # facet_by = c("Model", "Antibiotic", "Antibiotic_mg.mL", "Treatment2")) +
  # group_by = c("Model", "Antibiotic", "Antibiotic_mg.mL", "Treatment2"))
  coord_flip() + theme_light() + 
  scale_y_continuous(labels=percent) + 
  # facet_grid(Model2 ~ Antibiotic + Antibiotic_mg.L + Treatment2, scales = "free_y", space = "free") +
  facet_wrap(Model2 ~ Antibiotic + Antibiotic_mg.L + Treatment2, scales = "free",  ncol = 6, nrow = 4) +
  # scale_y_log10() +
  guides(fill=guide_legend(ncol = 1)) -> pAMR

pAMR

pAMR %>%
  export::graph2ppt(append = TRUE, width = 317.48031496 * 4 , height = 0.618 * 317.48031496 * 2.25, paper = "A3",  scaling = 2,
                    file = out_pptx)

pAMR$data %>% 
  data.frame() %>% 
  select(OTU, Sample, Abundance,Model2, Reactor_Treatment, Day_of_Treatment, Antibiotic_mg.L) %>% 
  xlsx::write.xlsx(file = out_xlsx,
                   sheetName = "AMRg_plasmid", append = TRUE)
```

No need to melt for this one since every otu is a gene with AMR hit and MGE / MAG ...

```{r}
ps_AMR %>%
  microViz::ps_mutate(Reactor_Treatment_Dose = paste0(Model2, Reactor, Treatment, Antibiotic_mg.L)) %>% 
  physeq_sel_tax_table(c("AMR_sub_class", "AMR_sub_class_multi",
                         "Resistance_mechanism","Resistance_mechanism_multi","MGE_prediction_clean", "t_phylum","t_class", "t_order","t_family","t_genus"  ,                  "t_species" ,"X..completion","X..redundancy","total.length" ,"bin_name_clean", "ARG"))  %>%
  # "viralverify_Prediction", "isescan_type", "isescan_count_IS_contig" , "ANVIO_CONCOCT_HQ_clean", "CONCOCT_bin",
  # "GTDB_tax_CONCOCT_clean", "PathoFact_AMR_MGE_prediction_clean_2")) %>% 
  subset_samples(Model == "Chicken") %>% 
  subset_samples((Reactor != "CR" & Model2 == "Chicken1") | Model == "Human" | Model2 == "Chicken2") %>% 
  ps_mutate(Antibiotic_mg.L = factor(Antibiotic_mg.L, levels = c(NA, 20, 200,90, 600))) %>% 
  ps_mutate(Antibiotic = ifelse(Antibiotic_mg.L == 90, "VAN", Antibiotic)) %>% 
  ps_mutate(Treatment2 = factor(Treatment2, levels = c("DONOR","UNTREATED","AB",   "AB+E. coli" ,"E. coli", "AB+E. faecium","E. faecium" ), exclude = NULL)) %>% 
  ps_mutate(Treatment = factor(Treatment, levels = c("DONOR", "UNTREATED", "CTX", "CTX+HV292.1", "HV292.1", "VAN" , "VAN+CCUG59168", "CCUG59168" ), exclude = NULL)) %>% 
  ps_mutate(Antibiotic = factor(Antibiotic, levels = c(NA, "CTX", "VAN"), exclude = NULL)) %>% 
  ps_mutate(Model2 = factor(Model2, levels = c("Chicken1", "Chicken2"), exclude = NULL)) -> ps_AMR_all

ps_AMR_all
```


```{r}
ps_AMR_all %>%
  sample_data() %>% 
  data.frame() %>% 
  rownames_to_column("sample_id") -> metadata_mapping

# metadata_mapping %>% 
#   DT::datatable()
```

```{r}
ps_AMR %>% 
  tax_table() %>% 
  data.frame() %>% 
  rownames_to_column("ORF_ID") -> tax_mapping
```

```{r}
as(tax_table(ps_AMR_all), "matrix") %>% 
  as.data.frame() %>%
  rownames_to_column('orf') -> tax_table

as(otu_table(ps_AMR_all), "matrix") %>% 
  as.data.frame() %>%
  rownames_to_column('orf') -> otu_table

otu_table %>%
  # left_join(refseq_df, by = 'ASV') %>%
  left_join(tax_table, by = c("orf" = "orf")) %>%
  dplyr::select(orf, everything()) -> merged_table


merged_table %>% 
  write_tsv("~/Desktop/AMR_ps_chick1-2.tsv")
```



Looks good, let's go.

# Total Resistome :

## Abundance - sum Rnum_Gi:

```{r}
ps_AMR_all %>%
  ps_mutate(Day_of_Treatment = ifelse(Treatment == "DONOR", -3, Day_of_Treatment)) %>% 
  # physeq_sel_tax_table(c("PathoFact_AMR_AMR_sub_class", "PathoFact_AMR_AMR_sub_class_multi", "PathoFact_AMR_ARG"))  %>% 
  physeq_sel_tax_table(c("AMR_sub_class", "AMR_sub_class_multi", "ARG"))  %>% 
  speedyseq::tax_glom(., taxrank = "ARG")  %>% 
  filter_taxa(function(x) sum(x > 0) > 0, TRUE) %>% 
  speedyseq::psmelt() %>% 
  select(-Lactose_mM:-Total_SCFA_mM) %>% 
  # filter(Day_of_Treatment < 30 &
  # !Sample %in% c("D3", "D19")) %>% 
  # filter((Reactor != "CR" & Model2 == "Chicken1") | Model == "Human" | Model2 == "Chicken2")  %>% 
  # filter(Reactor != "CR" & Model2 == "Chicken1" | Model == "Human" & Reactor %in% c("TR1", "TR2", "TR3", "TR4", "TR5", "TR6", "CR"))  %>% 
  group_by(Sample) %>% 
  summarise(sum_abundance = sum(Abundance), .groups = "drop") %>% 
  mutate(sum_abundance = na_if(sum_abundance, 0)) %>%
  left_join(metadata_mapping,
            by = c("Sample" = "sample_id")) %>%
  # mutate(Antibiotic_mg.mL = factor(Antibiotic_mg.mL, levels = c(NA, 20, 200,90, 600))) %>%
  ggplot(aes_string(x = "Day_of_Treatment", y = "sum_abundance", color = "Treatment", fill  = "Treatment", shape = "Antibiotic_mg.L")) + #shape = Fermentation
  geom_point(size = 2.5) + 
  geom_line(linetype = 2,  size = 0.5,  aes_string(group = "interaction(Model, Reactor_Treatment_Dose)")) +
  labs(y = "sum Rnum_Gi", x = "Day_of_Treatment") +
  scale_color_manual(name = "", values = treat_col,
                     na.value = "black") +
  scale_fill_manual(name = "", values = treat_col,
                    na.value = "black") +
  scale_shape_manual(name = "" ,values = c(15,16,18,19), na.value =  17) +
  # scale_alpha_continuous(name = "", range=c(0.6, 1), na.value =  1) +
  geom_vline(xintercept = 0, size = 0.5, linetype = 1, color = "grey60") +
  theme_light() + 
  scale_y_continuous(labels=scientific) -> p1 # scale_y_continuous(labels = function(x) format(x, scientific = TRUE))
```


```{r, fig.asp = 0.618}
p1 + facet_grid(as.formula(paste0(".  ~  Model2  ")), scales = "free", space = "fixed", drop = TRUE) -> ptmp

ptmp
```


```{r, fig.asp = 0.618}
# ptmp %>% 
#   ggsave(plot = .,bg = "transparent",scale = 2,
#          units = "mm",
#          dpi = 600,
#          height = 0.618 * 84,
#          width = 84,
#          filename = "~/Desktop/Resistome_abundance.eps")

ptmp %>%
  export::graph2ppt(append = TRUE, width = 317.48031496 , height = 0.618 * 317.48031496, paper = "A4",  scaling = 2,
                    file = out_pptx)
```
Export values displayed:

```{r}
ptmp$data %>%
  select(Sample, sum_abundance, Day_of_Treatment, Treatment, Model2) %>%
  arrange(Day_of_Treatment) %>%
  xlsx::write.xlsx(file = out_xlsx,
                   sheetName = "resistome_density", append = TRUE)
```

Also retrieve the samples we now keep for the final analysis:

```{r}
p1$data %>% distinct(Sample)  %>% pull %>%  as.character() -> samples

samples <- c(samples, "D43")
```


baseline normalized:

```{r}
#https://stackoverflow.com/questions/52581469/dynamically-normalize-all-rows-with-first-element-within-a-group
#<https://blog.exploratory.io/introducing-time-series-analysis-with-dplyr-60683587cf8a>

ps_AMR_all %>% 
  physeq_sel_tax_table(c("AMR_sub_class", "AMR_sub_class_multi", "ARG"))  %>% 
  speedyseq::tax_glom(., taxrank = "ARG")  %>% 
  # prune_samples(samples, .) %>% 
  filter_taxa(function(x) sum(x > 0) > 0, TRUE) %>% 
  speedyseq::psmelt() %>% 
  select(-Lactose_mM:-Total_SCFA_mM) %>% 
  group_by(Sample) %>% 
  summarise(sum_abundance = sum(Abundance), .groups = "drop") %>% 
  mutate(sum_abundance = na_if(sum_abundance, 0)) %>%
  left_join(metadata_mapping,
            by = c("Sample" = "sample_id")) %>% 
  ungroup() %>% 
  group_by(Model2, Reactor_Treatment_Dose) %>% 
  arrange(Day_of_Treatment) %>% 
  mutate(per_cent_change = sum_abundance / first(sum_abundance)) %>% 
  ggplot(aes_string(x = "Day_of_Treatment", y = "per_cent_change", color = "Treatment", fill  = "Treatment", shape = "Antibiotic_mg.L")) +
  geom_point(size = 2.5) + 
  geom_line(linetype = 2,  size = 0.5,  aes_string(group = "interaction(Model, Reactor_Treatment_Dose)")) +
  labs(y = "sum Rnum_Gi / baseline", x = "Day_of_Treatment") +
  scale_color_manual(name = "", values = treat_col,
                     na.value = "black") +
  scale_fill_manual(name = "", values = treat_col,
                    na.value = "black") +
  scale_shape_manual(name = "" ,values = c(15,16,18,19), na.value =  17) +
  geom_vline(xintercept = 0, size = 0.5, linetype = 1, color = "grey60") +
  geom_hline(yintercept = 1, size = 0.5, linetype = 2, color = "grey60") +
  theme_light() + facet_grid(as.formula(paste0(". ~  Model2  ")), 
                             scales = "free_x", space = "fixed", drop = TRUE) -> p1

p1


p1 %>%
  export::graph2ppt(append = TRUE, width = 317.48031496 , height = 0.618 * 317.48031496, paper = "A4",  scaling = 2,
                    file = out_pptx)
```

# AMR drug classes

```{r}

pdrugclass = NULL


ps_AMR_all %>% 
  physeq_sel_tax_table(c("AMR_sub_class", "AMR_sub_class_multi", "ARG"))  %>%
  speedyseq::tax_glom(., taxrank = "AMR_sub_class_multi")  %>% 
  physeq_sel_tax_table(c("AMR_sub_class_multi")) %>% 
  # prune_samples(samples, .) %>% 
  filter_taxa(function(x) sum(x > 0) > 0, TRUE) %>% 
  # physeq_add_metadata(metadata_mapping,
  #                     sample_column = "sample_id") %>% 
  transform_sample_counts(function(x) x/sum(x) * 100) %>% 
  subset_taxa(AMR_sub_class_multi %!in% c("Unclassified", "multidrug (unclassified)")) %>% 
  subset_samples(Model2 == "Chicken1") %>% 
  # ps_mutate(Day_of_Treatment = as.numeric(Day_of_Treatment)) %>% 
  ps_arrange(Day_of_Treatment) %>% 
  comp_barplot(
    palette = distinct_palette(12, pal = "kelly"),
    tax_level = "AMR_sub_class_multi",
    label = "Day_of_Treatment", # name an alternative variable to label axes
    n_taxa = 12, # give more taxa unique colours
    merge_other = FALSE, # split the "other" category to display alpha diversity
    bar_width = 0.9, # reduce the bar width to 70% of one row
    bar_outline_colour = "grey5",
    sample_order = "default") +#,# is the default (use NA to remove outlines)
  # facet_by = c("Model", "Antibiotic", "Antibiotic_mg.mL", "Treatment2")) +
  # group_by = c("Model", "Antibiotic", "Antibiotic_mg.mL", "Treatment2"))
  # coord_flip() + 
  theme_light() + 
  ylab("Proportion - %") +
  scale_y_continuous(labels=percent) + 
  scale_fill_manual(values = drug_classe_multi) +
  scale_color_manual(values = drug_classe_multi) +
  facet_grid(Model2 ~ Antibiotic + Treatment2  + Antibiotic_mg.L, scales = "free", space = "free") -> pdrugclass$Chicken1
# facet_wrap(Model2 ~ Antibiotic + Antibiotic_mg.L + Treatment2, scales = "free", ncol = 7, nrow = 3) -> pdrugclass$Chicken1

# # Plot them side by side with the patchwork package.
# patch <- patchwork::wrap_plots(plot_list, nrow = 1, guides = "collect")
# patch & coord_flip() # make all plots horizontal (note: use & instead of +)

pdrugclass$Chicken1
```



```{r}
ps_AMR_all %>% 
  physeq_sel_tax_table(c("AMR_sub_class", "AMR_sub_class_multi", "ARG"))  %>%
  speedyseq::tax_glom(., taxrank = "AMR_sub_class_multi")  %>% 
  physeq_sel_tax_table(c("AMR_sub_class_multi")) %>% 
  # prune_samples(samples, .) %>% 
  filter_taxa(function(x) sum(x > 0) > 0, TRUE) %>% 
  # physeq_add_metadata(metadata_mapping,
  #                     sample_column = "sample_id") %>% 
  transform_sample_counts(function(x) x/sum(x) * 100) %>% 
  subset_taxa(AMR_sub_class_multi %!in% c("Unclassified", "multidrug (unclassified)")) %>% 
  subset_samples(Model2 == "Chicken2") %>% 
  # ps_mutate(Day_of_Treatment = as.numeric(Day_of_Treatment)) %>% 
  ps_arrange(Day_of_Treatment) %>% 
  comp_barplot(
    # palette = distinct_palette(12, pal = "kelly"),
    tax_level = "AMR_sub_class_multi",
    label = "Day_of_Treatment", # name an alternative variable to label axes
    n_taxa = 12, # give more taxa unique colours
    merge_other = FALSE, # split the "other" category to display alpha diversity
    bar_width = 0.9, # reduce the bar width to 70% of one row
    bar_outline_colour = "grey5",
    sample_order = "default") +#,# is the default (use NA to remove outlines)
  # facet_by = c("Model", "Antibiotic", "Antibiotic_mg.mL", "Treatment2")) +
  # group_by = c("Model", "Antibiotic", "Antibiotic_mg.mL", "Treatment2"))
  # coord_flip() + 
  theme_light() + 
  ylab("Proportion - %") +
  scale_y_continuous(labels=percent) + 
  scale_color_manual(values = drug_classe_multi) +
  scale_fill_manual(values = drug_classe_multi) +
  facet_grid(Model2 ~ Antibiotic  + Treatment2 + Antibiotic_mg.L, scales = "free", space = "free") -> pdrugclass$Chicken2
# facet_wrap(Model2 ~ Antibiotic + Antibiotic_mg.L + Treatment2, scales = "free", ncol = 7, nrow = 3) -> pdrugclass$Chicken2

# # Plot them side by side with the patchwork package.
# patch <- patchwork::wrap_plots(plot_list, nrow = 1, guides = "collect")
# patch & coord_flip() # make all plots horizontal (note: use & instead of +)

pdrugclass$Chicken2
```

```{r}
# plot_list <- (list(Chicken1 = pdrugclass$Chicken1,
#                    Chicken2 = pdrugclass$Chicken2))
# 
# patchwork::wrap_plots(plot_list, nrow = 2, guides = "collect")
```


```{r}
combined <- ggpubr::ggarrange(pdrugclass$Chicken1, pdrugclass$Chicken2, ncol=1, nrow=2, common.legend = TRUE, legend="right")
# patch_drugclass <- patchwork::wrap_plots(pdrugclass, nrow = 2, guides = "collect")
# require(patchwork)

combined
```


```{r}
# p2 %>% 
#   ggsave(plot = .,bg = "transparent",scale = 2,
#          units = "mm",
#          dpi = 600,
#          height = 0.618 * 84 * 2,
#          width = 84 * 2,
#          filename = "~/Desktop/Resistome_sub_class_multi_prop.eps")

combined %>%
  export::graph2ppt(append = TRUE, width = 317.48031496 * 6  , height = 0.618 * 317.48031496 * 6 , paper = "A4",  scaling = 2,
                    file = out_pptx)
```

Export values displayed:

```{r}
rbind(pdrugclass$Chicken1$data, pdrugclass$Chicken2$data) %>% 
  data.frame() %>% 
  select(Sample, Abundance, Day_of_Treatment, Treatment, Model2, OTU) %>% 
  # arrange(Day_of_Treatment) %>% 
  xlsx::write.xlsx(x = . ,
                   append = TRUE,
                   file = out_xlsx,
                   sheetName = "drug_class")
```

same but heatmap:

```{r}
ps_AMR_all %>% 
  physeq_sel_tax_table(c("AMR_sub_class_multi"))  %>%
  speedyseq::tax_glom(., taxrank = "AMR_sub_class_multi")  %>% 
  physeq_sel_tax_table(c("AMR_sub_class_multi")) %>% 
  prune_samples(samples, .) %>% 
  filter_taxa(function(x) sum(x > 0) > 0, TRUE) %>% 
  # physeq_add_metadata(metadata_mapping,
  #                     sample_column = "sample_id") %>% 
  transform_sample_counts(function(x) x/sum(x) * 100) %>% 
  subset_taxa(AMR_sub_class_multi %!in% c("Unclassified", "multidrug (unclassified)", "MLS (unclassified)")) %>% 
  filter_taxa(function(x) sum(x > 0) > 0, TRUE) %>% 
  # ps_arrange(-Day_of_Treatment) %>% 
  speedyseq::psmelt() %>%
  # left_join(tax_mapping,
  #           by = c("Best_Hit_ARO" = "Best_Hit_ARO"),
  #           suffix = c("_x", "")) %>%
  # mutate(Sample = fct_reorder(Sample, Model,Treatment,Fermentation,Antibiotic_mg.mL, Day_of_Treatment )) %>%
  mutate(Abundance = na_if(Abundance, 0)) %>%
  mutate(logTPM = log10(Abundance + 1)) -> pheatmap_df


pheatmap_df %>%
  # replace(is.na(.), 0)  %>%
  mutate(log10Abundance = log10(Abundance)) %>% 
  select(Sample, AMR_sub_class_multi, Abundance) %>%
  group_by(Sample, AMR_sub_class_multi) %>%
  pivot_wider(names_from =  Sample, AMR_sub_class_multi,
              values_from = Abundance) %>%
  column_to_rownames("AMR_sub_class_multi") -> pheatmap_samples


pheatmap_df %>%
  # replace(is.na(.), 0)  %>%
  mutate(log10Abundance = log10(Abundance)) %>% 
  select(Sample, AMR_sub_class_multi, Abundance) %>%
  group_by(AMR_sub_class_multi) %>% 
  summarise(meanprop = mean(Abundance, na.rm = TRUE)) -> mean_ab


# we want to knownumber of AMR gene behind each category:

ps_AMR_all %>% 
  physeq_sel_tax_table(c("AMR_sub_class_multi", "ARG"))  %>%
  speedyseq::tax_glom(., taxrank = "ARG")  %>% 
  filter_taxa(function(x) sum(x > 0) > 0, TRUE) %>%
  transform_sample_counts(function(x) x/sum(x) * 100) %>% 
  subset_taxa(AMR_sub_class_multi %!in% c("Unclassified", "multidrug (unclassified)", "MLS (unclassified)")) %>% 
  # physeq_sel_tax_table(c("PathoFact_AMR_AMR_sub_class_multi")) %>% 
  # prune_samples(samples, .) %>% 
  filter_taxa(function(x) sum(x > 0) > 0, TRUE) %>% 
  # physeq_add_metadata(metadata_mapping,
  #                     sample_column = "sample_id") %>% 
  # ps_mutate(Day_of_Treatment = as.numeric(Day_of_Treatment)) %>% 
  # ps_arrange(-Day_of_Treatment) %>% 
  speedyseq::psmelt() %>% 
  select(Sample, AMR_sub_class_multi, ARG, Abundance) %>%
  filter(Abundance > 0) %>% 
  select(-Sample, -Abundance) %>% 
  distinct(ARG, .keep_all = TRUE) %>% 
  group_by(AMR_sub_class_multi) %>%
  summarise(nAMR = length(ARG)) -> num_AMR

# we want to knownumber of gene behind each AMR gene name:

ps_AMR_all %>% 
  physeq_sel_tax_table(c("AMR_sub_class_multi", "ARG"))  %>%
  # speedyseq::tax_glom(., taxrank = "PathoFact_AMR_ARG")  %>% 
  filter_taxa(function(x) sum(x > 0) > 0, TRUE) %>% 
  transform_sample_counts(function(x) x/sum(x) * 100) %>% 
  subset_taxa(AMR_sub_class_multi %!in% c("Unclassified", "multidrug (unclassified)", "MLS (unclassified)")) %>% 
  # physeq_sel_tax_table(c("PathoFact_AMR_AMR_sub_class_multi")) %>% 
  prune_samples(samples, .) %>% 
  filter_taxa(function(x) sum(x > 0) > 0, TRUE) %>% 
  # physeq_add_metadata(metadata_mapping,
  #                     sample_column = "sample_id") %>% 
  # ps_mutate(Day_of_Treatment = as.numeric(Day_of_Treatment)) %>% 
  # ps_arrange(-Day_of_Treatment) %>% 
  speedyseq::psmelt() %>% 
  select(Sample, AMR_sub_class_multi, OTU, Abundance) %>%
  filter(Abundance > 0) %>% 
  select(-Sample, -Abundance) %>% 
  distinct(OTU, .keep_all = TRUE) %>% 
  group_by(AMR_sub_class_multi) %>%
  summarise(ngn = length(OTU)) -> num_gn


cbind(mean_ab,
      nGN = num_gn$ngn,
      nAMR = num_AMR$nAMR) -> row_anot
```


dendrogram gN:

```{r}
ps_AMR_all %>% 
  physeq_sel_tax_table(c("AMR_sub_class_multi", "ARG"))  %>%
  # speedyseq::tax_glom(., taxrank = "PathoFact_AMR_ARG")  %>% 
  filter_taxa(function(x) sum(x > 0) > 0, TRUE) %>% 
  transform_sample_counts(function(x) x/sum(x) * 100) %>% 
  subset_taxa(AMR_sub_class_multi %!in% c("Unclassified", "multidrug (unclassified)", "MLS (unclassified)")) %>% 
  # physeq_sel_tax_table(c("PathoFact_AMR_AMR_sub_class_multi")) %>% 
  # prune_samples(samples, .) %>% 
  filter_taxa(function(x) sum(x > 0) > 0, TRUE) %>% 
  # physeq_add_metadata(metadata_mapping,
  #                     sample_column = "sample_id") %>% 
  # ps_mutate(Day_of_Treatment = as.numeric(Day_of_Treatment)) %>% 
  # ps_arrange(-Day_of_Treatment) %>% 
  speedyseq::psmelt() %>%
  # left_join(tax_mapping,
  #           by = c("Best_Hit_ARO" = "Best_Hit_ARO"),
  #           suffix = c("_x", "")) %>%
  # mutate(Sample = fct_reorder(Sample, Model,Treatment,Fermentation,Antibiotic_mg.mL, Day_of_Treatment )) %>%
  mutate(Abundance = na_if(Abundance, 0)) %>%
  # mutate(logTPM = log10(Abundance + 1)) %>% 
  # mutate(log10Abundance = log10(Abundance)) %>% 
  select(Sample, OTU, Abundance) %>%
  group_by(Sample, OTU) %>%
  pivot_wider(names_from =  Sample, OTU,
              values_from = Abundance) %>%
  column_to_rownames("OTU") %>% 
  replace(is.na(.), 0)  %>%
  t() %>%
  vegan::vegdist(na.rm = TRUE) %>%
  hclust(method = "ward.D2") -> clust_t_OTU

```

dendrogram AMRgN:

```{r}
ps_AMR_all %>% 
  physeq_sel_tax_table(c("AMR_sub_class_multi", "ARG"))  %>%
  speedyseq::tax_glom(., taxrank = "ARG")  %>% 
  filter_taxa(function(x) sum(x > 0) > 0, TRUE) %>% 
  transform_sample_counts(function(x) x/sum(x) * 100) %>% 
  subset_taxa(AMR_sub_class_multi %!in% c("Unclassified", "multidrug (unclassified)", "MLS (unclassified)")) %>% 
  # physeq_sel_tax_table(c("PathoFact_AMR_AMR_sub_class_multi")) %>% 
  # prune_samples(samples, .) %>% 
  filter_taxa(function(x) sum(x > 0) > 0, TRUE) %>% 
  # physeq_add_metadata(metadata_mapping,
  #                     sample_column = "sample_id") %>% 
  # ps_mutate(Day_of_Treatment = as.numeric(Day_of_Treatment)) %>% 
  # ps_arrange(-Day_of_Treatment) %>% 
  speedyseq::psmelt() %>%
  # left_join(tax_mapping,
  #           by = c("Best_Hit_ARO" = "Best_Hit_ARO"),
  #           suffix = c("_x", "")) %>%
  # mutate(Sample = fct_reorder(Sample, Model,Treatment,Fermentation,Antibiotic_mg.mL, Day_of_Treatment )) %>%
  mutate(Abundance = na_if(Abundance, 0)) %>%
  # mutate(logTPM = log10(Abundance + 1)) %>% 
  # mutate(log10Abundance = log10(Abundance)) %>% 
  select(Sample, ARG, Abundance) %>%
  group_by(Sample, ARG) %>%
  pivot_wider(names_from =  Sample, ARG,
              values_from = Abundance) %>%
  column_to_rownames("ARG") %>% 
  replace(is.na(.), 0)  %>%
  t() %>%
  # sqrt() %>% 
  vegan::vegdist(na.rm = TRUE, method = "robust.aitchison") %>%
  hclust(method = "ward.D2") -> clust_t_AMR

# NO : median, centroid
# "ward.D", "ward.D2", "single", "complete", "average" (= UPGMA), "mcquitty" (= WPGMA), "median" (= WPGMC) or "centroid" (= UPGMC)
```


```{r}
pheatmap_samples %>%
  replace(is.na(.), 0)  %>%
  t() %>%
  vegan::vegdist(na.rm = TRUE, method = "robust.aitchison") %>%
  hclust(method = "ward.D2") -> clust_t


pheatmap_samples %>%
  replace(is.na(.), 0)  %>%
  vegan::vegdist(na.rm = TRUE, method = "euc") %>%
  # cor() %>%
  hclust(method = "ward.D2") -> clust
```


Pheatmap:

dendro class

```{r}
pheatmap_samples %>%
  pheatmap::pheatmap(cluster_rows = clust ,
                     scale = "row",
                     cluster_cols = clust_t,
                     na_col = "transparent",
                     annotation_row = tax_mapping %>% rownames_to_column('tmp_col') %>%
                       filter(AMR_sub_class_multi %in% rownames(pheatmap_samples)) %>%
                       distinct(AMR_sub_class_multi, .keep_all = TRUE) %>%
                       left_join(., 
                                 row_anot,
                                 by = c("AMR_sub_class_multi" = "AMR_sub_class_multi")) %>% 
                       column_to_rownames("AMR_sub_class_multi") %>%
                       select(Resistance_mechanism_multi, meanprop, nGN, nAMR) %>% 
                       rename(mechanism = Resistance_mechanism_multi),
                     annotation_colors = list("mechanism" = resistance_type,
                                              "Treatment" = treat_col),
                     annotation_col = pheatmap_df %>%
                       select(Day_of_Treatment, Antibiotic_mg.L, Treatment, Sample, Model2) %>% 
                       mutate(Day_of_Treatment = ifelse(Day_of_Treatment < 0, NA,Day_of_Treatment)) %>% 
                       distinct(Sample, .keep_all = TRUE) %>% 
                       column_to_rownames("Sample"),
                     annotation_legend = TRUE)  -> p_pheat

p_pheat

# p_pheat %>% 
#   ggsave(plot = .,bg = "transparent",scale = 2,
#          units = "mm",
#          dpi = 600,
#          height = 0.618 * 84 * 2,
#          width = 84  * 1.5,
#          filename = "~/Desktop/Resistome_heatmap_class.eps")

p_pheat %>%
  export::graph2ppt(append = TRUE, width = 317.48031496 * 4, height = 0.618 * 317.48031496 * 6 , paper = "A4",  scaling = 1,
                    file = out_pptx)
```

dendro AMR

```{r}
pheatmap_samples %>%
  pheatmap::pheatmap(cluster_rows = clust ,
                     scale = "row",
                     cluster_cols = clust_t_AMR,
                     na_col = "transparent",
                     annotation_row = tax_mapping %>% rownames_to_column('tmp_col') %>%
                       filter(AMR_sub_class_multi %in% rownames(pheatmap_samples)) %>%
                       distinct(AMR_sub_class_multi, .keep_all = TRUE) %>%
                       left_join(., 
                                 row_anot,
                                 by = c("AMR_sub_class_multi" = "AMR_sub_class_multi")) %>% 
                       column_to_rownames("AMR_sub_class_multi") %>%
                       select(Resistance_mechanism_multi, meanprop, nGN, nAMR) %>% 
                       rename(mechanism = Resistance_mechanism_multi),
                     annotation_colors = list("mechanism" = resistance_type,
                                              "Treatment" = treat_col),
                     annotation_col = pheatmap_df %>%
                       select(Day_of_Treatment, Antibiotic_mg.L, Treatment, Sample, Model2) %>% 
                       mutate(Day_of_Treatment = ifelse(Day_of_Treatment < 0, NA,Day_of_Treatment)) %>% 
                       distinct(Sample, .keep_all = TRUE) %>% 
                       column_to_rownames("Sample"),
                     annotation_legend = TRUE)

p_pheat

# p_pheat %>% 
#   ggsave(plot = .,bg = "transparent",scale = 2,
#          units = "mm",
#          dpi = 600,
#          height = 0.618 * 84 * 2,
#          width = 84  * 1.5,
#          filename = "~/Desktop/Resistome_heatmap_class.eps")

p_pheat %>%
  export::graph2ppt(append = TRUE, width = 317.48031496 * 4, height = 0.618 * 317.48031496 * 6 , paper = "A4",  scaling = 1,
                    file = out_pptx)
```


dendro gN

```{r}
pheatmap_samples %>%
  pheatmap::pheatmap(cluster_rows = clust ,
                     scale = "row",
                     cluster_cols = clust_t_OTU,
                     na_col = "transparent",
                     annotation_row = tax_mapping %>% rownames_to_column('tmp_col') %>%
                       filter(AMR_sub_class_multi %in% rownames(pheatmap_samples)) %>%
                       distinct(AMR_sub_class_multi, .keep_all = TRUE) %>%
                       left_join(., 
                                 row_anot,
                                 by = c("AMR_sub_class_multi" = "AMR_sub_class_multi")) %>% 
                       column_to_rownames("AMR_sub_class_multi") %>%
                       select(Resistance_mechanism_multi, meanprop, nGN, nAMR) %>% 
                       rename(mechanism = Resistance_mechanism_multi),
                     annotation_colors = list("mechanism" = resistance_type,
                                              "Treatment" = treat_col),
                     annotation_col = pheatmap_df %>%
                       select(Day_of_Treatment, Antibiotic_mg.L, Treatment, Sample) %>% 
                       mutate(Day_of_Treatment = ifelse(Day_of_Treatment < 0, NA,Day_of_Treatment)) %>% 
                       distinct(Sample, .keep_all = TRUE) %>% 
                       column_to_rownames("Sample"),
                     annotation_legend = TRUE)  -> p_pheat

p_pheat

# p_pheat %>% 
#   ggsave(plot = .,bg = "transparent",scale = 2,
#          units = "mm",
#          dpi = 600,
#          height = 0.618 * 84 * 2,
#          width = 84  * 1.5,
#          filename = "~/Desktop/Resistome_heatmap_class.eps")

p_pheat %>%
  export::graph2ppt(append = TRUE, width = 317.48031496 * 8, height = 0.618 * 317.48031496 * 10 , paper = "A4",  scaling = 1,
                    file = out_pptx)
```

export underlying data:

```{r}
tax_mapping %>% rownames_to_column('tmp_col') %>%
  filter(AMR_sub_class_multi %in% rownames(pheatmap_samples)) %>%
  distinct(AMR_sub_class_multi, .keep_all = TRUE) %>%
  left_join(., 
            row_anot,
            by = c("AMR_sub_class_multi" = "AMR_sub_class_multi")) %>% 
  column_to_rownames("AMR_sub_class_multi") %>%
  select(-ARG) %>% 
  # select(Resistance_mechanism_multi, meanprop, nGN, nAMR) %>%
  rename(mechanism = Resistance_mechanism_multi) %>% 
  data.frame() %>% 
  # arrange(Day_of_Treatment) %>% 
  xlsx::write.xlsx(x = . ,
                   append = TRUE,
                   file = out_xlsx,
                   sheetName = "drug_class_heatmap")
```

only cephalosporin:

```{r}
ps_AMR_all %>% 
  physeq_sel_tax_table(c("AMR_sub_class", "AMR_sub_class_multi", "ARG")) %>% 
  subset_taxa(AMR_sub_class_multi == "cephalosporin") %>% 
  speedyseq::tax_glom(., taxrank = "ARG")  %>% 
  physeq_sel_tax_table(c("ARG")) %>% 
  # prune_samples(samples, .) %>% 
  # physeq_add_metadata(metadata_mapping,
  #                     sample_column = "sample_id") %>% 
  filter_taxa(function(x) sum(x > 0) > 0, TRUE) %>%
  transform_sample_counts(function(x) x/sum(x) * 100) %>% 
  # filter_taxa(function(x) sum(x > 0) > 0, TRUE) %>%
  # subset_taxa(PathoFact_AMR_AMR_sub_class_multi %!in% c("Unclassified", "multidrug (unclassified)", "MLS (unclassified)")) %>% 
  # ps_mutate(Day_of_Treatment = as.numeric(Day_of_Treatment)) %>% 
  ps_arrange(-Day_of_Treatment) %>% 
  speedyseq::psmelt() %>%
  # left_join(tax_mapping,
  #           by = c("Best_Hit_ARO" = "Best_Hit_ARO"),
  #           suffix = c("_x", "")) %>%
  # mutate(Sample = fct_reorder(Sample, Model,Treatment,Fermentation,Antibiotic_mg.mL, Day_of_Treatment )) %>%
  mutate(Abundance = na_if(Abundance, 0)) %>%
  mutate(logAbundance = log10(Abundance + 1)) %>% 
  mutate(ARG = ARG %>% stringr::str_trunc(45,  side ="center")) -> pheatmap_df


pheatmap_df %>%
  # replace(is.na(.), 0)  %>%
  mutate(log10Abundance = log10(Abundance)) %>% 
  select(Sample, ARG, Abundance) %>%
  filter(Abundance > 0) %>%
  # filter(Abundance > 1) %>%
  group_by(Sample, ARG) %>%
  pivot_wider(names_from =  Sample, ARG,
              values_from = Abundance) %>%
  # mutate(PathoFact_AMR_ARG = PathoFact_AMR_ARG %>%  str_to_lower(.)) %>% 
  # %>% str_to_lower(., locale = "en")
  column_to_rownames("ARG") -> pheatmap_samples

pheatmap_df %>%
  # replace(is.na(.), 0)  %>%
  mutate(log10Abundance = log10(Abundance)) %>% 
  select(Sample, ARG, Abundance) %>%
  group_by(ARG) %>% 
  summarise(meanAbundance = mean(Abundance, na.rm = TRUE)) -> mean_ab


pheatmap_samples %>%
  replace(is.na(.), 0)  %>%
  t() %>%
  vegan::vegdist(na.rm = TRUE) %>%
  hclust(method = "ward.D2") -> clust_t

pheatmap_samples %>%
  replace(is.na(.), 0)  %>%
  vegan::vegdist(na.rm = TRUE, method = "euc") %>%
  # cor() %>%
  hclust(method = "ward.D2") -> clust

pheatmap_samples %>%
  pheatmap::pheatmap(fontsize_col = 6,
                     fontsize = 6,
                     cluster_rows = FALSE ,
                     scale = "row",
                     cluster_cols = clust_t,
                     na_col = "transparent",
                     annotation_row = tax_mapping %>% rownames_to_column('tmp_col') %>%
                       filter(ARG %in% rownames(pheatmap_samples)) %>%
                       distinct(ARG, .keep_all = TRUE) %>%
                       left_join(., 
                                 mean_ab,
                                 by = c("ARG" = "ARG")) %>% 
                       column_to_rownames("ARG") %>%
                       select(meanAbundance),
                     annotation_colors = list("Treatment" = treat_col),
                     annotation_col = pheatmap_df %>%
                       select(Antibiotic_mg.L,Day_of_Treatment, Treatment, Model2, Sample) %>% 
                       distinct(Sample, .keep_all = TRUE) %>% 
                       column_to_rownames("Sample"))  -> p_pheat


p_pheat

# p_pheat %>% 
#   ggsave(plot = .,bg = "transparent",scale = 2,
#          units = "mm",
#          dpi = 600,
#          height = 0.618 * 84,
#          width = 84  ,
#          filename = "~/Desktop/Resistome_heatmap_CTX.eps")


p_pheat %>%
  export::graph2ppt(append = TRUE, width = 317.48031496 * 1.5, height = 0.618 * 317.48031496 * 2 , paper = "A4",  scaling = 1,
                    file = out_pptx)
```


same but only ....:

```{r}
ps_AMR_all %>% 
  physeq_sel_tax_table(c("AMR_sub_class", "AMR_sub_class_multi", "ARG")) %>% 
  subset_taxa(AMR_sub_class_multi == "glycopeptide antibiotic") %>% 
  speedyseq::tax_glom(., taxrank = "ARG")  %>% 
  physeq_sel_tax_table(c("ARG")) %>% 
  # prune_samples(samples, .) %>% 
  # physeq_add_metadata(metadata_mapping,
  #                     sample_column = "sample_id") %>% 
  transform_sample_counts(function(x) x/sum(x) * 100) %>% 
  # filter_taxa(function(x) sum(x > 0) > 0, TRUE) %>% 
  # subset_taxa(PathoFact_AMR_AMR_sub_class_multi %!in% c("Unclassified", "multidrug (unclassified)", "MLS (unclassified)")) %>% 
  # ps_mutate(Day_of_Treatment = as.numeric(Day_of_Treatment)) %>% 
  ps_arrange(-Day_of_Treatment) %>% 
  speedyseq::psmelt() %>%
  # left_join(tax_mapping,
  #           by = c("Best_Hit_ARO" = "Best_Hit_ARO"),
  #           suffix = c("_x", "")) %>%
  # mutate(Sample = fct_reorder(Sample, Model,Treatment,Fermentation,Antibiotic_mg.mL, Day_of_Treatment )) %>%
  mutate(Abundance = na_if(Abundance, 0)) %>%
  mutate(logAbundance = log10(Abundance + 1)) %>% 
  mutate(ARG = ARG %>% stringr::str_trunc(45,  side ="center")) -> pheatmap_df


pheatmap_df %>%
  # replace(is.na(.), 0)  %>%
  mutate(log10Abundance = log10(Abundance)) %>% 
  select(Sample, ARG, Abundance) %>%
  filter(Abundance > 0) %>%
  # filter(Abundance > 1) %>%
  group_by(Sample, ARG) %>%
  pivot_wider(names_from =  Sample, ARG,
              values_from = Abundance) %>%
  # mutate(PathoFact_AMR_ARG = PathoFact_AMR_ARG %>%  str_to_lower(.)) %>% 
  # %>% str_to_lower(., locale = "en")
  column_to_rownames("ARG") -> pheatmap_samples

pheatmap_df %>%
  # replace(is.na(.), 0)  %>%
  mutate(log10Abundance = log10(Abundance)) %>% 
  select(Sample, ARG, Abundance) %>%
  group_by(ARG) %>% 
  summarise(meanAbundance = mean(Abundance, na.rm = TRUE)) -> mean_ab


pheatmap_samples %>%
  replace(is.na(.), 0)  %>%
  t() %>%
  vegan::vegdist(na.rm = TRUE) %>%
  hclust(method = "ward.D2") -> clust_t

pheatmap_samples %>%
  replace(is.na(.), 0)  %>%
  vegan::vegdist(na.rm = TRUE, method = "euc") %>%
  # cor() %>%
  hclust(method = "ward.D2") -> clust

pheatmap_samples %>%
  pheatmap::pheatmap(fontsize_col = 6,
                     fontsize = 6,
                     cluster_rows = FALSE ,
                     scale = "row",
                     cluster_cols = clust_t,
                     na_col = "transparent",
                     annotation_row = tax_mapping %>% rownames_to_column('tmp_col') %>%
                       filter(ARG %in% rownames(pheatmap_samples)) %>%
                       distinct(ARG, .keep_all = TRUE) %>%
                       left_join(., 
                                 mean_ab,
                                 by = c("ARG" = "ARG")) %>% 
                       column_to_rownames("ARG") %>%
                       select(meanAbundance),
                     annotation_colors = list("Treatment" = treat_col),
                     annotation_col = pheatmap_df %>%
                       select(Antibiotic_mg.L,Day_of_Treatment, Treatment, Model2, Sample) %>% 
                       distinct(Sample, .keep_all = TRUE) %>% 
                       column_to_rownames("Sample"))  -> p_pheat


p_pheat

# p_pheat %>% 
#   ggsave(plot = .,bg = "transparent",scale = 2,
#          units = "mm",
#          dpi = 600,
#          height = 0.618 * 84 * 1.5,
#          width = 84  ,
#          filename = "~/Desktop/Resistome_heatmap_VAN.eps")


p_pheat %>%
  export::graph2ppt(append = TRUE, width = 317.48031496 * 1.5 , height = 0.618 * 317.48031496 * 3 , paper = "A4",  scaling = 1,
                    file = out_pptx)
```



all:


```{r}
ps_AMR_all %>% 
  physeq_sel_tax_table(c("AMR_sub_class", "AMR_sub_class_multi", "ARG")) %>% 
  speedyseq::tax_glom(., taxrank = "ARG")  %>% 
  physeq_sel_tax_table(c("ARG")) %>% 
  # prune_samples(samples, .) %>% 
  # physeq_add_metadata(metadata_mapping,
  #                     sample_column = "sample_id") %>% 
  transform_sample_counts(function(x) x/sum(x) * 100) %>% 
  # filter_taxa(function(x) sum(x > 0) > 0, TRUE) %>% 
  # subset_taxa(PathoFact_AMR_AMR_sub_class_multi %!in% c("Unclassified", "multidrug (unclassified)", "MLS (unclassified)")) %>% 
  # ps_mutate(Day_of_Treatment = as.numeric(Day_of_Treatment)) %>% 
  ps_arrange(-Day_of_Treatment) %>% 
  speedyseq::psmelt() %>%
  # left_join(tax_mapping,
  #           by = c("Best_Hit_ARO" = "Best_Hit_ARO"),
  #           suffix = c("_x", "")) %>%
  # mutate(Sample = fct_reorder(Sample, Model,Treatment,Fermentation,Antibiotic_mg.mL, Day_of_Treatment )) %>%
  mutate(Abundance = na_if(Abundance, 0)) %>%
  mutate(logAbundance = log10(Abundance + 1)) %>% 
  mutate(ARG = ARG %>% stringr::str_trunc(45,  side ="center")) -> pheatmap_df


pheatmap_df %>%
  # replace(is.na(.), 0)  %>%
  mutate(log10Abundance = log10(Abundance)) %>% 
  select(Sample, ARG, Abundance) %>%
  filter(Abundance > 0) %>% 
  # filter(Abundance > 1) %>% 
  group_by(Sample, ARG) %>%
  pivot_wider(names_from =  Sample, ARG,
              values_from = Abundance) %>%
  # mutate(PathoFact_AMR_ARG = PathoFact_AMR_ARG %>%  str_to_lower(.)) %>% 
  # %>% str_to_lower(., locale = "en")
  column_to_rownames("ARG") -> pheatmap_samples

pheatmap_df %>%
  # replace(is.na(.), 0)  %>%
  mutate(log10Abundance = log10(Abundance)) %>% 
  select(Sample, ARG, Abundance) %>%
  group_by(ARG) %>% 
  summarise(meanAbundance = mean(Abundance, na.rm = TRUE)) -> mean_ab


pheatmap_samples %>%
  replace(is.na(.), 0)  %>%
  t() %>%
  vegan::vegdist(na.rm = TRUE) %>%
  hclust(method = "ward.D2") -> clust_t

pheatmap_samples %>% 
  replace(is.na(.), 0)  %>%
  vegan::vegdist(na.rm = TRUE, method = "robust.aitchison") %>%
  # cor() %>%
  hclust(method = "ward.D2") -> clust

pheatmap_samples %>%
  pheatmap::pheatmap(fontsize_col = 1,
                     fontsize = 1,
                     cluster_rows = FALSE ,
                     scale = "row",
                     cluster_cols = clust_t,
                     na_col = "transparent",
                     annotation_row = tax_mapping %>% rownames_to_column('tmp_col') %>%
                       filter(ARG %in% rownames(pheatmap_samples)) %>%
                       distinct(ARG, .keep_all = TRUE) %>%
                       left_join(., 
                                 mean_ab,
                                 by = c("ARG" = "ARG")) %>% 
                       column_to_rownames("ARG") %>%
                       select(AMR_sub_class_multi, meanAbundance),
                     # annotation_colors = list("AMR_sub_class_multi" = drug_classe_multi,
                     # "Treatment" = treat_col),
                     annotation_col = pheatmap_df %>%
                       select(Antibiotic_mg.L,Day_of_Treatment, Treatment, Model2, Sample) %>% 
                       distinct(Sample, .keep_all = TRUE) %>% 
                       column_to_rownames("Sample"))  -> p_pheat


p_pheat

# p_pheat %>% 
#   ggsave(plot = .,bg = "transparent",scale = 2,
#          units = "mm",
#          dpi = 600,
#          height = 0.618 * 84 * 2.5,
#          width = 84  * 0.5,
#          filename = "~/Desktop/Resistome_heatmap_all_gn.eps")


p_pheat %>%
  export::graph2ppt(append = TRUE, width = 317.48031496 * 12, height = 0.618 * 317.48031496 * 60 , paper = "A2",  scaling = 2,
                    file = out_pptx)
```

# Beta-div resistome:

Generate a temporary phyloseq object first - containing PathoFact_AMR_ARG - which will be used for the 2 models:
*TODO*: filter based on BP's expertise? Unclassified... ?

## AMR gNs:

```{r}
ps_AMR_all %>% 
  physeq_sel_tax_table(c("AMR_sub_class", "AMR_sub_class_multi", "ARG"))  %>%
  speedyseq::tax_glom(., taxrank = "ARG")  %>%
  filter_taxa(function(x) sum(x > 0) > 0, TRUE) -> ps_tmp
# %>% 
#   prune_samples(samples, .) %>% 
#   physeq_add_metadata(metadata_mapping,
#                       sample_column = "sample_id") -> ps_tmp

ps_tmp
```


### Generate Aitchinson PCA:

```{r}
ps_tmp %>%
  subset_samples(Model == "Chicken") %>% 
  filter_taxa(function(x) sum(x > 0) > 0, TRUE) %>%
  # microbiome::transform(transform = "compositional") %>% 
  phyloseq_plot_bdiv(dlist = NULL,
                     m = "CoDa",
                     seed = 123,
                     axis1 = 1,
                     axis2 = 2) -> coda_resistome


ps_tmp %>%
  subset_samples(Model == "Chicken") %>%
  filter_taxa(function(x) sum(x > 0) > 0, TRUE) %>%
  microbiome::transform(transform = "compositional") %>%
  speedyseq::psmelt() %>%
  pivot_wider(names_from =  Sample, ARG,
              values_from = Abundance) %>%
  # mutate(PathoFact_AMR_ARG = PathoFact_AMR_ARG %>%  str_to_lower(.)) %>%
  # %>% str_to_lower(., locale = "en")
  column_to_rownames("ARG") %>%
  t() %>%
  replace(is.na(.), 0)  %>%
  vegan::vegdist(na.rm = TRUE, method = "robust.aitchison") %>% 
  magrittr::divide_by(100) -> robust.aitchison
#   # cor() %>%
#   # hclust(method = "ward.D2") -> clust
# 
# 
plot(
  robust.aitchison ,
  as.matrix(coda_resistome$aidist)[colnames(robust.aitchison %>% as.matrix()), rownames(robust.aitchison %>% as.matrix())] %>% as.dist()
)
```


```{r}
list = NULL

list$robust.aitchison <- robust.aitchison

ps_tmp %>%
  subset_samples(Model == "Chicken") %>% 
  filter_taxa(function(x) sum(x > 0) > 0, TRUE) %>%
  # microbiome::transform(transform = "compositional") %>% 
  phyloseq_plot_bdiv(dlist = list,
                     m = "PCoA",
                     seed = 123,
                     axis1 = 1,
                     axis2 = 2) -> robust.aitchison.pcoa


```


### Initial PCA:

```{r}
coda_resistome$PCA$layers[[1]] = NULL

coda_resistome$PCA + geom_point(size= 3.5,
                                aes(colour = Treatment ,
                                    shape = Antibiotic_mg.L,
                                    alpha = NULL)) +
  geom_path(data = coda_resistome$PCA$data %>%
              arrange(Day_of_Treatment) ,
            aes(colour = Treatment, group = Reactor_Treatment_Dose), #interaction(Model2, Antibiotic, Reactor_Treatment_Dose)),
            arrow = arrow(
              angle = 30, length = unit(0.15, "inches"),
              ends = "last", type = "open"
            ), linetype = "longdash", size = 0.1) +
  theme_light() +
  scale_color_manual(name = "", values = treat_col,
                     na.value = "black") +
  scale_fill_manual(name = "", values = treat_col,
                    na.value = "black") +
  scale_shape_manual(name = "" ,values = c(15,16,18,19), na.value =  17) +
  theme(legend.position = "bottom") +
  facet_grid(. ~ Model) -> pca_treat


pca_treat
```

```{r}
robust.aitchison.pcoa$robust.aitchison$layers[[1]] = NULL
robust.aitchison.pcoa$robust.aitchison$layers[[1]] = NULL
robust.aitchison.pcoa$robust.aitchison$layers[[2]] = NULL

robust.aitchison.pcoa$robust.aitchison + geom_point(size = 3.5,
                                                    aes(colour = Treatment ,
                                                        shape = Antibiotic_mg.L,
                                                        alpha = NULL)) +
  geom_path(data = robust.aitchison.pcoa$robust.aitchison$data %>% 
              arrange(Day_of_Treatment) ,
            aes(colour = Treatment, group = Reactor_Treatment_Dose), #interaction(Model2, Antibiotic, Reactor_Treatment_Dose)),
            arrow = arrow(
              angle = 30, length = unit(0.15, "inches"),
              ends = "last", type = "open"
            ), linetype = "longdash", size = 0.1) +
  theme_light() +
  scale_color_manual(name = "", values = treat_col,
                     na.value = "black") +
  scale_fill_manual(name = "", values = treat_col,
                    na.value = "black") +
  scale_shape_manual(name = "" ,values = c(15,16,18,19), na.value =  17) +
  theme(legend.position = "bottom") + 
  facet_grid(. ~ Model) -> pca_treat

pca_treat
```

```{r}
# pca_treat%>% 
#   ggsave(plot = .,bg = "transparent",scale = 2,
#          units = "mm",
#          dpi = 600,
#          height = 0.618 * 84,
#          width = 84,
#          filename = "~/Desktop/Resistome_beta_chick.eps")

pca_treat%>%
  export::graph2ppt(append = TRUE, width = 317.48031496  , height = 0.618 * 317.48031496  , paper = "A4",  scaling = 2,
                    file = out_pptx)
```

We generate a empty plot for vectors: 

```{r, warning = FALSE, results='hide'}
pca_treat + 
  scale_fill_manual(values = c("transparent")) + 
  scale_color_manual(values = c(rep("transparent", 8))) + 
  theme(panel.border = element_blank(), panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(), axis.line = element_line(colour = "black")) -> empty_plot_tmp
```

### Next, we add AM drug classes info as envfit vectors on the generated 'empty' plot:

```{r, warning = FALSE, results='hide'}
phyloseq_add_taxa_vector_fix(dist = list$robust.aitchison,
                             phyloseq = ps_tmp %>%  
                               physeq_sel_tax_table(c("AMR_sub_class_multi")) %>% 
                               subset_samples(Model == "Chicken") %>% 
                               filter_taxa(function(x) sum(x > 0) > 0, TRUE),
                             figure_ord = empty_plot_tmp,
                             taxrank_glom = "AMR_sub_class_multi") -> envfit_class

envfit_class$signenvfit %>% 
  mutate_if(is.numeric, ~round(., 4)) %>% 
  DT::datatable()
```

```{r}
envfit_class$plot


# ppp$plot + 
#   ggrepel::geom_text_repel(cex=2,
#                            aes(label= Phase),
#                            segment.color = 'black',
#                            segment.size = 0.5,
#                            ppp$plot$data %>% filter(Phase %in% c("Stab")) %>% unique()
#   ) -> pbeta

envfit_class$plot %>%
  export::graph2ppt(append = TRUE, width = 317.48031496, height = 0.618 * 317.48031496, paper = "A4",  scaling = 2,
                    file = out_pptx)


# envfit_class$plot %>% 
#   ggsave(plot = .,bg = "transparent",scale = 2,
#          units = "mm",
#          dpi = 600,
#          height = 0.618 * 84,
#          width = 84,
#          filename = "~/Desktop/Resistome_beta_chick_class.eps")
```

### Next, we add MGE prediction genes abundance as envfit vectors on the generated 'empty' plot:

```{r}
ps_AMR_all %>%
  physeq_sel_tax_table(c("MGE_prediction_clean"))  %>%
  speedyseq::tax_glom(., taxrank = "MGE_prediction_clean") %>%
  # physeq_sel_tax_table(c("PathoFact_AMR_AMR_sub_class_multi")) %>%
  # prune_samples(samples, .) %>%
  subset_samples(Model == "Chicken") %>%
  filter_taxa(function(x) sum(x > 0) > 0, TRUE) %>% 
  microViz::tax_mutate(MGE_prediction_clean = case_when(MGE_prediction_clean %in% c("MAG", "chromosome")  ~ "MAG - chromosome", TRUE ~ MGE_prediction_clean)) -> ps_tmp_envfit
# physeq_add_metadata(metadata_mapping,
#                     sample_column = "sample_id") -> ps_tmp_envfit
# 
# # custom PathoFact_AMR_MGE_prediction cleaning- should be done once at the begining...
# 
# ps_tmp_envfit %>%
#   tax_table() %>%
#   data.frame() %>%
#   # mutate(MGE_prediction_clean = case_when(grepl("ambiguous", MGE_prediction_clean) ~ "ambiguous", TRUE ~ MGE_prediction_clean)) %>%
#   mutate(MGE_prediction_clean = case_when(MGE_prediction_clean %in% c("MAG", "chromosome")  ~ "MAG - chromosome", TRUE ~ MGE_prediction_clean)) %>%
#   as.matrix() -> tax_table(ps_tmp_envfit)

ps_tmp_envfit %>%
  phyloseq_add_taxa_vector_fix(dist = coda_resistome$aidist,
                               phyloseq = . ,
                               taxnames_rm = c("unclassified", "ambiguous"),
                               vector_color = "tomato",
                               figure_ord = empty_plot_tmp,
                               taxrank_glom = "MGE_prediction_clean") -> envfit_MGE



envfit_MGE$signenvfit %>%
  mutate_if(is.numeric, ~round(., 4)) %>%
  DT::datatable()
```

```{r}
# 
envfit_MGE$plot
# 
# # ppp$plot + 
# #   ggrepel::geom_text_repel(cex=2,
# #                            aes(label= Phase),
# #                            segment.color = 'black',
# #                            segment.size = 0.5,
# #                            ppp$plot$data %>% filter(Phase %in% c("Stab")) %>% unique()
# #   ) -> pbeta
# 
# envfit_MGE$plot %>% 
#   ggsave(plot = .,bg = "transparent",scale = 2,
#          units = "mm",
#          dpi = 600,
#          height = 0.618 * 84,
#          width = 84,
#          filename = "~/Desktop/Resistome_beta_chick_MGE.eps")
# 
envfit_MGE$plot %>%
  export::graph2ppt(append = TRUE, width = 317.48031496, height = 0.618 * 317.48031496, paper = "A4",  scaling = 2,
                    file = out_pptx)
```

### Finally, we add AMR genes abunbundance as envfit vectors - similarly as before:

```{r}
phyloseq_add_taxa_vector_fix(dist = coda_resistome$aidist,
                             ggrepel_force = 30,
                             phyloseq = ps_tmp %>%
                               physeq_sel_tax_table(c("ARG")) %>%
                               subset_samples(Model == "Chicken") %>%
                               filter_taxa(function(x) sum(x > 0) > 0, TRUE) %>%
                               filter_taxa(., function(x){sum(x > 0.000005) > 4}, prune = TRUE), # only keeping dominant and occurent ones
                             figure_ord = empty_plot_tmp,
                             top_r = 20,
                             vector_color = "chartreuse",
                             taxrank_glom = "ARG") -> pca_env_ARG
# 
pca_env_ARG$signenvfit %>%
  mutate_if(is.numeric, ~round(., 4)) %>%
  DT::datatable()
```

```{r}
# pca_env_ARG$plot
```

```{r}
# pca_env_ARG$vectors
# 
# pca_env_ARG$vectors %>% 
#   ggsave(plot = .,bg = "transparent",scale = 2,
#          units = "mm",
#          dpi = 600,
#          height = 0.618 * 84,
#          width = 84,
#          filename = "~/Desktop/Resistome_beta_chick_ARG.eps")
# 
# pca_env_ARG$vectors  %>%
#   export::graph2ppt(append = TRUE, width = 317.48031496  , height = 0.618 * 317.48031496  , paper = "A4",  scaling = 2,
#                     file = out_pptx)
```

## Gn

```{r}
# ps_AMR_all %>% 
#   physeq_sel_tax_table(c("AMR_sub_class", "AMR_sub_class_multi", "ARG"))  %>%
#   # speedyseq::tax_glom(., taxrank = "PathoFact_AMR_ARG")  %>%
#   filter_taxa(function(x) sum(x > 0) > 0, TRUE) -> ps_tmp
# # prune_samples(samples, .) %>% 
# # physeq_add_metadata(metadata_mapping,
# #                     sample_column = "sample_id") -> ps_tmp
# 
# ps_tmp
```


### Generate Aitchinson PCA:

```{r}
# ps_tmp %>%
#   subset_samples(Model == "Chicken") %>% 
#   filter_taxa(function(x) sum(x > 0) > 0, TRUE) %>%
#   # microbiome::transform(transform = "compositional") %>% 
#   phyloseq_plot_bdiv(dlist = NULL,
#                      m = "CoDa",
#                      seed = 123,
#                      axis1 = 1,
#                      axis2 = 2) -> coda_resistome
# 
# 
# ps_tmp %>%
#   subset_samples(Model == "Chicken") %>%
#   filter_taxa(function(x) sum(x > 0) > 0, TRUE) %>%
#   microbiome::transform(transform = "compositional") %>%
#   speedyseq::psmelt() %>%
#   pivot_wider(names_from =  Sample, OTU,
#               values_from = Abundance) %>%
#   # mutate(PathoFact_AMR_ARG = PathoFact_AMR_ARG %>%  str_to_lower(.)) %>%
#   # %>% str_to_lower(., locale = "en")
#   column_to_rownames("OTU") %>%
#   t() %>%
#   replace(is.na(.), 0)  %>%
#   vegan::vegdist(na.rm = TRUE, method = "robust.aitchison") %>% 
#   magrittr::divide_by(100) -> robust.aitchison
# #   # cor() %>%
# #   # hclust(method = "ward.D2") -> clust
# # 
# # 
# plot(
#   robust.aitchison ,
#   as.matrix(coda_resistome$aidist)[colnames(robust.aitchison %>% as.matrix()), rownames(robust.aitchison %>% as.matrix())] %>% as.dist()
# )
```


```{r}
# list = NULL
# 
# list$robust.aitchison <- robust.aitchison
# 
# ps_tmp %>%
#   subset_samples(Model == "Chicken") %>% 
#   filter_taxa(function(x) sum(x > 0) > 0, TRUE) %>%
#   # microbiome::transform(transform = "compositional") %>% 
#   phyloseq_plot_bdiv(dlist = list,
#                      m = "PCoA",
#                      seed = 123,
#                      axis1 = 1,
#                      axis2 = 2) -> robust.aitchison.pcoa
# 

```


### Initial PCA:

```{r}
# coda_resistome$PCA$layers[[1]] = NULL
# 
# coda_resistome$PCA + geom_point(size= 3.5,
#                                 aes(colour = Treatment ,
#                                     shape = Antibiotic_mg.L,
#                                     alpha = NULL)) +
#   geom_path(data = coda_resistome$PCA$data %>%
#               arrange(Day_of_Treatment) ,
#             aes(colour = Treatment, group = Reactor_Treatment_Dose), #interaction(Model, Antibiotic, Reactor_Treatment_Dose)),
#             arrow = arrow(
#               angle = 30, length = unit(0.15, "inches"),
#               ends = "last", type = "open"
#             ), linetype = "longdash", size = 0.1) +
#   theme_light() +
#   scale_color_manual(name = "", values = treat_col,
#                      na.value = "black") +
#   scale_fill_manual(name = "", values = treat_col,
#                     na.value = "black") +
#   scale_shape_manual(name = "" ,values = c(15,16,18,19), na.value =  17) +
#   theme(legend.position = "bottom") +
#   facet_grid(. ~ Model) -> pca_treat
# 
# 
# pca_treat
```

```{r}
# robust.aitchison.pcoa$robust.aitchison$layers[[1]] = NULL
# robust.aitchison.pcoa$robust.aitchison$layers[[1]] = NULL
# robust.aitchison.pcoa$robust.aitchison$layers[[2]] = NULL
# 
# robust.aitchison.pcoa$robust.aitchison + geom_point(size = 3.5,
#                                                     aes(colour = Treatment ,
#                                                         shape = Antibiotic_mg.L,
#                                                         alpha = NULL)) +
#   geom_path(data = robust.aitchison.pcoa$robust.aitchison$data %>% 
#               arrange(Day_of_Treatment) ,
#             aes(colour = Treatment, group = Reactor_Treatment_Dose), #interaction(Model, Antibiotic, Reactor_Treatment_Dose)),         
#             arrow = arrow(
#               angle = 30, length = unit(0.15, "inches"),
#               ends = "last", type = "open"
#             ), linetype = "longdash", size = 0.1) +
#   theme_light() +
#   scale_color_manual(name = "", values = treat_col,
#                      na.value = "black") +
#   scale_fill_manual(name = "", values = treat_col,
#                     na.value = "black") +
#   scale_shape_manual(name = "" ,values = c(15,16,18,19), na.value =  17) +
#   theme(legend.position = "bottom") + 
#   facet_grid(. ~ Model) -> pca_treat
# 
# pca_treat
```

```{r}
# pca_treat%>% 
#   ggsave(plot = .,bg = "transparent",scale = 2,
#          units = "mm",
#          dpi = 600,
#          height = 0.618 * 84,
#          width = 84,
#          filename = "~/Desktop/Resistome_beta_chick_gn.eps")

# pca_treat%>%
#   export::graph2ppt(append = TRUE, width = 317.48031496  , height = 0.618 * 317.48031496  , paper = "A4",  scaling = 2,
#                     file = out_pptx)
```

We generate a empty plot for vectors: 

```{r, warning = FALSE, results='hide'}
# pca_treat + 
#   scale_fill_manual(values = c("transparent")) + 
#   scale_color_manual(values = c(rep("transparent", 8))) + 
#   theme(panel.border = element_blank(), panel.grid.major = element_blank(),
#         panel.grid.minor = element_blank(), axis.line = element_line(colour = "black")) -> empty_plot_tmp
```

### Next, we add AM drug classes info as envfit vectors on the generated 'empty' plot:

```{r, warning = FALSE, results='hide'}
# phyloseq_add_taxa_vector_fix(dist = list$robust.aitchison,
#                              phyloseq = ps_tmp %>%  
#                                physeq_sel_tax_table(c("AMR_sub_class_multi")) %>% 
#                                subset_samples(Model == "Chicken") %>% 
#                                filter_taxa(function(x) sum(x > 0) > 0, TRUE),
#                              figure_ord = empty_plot_tmp,
#                              taxrank_glom = "AMR_sub_class_multi") -> envfit_class
# 
# envfit_class$signenvfit %>% 
#   mutate_if(is.numeric, ~round(., 4)) %>% 
#   DT::datatable()
```

```{r}
# envfit_class$plot


# ppp$plot + 
#   ggrepel::geom_text_repel(cex=2,
#                            aes(label= Phase),
#                            segment.color = 'black',
#                            segment.size = 0.5,
#                            ppp$plot$data %>% filter(Phase %in% c("Stab")) %>% unique()
#   ) -> pbeta

# envfit_class$plot %>%
#   export::graph2ppt(append = TRUE, width = 317.48031496, height = 0.618 * 317.48031496, paper = "A4",  scaling = 2,
#                     file = out_pptx)


# envfit_class$plot %>% 
#   ggsave(plot = .,bg = "transparent",scale = 2,
#          units = "mm",
#          dpi = 600,
#          height = 0.618 * 84,
#          width = 84,
#          filename = "~/Desktop/Resistome_beta_chick_class_gn.eps")
```

### Next, we add MGE prediction genes abundance as envfit vectors on the generated 'empty' plot:

```{r}
# ps_AMR_all %>% 
#   physeq_sel_tax_table(c("PathoFact_AMR_MGE_prediction_clean_2"))  %>%
#   speedyseq::tax_glom(., taxrank = "PathoFact_AMR_MGE_prediction_clean_2") %>% 
#   # physeq_sel_tax_table(c("PathoFact_AMR_AMR_sub_class_multi")) %>% 
#   prune_samples(samples, .) %>% 
#   subset_samples(Model == "Chicken") %>% 
#   filter_taxa(function(x) sum(x > 0) > 0, TRUE) %>% 
#   physeq_add_metadata(metadata_mapping,
#                       sample_column = "sample_id") -> ps_tmp_envfit
# 
# # custom PathoFact_AMR_MGE_prediction cleaning- should be done once at the begining...
# 
# ps_tmp_envfit %>% 
#   tax_table() %>% 
#   data.frame() %>% 
#   mutate(PathoFact_AMR_MGE_prediction_clean_2 = case_when(grepl("ambiguous", PathoFact_AMR_MGE_prediction_clean_2) ~ "ambiguous", TRUE ~ PathoFact_AMR_MGE_prediction_clean_2)) %>% 
#   mutate(PathoFact_AMR_MGE_prediction_clean_2 = case_when(PathoFact_AMR_MGE_prediction_clean_2 %in% c("MAG", "chromosome")  ~ "MAG - chromosome", TRUE ~ PathoFact_AMR_MGE_prediction_clean_2)) %>% 
#   as.matrix() -> tax_table(ps_tmp_envfit)
# 
# ps_tmp_envfit %>% 
#   phyloseq_add_taxa_vector_fix(dist = coda_resistome$aidist,
#                                phyloseq = . ,
#                                taxnames_rm = c("unclassified", "ambiguous"),
#                                vector_color = "tomato",
#                                figure_ord = empty_plot_tmp,
#                                taxrank_glom = "PathoFact_AMR_MGE_prediction_clean_2") -> envfit_MGE
# 
# 
# 
# envfit_MGE$signenvfit %>% 
#   mutate_if(is.numeric, ~round(., 4)) %>% 
#   DT::datatable()
```

```{r}
# 
# envfit_MGE$plot 
# 
# # ppp$plot + 
# #   ggrepel::geom_text_repel(cex=2,
# #                            aes(label= Phase),
# #                            segment.color = 'black',
# #                            segment.size = 0.5,
# #                            ppp$plot$data %>% filter(Phase %in% c("Stab")) %>% unique()
# #   ) -> pbeta
# 
# envfit_MGE$plot %>% 
#   ggsave(plot = .,bg = "transparent",scale = 2,
#          units = "mm",
#          dpi = 600,
#          height = 0.618 * 84,
#          width = 84,
#          filename = "~/Desktop/Resistome_beta_chick_MGE_gn.eps")
# 
# envfit_MGE$plot %>%
#   export::graph2ppt(append = TRUE, width = 317.48031496, height = 0.618 * 317.48031496, paper = "A4",  scaling = 2,
#                     file = out_pptx)
```

### Finally, we add AMR genes abunbundance as envfit vectors - similarly as before:

```{r}
# phyloseq_add_taxa_vector_fix(dist = coda_resistome$aidist, 
#                              ggrepel_force = 30,
#                              phyloseq = ps_tmp %>% 
#                                physeq_sel_tax_table(c("PathoFact_AMR_ARG")) %>% 
#                                subset_samples(Model == "Chicken") %>% 
#                                filter_taxa(function(x) sum(x > 0) > 0, TRUE) %>% 
#                                filter_taxa(., function(x){sum(x > 0.0000001) > 3}, prune = TRUE), # only keeping dominant and occurent ones
#                              figure_ord = empty_plot_tmp, 
#                              top_r = 20,
#                              vector_color = "chartreuse",
#                              taxrank_glom = "PathoFact_AMR_ARG") -> pca_env_ARG
# 
# pca_env_ARG$signenvfit %>% 
#   mutate_if(is.numeric, ~round(., 4)) %>% 
#   DT::datatable()  
```

```{r}
# pca_env_ARG$plot
```

```{r}
# pca_env_ARG$vectors
# 
# pca_env_ARG$vectors %>% 
#   ggsave(plot = .,bg = "transparent",scale = 2,
#          units = "mm",
#          dpi = 600,
#          height = 0.618 * 84,
#          width = 84,
#          filename = "~/Desktop/Resistome_beta_chick_ARG.eps")
# 
# pca_env_ARG$vectors  %>% 
#   export::graph2ppt(append = TRUE, width = 317.48031496  , height = 0.618 * 317.48031496  , paper = "A4",  scaling = 2,
#                     file = out_pptx)
```


# Alpha-diversity resistome:

##  gN:

```{r}
# ps_AMR_all %>% 
#   physeq_sel_tax_table(c("AMR_sub_class", "AMR_sub_class_multi", "ARG"))  %>%
#   # speedyseq::tax_glom(., taxrank = "PathoFact_AMR_ARG")  %>%
#   filter_taxa(function(x) sum(x > 0) > 0, TRUE)-> ps_tmp # %>% 
# # prune_samples(samples, .) %>% 
# # physeq_add_metadata(metadata_mapping,
# #                     sample_column = "sample_id") -> ps_tmp
# 
# ps_tmp
```


```{r}
# sample_data(ps_tmp)$Read_nrom = 1000000
# 
# ps_tmp %>% 
#   phyloseq_density_normalize(physeq = .,
#                              # sample_idx = "Reactor",
#                              value_idx = "Read_nrom") %>% 
#   # transform_sample_counts(function(x) (x*sum(x)) *100) %>% 
#   phyloseq_alphas(phylo = FALSE) -> resistome_alpha
# 
# 
# # resistome_alpha %>% 
# #   glimpse()
```

```{r}
plot_alpha_div_NRP72 <- function(df = resistome_alpha, x = "Day_of_Treatment", y = "value",  point_size = 2.5, color = "Treatment", fill  = "Treatment", shape = "Antibiotic", alpha = "Antibiotic_mg.L", facet_formula = paste0("alphadiversiy ~  Model2 "), ylab = "Resistome alpha-diversity", xlab = "Days (Treatment)", measures = c("observed", "diversity_shannon", "evenness_pielou"), path_group = "interaction(Model, Reactor_Treatment_Dose)"){
  
  df %>% 
    pivot_longer(cols = all_of(measures), values_to = "value", names_to = 'alphadiversiy', values_drop_na  = TRUE) %>%
    mutate(alphadiversiy = fct_relevel(alphadiversiy, measures)) -> df_ready
  
  df_ready %>% 
    ggplot(aes_string(x = x, y = y, color = color, fill = fill, shape = shape, alpha = alpha)) + #shape = Fermentation
    geom_point(size = point_size) + 
    geom_line(linetype = 2,  size = 0.5,  aes_string(group = path_group)) +
    labs(y = ylab, x = xlab) +
    # scale_color_manual(name = "", values = treat_col,
    #                    na.value = "black") +
    # scale_fill_manual(name = "", values = treat_col,
    #                   na.value = "black") +
    facet_grid(as.formula(facet_formula), scales = "free", space = "fixed", drop = TRUE) +
    # scale_shape_manual(name = "" ,values = antibio_shape, na.value =  17) +
    # scale_alpha_discrete(name = "", range=c(0.6, 1), na.value =  0.6) + 
    geom_vline(xintercept = 0, size = 0.5, linetype = 1, color = "grey60") -> p_alpha_both
  
  return(p_alpha_both)
}
```


```{r}
# resistome_alpha %>% 
#   mutate(Day_of_Treatment = ifelse(Treatment == "DONOR", -4, Day_of_Treatment)) %>% 
#   plot_alpha_div_NRP72( shape = "as.factor(Antibiotic_mg.L)", alpha = NULL) -> p_alpha 
# 
# p_alpha +  scale_color_manual(name = "", values = treat_col,
#                               na.value = "black") +
#   scale_fill_manual(name = "", values = treat_col,
#                     na.value = "black") +
#   scale_shape_manual(name = "" ,values = c(15,16,18,19), na.value =  17) + 
#   theme_light() + facet_null() + facet_grid( alphadiversiy  ~ Model2, scales = "free", drop = TRUE) -> p_alpha #+
# # ggh4x::facetted_pos_scales(
# #   y = list(
# #     alphadiversiy == "Observed" ~ scale_y_continuous(limits = c(120,270)),
# #     alphadiversiy == "evenness_pielou" ~ scale_y_continuous(limits = c(0.6, 1))
# #   )
# # ) -> p_alpha
# 
# p_alpha
# 
# # p_alpha %>% 
# #   ggsave(plot = .,bg = "transparent",scale = 2,
# #          units = "mm",
# #          dpi = 600,
# #          height = 0.618 * 84,
# #          width = 84,
# #          filename = "~/Desktop/alpha_gn.eps")
# 
# p_alpha %>%
#   export::graph2ppt(append = TRUE, width = 317.48031496  , height = 0.618 * 317.48031496  , paper = "A4",  scaling = 2,
#                     file = out_pptx)
# 
# 
# 
# p_alpha$data %>% 
#   data.frame() %>% 
#   select(sample_id_tmp, Reactor, Treatment, Day_of_Treatment, Model2, Antibiotic_mg.L, Reactor_Treatment_Dose, alphadiversiy, value) %>% 
#   xlsx::write.xlsx(file = out_xlsx,
#                    sheetName = "alpha-gN", append = TRUE)

```

## AMR gNs:

```{r}
ps_AMR_all %>% 
  physeq_sel_tax_table(c("AMR_sub_class", "AMR_sub_class_multi", "ARG"))  %>%
  speedyseq::tax_glom(., taxrank = "ARG")  %>%
  filter_taxa(function(x) sum(x > 0) > 0, TRUE) -> ps_tmp# %>% 
# prune_samples(samples, .) %>% 
# physeq_add_metadata(metadata_mapping,
#                     sample_column = "sample_id") -> ps_tmp

ps_tmp
```


```{r}
sample_data(ps_tmp)$Read_nrom = 1000000

ps_tmp %>% 
  phyloseq_density_normalize(physeq = .,
                             # sample_idx = "Reactor",
                             value_idx = "Read_nrom") %>% 
  # transform_sample_counts(function(x) (x*sum(x)) *100) %>% 
  phyloseq_alphas(phylo = FALSE) -> resistome_alpha
```

```{r}
resistome_alpha %>% 
  mutate(Day_of_Treatment = ifelse(Treatment == "DONOR", -4, Day_of_Treatment)) %>% 
  plot_alpha_div_NRP72( shape = "as.factor(Antibiotic_mg.L)", alpha = NULL) -> p_alpha 

p_alpha +  scale_color_manual(name = "", values = treat_col,
                              na.value = "black") +
  scale_fill_manual(name = "", values = treat_col,
                    na.value = "black") +
  scale_shape_manual(name = "" ,values = c(15,16,18,19), na.value =  17) + 
  theme_light() + facet_null() + facet_grid(alphadiversiy  ~ Model2, scales = "free", drop = TRUE) -> p_alpha #+
# ggh4x::facetted_pos_scales(
#   y = list(
#     alphadiversiy == "Observed" ~ scale_y_continuous(limits = c(120,270)),
#     alphadiversiy == "evenness_pielou" ~ scale_y_continuous(limits = c(0.6, 1))
#   )
# ) -> p_alpha

p_alpha

# p_alpha %>% 
#   ggsave(plot = .,bg = "transparent",scale = 2,
#          units = "mm",
#          dpi = 600,
#          height = 0.618 * 84,
#          width = 84,
#          filename = "~/Desktop/alpha.eps")

p_alpha %>%
  export::graph2ppt(append = TRUE, width = 317.48031496  , height = 0.618 * 317.48031496  , paper = "A4",  scaling = 2,
                    file = out_pptx)


p_alpha$data %>% 
  data.frame() %>% 
  select(sample_id_tmp, Reactor, Treatment, Day_of_Treatment, Model2, Antibiotic_mg.L, Reactor_Treatment_Dose, alphadiversiy, value) %>% 
  xlsx::write.xlsx(file = out_xlsx,
                   sheetName = "alpha-AMRgN", append = TRUE)

```

# VAN A vs other gene in the plasmid.

```{r}

```


# MGE :

```{r}
ps_AMR_all %>%
  physeq_sel_tax_table(c("AMR_sub_class", "AMR_sub_class_multi", "ARG", "MGE_prediction_clean"))  %>%
  speedyseq::tax_glom(., taxrank = "MGE_prediction_clean")  %>%
  physeq_sel_tax_table(c("MGE_prediction_clean")) %>%
  # physeq_sel_tax_table(c("PathoFact_AMR_AMR_sub_class_multi")) %>%
  # prune_samples(samples, .) %>%
  filter_taxa(function(x) sum(x > 0) > 0, TRUE)  -> ps_tmp
# 
# ps_tmp %>% 
#   tax_table() %>% 
#   data.frame() %>% 
#   mutate(PathoFact_AMR_MGE_prediction_clean_2 = case_when(grepl("ambiguous", PathoFact_AMR_MGE_prediction_clean_2) ~ "ambiguous", TRUE ~ PathoFact_AMR_MGE_prediction_clean_2)) %>% 
#   mutate(PathoFact_AMR_MGE_prediction_clean_2 = case_when(PathoFact_AMR_MGE_prediction_clean_2 %in% c("MAG", "chromosome")  ~ "MAG - chromosome", TRUE ~ PathoFact_AMR_MGE_prediction_clean_2)) %>% 
#   mutate(PathoFact_AMR_MGE_prediction_clean_2 = fct_relevel(PathoFact_AMR_MGE_prediction_clean_2, c("MAG", "chromosome", "plasmid", "phage", "ambiguous", "unclassified"))) %>% 
#   as.matrix() -> tax_table(ps_tmp)
# 
ps_tmp %>%
  # ps_mutate(Day_of_Treatment = as.double(Day_of_Treatment)) %>% 
  ps_arrange(Day_of_Treatment) %>%
  ps_filter(Model2 == "Chicken1") %>% 
  comp_barplot(
    tax_level = "MGE_prediction_clean",
    label = "Day_of_Treatment", # name an alternative variable to label axes
    n_taxa = 20, # give more taxa unique colours
    merge_other = FALSE, # split the "other" category to display alpha diversity
    bar_width = 0.9, # reduce the bar width to 70% of one row
    bar_outline_colour = "grey5",
    sample_order = "default") +#,# is the default (use NA to remove outlines)
  # facet_by = c("Model", "Antibiotic", "Antibiotic_mg.mL", "Treatment2")) +
  # group_by = c("Model", "Antibiotic", "Antibiotic_mg.mL", "Treatment2"))
  # coord_flip() +
  theme_light() +
  scale_y_continuous(labels=percent) -> p_tmp

plist = NULL

p_tmp +
  facet_grid(Model2 ~ Antibiotic + Antibiotic_mg.L + Treatment2, scales = "free") -> plist$chicken1

ps_tmp %>%
  # ps_mutate(Day_of_Treatment = as.double(Day_of_Treatment)) %>% 
  ps_arrange(Day_of_Treatment) %>%
  ps_filter(Model2 == "Chicken2") %>% 
  comp_barplot(
    tax_level = "MGE_prediction_clean",
    label = "Day_of_Treatment", # name an alternative variable to label axes
    n_taxa = 20, # give more taxa unique colours
    merge_other = FALSE, # split the "other" category to display alpha diversity
    bar_width = 0.9, # reduce the bar width to 70% of one row
    bar_outline_colour = "grey5",
    sample_order = "default") +#,# is the default (use NA to remove outlines)
  # facet_by = c("Model", "Antibiotic", "Antibiotic_mg.mL", "Treatment2")) +
  # group_by = c("Model", "Antibiotic", "Antibiotic_mg.mL", "Treatment2"))
  # coord_flip() +
  theme_light() +
  scale_y_continuous(labels=percent) -> p_tmp

p_tmp +
  facet_grid(Model2 ~ Antibiotic + Antibiotic_mg.L + Treatment2, scales = "free") -> plist$chicken2


# p_tmp +
#   facet_wrap(Model2 ~ Antibiotic + Antibiotic_mg.L + Treatment2, scales = "free", ncol = 7, nrow = 4) -> p2
#
# # # Plot them side by side with the patchwork package.
patch <- patchwork::wrap_plots(plist, nrow = 2, guides = "collect")
# # patch & coord_flip() # make all plots horizontal (note: use & instead of +)
# #   
# 
# p2 %>%
#   ggpubr::set_palette("npg") -> p2
# 
# p2
# 
# p2 %>% 
#   ggsave(plot = .,bg = "transparent",scale = 2,
#          units = "mm",
#          dpi = 600,
#          height = 0.618 * 84 * 2,
#          width = 84 * 2,
#          filename = "~/Desktop/AMR_carrier.eps")
# 

patch


patch %>%
  export::graph2ppt(append = TRUE, width = 317.48031496 * 3  , height = 0.618 * 317.48031496 * 5  , paper = "A4",  scaling = 2,
                    file = out_pptx)
```

using the data stored in p2 object:

```{r}
patch$data %>% 
  data.frame() %>% 
  select(OTU,Sample, Abundance, Reactor, Treatment, Day_of_Treatment, Model2, Antibiotic_mg.L, Reactor_Treatment_Dose) %>% 
  xlsx::write.xlsx(file = out_xlsx,
                   sheetName = "AMR-carrier", append = TRUE)
```

# Contributing MAGs:

```{r}
pMAG = NULL
pMAG2 = NULL

pMAG_no_leg = NULL

ps_AMR_all %>%
  # physeq_sel_tax_table(c("AMR_sub_class", "AMR_sub_class_multi",  "MGE_prediction_clean", "t_phylum","t_class", "t_order","t_family","t_genus"  ,                  "t_species" ,"X..completion","X..redundancy","total.length" ,"bin_name_clean", "ARG"))  %>%
  physeq_sel_tax_table(c("MGE_prediction_clean", "t_phylum","t_class", "t_order","t_family","t_genus"  ,                  "t_species" ,"X..completion","X..redundancy","total.length" ,"bin_name_clean", "ARG"))  %>%
  microViz::tax_mutate(t_species = ifelse(t_species == "None", "", t_species),
                       t_genus = ifelse(t_genus == "None", "", t_genus)) %>% 
  microViz::tax_mutate(bin_tax = case_when(MGE_prediction_clean == "MAG" ~  paste0(bin_name_clean, "_o_", t_order, "_g_", t_genus, "_s_", t_species),
                                           TRUE ~ NA_character_),
                       bin_tax_MGE = case_when(MGE_prediction_clean == "MAG" ~ bin_tax,
                                               TRUE ~ MGE_prediction_clean))  %>% 
  # tax_table() %>%  data.frame() %>% write_tsv("~/Desktop/tttt.tsv")
  physeq_sel_tax_table("bin_tax_MGE") %>%  
  physeq_glom_rename(speedyseq = TRUE, taxrank = "bin_tax_MGE", rename_ASV = "bin_tax_MGE") -> ps_tmp
```


```{r}
ps_tmp %>% 
  # speedyseq::tax_glom(., taxrank = "bin_tax_MGE")  %>%
  # physeq_sel_tax_table(c("MGE_prediction_clean")) %>%
  # physeq_sel_tax_table(c("PathoFact_AMR_AMR_sub_class_multi")) %>%
  # prune_samples(samples, .) %>%
  subset_samples(Model2 == "Chicken1") %>% 
  filter_taxa(function(x) sum(x > 0) > 0, TRUE) %>% 
  microbiome::transform("compositional") %>% 
  subset_taxa(!bin_tax_MGE %in% c("unclassified", "ambiguous", "chromosome", "phage", "plasmid") ) %>% 
  ps_arrange(Day_of_Treatment) %>%
  comp_barplot(
    tax_transform_for_plot = "identity",
    tax_level = "bin_tax_MGE",
    label = "Day_of_Treatment", # name an alternative variable to label axes
    palette = distinct_palette(18, pal = "greenArmytage"),
    n_taxa = 18, # give more taxa unique colours
    merge_other = FALSE, # split the "other" category to display alpha diversity
    bar_width = 0.9, # reduce the bar width to 70% of one row
    bar_outline_colour = "grey5",
    sample_order = "default") +#,# is the default (use NA to remove outlines)
  # facet_by = c("Model", "Antibiotic", "Antibiotic_mg.mL", "Treatment2")) +
  # group_by = c("Model", "Antibiotic", "Antibiotic_mg.mL", "Treatment2"))
  # coord_flip() +
  theme_light() -> plot

plot +
  facet_grid(Model2 ~ Antibiotic + Antibiotic_mg.L + Treatment2, scales = "free") -> pMAG$chicken1

pMAG$chicken1 +  guides(fill=guide_legend(ncol=2)) + theme(legend.position = "bottom")  -> ptmp

ptmp
```


```{r}
ptmp %>%
  export::graph2ppt(append = TRUE, width = 317.48031496 * 1.5  , height = 0.618 * 317.48031496 * 1.25   , paper = "A4",  scaling = 2,
                    file = out_pptx)
```


```{r}
ps_tmp %>% 
  # speedyseq::tax_glom(., taxrank = "bin_tax_MGE")  %>%
  # physeq_sel_tax_table(c("MGE_prediction_clean")) %>%
  # physeq_sel_tax_table(c("PathoFact_AMR_AMR_sub_class_multi")) %>%
  # prune_samples(samples, .) %>%
  subset_samples(Model2 == "Chicken1") %>% 
  filter_taxa(function(x) sum(x > 0) > 0, TRUE) %>% 
  microbiome::transform("compositional") %>% 
  subset_taxa(!bin_tax_MGE %in% c("unclassified", "ambiguous", "chromosome", "phage", "plasmid") ) %>% 
  ps_arrange(Day_of_Treatment) %>%
  comp_barplot(
    # tax_transform_for_plot = "identity",
    tax_level = "bin_tax_MGE",
    label = "Day_of_Treatment", # name an alternative variable to label axes
    palette = distinct_palette(18, pal = "greenArmytage"),
    n_taxa = 18, # give more taxa unique colours
    merge_other = FALSE, # split the "other" category to display alpha diversity
    bar_width = 0.9, # reduce the bar width to 70% of one row
    bar_outline_colour = "grey5",
    sample_order = "default") +#,# is the default (use NA to remove outlines)
  # facet_by = c("Model", "Antibiotic", "Antibiotic_mg.mL", "Treatment2")) +
  # group_by = c("Model", "Antibiotic", "Antibiotic_mg.mL", "Treatment2"))
  # coord_flip() +
  theme_light() -> plot

plot +
  facet_grid(Model2 ~ Antibiotic + Antibiotic_mg.L + Treatment2, scales = "free") -> pMAG2$chicken1

pMAG2$chicken1 +  guides(fill=guide_legend(ncol=2)) + theme(legend.position = "bottom")  -> ptmp

ptmp
```


```{r}
ptmp %>%
  export::graph2ppt(append = TRUE, width = 317.48031496 * 1.5  , height = 0.618 * 317.48031496 * 1.25   , paper = "A4",  scaling = 2,
                    file = out_pptx)
```


```{r}
ps_tmp %>% 
  # speedyseq::tax_glom(., taxrank = "bin_tax_MGE")  %>%
  # physeq_sel_tax_table(c("MGE_prediction_clean")) %>%
  # physeq_sel_tax_table(c("PathoFact_AMR_AMR_sub_class_multi")) %>%
  # prune_samples(samples, .) %>%
  subset_samples(Model2 == "Chicken2") %>% 
  filter_taxa(function(x) sum(x > 0) > 0, TRUE) %>% 
  microbiome::transform("compositional") %>% 
  subset_taxa(!bin_tax_MGE %in% c("unclassified", "ambiguous", "chromosome", "phage", "plasmid") ) %>% 
  ps_arrange(Day_of_Treatment) %>%
  comp_barplot(
    tax_transform_for_plot = "identity",
    tax_level = "bin_tax_MGE",
    label = "Day_of_Treatment", # name an alternative variable to label axes
    palette = distinct_palette(18, pal = "kelly"),
    n_taxa = 18, # give more taxa unique colours
    merge_other = FALSE, # split the "other" category to display alpha diversity
    bar_width = 0.9, # reduce the bar width to 70% of one row
    bar_outline_colour = "grey5",
    sample_order = "default") +#,# is the default (use NA to remove outlines)
  # facet_by = c("Model", "Antibiotic", "Antibiotic_mg.mL", "Treatment2")) +
  # group_by = c("Model", "Antibiotic", "Antibiotic_mg.mL", "Treatment2"))
  # coord_flip() +
  theme_light() -> plot

plot +
  facet_grid(Model2 ~ Antibiotic + Antibiotic_mg.L + Treatment2, scales = "free") -> pMAG$chicken2

pMAG$chicken2 +  guides(fill=guide_legend(ncol=2)) + theme(legend.position = "bottom")  -> ptmp

ptmp
```


```{r}
ptmp %>%
  export::graph2ppt(append = TRUE, width = 317.48031496 * 1.5  , height = 0.618 * 317.48031496 * 1.25   , paper = "A4",  scaling = 2,
                    file = out_pptx)
```


```{r}
ps_tmp %>% 
  # speedyseq::tax_glom(., taxrank = "bin_tax_MGE")  %>%
  # physeq_sel_tax_table(c("MGE_prediction_clean")) %>%
  # physeq_sel_tax_table(c("PathoFact_AMR_AMR_sub_class_multi")) %>%
  # prune_samples(samples, .) %>%
  subset_samples(Model2 == "Chicken2") %>% 
  filter_taxa(function(x) sum(x > 0) > 0, TRUE) %>% 
  microbiome::transform("compositional") %>% 
  subset_taxa(!bin_tax_MGE %in% c("unclassified", "ambiguous", "chromosome", "phage", "plasmid") ) %>% 
  ps_arrange(Day_of_Treatment) %>%
  comp_barplot(
    # tax_transform_for_plot = "identity",
    tax_level = "bin_tax_MGE",
    label = "Day_of_Treatment", # name an alternative variable to label axes
    palette = distinct_palette(18, pal = "kelly"),
    n_taxa = 18, # give more taxa unique colours
    merge_other = FALSE, # split the "other" category to display alpha diversity
    bar_width = 0.9, # reduce the bar width to 70% of one row
    bar_outline_colour = "grey5",
    sample_order = "default") +#,# is the default (use NA to remove outlines)
  # facet_by = c("Model", "Antibiotic", "Antibiotic_mg.mL", "Treatment2")) +
  # group_by = c("Model", "Antibiotic", "Antibiotic_mg.mL", "Treatment2"))
  # coord_flip() +
  theme_light() -> plot

plot +
  facet_grid(Model2 ~ Antibiotic + Antibiotic_mg.L + Treatment2, scales = "free") -> pMAG2$chicken2

pMAG2$chicken2 +  guides(fill=guide_legend(ncol=2)) + theme(legend.position = "bottom")  -> ptmp

ptmp
```


```{r}
ptmp %>%
  export::graph2ppt(append = TRUE, width = 317.48031496 * 1.5  , height = 0.618 * 317.48031496 * 1.25   , paper = "A4",  scaling = 2,
                    file = out_pptx)
```


Export values displayed:

```{r}
rbind( pMAG$chicken1$data,  pMAG$chicken2$data) %>% 
  data.frame() %>% 
  select(OTU,Sample, Abundance, Reactor, Treatment, Day_of_Treatment, Model2, Antibiotic_mg.L, Reactor_Treatment_Dose) %>% 
  # arrange(Day_of_Treatment) %>% 
  xlsx::write.xlsx(x = . ,
                   append = TRUE,
                   file = out_xlsx,
                   sheetName = "MAG_resistome_contribution")
```


```{r}
rbind( pMAG2$chicken1$data,  pMAG2$chicken2$data) %>% 
  data.frame() %>% 
  select(OTU,Sample, Abundance, Reactor, Treatment, Day_of_Treatment, Model2, Antibiotic_mg.L, Reactor_Treatment_Dose) %>% 
  # arrange(Day_of_Treatment) %>% 
  xlsx::write.xlsx(x = . ,
                   append = TRUE,
                   file = out_xlsx,
                   sheetName = "MAG_resistome_contribution_onlyMAG")
```

## chicken1 
```{r}
ps_AMR %>%
  # physeq_sel_tax_table(c("AMR_sub_class", "AMR_sub_class_multi",  "MGE_prediction_clean", "t_phylum","t_class", "t_order","t_family","t_genus"  ,                  "t_species" ,"X..completion","X..redundancy","total.length" ,"bin_name_clean", "ARG"))  %>%
  physeq_sel_tax_table(c("MGE_prediction_clean", "t_phylum","t_class", "t_order","t_family","t_genus"  ,                  "t_species" ,"X..completion","X..redundancy","total.length" ,"bin_name_clean", "ARG"))  %>%
  microViz::tax_mutate(t_species = ifelse(t_species == "None", "", t_species),
                       t_genus = ifelse(t_genus == "None", "", t_genus)) %>% 
  microViz::tax_mutate(bin_tax = paste0(bin_name_clean, "_o_", t_order, "_g_", t_genus, "_s_", t_species),
                       bin_tax_MGE = case_when(MGE_prediction_clean == "MAG" ~ bin_tax,
                                               TRUE ~ MGE_prediction_clean)) %>% 
  physeq_sel_tax_table("bin_tax_MGE") %>% 
  physeq_glom_rename(speedyseq = TRUE, taxrank = "bin_tax_MGE", rename_ASV = "bin_tax_MGE") %>%
  # speedyseq::tax_glom(., taxrank = "bin_tax_MGE")  %>%
  # physeq_sel_tax_table(c("MGE_prediction_clean")) %>%
  # physeq_sel_tax_table(c("PathoFact_AMR_AMR_sub_class_multi")) %>%
  # prune_samples(samples, .) %>%
  subset_samples(Model2 == "Chicken1") %>% 
  filter_taxa(function(x) sum(x > 0) > 0, TRUE) %>% 
  microbiome::transform("compositional") %>% 
  ps_arrange(-Day_of_Treatment) %>% 
  speedyseq::psmelt() %>%
  # left_join(tax_mapping,
  #           by = c("Best_Hit_ARO" = "Best_Hit_ARO"),
  #           suffix = c("_x", "")) %>%
  # mutate(Sample = fct_reorder(Sample, Model,Treatment,Fermentation,Antibiotic_mg.mL, Day_of_Treatment )) %>%
  mutate(Abundance = na_if(Abundance, 0)) %>%
  mutate(logAbundance = log10(Abundance)) -> pheatmap_df #%>% 
# mutate(bin_tax_MGE = bin_tax_MGE %>% stringr::str_trunc(26,  side ="left")) -> pheatmap_df


pheatmap_df %>%
  # replace(is.na(.), 0)  %>%
  # mutate(log10Abundance = log10(Abundance)) %>% 
  filter(Abundance > 0) %>%
  select(Sample, bin_tax_MGE, logAbundance) %>%
  # filter(Abundance > 1) %>%
  group_by(Sample, bin_tax_MGE) %>%
  pivot_wider(names_from =  Sample, bin_tax_MGE,
              values_from = logAbundance) %>%
  # mutate(PathoFact_AMR_ARG = PathoFact_AMR_ARG %>%  str_to_lower(.)) %>% 
  # %>% str_to_lower(., locale = "en")
  column_to_rownames("bin_tax_MGE") -> pheatmap_samples

pheatmap_df %>%
  # replace(is.na(.), 0)  %>%
  mutate(log10Abundance = log10(Abundance)) %>% 
  select(Sample, bin_tax_MGE, Abundance) %>%
  group_by(bin_tax_MGE) %>% 
  summarise(meanAbundance = mean(Abundance, na.rm = TRUE)) -> mean_ab


pheatmap_df %>%
  # replace(is.na(.), 0)  %>%
  # mutate(log10Abundance = log10(Abundance)) %>% 
  filter(Abundance > 0) %>%
  select(Sample, bin_tax_MGE, Abundance) %>%
  # filter(Abundance > 1) %>%
  group_by(Sample, bin_tax_MGE) %>%
  pivot_wider(names_from =  Sample, bin_tax_MGE,
              values_from = Abundance) %>%
  # mutate(PathoFact_AMR_ARG = PathoFact_AMR_ARG %>%  str_to_lower(.)) %>% 
  # %>% str_to_lower(., locale = "en")
  column_to_rownames("bin_tax_MGE") %>% 
  replace(is.na(.), 0)  %>%
  t() %>%
  vegan::vegdist(na.rm = TRUE) %>%
  hclust(method = "ward.D2") -> clust_t

pheatmap_samples %>%
  replace(is.na(.), 0)  %>%
  vegan::vegdist(na.rm = TRUE, method = "euc") %>%
  # cor() %>%
  hclust(method = "ward.D2") -> clust

pheatmap_samples %>%
  pheatmap::pheatmap(fontsize_col = 6,
                     display_numbers = FALSE,
                     fontsize = 6,
                     cluster_rows = clust ,
                     scale = "none",# scale = "row",
                     cluster_cols = clust_t,
                     na_col = "transparent",
                     # annotation_row = tax_mapping %>% rownames_to_column('tmp_col') %>%
                     #   filter(ARG %in% rownames(pheatmap_samples)) %>%
                     #   distinct(ARG, .keep_all = TRUE) %>%
                     #   left_join(., 
                     #             mean_ab,
                     #             by = c("ARG" = "ARG")) %>% 
                     #   column_to_rownames("ARG") %>%
                     #   select(meanAbundance),
                     annotation_colors = list("Treatment" = treat_col),
                     annotation_col = pheatmap_df %>%
                       select(Antibiotic_mg.L,Day_of_Treatment, Treatment, Model2, Sample) %>% 
                       left_join(patch$data %>% 
                                   filter(Model2 == "Chicken2") %>% 
                                   mutate(Sample = str_replace_all(Sample, "_", "")) %>% 
                                   select(Sample, Abundance, OTU)
                       )
                     # mutate(Sample = gsub("_",Sample, "")) %>% 
                     select(OTU,Sample, Abundance)) %>% 
  distinct(Sample, .keep_all = TRUE) %>% 
  column_to_rownames("Sample"))  -> p_pheat


p_pheat

# p_pheat %>% 
#   ggsave(plot = .,bg = "transparent",scale = 2,
#          units = "mm",
#          dpi = 600,
#          height = 0.618 * 84 * 1.5,
#          width = 84  ,
#          filename = "~/Desktop/Resistome_heatmap_VAN.eps")


p_pheat %>%
  export::graph2ppt(append = TRUE, width = 317.48031496 * 1.5 , height = 0.618 * 317.48031496 * 3 , paper = "A4",  scaling = 1,
                    file = out_pptx)
```

```{r}
# ps_AMR_all %>% 
#   physeq_sel_tax_table(c("PathoFact_AMR_AMR_sub_class", "PathoFact_AMR_AMR_sub_class_multi", "PathoFact_AMR_ARG", "PathoFact_AMR_MGE_prediction_clean_2"))  %>%
#   speedyseq::tax_glom(., taxrank = "PathoFact_AMR_MGE_prediction_clean_2")  %>% 
#   physeq_sel_tax_table(c("PathoFact_AMR_MGE_prediction_clean_2")) %>% 
#   # physeq_sel_tax_table(c("PathoFact_AMR_AMR_sub_class_multi")) %>% 
#   prune_samples(samples, .) %>% 
#   filter_taxa(function(x) sum(x > 0) > 0, TRUE) %>% 
#   physeq_add_metadata(metadata_mapping,
#                       sample_column = "sample_id") -> ps_tmp
# 
# ps_tmp %>% 
#   tax_table() %>% 
#   data.frame() %>% 
#   mutate(PathoFact_AMR_MGE_prediction_clean_2 = case_when(grepl("ambiguous", PathoFact_AMR_MGE_prediction_clean_2) ~ "ambiguous", TRUE ~ PathoFact_AMR_MGE_prediction_clean_2)) %>% 
#   # mutate(PathoFact_AMR_MGE_prediction_clean_2 = case_when(PathoFact_AMR_MGE_prediction_clean_2 %in% c("MAG", "chromosome")  ~ "MAG - chromosome", TRUE ~ PathoFact_AMR_MGE_prediction_clean_2)) %>% 
#   mutate(PathoFact_AMR_MGE_prediction_clean_2 = fct_relevel(PathoFact_AMR_MGE_prediction_clean_2, c("MAG", "chromosome", "plasmid", "phage", "ambiguous", "unclassified"))) %>% 
#   as.matrix() -> tax_table(ps_tmp)
# 
# ps_tmp %>% 
#   ps_arrange(-Day_of_Treatment) %>% 
#   comp_barplot(
#     tax_level = "PathoFact_AMR_MGE_prediction_clean_2",
#     label = "Day_of_Treatment", # name an alternative variable to label axes
#     n_taxa = 20, # give more taxa unique colours
#     merge_other = FALSE, # split the "other" category to display alpha diversity
#     bar_width = 0.9, # reduce the bar width to 70% of one row
#     bar_outline_colour = "grey5",
#     sample_order = "default") +#,# is the default (use NA to remove outlines)
#   # facet_by = c("Model", "Antibiotic", "Antibiotic_mg.mL", "Treatment2")) +
#   # group_by = c("Model", "Antibiotic", "Antibiotic_mg.mL", "Treatment2"))
#   coord_flip() +  theme_light() +
#   scale_y_continuous(labels=percent) + 
#   facet_wrap(Model ~ Antibiotic + Antibiotic_mg.mL + Treatment2, scales = "free", ncol = 6, nrow = 4) -> p2
# 
# p_carrier_all <- p2
# # # Plot them side by side with the patchwork package.
# # patch <- patchwork::wrap_plots(plot_list, nrow = 1, guides = "collect")
# # patch & coord_flip() # make all plots horizontal (note: use & instead of +)
# #   
# 
# p2 %>%
#   ggpubr::set_palette("npg") -> p2
# 
# p2
# 
# p2 %>% 
#   ggsave(plot = .,bg = "transparent",scale = 2,
#          units = "mm",
#          dpi = 600,
#          height = 0.618 * 84 * 2,
#          width = 84 * 2,
#          filename = "~/Desktop/AMR_carrier_3.eps")
# 
# p2 %>%
#   export::graph2ppt(append = TRUE, width = 317.48031496 * 2  , height = 0.618 * 317.48031496 * 0.75 , paper = "A4",  scaling = 2,
#                     file = out_pptx)
```

### Fig. 2b

```{r}
ps_AMR_all %>% 
  physeq_sel_tax_table(c("PathoFact_AMR_AMR_sub_class_multi",
                         "PathoFact_AMR_ARG"))  %>%
  # physeq_sel_tax_table(c("PathoFact_AMR_ARG"))  %>% 
  speedyseq::tax_glom(., taxrank = "PathoFact_AMR_ARG")  %>% 
  prune_samples(samples, .) %>% 
  physeq_add_metadata(metadata_mapping,
                      sample_column = "sample_id") %>% 
  filter_taxa(function(x) sum(x > 0) > 0, TRUE) %>% 
  speedyseq::psmelt()  %>% 
  filter(! grepl('nclassified', PathoFact_AMR_AMR_sub_class_multi)) -> df

results <- vector("list", length(df$PathoFact_AMR_AMR_sub_class_multi %>%  unique()))
names(results) = c(df$PathoFact_AMR_AMR_sub_class_multi %>%  unique())

for (dd in df$PathoFact_AMR_AMR_sub_class_multi %>%  unique())
{
  print(dd)
  df %>% 
    filter(PathoFact_AMR_AMR_sub_class_multi == !!dd) %>% 
    select(-Lactose_mM:-Total_SCFA_mM) %>% 
    group_by(Sample, PathoFact_AMR_AMR_sub_class_multi) %>% 
    summarise(sum_abundance = sum(Abundance), .groups = "drop") %>% 
    mutate(sum_abundance = na_if(sum_abundance, 0)) %>%
    left_join(metadata_mapping,
              by = c("Sample" = "sample_id")) %>% 
    # filter(Model == "Chicken") %>% 
    # mutate(Day_of_Treatment_num = as.double(Day_of_Treatment)) %>% 
    # mutate(Day_of_Treatment = fct_relevel(Day_of_Treatment, c(NA, "-3", "-1", "1" , "3", "4", "5","7", "30", "47", "48")))  %>% 
    mutate(Antibiotic_mg.mL = factor(Antibiotic_mg.mL, levels = c(NA, 20, 200,90, 600))) %>%
    ggplot(aes_string(x = "Day_of_Treatment", y = "sum_abundance", color = "Treatment", fill  = "Treatment", shape = "Antibiotic_mg.mL")) + #shape = Fermentation
    geom_point(size = 2.5) + 
    geom_line(linetype = 2,  size = 0.5,  aes_string(group = "interaction(Model, Reactor_Treatment_Dose)")) +
    labs(y = paste0("sum Rnum_Gi ") , x = "Day_of_Treatment") +
    scale_color_manual(name = "", values = treat_col,
                       na.value = "black") +
    scale_fill_manual(name = "", values = treat_col,
                      na.value = "black") +
    scale_shape_manual(name = "" ,values = c(15,16,18,19), na.value =  17) +
    # scale_alpha_continuous(name = "", range=c(0.6, 1), na.value =  1) +
    geom_vline(xintercept = 0, size = 0.5, linetype = 1, color = "grey60") +
    ggtitle(dd) +
    theme_light() + 
    facet_grid(Antibiotic  ~ Model , scales = "free", drop = TRUE) +  scale_y_continuous(labels=scientific) -> p1
  # scale_y_continuous(labels = function(x) format(x, scientific = TRUE)) -> p1
  
  # p1 + facet_grid(as.formula(paste0("Antibiotic_mg.mL + PathoFact_AMR_AMR_sub_class_multi ~  Model  ")), scales = "fixed", space = "fixed", drop = TRUE)
  # 
  # p1 + facet_wrap(as.formula(paste0("PathoFact_AMR_AMR_sub_class_multi ~  Model  ")),scales = "free"
  #)
  results[[dd]] <- p1
  
  
  results[[dd]]  %>% 
    ggsave(plot = .,bg = "transparent",scale = 2,
           units = "mm",
           dpi = 600,
           height = 0.618 * 84,
           width = 84,
           filename = paste0("~/Desktop/AMR_class_", dd, ".eps"))
  
  results[[dd]]  %>%
    export::graph2ppt(append = TRUE, width = 317.48031496 , height = 0.618 * 317.48031496, paper = "A4",  scaling = 2,
                      file = out_pptx)
  
}
```

Vancomycin is a glycopeptide antibiotic used in the prophylaxis and treatment of infections caused by Gram-positive bacteria. Vancomycin inhibits the synthesis of peptidoglycan, the major component of the cell wall of gram-positive bacteria. Its mechanism of action is unusual in that it acts by binding precursors of peptidoglycan, rather than by interacting with an enzyme. -> glycopeptide antibiotic

```{r}
results$`glycopeptide antibiotic`

# results$`glycopeptide antibiotic` %>% 
#   export::graph2ppt(append = TRUE, width = 8, height = 4,
#                     file = out_pptx)
```

Cefotaxime is a semisynthetic cephalosporin taken parenterally. It is resistant to most beta-lactamases and active against Gram-negative rods and cocci due to its aminothiazoyl and methoximino functional groups. -> cephalosporin

```{r}
results$cephalosporin 

# results$cephalosporin %>% 
#   export::graph2ppt(append = TRUE, width = 8, height = 4,
#                     file = out_pptx)
```

same but normalized by baseline:

```{r}
# ps_AMR_all %>% 
#   physeq_sel_tax_table(c("PathoFact_AMR_AMR_sub_class_multi",
#                          "PathoFact_AMR_ARG"))  %>%
#   # physeq_sel_tax_table(c("PathoFact_AMR_ARG"))  %>% 
#   speedyseq::tax_glom(., taxrank = "PathoFact_AMR_ARG")  %>% 
#   prune_samples(samples, .) %>% 
#   physeq_add_metadata(metadata_mapping,
#                       sample_column = "sample_id") %>% 
#   filter_taxa(function(x) sum(x > 0) > 0, TRUE) %>% 
#   speedyseq::psmelt() %>% 
#   filter(! grepl('nclassified', PathoFact_AMR_AMR_sub_class_multi)) -> df
# 
# results <- vector("list", length(df$PathoFact_AMR_AMR_sub_class_multi %>%  unique()))
# names(results) = c(df$PathoFact_AMR_AMR_sub_class_multi %>%  unique())
# 
# for (dd in df$PathoFact_AMR_AMR_sub_class_multi %>%  unique())
# {
#   print(dd)
#   
#   df %>% 
#     filter(PathoFact_AMR_AMR_sub_class_multi == !!dd) %>% 
#     select(-Lactose_mM:-Total_SCFA_mM) %>% 
#     group_by(Sample, PathoFact_AMR_AMR_sub_class_multi) %>% 
#     summarise(sum_abundance = sum(Abundance), .groups = "drop") %>% 
#     mutate(sum_abundance = na_if(sum_abundance, 0)) %>%
#     left_join(metadata_mapping,
#               by = c("Sample" = "sample_id")) %>% 
#     ungroup() %>% 
#     # mutate(Reactor_Treatment_Dose = paste0(Period, Reactor_Treatment_Dose)) %>% 
#     group_by(Model, Fermentation, 
#              Reactor_Treatment_Dose) %>% 
#     arrange(Day_of_Treatment) %>% 
#     mutate(per_cent_change = sum_abundance / first(sum_abundance)) %>% 
#     # filter(Model == "Chicken") %>% 
#     # mutate(Day_of_Treatment_num = as.double(Day_of_Treatment)) %>% 
#     # mutate(Day_of_Treatment = fct_relevel(Day_of_Treatment, c(NA, "-3", "-1", "1" , "3", "4", "5","7", "30", "47", "48")))  %>% 
#     mutate(Antibiotic_mg.mL = factor(Antibiotic_mg.mL, levels = c(NA, 20, 200,90, 600))) %>%
#     ggplot(aes_string(x = "Day_of_Treatment", y = "per_cent_change", color = "Treatment", fill  = "Treatment", shape = "Antibiotic_mg.mL")) + #shape = Fermentation
#     geom_point(size = 2.5) + 
#     geom_line(linetype = 2,  size = 0.5,  aes_string(group = "interaction(Model, Reactor_Treatment_Dose)")) +
#     labs(y = paste0("sum Rnum_Gi / baseline ") , x = "Day_of_Treatment") +
#     scale_color_manual(name = "", values = treat_col,
#                        na.value = "black") +
#     scale_fill_manual(name = "", values = treat_col,
#                       na.value = "black") +
#     scale_shape_manual(name = "" ,values = c(15,16,18,19), na.value =  17) +
#     # scale_alpha_continuous(name = "", range=c(0.6, 1), na.value =  1) +
#     geom_vline(xintercept = 0, size = 0.5, linetype = 1, color = "grey60") +
#     ggtitle(dd) +
#     theme_light() + 
#     facet_grid(Antibiotic  ~ Model , scales = "free_x", drop = TRUE) -> p1
#   # scale_y_continuous(labels = function(x) format(x, scientific = TRUE)) -> p1
#   
#   # p1 + facet_grid(as.formula(paste0("Antibiotic_mg.mL + PathoFact_AMR_AMR_sub_class_multi ~  Model  ")), scales = "fixed", space = "fixed", drop = TRUE)
#   # 
#   # p1 + facet_wrap(as.formula(paste0("PathoFact_AMR_AMR_sub_class_multi ~  Model  ")),scales = "free"
#   #)
#   results[[dd]] <- p1
#   
#   
#   results[[dd]]  %>% 
#     ggsave(plot = .,bg = "transparent",scale = 2,
#            units = "mm",
#            dpi = 600,
#            height = 0.618 * 84,
#            width = 84,
#            filename = paste0("~/Desktop/AMR_class_", dd, "_baseline.eps"))
#   
#   results[[dd]]  %>%
#     export::graph2ppt(append = TRUE, width = 317.48031496 , height = 0.618 * 317.48031496, paper = "A4",  scaling = 2,
#                       file = out_pptx)
#   
# }
```



Vancomycin is a glycopeptide antibiotic used in the prophylaxis and treatment of infections caused by Gram-positive bacteria. Vancomycin inhibits the synthesis of peptidoglycan, the major component of the cell wall of gram-positive bacteria. Its mechanism of action is unusual in that it acts by binding precursors of peptidoglycan, rather than by interacting with an enzyme. -> glycopeptide antibiotic

```{r}
# results$`glycopeptide antibiotic`

# results$`glycopeptide antibiotic` %>% 
#   export::graph2ppt(append = TRUE, width = 8, height = 4,
#                     file = out_pptx)
```

Cefotaxime is a semisynthetic cephalosporin taken parenterally. It is resistant to most beta-lactamases and active against Gram-negative rods and cocci due to its aminothiazoyl and methoximino functional groups. -> cephalosporin

```{r}
# results$cephalosporin 

# results$cephalosporin %>% 
#   export::graph2ppt(append = TRUE, width = 8, height = 4,
#                     file = out_pptx)
```

- line plot MAG-Chrosmosome - Plasmid - phage - unknown facet | quant & % 

-- quant

```{r}
# ps_AMR_all %>% 
#   physeq_sel_tax_table(c("PathoFact_AMR_AMR_sub_class_multi",
#                          "PathoFact_AMR_MGE_prediction_clean_2",
#                          "PathoFact_AMR_ARG"))  %>%
#   # physeq_sel_tax_table(c("PathoFact_AMR_ARG"))  %>% 
#   speedyseq::tax_glom(., taxrank = "PathoFact_AMR_ARG")  %>% 
#   physeq_sel_tax_table(c("PathoFact_AMR_AMR_sub_class_multi",
#                          "PathoFact_AMR_ARG",
#                          "PathoFact_AMR_MGE_prediction_clean_2"))  %>%
#   speedyseq::tax_glom(., taxrank = "PathoFact_AMR_MGE_prediction_clean_2")  %>% 
#   prune_samples(samples, .) %>% 
#   physeq_add_metadata(metadata_mapping,
#                       sample_column = "sample_id") %>% 
#   filter_taxa(function(x) sum(x > 0) > 0, TRUE) %>% 
#   speedyseq::psmelt() %>% 
#   filter(!grepl('nclassified', PathoFact_AMR_AMR_sub_class_multi))  -> df
# 
# results <- vector("list", length(df$PathoFact_AMR_AMR_sub_class_multi %>%  unique()))
# names(results) = c(df$PathoFact_AMR_AMR_sub_class_multi %>%  unique())
# 
# for (dd in df %>% filter(PathoFact_AMR_AMR_sub_class_multi != "pleuromutilin antibiotic") %>% pull(PathoFact_AMR_AMR_sub_class_multi) %>%  unique()
# ){
#   
#   print(dd)
#   
#   df %>% 
#     filter(PathoFact_AMR_AMR_sub_class_multi == !!dd) %>% 
#     select(-Lactose_mM:-Total_SCFA_mM) %>% 
#     mutate(PathoFact_AMR_MGE_prediction_clean_2 = case_when(PathoFact_AMR_MGE_prediction_clean_2 %in% c("MAG", "chromosome")  ~ "MAG - chromosome", TRUE ~ PathoFact_AMR_MGE_prediction_clean_2)) %>% 
#     mutate(PathoFact_AMR_MGE_prediction_clean_2 = fct_relevel(PathoFact_AMR_MGE_prediction_clean_2, c("MAG - chromosome", "plasmid", "phage", "ambiguous", "unclassified"))) %>% 
#     group_by(Sample, PathoFact_AMR_MGE_prediction_clean_2, PathoFact_AMR_AMR_sub_class_multi) %>% 
#     summarise(sum_abundance = sum(Abundance), .groups = "drop") %>% 
#     mutate(sum_abundance = na_if(sum_abundance, 0)) %>%
#     left_join(metadata_mapping,
#               by = c("Sample" = "sample_id")) %>% 
#     filter(! PathoFact_AMR_MGE_prediction_clean_2  %in% c("unclassified", "ambiguous")) %>% 
#     # filter(Model == "Chicken") %>% 
#     # mutate(Day_of_Treatment_num = as.double(Day_of_Treatment)) %>% 
#     # mutate(Day_of_Treatment = fct_relevel(Day_of_Treatment, c(NA, "-3", "-1", "1" , "3", "4", "5","7", "30", "47", "48")))  %>% 
#     mutate(Antibiotic_mg.mL = factor(Antibiotic_mg.mL, levels = c(NA, 20, 200,90, 600))) %>%
#     ggplot(aes_string(x = "Day_of_Treatment", y = "sum_abundance", color = "Treatment", fill  = "Treatment", shape = "Antibiotic_mg.mL")) + #shape = Fermentation
#     geom_point(size = 2.5) + 
#     geom_line(linetype = 2,  size = 0.5,  aes_string(group = "interaction(Model, Reactor_Treatment_Dose)")) +
#     labs(y = paste0("sum Rnum_Gi ") , x = "Day_of_Treatment") +
#     scale_color_manual(name = "", values = treat_col,
#                        na.value = "black") +
#     scale_fill_manual(name = "", values = treat_col,
#                       na.value = "black") +
#     scale_shape_manual(name = "" ,values = c(15,16,18,19), na.value =  17) +
#     # scale_alpha_continuous(name = "", range=c(0.6, 1), na.value =  1) +
#     geom_vline(xintercept = 0, size = 0.5, linetype = 1, color = "grey60") +
#     ggtitle(dd) +
#     theme_light() + 
#     facet_grid(Antibiotic + PathoFact_AMR_MGE_prediction_clean_2 ~ Model, scales = "free", drop = TRUE) +  scale_y_continuous(labels=scientific) -> p1
#   
#   # p1 + facet_grid(as.formula(paste0("Antibiotic_mg.mL + PathoFact_AMR_AMR_sub_class_multi ~  Model  ")), scales = "fixed", space = "fixed", drop = TRUE)
#   # 
#   # p1 + facet_wrap(as.formula(paste0("PathoFact_AMR_AMR_sub_class_multi ~  Model  ")),scales = "free"
#   #)
#   results[[dd]] <- p1
#   
#   results[[dd]]  %>% 
#     ggsave(plot = .,bg = "transparent",scale = 2,
#            units = "mm",
#            dpi = 600,
#            height = 0.618 * 84 * 1.5,
#            width = 84,
#            filename = paste0("~/Desktop/lineplot_", dd, ".eps"))
#   
#   results[[dd]]  %>%
#     export::graph2ppt(append = TRUE, width = 317.48031496 , height = 0.618 * 317.48031496, paper = "A4",  scaling = 2,
#                       file = out_pptx)
#   
# }
```

-- then %

```{r}
# ps_AMR_all %>% 
#   physeq_sel_tax_table(c("PathoFact_AMR_AMR_sub_class_multi",
#                          "PathoFact_AMR_MGE_prediction_clean_2",
#                          "PathoFact_AMR_ARG"))  %>%
#   # physeq_sel_tax_table(c("PathoFact_AMR_ARG"))  %>% 
#   speedyseq::tax_glom(., taxrank = "PathoFact_AMR_ARG")  %>% 
#   physeq_sel_tax_table(c("PathoFact_AMR_AMR_sub_class_multi",
#                          "PathoFact_AMR_ARG",
#                          "PathoFact_AMR_MGE_prediction_clean_2"))  %>%
#   speedyseq::tax_glom(., taxrank = "PathoFact_AMR_MGE_prediction_clean_2")  %>% 
#   prune_samples(samples, .) %>% 
#   physeq_add_metadata(metadata_mapping,
#                       sample_column = "sample_id") %>% 
#   filter_taxa(function(x) sum(x > 0) > 0, TRUE) %>% 
#   speedyseq::psmelt() %>% 
#   select(-Lactose_mM:-Total_SCFA_mM) %>% 
#   mutate(PathoFact_AMR_MGE_prediction_clean_2 = case_when(PathoFact_AMR_MGE_prediction_clean_2 %in% c("MAG", "chromosome")  ~ "MAG - chromosome", TRUE ~ PathoFact_AMR_MGE_prediction_clean_2)) %>% 
#   mutate(PathoFact_AMR_MGE_prediction_clean_2 = fct_relevel(PathoFact_AMR_MGE_prediction_clean_2, c("MAG - chromosome", "plasmid", "phage", "ambiguous", "unclassified"))) -> df  #%>% 
# # filter(grepl('unclassified', PathoFact_AMR_AMR_sub_class_multi))  -> df
# 
# results <- vector("list", length(df$PathoFact_AMR_AMR_sub_class_multi %>%  unique()))
# names(results) = c(df$PathoFact_AMR_AMR_sub_class_multi %>%  unique())
# 
# for (dd in df %>% filter(PathoFact_AMR_AMR_sub_class_multi %!in% c("pleuromutilin antibiotic", "MLS (unclassified)", "Unclassified", "multidrug (unclassified)")) %>% pull(PathoFact_AMR_AMR_sub_class_multi) %>%  unique()
# )
# {
#   print(dd)
#   
#   df %>% 
#     filter(PathoFact_AMR_AMR_sub_class_multi == !!dd) %>% 
#     select(Sample, PathoFact_AMR_AMR_sub_class_multi, PathoFact_AMR_MGE_prediction_clean_2, Abundance) %>%
#     group_by(PathoFact_AMR_AMR_sub_class_multi, PathoFact_AMR_MGE_prediction_clean_2, Sample) -> df_tmp
#   
#   df_tmp %>% 
#     summarise(sum_MGE = sum(Abundance), .groups = "drop") %>%
#     mutate(sum_MGE = na_if(sum_MGE, 0)) -> df_MGE_sum
#   
#   df_tmp %>% 
#     group_by(Sample) %>% 
#     summarise(sample_sum = sum(Abundance), .groups = "drop") %>% 
#     mutate(sample_sum = na_if(sample_sum, 0)) -> df_sample_sum 
#   
#   df_MGE_sum %>% 
#     left_join(df_sample_sum,
#               by = c("Sample" = "Sample")) %>% 
#     mutate(percent_MGE = sum_MGE / sample_sum) %>% 
#     # left_join()
#     left_join(metadata_mapping,
#               by = c("Sample" = "sample_id")) %>% 
#     filter(! PathoFact_AMR_MGE_prediction_clean_2  %in% c("unclassified", "ambiguous")) %>%
#     # filter(Model == "Chicken") %>% 
#     # mutate(Day_of_Treatment_num = as.double(Day_of_Treatment)) %>% 
#     # mutate(Day_of_Treatment = fct_relevel(Day_of_Treatment, c(NA, "-3", "-1", "1" , "3", "4", "5","7", "30", "47", "48")))  %>% 
#     mutate(Antibiotic_mg.mL = factor(Antibiotic_mg.mL, levels = c(NA, 20, 200,90, 600))) %>%
#     ggplot(aes_string(x = "Day_of_Treatment", y = "percent_MGE", color = "Treatment", fill  = "Treatment", shape = "Antibiotic_mg.mL")) + #shape = Fermentation
#     geom_point(size = 2.5) + 
#     geom_line(linetype = 2,  size = 0.5,  aes_string(group = "interaction(Model, Reactor_Treatment_Dose)")) +
#     labs(y = paste0(" proportion % ") , x = "Day_of_Treatment") +
#     scale_color_manual(name = "", values = treat_col,
#                        na.value = "black") +
#     scale_fill_manual(name = "", values = treat_col,
#                       na.value = "black") +
#     scale_shape_manual(name = "" ,values = c(15,16,18,19), na.value =  17) +
#     # scale_alpha_continuous(name = "", range=c(0.6, 1), na.value =  1) +
#     geom_vline(xintercept = 0, size = 0.5, linetype = 1, color = "grey60") +
#     ggtitle(dd) +
#     theme_light() + 
#     facet_grid(Antibiotic + PathoFact_AMR_MGE_prediction_clean_2 ~ Model, scales = "free", drop = TRUE) +  scale_y_continuous(labels=percent) -> p1
#   # scale_y_continuous(labels = function(x) format(x, scientific = TRUE)) -> p1
#   
#   # p1 + facet_grid(as.formula(paste0("Antibiotic_mg.mL + PathoFact_AMR_AMR_sub_class_multi ~  Model  ")), scales = "fixed", space = "fixed", drop = TRUE)
#   # 
#   # p1 + facet_wrap(as.formula(paste0("PathoFact_AMR_AMR_sub_class_multi ~  Model  ")),scales = "free"
#   #)
#   results[[dd]] <- p1
#   
#   
#   results[[dd]]  %>% 
#     ggsave(plot = .,bg = "transparent",scale = 2,
#            units = "mm",
#            dpi = 600,
#            height = 0.618 * 84 * 1.5,
#            width = 84,
#            filename = paste0("~/Desktop/lineplot_", dd, "_percent.eps"))
#   
#   results[[dd]]  %>%
#     export::graph2ppt(append = TRUE, width = 317.48031496 , height = 0.618 * 317.48031496, paper = "A4",  scaling = 2,
#                       file = out_pptx)
#   
# }
```

and then baseline ratio.

```{r}
# ps_AMR_all %>% 
#   physeq_sel_tax_table(c("PathoFact_AMR_AMR_sub_class_multi",
#                          "PathoFact_AMR_MGE_prediction_clean_2",
#                          "PathoFact_AMR_ARG"))  %>%
#   # physeq_sel_tax_table(c("PathoFact_AMR_ARG"))  %>% 
#   speedyseq::tax_glom(., taxrank = "PathoFact_AMR_ARG")  %>% 
#   physeq_sel_tax_table(c("PathoFact_AMR_AMR_sub_class_multi",
#                          "PathoFact_AMR_ARG",
#                          "PathoFact_AMR_MGE_prediction_clean_2"))  %>%
#   speedyseq::tax_glom(., taxrank = "PathoFact_AMR_MGE_prediction_clean_2")  %>% 
#   prune_samples(samples, .) %>% 
#   physeq_add_metadata(metadata_mapping,
#                       sample_column = "sample_id") %>% 
#   filter_taxa(function(x) sum(x > 0) > 0, TRUE) %>% 
#   speedyseq::psmelt() %>% 
#   filter(! grepl('nclassified', PathoFact_AMR_AMR_sub_class_multi))  -> df
# 
# results <- vector("list", length(df$PathoFact_AMR_AMR_sub_class_multi %>%  unique()))
# names(results) = c(df$PathoFact_AMR_AMR_sub_class_multi %>%  unique())
# 
# for (dd in df %>% filter(PathoFact_AMR_AMR_sub_class_multi != "pleuromutilin antibiotic") %>% pull(PathoFact_AMR_AMR_sub_class_multi) %>%  unique()
# )
# {
#   print(dd)
#   
#   df %>% 
#     filter(PathoFact_AMR_AMR_sub_class_multi == !!dd) %>% 
#     select(-Lactose_mM:-Total_SCFA_mM) %>% 
#     mutate(PathoFact_AMR_MGE_prediction_clean_2 = case_when(PathoFact_AMR_MGE_prediction_clean_2 %in% c("MAG", "chromosome")  ~ "MAG - chromosome", TRUE ~ PathoFact_AMR_MGE_prediction_clean_2)) %>% 
#     mutate(PathoFact_AMR_MGE_prediction_clean_2 = fct_relevel(PathoFact_AMR_MGE_prediction_clean_2, c("MAG - chromosome", "plasmid", "phage", "ambiguous", "unclassified"))) %>% 
#     group_by(Sample, PathoFact_AMR_AMR_sub_class_multi, PathoFact_AMR_MGE_prediction_clean_2) %>% 
#     summarise(sum_abundance = sum(Abundance), .groups = "drop") %>% 
#     mutate(sum_abundance = na_if(sum_abundance, 0)) %>%
#     ungroup() %>% 
#     left_join(metadata_mapping,
#               by = c("Sample" = "sample_id")) %>% 
#     filter(! PathoFact_AMR_MGE_prediction_clean_2  %in% c("unclassified", "ambiguous")) %>%
#     group_by(PathoFact_AMR_AMR_sub_class_multi, PathoFact_AMR_MGE_prediction_clean_2, Model, Fermentation, Reactor_Treatment_Dose) %>% 
#     arrange(Day_of_Treatment) %>% 
#     mutate(per_cent_change = sum_abundance / first(sum_abundance)) %>% 
#     ggplot(aes_string(x = "Day_of_Treatment", y = "per_cent_change", color = "Treatment", fill  = "Treatment", shape = "Antibiotic_mg.L")) + #shape = Fermentation
#     geom_point(size = 2.5) + 
#     geom_line(linetype = 2,  size = 0.5,  aes_string(group = "interaction(Model, Reactor_Treatment_Dose)")) +
#     labs(y = "sum Rnum_Gi / baseline", x = "Day_of_Treatment") +
#     scale_color_manual(name = "", values = treat_col,
#                        na.value = "black") +
#     scale_fill_manual(name = "", values = treat_col,
#                       na.value = "black") +
#     scale_shape_manual(name = "" ,values = c(15,16,18,19), na.value =  17) +
#     # scale_alpha_continuous(name = "", range=c(0.6, 1), na.value =  1) +
#     geom_vline(xintercept = 0, size = 0.5, linetype = 1, color = "grey60") +
#     geom_hline(yintercept = 1, size = 0.5, linetype = 2, color = "grey60") +
#     ggtitle(dd) +
#     theme_light() + facet_grid(Antibiotic + PathoFact_AMR_MGE_prediction_clean_2 ~ Model, scales = "free", drop = TRUE) -> p1
#   
#   p1
#   
#   results[[dd]] <- p1
#   
#   
#   results[[dd]]  %>% 
#     ggsave(plot = .,bg = "transparent",scale = 2,
#            units = "mm",
#            dpi = 600,
#            height = 0.618 * 84 * 1.5,
#            width = 84,
#            filename = paste0("~/Desktop/lineplot_", dd, "_baseline_ratio.eps"))
#   
#   results[[dd]]  %>%
#     export::graph2ppt(append = TRUE, width = 317.48031496 , height = 0.618 * 317.48031496, paper = "A4",  scaling = 2,
#                       file = out_pptx)
#   
# }

```

- glycopeptide antibiotic:

```{r}
ps_AMR_all %>% 
  physeq_sel_tax_table(c("PathoFact_AMR_AMR_sub_class", "PathoFact_AMR_AMR_sub_class_multi", "PathoFact_AMR_ARG", "PathoFact_AMR_MGE_prediction_clean_2"))  %>%
  prune_samples(samples, .) %>% 
  physeq_add_metadata(metadata_mapping,
                      sample_column = "sample_id") %>% 
  subset_taxa(PathoFact_AMR_AMR_sub_class_multi %in% c("glycopeptide antibiotic")) %>% 
  filter_taxa(function(x) sum(x > 0) > 0, TRUE) -> ps_tmp

ps_tmp %>% 
  tax_table() %>% 
  data.frame() %>% 
  mutate(PathoFact_AMR_MGE_prediction_clean_2 = case_when(grepl("ambiguous", PathoFact_AMR_MGE_prediction_clean_2) ~ "ambiguous", TRUE ~ PathoFact_AMR_MGE_prediction_clean_2)) %>% 
  # mutate(PathoFact_AMR_MGE_prediction_clean_2 = case_when(PathoFact_AMR_MGE_prediction_clean_2 %in% c("MAG", "chromosome")  ~ "MAG - chromosome", TRUE ~ PathoFact_AMR_MGE_prediction_clean_2)) %>% 
  mutate(PathoFact_AMR_MGE_prediction_clean_2 = fct_relevel(PathoFact_AMR_MGE_prediction_clean_2, c("MAG", "chromosome", "plasmid", "phage", "ambiguous", "unclassified"))) %>% 
  as.matrix() -> tax_table(ps_tmp)

ps_tmp %>% 
  ps_arrange(-Day_of_Treatment) %>% 
  comp_barplot(
    tax_transform_for_plot = "identity",
    tax_level = "PathoFact_AMR_ARG",
    label = "Day_of_Treatment", # name an alternative variable to label axes
    n_taxa = 20, # give more taxa unique colours
    merge_other = FALSE, # split the "other" category to display alpha diversity
    bar_width = 0.9, # reduce the bar width to 70% of one row
    bar_outline_colour = "grey5",
    sample_order = "default") +#,# is the default (use NA to remove outlines)
  # facet_by = c("Model", "Antibiotic", "Antibiotic_mg.mL", "Treatment2")) +
  # group_by = c("Model", "Antibiotic", "Antibiotic_mg.mL", "Treatment2"))
  coord_flip() +  theme_light() +
  scale_y_continuous(labels=scientific) -> p2
# 
# # Plot them side by side with the patchwork package.
# patch <- patchwork::wrap_plots(plot_list, nrow = 1, guides = "collect")
# patch & coord_flip() # make all plots horizontal (note: use & instead of +)
#   



p2 + facet_wrap(Model ~ Antibiotic + Antibiotic_mg.mL + Treatment2, scales = "free", ncol = 6, nrow = 4) -> p2_1

p2_1

p2 + facet_wrap(Model ~ Antibiotic + Antibiotic_mg.mL + Treatment2, scales = "free_y", ncol = 6, nrow = 4) -> p2_2

p2_2

p2_1 %>% 
  ggsave(plot = .,bg = "transparent",scale = 2,
         units = "mm",
         dpi = 600,
         height = 0.618 * 84 * 2,
         width = 84 * 2,
         filename = "~/Desktop/glyco_1-1.eps")

p2_2 %>% 
  ggsave(plot = .,bg = "transparent",scale = 2,
         units = "mm",
         dpi = 600,
         height = 0.618 * 84 * 2,
         width = 84 * 2,
         filename = "~/Desktop/glyco_1-2.eps")


p2_1 %>%
  export::graph2ppt(append = TRUE, width = 317.48031496 * 2  , height = 0.618 * 317.48031496 * 0.75 , paper = "A4",  scaling = 2,
                    file = out_pptx)

p2_2 %>%
  export::graph2ppt(append = TRUE, width = 317.48031496 * 2  , height = 0.618 * 317.48031496 * 0.75 , paper = "A4",  scaling = 2,
                    file = out_pptx)

```

- cephalosporin:

```{r}
ps_AMR_all %>% 
  physeq_sel_tax_table(c("PathoFact_AMR_AMR_sub_class", "PathoFact_AMR_AMR_sub_class_multi", "PathoFact_AMR_ARG", "PathoFact_AMR_MGE_prediction_clean_2"))  %>%
  prune_samples(samples, .) %>% 
  physeq_add_metadata(metadata_mapping,
                      sample_column = "sample_id") %>% 
  subset_taxa(PathoFact_AMR_AMR_sub_class_multi %in% c("cephalosporin")) %>% 
  filter_taxa(function(x) sum(x > 0) > 0, TRUE) -> ps_tmp

ps_tmp %>% 
  tax_table() %>% 
  data.frame() %>% 
  mutate(PathoFact_AMR_MGE_prediction_clean_2 = case_when(grepl("ambiguous", PathoFact_AMR_MGE_prediction_clean_2) ~ "ambiguous", TRUE ~ PathoFact_AMR_MGE_prediction_clean_2)) %>% 
  # mutate(PathoFact_AMR_MGE_prediction_clean_2 = case_when(PathoFact_AMR_MGE_prediction_clean_2 %in% c("MAG", "chromosome")  ~ "MAG - chromosome", TRUE ~ PathoFact_AMR_MGE_prediction_clean_2)) %>% 
  mutate(PathoFact_AMR_MGE_prediction_clean_2 = fct_relevel(PathoFact_AMR_MGE_prediction_clean_2, c("MAG", "chromosome", "plasmid", "phage", "ambiguous", "unclassified"))) %>% 
  as.matrix() -> tax_table(ps_tmp)

ps_tmp %>% 
  ps_arrange(-Day_of_Treatment) %>% 
  comp_barplot(
    tax_transform_for_plot = "identity",
    tax_level = "PathoFact_AMR_ARG",
    label = "Day_of_Treatment", # name an alternative variable to label axes
    n_taxa = 20, # give more taxa unique colours
    merge_other = FALSE, # split the "other" category to display alpha diversity
    bar_width = 0.9, # reduce the bar width to 70% of one row
    bar_outline_colour = "grey5",
    sample_order = "default") +#,# is the default (use NA to remove outlines)
  # facet_by = c("Model", "Antibiotic", "Antibiotic_mg.mL", "Treatment2")) +
  # group_by = c("Model", "Antibiotic", "Antibiotic_mg.mL", "Treatment2"))
  coord_flip() +  theme_light() +
  scale_y_continuous(labels=scientific) -> p2
# 
# # Plot them side by side with the patchwork package.
# patch <- patchwork::wrap_plots(plot_list, nrow = 1, guides = "collect")
# patch & coord_flip() # make all plots horizontal (note: use & instead of +)
#   



p2 + facet_wrap(Model ~ Antibiotic + Antibiotic_mg.mL + Treatment2, scales = "free", ncol = 6, nrow = 4) -> p2_1

p2_1

p2 + facet_wrap(Model ~ Antibiotic + Antibiotic_mg.mL + Treatment2, scales = "free_y", ncol = 6, nrow = 4) -> p2_2

p2_2

p2_1 %>% 
  ggsave(plot = .,bg = "transparent",scale = 2,
         units = "mm",
         dpi = 600,
         height = 0.618 * 84 * 2,
         width = 84 * 2,
         filename = "~/Desktop/cephalosporin_1-1.eps")

p2_2 %>% 
  ggsave(plot = .,bg = "transparent",scale = 2,
         units = "mm",
         dpi = 600,
         height = 0.618 * 84 * 2,
         width = 84 * 2,
         filename = "~/Desktop/cephalosporin_1-2.eps")


p2_1 %>%
  export::graph2ppt(append = TRUE, width = 317.48031496 * 2  , height = 0.618 * 317.48031496 * 0.75 , paper = "A4",  scaling = 2,
                    file = out_pptx)

p2_2 %>%
  export::graph2ppt(append = TRUE, width = 317.48031496 * 2  , height = 0.618 * 317.48031496 * 0.75 , paper = "A4",  scaling = 2,
                    file = out_pptx)

```


- VANA

```{r, eval = FALSE}
ps_AMR_all %>% 
  physeq_sel_tax_table(c("PathoFact_AMR_AMR_sub_class", "PathoFact_AMR_AMR_sub_class_multi", "PathoFact_AMR_MGE_prediction_clean_2", "PathoFact_AMR_ARG"))  %>%
  prune_samples(samples, .) %>% 
  physeq_add_metadata(metadata_mapping,
                      sample_column = "sample_id") %>% 
  subset_taxa(PathoFact_AMR_ARG %in% c("VANA")) %>% 
  filter_taxa(function(x) sum(x > 0) > 0, TRUE) -> ps_tmp

ps_tmp %>% 
  tax_table() %>% 
  data.frame() %>% 
  rownames_to_column('GN_id') %>% 
  mutate(PathoFact_AMR_MGE_prediction_clean_2 = case_when(grepl("ambiguous", PathoFact_AMR_MGE_prediction_clean_2) ~ "ambiguous", TRUE ~ PathoFact_AMR_MGE_prediction_clean_2)) %>% 
  # mutate(PathoFact_AMR_MGE_prediction_clean_2 = case_when(PathoFact_AMR_MGE_prediction_clean_2 %in% c("MAG", "chromosome")  ~ "MAG - chromosome", TRUE ~ PathoFact_AMR_MGE_prediction_clean_2)) %>% 
  mutate(PathoFact_AMR_MGE_prediction_clean_2 = fct_relevel(PathoFact_AMR_MGE_prediction_clean_2, c("MAG", "chromosome", "plasmid", "phage", "ambiguous", "unclassified"))) %>% 
  mutate(GN = paste(GN_id, PathoFact_AMR_MGE_prediction_clean_2, PathoFact_AMR_ARG, sep  = "_")) %>% 
  column_to_rownames("GN_id") %>% 
  as.matrix() -> tax_table(ps_tmp)

ps_tmp %>% 
  tax_filter(undetected = 0, use_counts = FALSE) %>% 
  # microViz::phyloseq_validate(use_counts = FALSE) %>% 
  microViz::ps_arrange(-Day_of_Treatment) %>% 
  microViz::comp_barplot(
    tax_transform_for_plot = "identity",
    tax_level = "GN",
    label = "Day_of_Treatment", # name an alternative variable to label axes
    n_taxa = 15, # give more taxa unique colours
    merge_other = FALSE, # split the "other" category to display alpha diversity
    bar_width = 0.9, # reduce the bar width to 70% of one row
    bar_outline_colour = "grey5",
    sample_order = "default") +#,# is the default (use NA to remove outlines)
  # facet_by = c("Model", "Antibiotic", "Antibiotic_mg.mL", "Treatment2")) +
  # group_by = c("Model", "Antibiotic", "Antibiotic_mg.mL", "Treatment2"))
  coord_flip() +  theme_light() +
  scale_y_continuous(labels=percent) -> p2


p2 + facet_wrap(Model ~ Antibiotic + Antibiotic_mg.mL + Treatment2, scales = "free", ncol = 6, nrow = 4) -> p2_1

p2_1

p2 + facet_wrap(Model ~ Antibiotic + Antibiotic_mg.mL + Treatment2, scales = "free_y", ncol = 6, nrow = 4) -> p2_2

p2_2

p2_1 %>% 
  ggsave(plot = .,bg = "transparent",scale = 2,
         units = "mm",
         dpi = 600,
         height = 0.618 * 84 * 2,
         width = 84 * 2,
         filename = "~/Desktop/vana_2-1.eps")

p2_2 %>% 
  ggsave(plot = .,bg = "transparent",scale = 2,
         units = "mm",
         dpi = 600,
         height = 0.618 * 84 * 2,
         width = 84 * 2,
         filename = "~/Desktop/vana_2-2.eps")


p2_1 %>%
  export::graph2ppt(append = TRUE, width = 317.48031496 * 2  , height = 0.618 * 317.48031496 * 0.75 , paper = "A4",  scaling = 2,
                    file = out_pptx)

p2_2 %>%
  export::graph2ppt(append = TRUE, width = 317.48031496 * 2  , height = 0.618 * 317.48031496 * 0.75 , paper = "A4",  scaling = 2,
                    file = out_pptx)

```

- CTX-M-1

```{r, eval = FALSE}
# ps_AMR_all %>% 
#   physeq_sel_tax_table(c("PathoFact_AMR_AMR_sub_class", "PathoFact_AMR_AMR_sub_class_multi", "PathoFact_AMR_MGE_prediction_clean_2", "PathoFact_AMR_ARG"))  %>%
#   prune_samples(samples, .) %>% 
#   physeq_add_metadata(metadata_mapping,
#                       sample_column = "sample_id") %>% 
#   subset_taxa(PathoFact_AMR_ARG %in% c("CTX-M-1")) %>% 
#   filter_taxa(function(x) sum(x > 0) > 0, TRUE) -> ps_tmp
# 
# ps_tmp %>% 
#   tax_table() %>% 
#   data.frame() %>% 
#   rownames_to_column('GN_id') %>% 
#   mutate(PathoFact_AMR_MGE_prediction_clean_2 = case_when(grepl("ambiguous", PathoFact_AMR_MGE_prediction_clean_2) ~ "ambiguous", TRUE ~ PathoFact_AMR_MGE_prediction_clean_2)) %>% 
#   # mutate(PathoFact_AMR_MGE_prediction_clean_2 = case_when(PathoFact_AMR_MGE_prediction_clean_2 %in% c("MAG", "chromosome")  ~ "MAG - chromosome", TRUE ~ PathoFact_AMR_MGE_prediction_clean_2)) %>% 
#   mutate(PathoFact_AMR_MGE_prediction_clean_2 = fct_relevel(PathoFact_AMR_MGE_prediction_clean_2, c("MAG", "chromosome", "plasmid", "phage", "ambiguous", "unclassified"))) %>% 
#   mutate(GN = paste(GN_id, PathoFact_AMR_MGE_prediction_clean_2, PathoFact_AMR_ARG, sep  = "_")) %>% 
#   column_to_rownames("GN_id") %>% 
#   as.matrix() -> tax_table(ps_tmp)
# 
# ps_tmp %>% 
#   ps_arrange(-Day_of_Treatment) %>% 
#   comp_barplot(
#     tax_transform_for_plot = "identity",
#     tax_level = "GN",
#     label = "Day_of_Treatment", # name an alternative variable to label axes
#     n_taxa = 18, # give more taxa unique colours
#     merge_other = FALSE, # split the "other" category to display alpha diversity
#     bar_width = 0.9, # reduce the bar width to 70% of one row
#     bar_outline_colour = "grey5",
#     sample_order = "default") +#,# is the default (use NA to remove outlines)
#   # facet_by = c("Model", "Antibiotic", "Antibiotic_mg.mL", "Treatment2")) +
#   # group_by = c("Model", "Antibiotic", "Antibiotic_mg.mL", "Treatment2"))
#   coord_flip() +  theme_light() +
#   scale_y_continuous(labels=scientific) -> p2
# # 
# # # Plot them side by side with the patchwork package.
# # patch <- patchwork::wrap_plots(plot_list, nrow = 1, guides = "collect")
# # patch & coord_flip() # make all plots horizontal (note: use & instead of +)
# #   
# 
# p2 + facet_wrap(Model ~ Antibiotic + Antibiotic_mg.mL + Treatment2, scales = "free", ncol = 6, nrow = 4) -> p2_1
# 
# p2_1
# 
# p2 + facet_wrap(Model ~ Antibiotic + Antibiotic_mg.mL + Treatment2, scales = "free_y", ncol = 6, nrow = 4) -> p2_2
# 
# p2_2
# 
# p2_1 %>% 
#   ggsave(plot = .,bg = "transparent",scale = 2,
#          units = "mm",
#          dpi = 600,
#          height = 0.618 * 84 * 2,
#          width = 84 * 2,
#          filename = "~/Desktop/CTX_1.eps")
# 
# p2_2 %>% 
#   ggsave(plot = .,bg = "transparent",scale = 2,
#          units = "mm",
#          dpi = 600,
#          height = 0.618 * 84 * 2,
#          width = 84 * 2,
#          filename = "~/Desktop/CTX_2.eps")
# 
# 
# p2_1 %>%
#   export::graph2ppt(append = TRUE, width = 317.48031496 * 2  , height = 0.618 * 317.48031496 * 0.75 , paper = "A4",  scaling = 2,
#                     file = out_pptx)
# 
# p2_2 %>%
#   export::graph2ppt(append = TRUE, width = 317.48031496 * 2  , height = 0.618 * 317.48031496 * 0.75 , paper = "A4",  scaling = 2,
#                     file = out_pptx)
```

HQ MAG contributors:

```{r}
ps_AMR_all %>% 
  physeq_sel_tax_table(c("PathoFact_AMR_AMR_sub_class", "PathoFact_AMR_AMR_sub_class_multi", "PathoFact_AMR_MGE_prediction_clean_2", "ANVIO_CONCOCT_HQ_clean", "CONCOCT_bin",
                         "GTDB_tax_CONCOCT_clean" ,"PathoFact_AMR_ARG")) %>% 
  prune_samples(samples, .) %>% 
  physeq_add_metadata(metadata_mapping,
                      sample_column = "sample_id") %>% 
  # subset_taxa(!is.na("GTDB_tax_CONCOCT_clean")) %>% 
  # physeq_sel_tax_table(c("PathoFact_AMR_ARG"))  %>% 
  speedyseq::tax_glom(., taxrank = "GTDB_tax_CONCOCT_clean")  %>% 
  filter_taxa(function(x) sum(x > 0) > 0, TRUE) %>% 
  microbiome::transform("compositional") %>%
  speedyseq::psmelt() %>% 
  separate(GTDB_tax_CONCOCT_clean, into = c("domain", "phylum", "class", "order", "family", "genus", "species"), sep  = ";" , remove = FALSE) -> df
```


```{r}

df %>% 
  filter(genus %in% c("g__Enterococcus_B",
                      "g__Escherichia")) %>%
  mutate(Abundance = na_if(Abundance, 0)) %>%
  mutate(logTPM = log10(Abundance + 1))  %>% 
  # separate_rows(., Resistance_Mechanism,sep = '; ',convert = TRUE) %>% 
  # ggplot(aes(x = as.factor(Day_of_Treatment), y = Abundance, fill = fct_reorder(Strain, logTPM))) +   
  ggplot(aes(x = as.factor(Day_of_Treatment), y = Abundance, fill = species)) +#reorder(genus, Abundance, FUN=median, na.rm = TRUE))) +
  geom_bar( stat = "identity", colour="black", size=0.025) +
  facet_wrap(Model ~ Antibiotic + Antibiotic_mg.mL + Treatment2, scales = "free_x", ncol = 6, nrow = 4) +
  # facet_grid(Model ~ Treatment + Fermentation + Antibiotic_mg.mL , scales = "free", space = "free_x", drop = TRUE) +
  ggtitle("ARG Type Abundance per Sample") +
  xlab("Day (Treatment)")  +
  # ylab("Proportion") + 
  theme_light() +
  scale_y_continuous(labels=percent) + 
  # scale_fill_manual(name = "", values = col_strains) +
  theme(legend.position = "bottom") +
  ggpubr::rotate_x_text(60) +
  theme(legend.position = ("bottom")) -> p_mec

# p_mec %>% 
#   ggpubr::set_palette("jco") -> p_mec

p_mec


p_mec %>% 
  export::graph2ppt(append = TRUE, width = 317.48031496 * 2  , height = 0.618 * 317.48031496 * 2 , paper = "A4",  scaling = 2,
                    file = out_pptx)
```

Need to clarify resistome contributors since some gene are assigned to bins but are lacking GTDB taxonomy. updated collection?


https://github.com/fconstancias/NRP72-FBT/blob/2ad42987be250d5b4ab81758733571c57d0971b1/data/processed/16S-AMR-MGE-astetics.Rmd

mutate(ANVIO_CONCOCT_HQ_clean = ifelse(PathoFact_AMR_MGE_prediction == "plasmid", NA, ANVIO_CONCOCT_HQ)) %>%

contigs of interest: 
c1_megahit_8588
c1_megahit_219391


```{r}
devtools::session_info()
```

- the top gene -> where line plot facet localisation? heatmap?
- heatmap AMR class + genetic material + dendro
- heatmap AMR GN + dendro


Resistome Proportion
AMR categories
Gene heatmap
Diversity: alpha/beta
Contributing genetic material (MAGs/Chromosomes/plasmids/)
Number of AMR/ contigs / plasmids  length
Number of unmapped reads / sample
Prop/coverage or whatsoever per gene per mag per sample (rector/time)
Visualise contigs with ARG in anvo -> coverage variability. 

proof of concept with WAFLE METACHIP
proof of concept ANVIO -> population variability

MAG resistome?
donor strain chromosomal resistome?
