---
title: "NTP72 - 16S - taxa compostiion - chicken draft "
author: "Florentin Constancias"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: 
  html_document: 
    toc: yes
    keep_md: yes
---

```{r setup, include=FALSE}
rm(list = ls())
gc()
options(java.parameters = "-Xmx80000m")
```

```{r setup2, include=FALSE}

##knitr::opts_chunk$set(cache=TRUE)
##knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(dev = "ragg_png") # https://ragg.r-lib.org/index.html
##knitr::opts_chunk$set(fig.width = 6) #https://r4ds.had.co.nz/graphics-for-communication.html#figure-sizing
##knitr::opts_chunk$set(fig.asp = 0.618)
##knitr::opts_chunk$set(fig.show = "hold")
##knitr::opts_chunk$set(fig.show = "70%")
##knitr::opts_chunk$set(fig.align = "center")
```

```{r packages, results=FALSE, warning=FALSE}
#install.packages("tidyverse")
require(tidyverse); packageVersion("tidyverse")
require(phyloseq); packageVersion("phyloseq")
```

```{r sourcing, warning=FALSE, results=FALSE , message= FALSE}
source("https://raw.githubusercontent.com/fconstancias/DivComAnalyses/master/R/phyloseq_taxa_tests.R") 
source("https://raw.githubusercontent.com/fconstancias/DivComAnalyses/master/R/phyloseq_alpha.R") 
source("https://raw.githubusercontent.com/fconstancias/DivComAnalyses/master/R/phyloseq_beta.R") 
source("https://raw.githubusercontent.com/fconstancias/DivComAnalyses/master/R/phyloseq_heatmap.R")
source("https://raw.githubusercontent.com/fconstancias/metabaRpipe-source/master/Rscripts/functions.R")
source("https://raw.githubusercontent.com/fconstancias/DivComAnalyses/master/R/phyloseq_normalisation.R") 
source("https://raw.githubusercontent.com/fconstancias/DivComAnalyses/master/R/phyloseq_varia.R")


'%!in%' <- function(x,y)!('%in%'(x,y))

```


```{r}
out_pptx = "~/Desktop/16S-chick.pptx"
out_xlsx = "~/Desktop/16S-chick.xlsx"
```

```{r}
load(url("https://github.com/fconstancias/NRP72-FBT/blob/master/Figures/Rastetics.Rdata?raw=true"))
```

## load data:

```{r}
"data/processed/16S/16S_working_phyloseq.RDS" %>% 
  here::here() %>% 
  readRDS()  %>% 
  subset_samples(Enrichment == "NotEnriched") %>%
      subset_samples(Experiment %in% c("Continuous", "Cecum")) %>%
   subset_samples(Day_of_Treatment > -4 & Day_of_Treatment <= 30 | Reactor == "DONOR") %>% 
  subset_samples(Model == "Chicken") %>% 
   subset_samples(Reactor != "CR" & Model2 == "Chicken1" | Model2 == "Chicken2" & Reactor %in% c("TR2", "TR3", "TR4", "TR5", "TR6", "TR7", "CR", "DONOR")) %>% 
   subset_samples(Reactor != "IR") -> ps_filtered
```

```{r}
ps_filtered %>% 
  phyloseq_check_lib_size(data_color = "Reactor", data_facet = "Reactor", first_n = 50, nreads_display = 1000) -> out

out$df %>% 
  # rownames_to_column('sample_id') %>% 
  # filter(Treatment == "DONOR")  %>% 
  select(SampleID, Day_of_Treatment,Treatment, Reactor, Model2, Model, Period, LibrarySize) 
```

## rarefraction

```{r}
min_sample_size = 3738

ps_filtered %>%
  rarefy_even_depth(sample.size = min_sample_size, rngseed = 123)  -> ps_rare

ps_filtered %>%
  prune_samples(sample_sums(.)>= min_sample_size, .) -> ps_fil
```

Objectives:

-   compare donor microbiota and pre-treatment / CR - heatmap: v
-   evaluate effect treatment: -- alpha (shanon, simpson --\> hill d1, d2) -- beta (achinson, log -\> bray)
-   describe effect on taxa (Family, Genus, ASV): v -- heatmap v -- alluvial plot v

--> Ezequiel: process - co-selection of AMR.


```{r}
# ps_rare %>%  
#   sample_data() %>%
#   # dplyr::select(Model2)
#   # dplyr::filter(model2 == "Chicken2") %>%
#   data.frame()
```
## Before Treatment:

Obj: compare donor microbiota and pre-treatment / CR - heatmap

```{r}
ps_rare %>%
  subset_samples(Day_of_Treatment > -4 & Day_of_Treatment <= 0 | Reactor %in% c("DONOR", "CR")) -> before_treat # %>% 
# subset_samples(Reactor %!in%c("IR")) %>% 
# subset_samples() -> before_treat
```

```{r}
ps_rare %>% 
  sample_data() %>% 
  data.frame() %>% 
  filter(Reactor == "DONOR") 
```

```{r}
# before_treat %>% 
#   # microViz::ps_mutate(tp_var = )
#   physeq_most_abundant(group_var = "Model2",
#                        ntax = 24,
#                        tax_level = "Strain") -> top_taxa_per_group

before_treat %>%
  physeq_most_abundant(group_var = "Reactor_Treatment",
                       ntax = 10,
                       tax_level = "Strain") -> top_taxa_per_group

top_taxa_per_group
```

```{r}
before_treat %>%
  rarefy_even_depth(sample.size = min_sample_size, rngseed = 123) %>%
  # subset_samples(Reactor == "DONOR") %>%  sample_data() %>%  data.frame()
  # speedyseq::tax_glom(taxrank = "Family") %>% 
  transform_sample_counts(function(x) x/sum(x) * 100) %>%
  subset_taxa(Strain %in% top_taxa_per_group) %>% # extract only the taxa to display - after percentage normalisation
  phyloseq_ampvis_heatmap(physeq = ., tax_aggregate = "Species",
                          transform = FALSE,
                          plot_values = FALSE,
                          facet_by = c("Model2", "Fermentation","Reactor_Treatment_Dose"),
                          group_by = "Day_of_Treatment",#"sample_name",
                          ntax = Inf
  ) -> pre_speces

pre_speces$data %>% 
  mutate(Abundance = na_if(Abundance, 0)) %>% 
  mutate(Fermentation = Fermentation %>%  as.character() %>%  replace_na(., "NA")) %>% 
  mutate(Fermentation = factor(Fermentation, levels = c("NA", 1, 2))) -> pre_speces$data 

pre_speces + 
  facet_grid(. ~  Model2 + Fermentation + Reactor_Treatment_Dose, scales = "free", space = "free") +
  scale_color_manual(values = c("#31a354", "orange", 
                                "#f03b20"), drop = FALSE, na.value = 'transparent') -> pre_speces
# scale_fill_viridis_c(breaks = c(0,  0.01, 1, 10, 25, 50, 75, 100),
# labels = c(0,  0.01, 1, 10, 25,  50, 75, 100),
# trans = scales::pseudo_log_trans(sigma = 0.9),
# na.value = 'transparent') -> pre_speces


#pre_speces

```

```{r}
before_treat %>%
  rarefy_even_depth(sample.size = min_sample_size, rngseed = 123) %>%
  # subset_samples(Reactor == "DONOR") %>%  sample_data() %>%  data.frame()
  # speedyseq::tax_glom(taxrank = "Family") %>% 
  transform_sample_counts(function(x) x/sum(x) * 100) %>%
  phyloseq_ampvis_heatmap(physeq = ., tax_aggregate = "Family",
                          transform = FALSE,
                          plot_values = FALSE,
                          facet_by = c("Model2", "Fermentation","Reactor_Treatment_Dose"),
                          group_by = "Day_of_Treatment",#"sample_name",
                          ntax = 8
  ) -> pre_Family

pre_Family$data %>% 
  mutate(Abundance = na_if(Abundance, 0)) %>% 
  mutate(Fermentation = Fermentation %>%  as.character() %>%  replace_na(., "NA")) %>% 
  mutate(Fermentation = factor(Fermentation, levels = c("NA", 1, 2))) -> pre_Family$data 


pre_Family + 
  facet_grid(. ~  Model2 + Fermentation + Reactor_Treatment_Dose, scales = "free", space = "free") +
  scale_color_manual(values = c("#31a354", "orange", 
                                "#f03b20"), drop = FALSE, na.value = 'transparent') -> pre_Family


# pre_Family
```

```{r}
# cowplot::plot_grid(pre_speces,
#           pre_Family, 
#            ncol = 1,
#           align = "v", axis = "b",
#           rel_heights = c(3, 1)) -> pre_plots
# 
# pre_plots
```

```{r}

ggpubr::ggarrange(pre_Family + theme(legend.position = "null") + 
                    theme(axis.title.x=element_blank(),
                          axis.text.x=element_blank(),
                          axis.ticks.x=element_blank()),
                  pre_speces + theme(
                    strip.background = element_blank(),
                    strip.text.x = element_blank()
                  )  + theme(legend.position = "null"),
                  align = "v",
                  ncol = 1, 
                  heights = c(1, 2.40),
                  common.legend = FALSE) -> pre_plots

pre_plots

pre_plots %>% 
  export::graph2ppt(append = TRUE, 
                    width = 317.48031496 * 6, 
                    height = 0.618 * 317.48031496 * 8 , paper = "A4",  scaling = 2,
                    file = out_pptx)

pre_Family %>% 
  ggpubr::get_legend() %>% 
  ggpubr::as_ggplot() %>% 
  export::graph2ppt(append = TRUE, 
                    width = 317.48031496 * 1, 
                    height = 0.618 * 317.48031496 * 1 , paper = "A4",  scaling = 2,
                    file = out_pptx)


pre_Family$data %>% 
  select(Display, Sample, Abundance,Reactor_Treatment_Dose) %>% 
  pivot_wider(names_from = Display, values_from = Abundance) %>% 
  xlsx::write.xlsx(file = out_xlsx,
                   append = TRUE,
                   sheetName = "pre_family",
                   showNA = TRUE)
pre_speces$data %>% 
  select(Display, Sample, Abundance,Reactor_Treatment_Dose) %>% 
  pivot_wider(names_from = Display, values_from = Abundance) %>% 
  xlsx::write.xlsx(file = out_xlsx,
                   append = TRUE,
                   sheetName = "pre_ASV",
                   showNA = TRUE)
```
## Hannah's classification:

<https://github.com/haegih/MSc_Thesis-NRP72/blob/main/scripts/ASV_classification/chicken_asv-classification.md>

Species were categorized into three groups based on the following criteria: 

<https://www.tandfonline.com/doi/full/10.1080/19490976.2021.1900995>

1. Disappeared: the species undetected in all post-treatment samples (T4 to T6); 
2. New: the species undetected in the baseline, but persistently abun- dant (relative abundance > 0.1%) in all post-treatment samples (T4 to T6); 3. Proliferating: the species meeting any of the two criteria: 
  1) a species with a relative abundance < 0.1% in baseline and reached 5% in T2 or T3, or 
  2) a species with a relative abundance ≥ 0.1% in baseline and increased by at least 10-fold in the relative abundance in T2 or T3.

Challenges: 
- thresholds
- treatment phases vs days

Define function to make transform absolute into relative abundance:

```{r}

rel_abundance_transformation <- function(list){
    list %>%
    data.frame() %>%
    dplyr::filter(!is.na(Period)) %>% 
    select(-c(Model, Treatment)) %>%
    column_to_rownames("Period") %>%
    data.matrix() %>%
    funrar::make_relative() %>%
    data.frame() -> rel_abund_df

    
    return(rel_abund_df)
}
```

Define function for asv classification:

```{r}
asv_classification <- function(abund_per_period, class_table){


for (asv in rownames(class_table)){
  
  # define as new: 
  # TODO: lower threshold?

    # define as proliferating
if (abund_per_period['pret',asv] == 0 & 
      (abund_per_period['t1',asv] > 0.001 | 
      abund_per_period['t2',asv] > 0.001 |
      abund_per_period['t3',asv] > 0.001 |
      abund_per_period['t4',asv] > 0.001 |
      abund_per_period['t5',asv] > 0.001)){

     class_table[asv,] = "new"
    
} else if (abund_per_period['pret',asv] < 0.001 &
            (abund_per_period['t1',asv] > 0.05 |
             abund_per_period['t2',asv] > 0.05) &
             abund_per_period['t3',asv] > max(abund_per_period['t1',asv], abund_per_period['t2',asv]) &
             abund_per_period['t4',asv] > abund_per_period['t3',asv] &
             abund_per_period['t5',asv] > abund_per_period['t4',asv]
            ){
   class_table[asv,] = "prol"
   
 } else if (abund_per_period['pret',asv] < 0.001 & #including all the options increases amount of classified "proliferating"
             (abund_per_period['t1',asv] > 0.05 | 
              abund_per_period['t2',asv] > 0.05 |
              abund_per_period['t3',asv] > 0.05 |
              abund_per_period['t4',asv] > 0.05 | 
              abund_per_period['t5',asv] > 0.05)
             ){
    
    class_table[asv,] = "prol"
  
  } else if (abund_per_period['pret',asv] > 0 & ##changed
             (abund_per_period['t1',asv] >= 5*abund_per_period['pret',asv] &
              abund_per_period['t2',asv] >= abund_per_period['t1',asv] &
              abund_per_period['t3',asv] >= abund_per_period['t2',asv] &
              abund_per_period['t4',asv] >= abund_per_period['t3',asv] &
              abund_per_period['t5',asv] >= abund_per_period['t4',asv])
             ) {

    class_table[asv,] = "prol"
    
  } else if (abund_per_period['pret',asv] >= 0.001 &
             (abund_per_period['t1',asv] >= 5*abund_per_period['pret',asv] |
              abund_per_period['t2',asv] >= 5*abund_per_period['pret',asv] |
              abund_per_period['t3',asv] >= 5*abund_per_period['pret',asv] |
              abund_per_period['t4',asv] >= 5*abund_per_period['pret',asv] |
              abund_per_period['t5',asv] >= 5*abund_per_period['pret',asv])
             ) {

    class_table[asv,] = "prol"

 # Define as recovering
  } else if (abund_per_period['pret',asv] >= 0.01 &
             10*abund_per_period['t1',asv] < abund_per_period['pret',asv] &
             # # Case 1: Recovery only at t5
             # ((abund_per_period['t2',asv] <= abund_per_period['t1',asv] &
             #  abund_per_period['t3',asv] <= abund_per_period['t2',asv] &
             #  abund_per_period['t4',asv] <= abund_per_period['t3',asv])|
             #  # Case 2: Recovery at t4
             # (abund_per_period['t2',asv] <= abund_per_period['t1',asv] &
             #  abund_per_period['t3',asv] <= abund_per_period['t2',asv] &
             #  abund_per_period['t4',asv] > abund_per_period['t3',asv]) |
             # # Case 3: Recovery at t3
             # (abund_per_period['t2',asv] <= abund_per_period['t1',asv] &
             #  abund_per_period['t3',asv] > abund_per_period['t2',asv]  &
             #  abund_per_period['t4',asv] > abund_per_period['t3',asv]) |
             # # Case 4: Recovery at t2
             # (abund_per_period['t2',asv] > abund_per_period['t1',asv] &
             #  abund_per_period['t3',asv] > abund_per_period['t2',asv] &
             #  abund_per_period['t4',asv] > abund_per_period['t3',asv])) &
              abund_per_period['t5',asv] >= 0.01
             ) {

    class_table[asv,] = "recovering"
  
    # define as having temporal advantage
    
   } else if (abund_per_period['pret',asv] == 0 &
              abund_per_period['t1',asv] > 0.01 &
              abund_per_period['t5',asv] < 0.001 &
              abund_per_period['t5',asv] <= abund_per_period['t1',asv] # not sure about this one
              ) {
      class_table[asv,] = "temp adv"
      
   } else if (abund_per_period['pret',asv] > 0 &
              abund_per_period['t1',asv] > 5* abund_per_period['pret',asv] &
              5*abund_per_period['t5',asv] <= abund_per_period['t1',asv]
              ) {
      class_table[asv,] = "temp adv"
     
      # define as disappearing
   } else if (abund_per_period['pret',asv] > 0 & 
             abund_per_period['t1',asv] >= abund_per_period['t2',asv] &
             abund_per_period['t2',asv] >= abund_per_period['t3',asv] &
             abund_per_period['t3',asv] >= abund_per_period['t4',asv] &
             abund_per_period['t4',asv] >= abund_per_period['t5',asv] &
             abund_per_period['t5',asv] == 0){
    
    class_table[asv,] = "disappeared"     
     
   }  
     
    else {
    
     class_table[asv,] = "normal"
  }
  
}
  return(class_table)
}
```

Define function to plot the figures of each subcategory per model:

```{r}
get_plot_data <- function(list, cols){

  fig_title <- unique(list$Treatment)
  
  print(paste0("working on ", fig_title))
  # transform absolute into relative abundance
  rel_abund_df <- rel_abundance_transformation(list)
  
  # define category:
  data.frame(asv = colnames(rel_abund_df)) %>%
  mutate(cat = NA) %>%
  column_to_rownames('asv')-> asv_cat

  asv_cat <- asv_classification(rel_abund_df, asv_cat)
  
  # make df longer, such that every asv + period combination has its own entry
  rel_abund_df %>%
  # dplyr::filter(!is.na(Period)) %>% 
  rownames_to_column(var = "period") %>%
  pivot_longer(-period, names_to = "ASV", values_to = "rel_abundance") %>%
  filter(rel_abundance > 0) -> df_longer


  # combine df_longer with classification categories
  asv_cat %>%
  rownames_to_column(var = "ASV") %>%
  full_join(df_longer, by = "ASV") -> df
  
  # Generate plotting-ready data
  df %>%
  select(cat, period, rel_abundance) %>%
  group_by(period, cat) %>%
  # summarise(sum_abund = sum(rel_abundance)) %>%
  filter(!is.na(period)) -> plot_data

  # plot the data
  ggplot(plot_data, aes(fill = cat, y = rel_abundance, x = period)) +
    geom_bar(position = "stack", stat = "identity") +
    ggtitle(fig_title) +
    scale_fill_manual(name = "ASV category", values = cols,
                      na.value = "grey") +
    theme(strip.text.x = element_text(size=20),
        axis.title = element_text(size = 22),
        axis.text.x = element_text(size = 20),
        axis.text.y = element_text(size = 20),
        legend.position = "bottom",
        legend.text = element_text(size=20),
        legend.title = element_text(size=22),
        plot.title = element_text(size=26)) +
    ylab("Relative abundance") -> p1

  
  
  #print(p1)
  
  return(list(p1, df))
    
}
```


Define astethics:

```{r}
cat_col <- c(new='#3cb44b', prol="#6495ED", recovering='#FFB6C1', disappeared='#e6194B', `temp adv`='#F28C28', normal='#a9a9a9')

```

In a next step, the data is split into 6 tables: for each model (human and chicken) there are 3 tables for each antibioticum (CTX, untreated, VAN). Then, ASVs need to be classified and finally, the data is plotted:

chicken 1

```{r}
ps_tmp <- ps_rare

# #add Day_of_Treatment groups.
# ps_tmp %>%
#   sample_data() %>%
#   data.frame() %>%
# #   # mutate(Treatment = base::replace(Treatment, Experiment == "Cecum", "DONOR")) %>%
# #   # filter(Day_of_Treatment != -1) %>% # Excluded due to instability
#     dplyr::mutate(Period = ifelse(Day_of_Treatment < 0, "pret", Period))  -> sample_data(ps_tmp)
#   # filter(Model2 == "Chicken1") %>%
    # filter(Treatment != "DONOR") -> sample_data(ps_tmp)


# combine phyloseq tables and calculate relativ abundance across periods
ps_tmp %>%
  subset_samples(Model2 == "Chicken1") %>%
  subset_samples(Treatment != "DONOR") %>%
  psmelt() %>%
  select(OTU, Period, Abundance, Model, Treatment) %>%
  group_by(Model, Treatment, OTU, Period) %>%
  summarize(cum_abund = mean(Abundance)) %>%
  pivot_wider(names_from = OTU, values_from = cum_abund) -> ps_merged


# ps_merged$Treatment <- factor(ps_merged$Treatment, ordered = TRUE, levels = c("UNTREATED", "CTX+HV292.1", "CTX", "HV292.1", "VAN+CCUG59168", "VAN", "CCUG59168"))
# split lists
chicken_split <- split(ps_merged, ps_merged$Treatment)

# Get plot data
plot_data <- lapply(1:length(unique(ps_merged$Treatment)),
                    function(i) get_plot_data(chicken_split[[i]], cat_col))


plots <- lapply(plot_data, `[[`, 1)

# plots <- lapply(plot_data, `[[`, 2)

data_classification <- lapply(plot_data, `[[`, 2)
names(data_classification) <- names(chicken_split)

# plot the data
# ggpubr::ggarrange(plotlist = plots, ncol=7, nrow=1, common.legend = TRUE, legend="bottom") -> combined_plot

# combined_plot 
# 
# combined_plot %>% 
#     export::graph2ppt(append = TRUE, 
#                     width = 317.48031496 * 6, 
#                     height = 0.618 * 317.48031496 * 4 , paper = "A4",  scaling = 2,
#                     file = out_pptx)
```

export ASV level classification:

```{r}
data_classification %>% 
  bind_rows(.id = "reactor") %>%
   mutate("model2" = "Chicken1") -> classif_chicken1 
```

chicken 2

```{r}
ps_tmp <- ps_rare

# add Day_of_Treatment groups.
# ps_tmp %>% 
#   subset_samples(Model2 == "Chicken2") %>%
#   subset_samples(Treatment != "DONOR") %>%
#   # subset_samples(Treatment == "VAN+CCUG59168") %>%
#   sample_data() %>%
#   data.frame() %>% #select(Day_of_Treatment, Period)
#   # mutate(Treatment = base::replace(Treatment, Experiment == "Cecum", "DONOR")) %>%
#   # filter(Day_of_Treatment != -1) %>% # Excluded due to instability
#   # dplyr::mutate(Period = ifelse(Day_of_Treatment<= -1, "Pret", Period)) %>%
#   select(Treatment,Day_of_Treatment, Period) %>%  
#   group_by(Treatment, Period) %>% count(Period)#-> sample_data(ps_tmp)
 # mutate(case_when())

# combine phyloseq tables and calculate relativ abundance across periods
ps_tmp %>%
  subset_samples(Model2 == "Chicken2") %>%
  subset_samples(Treatment != "DONOR") %>% # subset_samples(Treatment == "UNTREATED") %>%  sample_data() %>%  data.frame()
  # subset_samples(Treatment != "UNTREATED") %>% 
  psmelt() %>%
  select(OTU, Period, Abundance, Model, Treatment) %>%
  group_by(Model, Treatment, OTU, Period) %>%
  summarize(cum_abund = mean(Abundance)) %>%
  pivot_wider(names_from = OTU, values_from = cum_abund) -> ps_merged
#filter(Treatment == "VAN+CCUG59168")

# ps_merged$Treatment <- factor(ps_merged$Treatment, ordered = TRUE, levels = c("UNTREATED", "CTX+HV292.1", "CTX", "HV292.1", "VAN+CCUG59168", "VAN", "CCUG59168"))
# split lists
chicken_split <- split(ps_merged, ps_merged$Treatment)


chicken_split$`VAN+CCUG59168` %>%
  # data.frame() %>% 
  mutate(Period = replace_na(Period,"pret")) -> chicken_split$`VAN+CCUG59168`


# Get plot data
plot_data <- lapply(1:length(unique(ps_merged$Treatment)),
                    function(i) get_plot_data(chicken_split[[i]], cat_col))


plots <- lapply(plot_data, `[[`, 1)

data_classification <- lapply(plot_data, `[[`, 2)
names(data_classification) <- names(chicken_split)

# plot the data
# ggpubr::ggarrange(plotlist = plots, ncol=7, nrow=1, common.legend = TRUE, legend="bottom")
```

export ASV level classification:

```{r}
data_classification %>% 
  bind_rows(.id = "reactor") %>%
   mutate("model2" = "Chicken2") -> classif_chicken2

```

```{r}
rbind(classif_chicken1,
      classif_chicken2) %>% 
  distinct(model2, reactor, ASV, .keep_all = TRUE) %>% 
  dplyr::filter(!is.na(period)) %>% 
  left_join(ps_tmp %>%
              tax_table() %>%
              data.frame() %>%
              rownames_to_column("ASV"),
            by = c("ASV" = "ASV")) %>% 
  dplyr::select(model2, reactor, ASV, cat, everything(), -Kingdom) %>% 
  dplyr::select(-period, -rel_abundance) %>%
xlsx::write.xlsx(file = "~/Desktop/16S-chick.xlsx",
                 append = TRUE,
                 sheetName = "ASV-classification",
                 showNA = TRUE)
```


```{r}
rbind(classif_chicken1,
      classif_chicken2) %>% 
  dplyr::filter(!is.na(period)) %>% 
  left_join(ps_tmp %>%
              tax_table() %>%
              data.frame() %>%
              rownames_to_column("ASV"),
            by = c("ASV" = "ASV")) %>% 
  mutate(reactor = fct_relevel(reactor,c("UNTREATED", "CTX", "CTX+HV292.1", "HV292.1", "VAN", "VAN+CCUG59168", "CCUG59168"))) %>%  # plot the data
  ggplot(., aes(fill = cat, y = rel_abundance, x = period)) +
    geom_bar(position = "stack", stat = "identity") +
    # ggtitle(fig_title) +
    scale_fill_manual(name = "ASV category", values = cat_col,
                      na.value = "grey") +
    # theme(strip.text.x = element_text(size=20),
    #     axis.title = element_text(size = 22),
    #     axis.text.x = element_text(size = 20),
    #     axis.text.y = element_text(size = 20),
    #     legend.position = "bottom",
    #     legend.text = element_text(size=20),
    #     legend.title = element_text(size=22),
    #     plot.title = element_text(size=26)) +
    ylab("Relative abundance") + 
  facet_grid( model2 ~ reactor) -> ASV_classification_plot

ASV_classification_plot
```

```{r}
ASV_classification_plot %>%
    export::graph2ppt(append = TRUE,
                    width = 317.48031496 * 1.25,
                    height = 0.618 * 317.48031496 * 1 , paper = "A4",  scaling = 2,
                    file = out_pptx)
```


```{r}
ASV_classification_plot$data %>% 
  xlsx::write.xlsx(file = "~/Desktop/16S-chick.xlsx",
                 append = TRUE,
                 sheetName = "ASV-classification2",
                 showNA = TRUE)
```


```{r}
ASV_classification_plot$data %>% 
   ggplot(., aes(fill = Phylum, y = rel_abundance, x = cat)) +
    geom_bar(position = "fill", stat = "identity") +
    # ggtitle(fig_title) +
    # scale_fill_manual(name = "ASV category", values = cat_col,
    #                   na.value = "grey") +
    # theme(strip.text.x = element_text(size=20),
    #     axis.title = element_text(size = 22),
    #     axis.text.x = element_text(size = 20),
    #     axis.text.y = element_text(size = 20),
    #     legend.position = "bottom",
    #     legend.text = element_text(size=20),
    #     legend.title = element_text(size=22),
    #     plot.title = element_text(size=26)) +
    ylab("Relative abundance") + 
  facet_grid( model2 ~ reactor) -> ASV_classification_plot_2
  
ASV_classification_plot_2

ASV_classification_plot_2 %>%   
      export::graph2ppt(append = TRUE,
                    width = 317.48031496 * 1.25,
                    height = 0.618 * 317.48031496 * 1 , paper = "A4",  scaling = 2,
                    file = out_pptx)
```

```{r}
ASV_classification_plot$data %>%
   ggplot(., aes(fill = Family, y = rel_abundance, x = cat)) +
    geom_bar(position = "fill", stat = "identity") +
    # ggtitle(fig_title) +
    scale_fill_manual(values = cat_col,
                      na.value = "grey") +
    # theme(strip.text.x = element_text(size=20),
    #     axis.title = element_text(size = 22),
    #     axis.text.x = element_text(size = 20),
    #     axis.text.y = element_text(size = 20),
    #     legend.position = "bottom",
    #     legend.text = element_text(size=20),
    #     legend.title = element_text(size=22),
    #     plot.title = element_text(size=26)) +
    ylab("Relative abundance") + 
  facet_grid( model2 ~ reactor) -> ASV_classification_plot_3
  
ASV_classification_plot_3

# ASV_classification_plot_2 %>%   
#       export::graph2ppt(append = TRUE,
#                     width = 317.48031496 * 1.25,
#                     height = 0.618 * 317.48031496 * 1 , paper = "A4",  scaling = 2,
#                     file = out_pptx)
```



other plot:

<https://github.com/haegih/MSc_Thesis-NRP72/blob/main/scripts/ASV_classification/human_taxanomy-with-classification.md>

```{r}
include_classification <- function(tax_df, asv_class_df){
  
  tax_df %>%
    left_join(asv_class_df %>% select(ASV, cat), by = "ASV") -> df
    
  return(df)
}


make_heatmap <- function(tax_df, asv_class_df, tax_level = Species){
  
  full_df <- include_classification(tax_df, asv_class_df)
  
  
  full_df %>% 
    unique %>%
    drop_na(Abundance) %>%
    ggplot(mapping = aes(x = Period,
                       y = {{tax_level}},
                       fill = Abundance,
  )) +
  facet_grid(cat ~ Treatment_Dose, scales = "free", space = "free", drop = TRUE,labeller = labeller(cat = label_wrap_gen(width = 15))) +
  #we adjust the vertical labels such that it is printed correctly. 
  geom_tile() +
  scale_fill_viridis_c(name = "Abundance",
                       na.value = "transparent", trans = scales::pseudo_log_trans(sigma = 0.1),
                       # trans =  scales::pseudo_log_trans(sigma = 1),
                       breaks = c(1, 50, 400, 10000), labels = c(1, 50, 400, 10000),
                       limits = c(0, 22500)) +
  theme_bw() -> p1
  
  return(p1)
}
```


```{r}
# combine phyloseq tables and calculate relativ abundance across periods
ps_rare %>%
  subset_samples(Model == "Chicken") %>% 
  psmelt() %>%
  select(OTU, Period, Abundance, Model2, Treatment_Dose, Kingdom:Strain) -> ps_melted



ps_melted %>%
  group_by(Model2, Treatment_Dose, OTU, Period) %>%
  summarize(cum_abund = sum(Abundance)) %>%
  pivot_wider(names_from = OTU, values_from = cum_abund) -> ps_merged_abund

ps_merged_abund %>%
  pivot_longer(cols = ASV0001:ASV0100, names_to = "ASV", values_to = "Abundance") %>%
  left_join(ps_melted %>% select(OTU, Kingdom:Strain), by = c("ASV" = "OTU")) %>%
  unique() %>%
  mutate(Abundance = na_if(Abundance, 0)) %>%
  mutate(Period = ordered(Period, levels = c("pret", factor(1:7))))-> ps_with_tax

ps_with_tax.split <- split(ps_with_tax, ps_with_tax$Treatment_Dose)


```

```{r}
abundance_lists <- list()
i <- 0
for (treat in names(ps_with_tax.split)){
  i = i + 1
  full_df <- include_classification(ps_with_tax.split[[treat]], asv_class[[treat]]) %>% filter(!is.na(cat))
  df <- full_df %>% select(Model, ASV, Treatment_Dose, cat, Kingdom:Strain) %>% group_by(Model, Treatment_Dose, cat) %>% count() %>% mutate(RelAbund = n/nrow(full_df)) %>% data.frame()
  abundance_lists[[i]] <- df
  
}

abundance_table <- bind_rows(abundance_lists) 

abundance_table %>% data.frame() %>% arrange(RelAbund) 
```



```{r}
tax_list <- list()

for (treat in names(ps_with_tax.split)){
  i = i + 1
  full_df <- include_classification(ps_with_tax.split[[treat]], asv_class[[treat]]) %>% filter(!is.na(cat))
  tax_list[[i]] <- full_df 
}

tax_cat_table <- bind_rows(tax_list) 



qual_col_pals = brewer.pal.info[brewer.pal.info$category == 'qual',]
col = unlist(mapply(brewer.pal, qual_col_pals$maxcolors, rownames(qual_col_pals)))
phyla_col <- col
names(phyla_col) <- tax_cat_table$Phylum %>% unique()


tax_cat_table %>%
  ggplot(aes(fill = Phylum, y = Abundance, x=cat)) +
  geom_bar(position="fill", stat="identity") +
  scale_color_manual(name = "Phylum", values = phyla_col,
                       na.value = "grey") +
  scale_fill_manual(name = "Phylum", values = phyla_col,
                      na.value = "grey") +
  theme(strip.text.x = element_text(size=18),
        axis.title = element_text(size = 18),
        axis.text.x = element_text(angle = 90,  size = 14, vjust = 0.5, hjust=1),
        axis.text.y = element_text(size = 18),
        legend.position = "bottom",
        legend.text = element_text(size=14),
        legend.title = element_text(size=14)) +
  facet_grid(~Treatment_Dose) +
  xlab("ASV category") + 
  ylab("Relative abundance")
```

human1 :


```{r}
# "data/processed/16S/16S_working_phyloseq.RDS" %>% 
#   here::here() %>% 
#   readRDS()  %>% 
#   subset_samples(Enrichment == "NotEnriched") %>%
#       subset_samples(Experiment %in% c("Continuous", "Cecum")) %>%
#    subset_samples(Day_of_Treatment > -6 & Day_of_Treatment <=30 | Reactor == "DONOR") %>% 
#   subset_samples(Model == "Human") %>% 
#    # subset_samples(Reactor != "CR" & Model2 == "Chicken1" | Model2 == "Chicken2" & Reactor %in% c("TR2", "TR3", "TR4", "TR5", "TR6", "TR7", "CR", "DONOR")) %>% 
#    subset_samples(Reactor != "IR") %>%
#   rarefy_even_depth(sample.size = 3738, rngseed = 123) -> ps_tmp
#   
# # add Day_of_Treatment groups.
# ps_tmp %>% 
#   sample_data() %>% 
#   data.frame() %>% 
#   # mutate(Treatment = base::replace(Treatment, Experiment == "Cecum", "DONOR")) %>%
#   filter(Day_of_Treatment != -1) %>% # Excluded due to instability
#   filter(Model2 == "Human1") %>%
#     filter(Treatment != "DONOR") %>% 
#   filter(!Treatment %in%  c("HV292.1", "CCUG59168")) -> sample_data(ps_tmp)
# 
# # combine phyloseq tables and calculate relativ abundance across periods
# ps_tmp %>%
#   psmelt() %>%
#   select(OTU, Period, Abundance, Model, Treatment) %>%
#   group_by(Model, Treatment, OTU, Period) %>%
#   summarize(cum_abund = mean(Abundance)) %>%
#   pivot_wider(names_from = OTU, values_from = cum_abund) -> ps_merged
# 
# 
# # ps_merged$Treatment <- factor(ps_merged$Treatment, ordered = TRUE, levels = c("UNTREATED", "CTX+HV292.1", "CTX", "HV292.1", "VAN+CCUG59168", "VAN", "CCUG59168"))
# # split lists
# chicken_split <- split(ps_merged, ps_merged$Treatment)
# 
# # Get plot data
# plot_data <- lapply(1:length(unique(ps_merged$Treatment)),
#                     function(i) get_plot_data(chicken_split[[i]], cat_col))
# 
# 
# plots <- lapply(plot_data, `[[`, 1)
# 
# data_classification <- lapply(plot_data, `[[`, 2)
# names(data_classification) <- names(chicken_split)
# 
# # plot the data
# ggpubr::ggarrange(plotlist = plots, ncol=7, nrow=1, common.legend = TRUE, legend="bottom")
```

human2 :


```{r}
# "data/processed/16S/16S_working_phyloseq.RDS" %>% 
#   here::here() %>% 
#   readRDS()  %>% 
#   subset_samples(Enrichment == "NotEnriched") %>%
#       subset_samples(Experiment %in% c("Continuous", "Cecum")) %>%
#    subset_samples(Day_of_Treatment > -6 & Day_of_Treatment <=30 | Reactor == "DONOR") %>% 
#   subset_samples(Model == "Human") %>% 
#    # subset_samples(Reactor != "CR" & Model2 == "Chicken1" | Model2 == "Chicken2" & Reactor %in% c("TR2", "TR3", "TR4", "TR5", "TR6", "TR7", "CR", "DONOR")) %>% 
#    subset_samples(Reactor != "IR") %>%
#   rarefy_even_depth(sample.size = 3738, rngseed = 123) -> ps_tmp
#   
# # add Day_of_Treatment groups.
# ps_tmp %>% 
#   sample_data() %>% 
#   data.frame() %>% 
#   # mutate(Treatment = base::replace(Treatment, Experiment == "Cecum", "DONOR")) %>%
#   filter(Day_of_Treatment != -1) %>% # Excluded due to instability
#   filter(Model2 == "Human2") %>%
#     filter(Treatment != "DONOR") %>% 
#   filter(!Treatment %in%  c("CTX+HV292.1")) -> sample_data(ps_tmp)
# 
# # combine phyloseq tables and calculate relativ abundance across periods
# ps_tmp %>%
#   psmelt() %>%
#   select(OTU, Period, Abundance, Model, Treatment) %>%
#   group_by(Model, Treatment, OTU, Period) %>%
#   summarize(cum_abund = mean(Abundance)) %>%
#   pivot_wider(names_from = OTU, values_from = cum_abund) -> ps_merged
# 
# 
# # ps_merged$Treatment <- factor(ps_merged$Treatment, ordered = TRUE, levels = c("UNTREATED", "CTX+HV292.1", "CTX", "HV292.1", "VAN+CCUG59168", "VAN", "CCUG59168"))
# # split lists
# chicken_split <- split(ps_merged, ps_merged$Treatment)
# 
# # Get plot data
# plot_data <- lapply(1:length(unique(ps_merged$Treatment)),
#                     function(i) get_plot_data(chicken_split[[i]], cat_col))
# 
# 
# plots <- lapply(plot_data, `[[`, 1)
# 
# data_classification <- lapply(plot_data, `[[`, 2)
# names(data_classification) <- names(chicken_split)
# 
# # plot the data
# ggpubr::ggarrange(plotlist = plots, ncol=7, nrow=1, common.legend = TRUE, legend="bottom")
```


## Taxa over time:

### Heatmap:

#### chicken1:

```{r}
ps_rare %>% 
  subset_samples(Model2 == "Chicken1" & Treatment != "DONOR" & Day_of_Treatment >= 1) %>% 
  physeq_most_abundant(group_var = "Reactor_Treatment_Dose",
                       ntax = 12,
                       tax_level = "Strain") -> top_taxa_per_group_h1
```

```{r}
ps_rare %>%
  subset_samples(Model2 == "Chicken1" & Treatment != "DONOR" & Day_of_Treatment >= -1) %>% 
  transform_sample_counts(function(x) x/sum(x) * 100) %>%
  subset_taxa(Strain %in% top_taxa_per_group_h1) %>% # extract only the taxa to display - after percentage normalisation
  phyloseq_ampvis_heatmap(physeq = ., tax_aggregate = "Species",
                          transform = FALSE,
                          plot_values = FALSE,
                          facet_by = c("Model2","Reactor_Treatment_Dose"),
                          group_by = "Day_of_Treatment",#"sample_name",
                          ntax = Inf
  ) -> treat_speces

treat_speces$data %>% 
  mutate(Abundance = na_if(Abundance, 0)) -> treat_speces$data 

treat_speces + 
  facet_grid(Model2 ~  Reactor_Treatment_Dose , scales = "free", space = "free") +
  scale_color_manual(values = c("#31a354", "orange", 
                                "#f03b20"), drop = FALSE, na.value = 'transparent') -> treat_speces
# scale_fill_viridis_c(breaks = c(0,  0.01, 1, 10, 25, 50, 75, 100),
# labels = c(0,  0.01, 1, 10, 25,  50, 75, 100),
# trans = scales::pseudo_log_trans(sigma = 0.9),
# na.value = 'transparent') -> pre_speces


treat_speces

# pre_speces %>%
#   export::graph2ppt(append = TRUE, width = 317.48031496 * 4  , height = 0.618 * 317.48031496 * 4 , paper = "A4",  scaling = 2,
#                     file = out_pptx)

```

```{r}
ps_rare %>%
  subset_samples(Model2 == "Chicken1" & Treatment != "DONOR" & Day_of_Treatment >= -1) %>% 
  transform_sample_counts(function(x) x/sum(x) * 100) %>%
  phyloseq_ampvis_heatmap(physeq = ., tax_aggregate = "Family",
                          transform = FALSE,
                          plot_values = FALSE,
                          facet_by = c("Model2","Reactor_Treatment_Dose"),
                          group_by = "Day_of_Treatment",#"sample_name",
                          ntax = 10
  ) -> treat_fam

treat_fam$data %>% 
  mutate(Abundance = na_if(Abundance, 0)) -> treat_fam$data 

treat_fam + 
  facet_grid(Model2 ~  Reactor_Treatment_Dose , scales = "free", space = "free") +
  scale_color_manual(values = c("#31a354", "orange", 
                                "#f03b20"), drop = FALSE, na.value = 'transparent') -> treat_fam
# scale_fill_viridis_c(breaks = c(0,  0.01, 1, 10, 25, 50, 75, 100),
# labels = c(0,  0.01, 1, 10, 25,  50, 75, 100),
# trans = scales::pseudo_log_trans(sigma = 0.9),
# na.value = 'transparent') -> pre_speces


treat_fam


# pre_speces %>%
#   export::graph2ppt(append = TRUE, width = 317.48031496 * 4  , height = 0.618 * 317.48031496 * 4 , paper = "A4",  scaling = 2,
#                     file = out_pptx)

```

```{r}



ggpubr::ggarrange(treat_fam + theme(legend.position = "null") + 
                    theme(axis.title.x=element_blank(),
                          axis.text.x=element_blank(),
                          axis.ticks.x=element_blank()),
                  treat_speces + theme(
                    strip.background = element_blank(),
                    strip.text.x = element_blank()
                  )  + theme(legend.position = "null"),
                  align = "v",
                  ncol = 1, 
                  heights = c(1, 2.60),
                  common.legend = FALSE) -> human1_heat

human1_heat

human1_heat %>% 
  export::graph2ppt(append = TRUE, 
                    width = 317.48031496 * 4.75, 
                    height = 0.618 * 317.48031496 * 6.5 , paper = "A4",  scaling = 2,
                    file = out_pptx)



treat_fam$data %>% 
  select(Display, Sample, Abundance,Reactor_Treatment_Dose) %>% 
  pivot_wider(names_from = Display, values_from = Abundance) %>% 
  xlsx::write.xlsx(file = out_xlsx,
                   append = TRUE,
                   sheetName = "treat_family",
                   showNA = TRUE)
treat_speces$data %>% 
  select(Display, Sample, Abundance,Reactor_Treatment_Dose) %>% 
  pivot_wider(names_from = Display, values_from = Abundance) %>% 
  xlsx::write.xlsx(file = out_xlsx,
                   append = TRUE,
                   sheetName = "treat_ASV",
                   showNA = TRUE)
```

#### Chicken2:

```{r}
ps_rare %>% 
  subset_samples(Model2 == "Chicken2" & Treatment != "DONOR" & Day_of_Treatment >= -1) %>% 
  physeq_most_abundant(group_var = "Reactor_Treatment_Dose",
                       ntax = 14,
                       tax_level = "Strain") -> top_taxa_per_group_h2
```

```{r}
ps_rare %>%
  subset_samples(Model2 == "Chicken2" & Treatment != "DONOR" & Day_of_Treatment >= -1) %>% 
  transform_sample_counts(function(x) x/sum(x) * 100) %>%
  subset_taxa(Strain %in% top_taxa_per_group_h2) %>% # extract only the taxa to display - after percentage normalisation
  phyloseq_ampvis_heatmap(physeq = ., tax_aggregate = "Species",
                          transform = FALSE,
                          plot_values = FALSE,
                          facet_by = c("Model2","Reactor_Treatment_Dose"),
                          group_by = "Day_of_Treatment",#"sample_name",
                          ntax = Inf
  ) -> treat_speces

treat_speces$data %>% 
  mutate(Abundance = na_if(Abundance, 0)) -> treat_speces$data 

treat_speces + 
  facet_grid(Model2 ~  Reactor_Treatment_Dose , scales = "free", space = "free") +
  scale_color_manual(values = c("#31a354", "orange", 
                                "#f03b20"), drop = FALSE, na.value = 'transparent') -> treat_speces
# scale_fill_viridis_c(breaks = c(0,  0.01, 1, 10, 25, 50, 75, 100),
# labels = c(0,  0.01, 1, 10, 25,  50, 75, 100),
# trans = scales::pseudo_log_trans(sigma = 0.9),
# na.value = 'transparent') -> pre_speces

treat_speces

# pre_speces %>%
#   export::graph2ppt(append = TRUE, width = 317.48031496 * 4  , height = 0.618 * 317.48031496 * 4 , paper = "A4",  scaling = 2,
#                     file = out_pptx)

```

```{r}
ps_rare %>%
  subset_samples(Model2 == "Chicken2" & Treatment != "DONOR" & Day_of_Treatment >= -1) %>% 
  transform_sample_counts(function(x) x/sum(x) * 100) %>%
  phyloseq_ampvis_heatmap(physeq = ., tax_aggregate = "Family",
                          transform = FALSE,
                          plot_values = FALSE,
                          facet_by = c("Model2","Reactor_Treatment_Dose"),
                          group_by = "Day_of_Treatment",#"sample_name",
                          ntax = 10
  ) -> treat_fam

treat_fam$data %>% 
  mutate(Abundance = na_if(Abundance, 0)) -> treat_fam$data 

treat_fam + 
  facet_grid(Model2 ~  Reactor_Treatment_Dose , scales = "free", space = "free") +
  scale_color_manual(values = c("#31a354", "orange", 
                                "#f03b20"), drop = FALSE, na.value = 'transparent') -> treat_fam
# scale_fill_viridis_c(breaks = c(0,  0.01, 1, 10, 25, 50, 75, 100),
# labels = c(0,  0.01, 1, 10, 25,  50, 75, 100),
# trans = scales::pseudo_log_trans(sigma = 0.9),
# na.value = 'transparent') -> pre_speces

treat_fam

# pre_speces %>%
#   export::graph2ppt(append = TRUE, width = 317.48031496 * 4  , height = 0.618 * 317.48031496 * 4 , paper = "A4",  scaling = 2,
#                     file = out_pptx)

```

```{r}
ggpubr::ggarrange(treat_fam + theme(legend.position = "null") + 
                    theme(axis.title.x=element_blank(),
                          axis.text.x=element_blank(),
                          axis.ticks.x=element_blank()),
                  treat_speces + theme(
                    strip.background = element_blank(),
                    strip.text.x = element_blank()
                  )  + theme(legend.position = "null"),
                  align = "v",
                  ncol = 1, 
                  heights = c(1, 2.60),
                  common.legend = FALSE) -> human2_heat


human2_heat

human2_heat %>% 
  export::graph2ppt(append = TRUE, 
                    width = 317.48031496 * 4.75, 
                    height = 0.618 * 317.48031496 * 5.5 , paper = "A4",  scaling = 2,
                    file = out_pptx)




treat_fam$data %>% 
  select(Display, Sample, Abundance,Reactor_Treatment_Dose) %>% 
  pivot_wider(names_from = Display, values_from = Abundance) %>% 
  xlsx::write.xlsx(file = out_xlsx,
                   append = TRUE,
                   sheetName = "treat_family_h2",
                   showNA = TRUE)

treat_speces$data %>% 
  select(Display, Sample, Abundance,Reactor_Treatment_Dose) %>% 
  pivot_wider(names_from = Display, values_from = Abundance) %>% 
  xlsx::write.xlsx(file = out_xlsx,
                   append = TRUE,
                   sheetName = "treat_ASV_h2",
                   showNA = TRUE)

```

Could be improved:

-clustering taxa based on correlation -\> response (pheatmap?) <https://btep.ccr.cancer.gov/docs/data-visualization-with-r/Lesson5_intro_to_ggplot/#make-this-plot-compatible-with-ggplot2-using-the-package-ggplotify>

<https://jcoliver.github.io/learn-r/008-ggplot-dendrograms-and-heatmaps.html>

-additional layer depicting the % of the community covered by the taxa displayed.

### barplot:

#### chicken1:

```{r}
ps_rare %>% 
  subset_samples(Model2 == "Chicken1" & Treatment != "DONOR" & Day_of_Treatment >= -1) %>% 
  microViz::ps_mutate(Reactor_Treatment_Dose = paste0(Reactor_Treatment_Dose, "_", Fermentation)) %>% 
  transform_sample_counts(function(x) x/sum(x) * 100) %>%
  subset_taxa(Strain %in% top_taxa_per_group_h1) %>% # extract only the taxa to display - after percentage normalisation
  microViz::phyloseq_validate() %>%
  microViz::tax_fix() %>% 
  microViz::ps_arrange(Day_of_Treatment) %>% 
  microViz::comp_barplot(
    tax_transform_for_plot = "identity",
    tax_level = "Strain",
    label = "Day_of_Treatment", # name an alternative variable to label axes
    n_taxa = length(top_taxa_per_group_h1), # give more taxa unique colours
    merge_other = FALSE, # split the "other" category to display alpha diversity
    bar_width = 0.9, # reduce the bar width to 70% of one row
    bar_outline_colour = "grey5",
    sample_order = "default") +
  ylab("Proportion - %") + xlab( " Day (Treatment) ") -> p_hist

p_hist + facet_grid(Phylum ~ Reactor_Treatment_Dose, scales = "free", space = "free_x") -> p_hist_human1_strain

p_hist_human1_strain  +  theme_light() + theme(legend.position = "none") -> p_hist_human1_strain_noleg

p_hist_human1_strain_noleg

p_hist_human1_strain %>%  
  ggpubr::get_legend() %>% 
  ggpubr::as_ggplot() -> p_hist_human1_strain_leg
```

```{r}
ps_rare %>% 
  subset_samples(Model2 == "Chicken1" & Treatment != "DONOR" & Day_of_Treatment >= -1) %>% 
  microViz::ps_mutate(Reactor_Treatment_Dose = paste0(Reactor_Treatment_Dose, "_", Fermentation)) %>% 
  transform_sample_counts(function(x) x/sum(x) * 100) %>%
  microViz::phyloseq_validate() %>%
  microViz::tax_fix() %>% 
  microViz::ps_arrange(Day_of_Treatment) %>% 
  microViz::comp_barplot(
    tax_transform_for_plot = "identity",
    tax_level = "Family",
    label = "Day_of_Treatment", # name an alternative variable to label axes
    n_taxa = 10, # give more taxa unique colours
    palette = microViz::distinct_palette(10, pal = "kelly"),
    merge_other = FALSE, # split the "other" category to display alpha diversity
    bar_width = 0.9, # reduce the bar width to 70% of one row
    bar_outline_colour = "grey5",
    sample_order = "default") +
  ylab("Proportion - %") + xlab( " Day (Treatment) ") -> p_hist

p_hist + facet_grid(Kingdom ~ Reactor_Treatment_Dose, scales = "free", space = "free_x", drop = TRUE) -> p_hist_human1_fam

p_hist_human1_fam  +  theme_light() + theme(legend.position = "none") -> p_hist_human1_fam_no_leg

p_hist_human1_fam_no_leg

p_hist_human1_fam %>%  
  ggpubr::get_legend() %>% 
  ggpubr::as_ggplot() -> p_hist_human1_fam_leg
```

#### Chicken2:

```{r}
ps_rare %>% 
  subset_samples(Model2 == "Chicken2" & Treatment != "DONOR" & Day_of_Treatment >= -1) %>% 
  microViz::ps_mutate(Reactor_Treatment_Dose = paste0(Reactor_Treatment_Dose, "_", Fermentation)) %>% 
  transform_sample_counts(function(x) x/sum(x) * 100) %>%
  subset_taxa(Strain %in% top_taxa_per_group_h2) %>% # extract only the taxa to display - after percentage normalisation
  microViz::phyloseq_validate() %>%
  microViz::tax_fix() %>% 
  microViz::ps_arrange(Day_of_Treatment) %>% 
  microViz::comp_barplot(
    tax_transform_for_plot = "identity",
    tax_level = "Strain",
    label = "Day_of_Treatment", # name an alternative variable to label axes
    n_taxa = length(top_taxa_per_group_h2), # give more taxa unique colours
    merge_other = FALSE, # split the "other" category to display alpha diversity
    bar_width = 0.9, # reduce the bar width to 70% of one row
    bar_outline_colour = "grey5",
    sample_order = "default") +
  ylab("Proportion - %") + xlab( " Day (Treatment) ") -> p_hist

p_hist + facet_grid(Phylum ~ Reactor_Treatment_Dose, scales = "free", space = "free_x") -> p_hist_human2_strain

p_hist_human2_strain +  theme_light()  + theme(legend.position = "none") -> p_hist_human2_strain_noleg

p_hist_human2_strain_noleg



p_hist_human2_strain %>%  
  ggpubr::get_legend() %>% 
  ggpubr::as_ggplot() -> p_hist_human2_strain_leg
```

```{r}
ps_rare %>% 
  subset_samples(Model2 == "Chicken2" & Treatment != "DONOR" & Day_of_Treatment >= -1) %>% 
  microViz::ps_mutate(Reactor_Treatment_Dose = paste0(Reactor_Treatment_Dose, "_", Fermentation)) %>% 
  transform_sample_counts(function(x) x/sum(x) * 100) %>%
  # subset_taxa(Strain %in% top_taxa_per_group_h2) %>% # extract only the taxa to display - after percentage normalisation
  microViz::phyloseq_validate() %>%
  microViz::tax_fix() %>% 
  microViz::ps_arrange(Day_of_Treatment) %>% 
  microViz::comp_barplot(
    tax_transform_for_plot = "identity",
    tax_level = "Family",
    label = "Day_of_Treatment", # name an alternative variable to label axes
    palette = microViz::distinct_palette(10, pal = "kelly"),
    n_taxa = 10, # give more taxa unique colours
    merge_other = FALSE, # split the "other" category to display alpha diversity
    bar_width = 0.9, # reduce the bar width to 70% of one row
    bar_outline_colour = "grey5",
    sample_order = "default") +
  ylab("Proportion - %") + xlab( " Day (Treatment) ") -> p_hist

p_hist + facet_grid(Kingdom ~ Reactor_Treatment_Dose, scales = "free", space = "free_x", drop = TRUE) -> p_hist_human2_fam

p_hist_human2_fam +  theme_light()  + theme(legend.position = "none") -> p_hist_human2_fam_no_leg

p_hist_human2_fam_no_leg

p_hist_human2_fam %>%  
  ggpubr::get_legend() %>% 
  ggpubr::as_ggplot() -> p_hist_human2_fam_leg


# g <- ggplot_build(p_hist_human2_fam)
# 
# data.frame(colours = unique(g$data[[1]]["fill"]), 
#               label = levels(g$plot$data[, g$plot$labels$unique]))
```

Unify colors:

```{r}
c(p_hist_human2_fam$data$Family, p_hist_human1_fam$data$Family) %>% 
  unique()  -> families
```

```{r}
# families %>%
#   ggpubr::get_palette(k = ., palette = "kelly") -> col

families[!families %in% c("other")] %>% 
  length() %>% 
  microViz::distinct_palette(n = ., pal = "kelly", add = NA) -> col_fam

names(col_fam) <- families[!families %in% c("other")]

c(col_fam, "lightgrey") -> col_fam

names(col_fam) = NULL

names(col_fam) <- c(families[!families %in% c("other")], families[families %in% c("other")])

scales::show_col(col_fam)
```

```{r}
p_hist_human1_fam  +  theme_light() + scale_fill_manual(values = col_fam) ->  p_hist_human1_fam

p_hist_human1_fam + theme(legend.position = "none") -> p_hist_human1_fam_no_leg

p_hist_human1_fam_no_leg
```

```{r}
p_hist_human1_fam %>%  
  ggpubr::get_legend() %>% 
  ggpubr::as_ggplot() -> p_hist_human1_fam_leg

p_hist_human1_fam_leg
```

```{r}
c(p_hist_human2_strain$data$Strain, p_hist_human1_strain$data$Strain) %>% 
  unique() -> ASVs

# ASVs
```

```{r}
# families %>%
#   ggpubr::get_palette(k = ., palette = "kelly") -> col
set.seed(1234)

ASVs[!ASVs %in% c("other")] %>% 
  length() %>% 
  randomcoloR::distinctColorPalette(k = ., altCol = FALSE, runTsne = TRUE) -> col_ASVs

names(col_ASVs) <- ASVs[!ASVs %in% c("other")]

c(col_ASVs, "lightgrey") -> col_ASVs

names(col_ASVs) = NULL

names(col_ASVs) <- c(ASVs[!ASVs %in% c("other")], ASVs[ASVs %in% c("other")])

scales::show_col(col_ASVs)
```

```{r}
p_hist_human1_strain + scale_fill_manual(values = col_ASVs) ->  p_hist_human1_strain

p_hist_human1_strain + theme_light() +
  theme(legend.position = "none") -> p_hist_human1_strain_no_leg

p_hist_human1_strain_no_leg
```

```{r}
p_hist_human1_strain %>%  
  ggpubr::get_legend() %>% 
  ggpubr::as_ggplot() -> p_hist_human1_strain_leg

p_hist_human1_strain_leg
```

```{r}
p_hist_human2_strain + scale_fill_manual(values = col_ASVs) ->  p_hist_human2_strain

p_hist_human2_strain + theme_light() +theme(legend.position = "none") ->p_hist_human2_strain_no_leg

p_hist_human2_strain_no_leg
```

```{r}
p_hist_human2_strain %>%  
  ggpubr::get_legend() %>% 
  ggpubr::as_ggplot() -> p_hist_human2_strain_leg

p_hist_human2_strain_leg
```

```{r}
ggpubr::ggarrange(p_hist_human1_fam_no_leg,
                  p_hist_human1_strain_no_leg ,
                  align = "v",
                  ncol = 1, 
                  heights = c(1, 1),
                  common.legend = FALSE) -> p_hist_human1_f1

p_hist_human1_f1

p_hist_human1_f1 %>% 
  export::graph2ppt(append = TRUE, 
                    width = 317.48031496 * 4.75, 
                    height = 0.618 * 317.48031496 * 6.5 , paper = "A4",  scaling = 2,
                    file = out_pptx)

p_hist_human1_fam_leg %>% 
  export::graph2ppt(append = TRUE, 
                    width = 317.48031496 * 1, 
                    height = 0.618 * 317.48031496 * 1 , paper = "A4",  scaling = 2,
                    file = out_pptx)

p_hist_human1_strain_leg %>% 
  export::graph2ppt(append = TRUE, 
                    width = 317.48031496 * 5, 
                    height = 0.618 * 317.48031496 * 2.5 , paper = "A3",  scaling = 2,
                    file = out_pptx)
```

```{r}
ggpubr::ggarrange(p_hist_human2_fam_no_leg,
                  p_hist_human2_strain_noleg,
                  align = "v",
                  ncol = 1, 
                  heights = c(1, 1),
                  common.legend = FALSE) -> p_hist_human2

p_hist_human2

p_hist_human2 %>% 
  export::graph2ppt(append = TRUE, 
                    width = 317.48031496 * 4.75, 
                    height = 0.618 * 317.48031496 * 6.5 , paper = "A4",  scaling = 2,
                    file = out_pptx)

p_hist_human2_fam_leg %>% 
  export::graph2ppt(append = TRUE, 
                    width = 317.48031496 * 1, 
                    height = 0.618 * 317.48031496 * 1 , paper = "A4",  scaling = 2,
                    file = out_pptx)



p_hist_human2_strain_leg %>% 
  export::graph2ppt(append = TRUE, 
                    width = 317.48031496 * 6, 
                    height = 0.618 * 317.48031496 * 2.5 , paper = "A3",  scaling = 2,
                    file = out_pptx)
```

### Alluvial:

```{r}
# p_hist_human2_fam$data %>%
#   filter(Model2 == "Human2") %>%
#   # microViz::ps_mutate(Reactor_Treatment_Dose = paste0(Reactor_Treatment_Dose, "_", Fermentation)) %>% 
#   # mutate(combo = paste0(Model2, "_" , Reactor_F1_F2)) %>%
#   # mutate(combo = as.factor(combo)) %>%
#   select(SAMPLE, top,OTU, Abundance, Day_of_Treatment, Reactor_Treatment_Dose) %>%  distinct(top)
```

```{r}
# ggpubr::ggarrange(p_hist_human2_fam_no_leg,
#                   p_hist_human2_strain_noleg ,
#                   align = "v",
#                   ncol = 1, 
#                   heights = c(1, 1),
#                   common.legend = FALSE) -> p_hist_human2
# 
# p_hist_human2
# 
# p_hist_human2 %>% 
#   export::graph2ppt(append = TRUE, 
#                     width = 317.48031496 * 4.75, 
#                     height = 0.618 * 317.48031496 * 6.5 , paper = "A4",  scaling = 2,
#                     file = out_pptx)
```

#### human1:

```{r}
require(ggalluvial)
```

##### Family

```{r}
p_hist_human1_fam$data %>%
  # filter(Model2 == "Human2") %>%
  # microViz::ps_mutate(Reactor_Treatment_Dose = paste0(Reactor_Treatment_Dose, "_", Fermentation)) %>% 
  # mutate(combo = paste0(Model2, "_" , Reactor_F1_F2)) %>%
  # mutate(combo = as.factor(combo)) %>%
  select(SAMPLE, Family,OTU, Abundance, Day_of_Treatment, Reactor_Treatment_Dose) %>% 
  ggplot(.,
         aes(x = Day_of_Treatment, 
             y = Abundance,
             stratum = Family, 
             alluvium = OTU,
             fill = Family,
             label = Family))+
  facet_grid(Reactor_Treatment_Dose  ~ .)+
  # scale_fill_manual(values = c("red", "yellow", "green3"))+
  geom_flow(stat = "alluvium", lode.guidance = "frontback", color = "darkgray")+
  geom_stratum() +
  theme_light() +
  # geom_text(stat = "alluvium", aes(label = top), lode.guidance = "frontback") +
  scale_fill_manual(values = col_fam) +
  # scale_color_manual(values  = microViz::distinct_palette(10, pal = "kelly")) +
  # scale_fill_manual(values  = microViz::distinct_palette(10, pal = "kelly")) +
  ylab("Proportion") -> p_allu_human1

p_allu_human1 + theme(legend.position = "none") -> p_allu_human1_noleg 

p_allu_human1_noleg

p_allu_human1 %>% 
  ggpubr::get_legend() %>% 
  ggpubr::as_ggplot() ->  p_allu_human1_leg
```

```{r}
p_allu_human1_noleg %>% 
  export::graph2ppt(append = TRUE, 
                    width = 317.48031496 * 1, 
                    height = 0.618 * 317.48031496 * 2 , paper = "A4",  scaling = 2,
                    file = out_pptx)

p_allu_human1_leg %>% 
  export::graph2ppt(append = TRUE, 
                    width = 317.48031496 * 1, 
                    height = 0.618 * 317.48031496 * 2 , paper = "A4",  scaling = 2,
                    file = out_pptx)
```

##### ASV:

```{r}
p_hist_human1_strain$data %>%
  # filter(Model2 == "Human2") %>%
  # microViz::ps_mutate(Reactor_Treatment_Dose = paste0(Reactor_Treatment_Dose, "_", Fermentation)) %>% 
  # mutate(combo = paste0(Model2, "_" , Reactor_F1_F2)) %>%
  # mutate(combo = as.factor(combo)) %>%
  select(SAMPLE, Strain,OTU, Abundance, Day_of_Treatment, Reactor_Treatment_Dose) %>% 
  ggplot(.,
         aes(x = Day_of_Treatment, 
             y = Abundance,
             stratum = Strain, 
             alluvium = OTU,
             fill = Strain,
             label = Strain))+
  facet_grid(Reactor_Treatment_Dose  ~ .)+
  # scale_fill_manual(values = c("red", "yellow", "green3"))+
  geom_flow(stat = "alluvium", lode.guidance = "frontback", color = "darkgray")+
  geom_stratum() +
  theme_light() +
  # geom_text(stat = "alluvium", aes(label = top), lode.guidance = "frontback") +
  scale_fill_manual(values = col_ASVs) +
  # scale_color_manual(values  = microViz::distinct_palette(10, pal = "kelly")) +
  # scale_fill_manual(values  = microViz::distinct_palette(10, pal = "kelly")) +
  ylab("Proportion") -> p_allu_human1_ASV

p_allu_human1_ASV + theme(legend.position = "none") -> p_allu_human1_ASV_noleg 

p_allu_human1_ASV_noleg

p_allu_human1_ASV %>% 
  ggpubr::get_legend() %>% 
  ggpubr::as_ggplot() ->  p_allu_human1_ASV_leg
```

```{r}
p_allu_human1_ASV_noleg %>% 
  export::graph2ppt(append = TRUE, 
                    width = 317.48031496 * 1, 
                    height = 0.618 * 317.48031496 * 2 , paper = "A4",  scaling = 2,
                    file = out_pptx)
p_allu_human1_ASV_leg %>% 
  export::graph2ppt(append = TRUE, 
                    width = 317.48031496 * 4, 
                    height = 0.618 * 317.48031496 * 2 , paper = "A3",  scaling = 2,
                    file = out_pptx)
```

#### human2

##### Family

```{r}

p_hist_human2_fam$data %>%
  # filter(Model2 == "Human2") %>%
  # microViz::ps_mutate(Reactor_Treatment_Dose = paste0(Reactor_Treatment_Dose, "_", Fermentation)) %>% 
  # mutate(combo = paste0(Model2, "_" , Reactor_F1_F2)) %>%
  # mutate(combo = as.factor(combo)) %>%
  select(SAMPLE, Family,OTU, Abundance, Day_of_Treatment, Reactor_Treatment_Dose) %>% 
  ggplot(.,
         aes(x = Day_of_Treatment, 
             y = Abundance,
             stratum = Family, 
             alluvium = OTU,
             fill = Family,
             label = Family))+
  facet_grid(Reactor_Treatment_Dose  ~ .)+
  # scale_fill_manual(values = c("red", "yellow", "green3"))+
  geom_flow(stat = "alluvium", lode.guidance = "frontback", color = "darkgray")+
  geom_stratum() +
  theme_light() +
  # geom_text(stat = "alluvium", aes(label = top), lode.guidance = "frontback") +
  scale_fill_manual(values = col_fam) +
  # scale_color_manual(values  = microViz::distinct_palette(10, pal = "kelly")) +
  # scale_fill_manual(values  = microViz::distinct_palette(10, pal = "kelly")) +
  ylab("Proportion") -> p_allu_human2

p_allu_human2 + theme(legend.position = "none") -> p_allu_human2_noleg 

p_allu_human2_noleg

```

```{r}
p_allu_human2 %>% 
  ggpubr::get_legend() %>% 
  ggpubr::as_ggplot() ->  p_allu_human2_leg

p_allu_human2_leg
```

```{r}
p_allu_human2_noleg %>% 
  export::graph2ppt(append = TRUE, 
                    width = 317.48031496 * 1, 
                    height = 0.618 * 317.48031496 * 2 , paper = "A4",  scaling = 2,
                    file = out_pptx)

p_allu_human2_leg %>% 
  export::graph2ppt(append = TRUE, 
                    width = 317.48031496 * 1, 
                    height = 0.618 * 317.48031496 * 2 , paper = "A4",  scaling = 2,
                    file = out_pptx)
```

##### ASV:

```{r}
p_hist_human2_strain$data %>%
  # filter(Model2 == "Human2") %>%
  # microViz::ps_mutate(Reactor_Treatment_Dose = paste0(Reactor_Treatment_Dose, "_", Fermentation)) %>% 
  # mutate(combo = paste0(Model2, "_" , Reactor_F1_F2)) %>%
  # mutate(combo = as.factor(combo)) %>%
  select(SAMPLE, Strain,OTU, Abundance, Day_of_Treatment, Reactor_Treatment_Dose) %>% 
  ggplot(.,
         aes(x = Day_of_Treatment, 
             y = Abundance,
             stratum = Strain, 
             alluvium = OTU,
             fill = Strain,
             label = Strain))+
  facet_grid(Reactor_Treatment_Dose  ~ .)+
  # scale_fill_manual(values = c("red", "yellow", "green3"))+
  geom_flow(stat = "alluvium", lode.guidance = "frontback", color = "darkgray")+
  geom_stratum() +
  theme_light() +
  # geom_text(stat = "alluvium", aes(label = top), lode.guidance = "frontback") +
  scale_fill_manual(values = col_ASVs) +
  # scale_color_manual(values  = microViz::distinct_palette(10, pal = "kelly")) +
  # scale_fill_manual(values  = microViz::distinct_palette(10, pal = "kelly")) +
  ylab("Proportion") -> p_allu_human2_ASV

p_allu_human2_ASV + theme(legend.position = "none") -> p_allu_human2_ASV_noleg 

p_allu_human2_ASV_noleg

p_allu_human2_ASV %>% 
  ggpubr::get_legend() %>% 
  ggpubr::as_ggplot() ->  p_allu_human2_ASV_leg
```

```{r}
p_allu_human2_ASV_noleg %>% 
  export::graph2ppt(append = TRUE, 
                    width = 317.48031496 * 1, 
                    height = 0.618 * 317.48031496 * 2 , paper = "A4",  scaling = 2,
                    file = out_pptx)

p_allu_human2_ASV_leg %>% 
  export::graph2ppt(append = TRUE, 
                    width = 317.48031496 * 4, 
                    height = 0.618 * 317.48031496 * 2 , paper = "A3",  scaling = 2,
                    file = out_pptx)
```

## gram plots:

histogram of gram+ gram- & intrinsicly resistant or not an annotation color in a heatmap would be cool as well.

```{r}
"data/processed/16S/16S_based_gram.RDS" %>% 
  here::here() %>% 
  readRDS() -> gram

gram %>% 
  summarise_all(~ sum(is.na(.)))
```

```{r}
dim(gram)

intersect(gram$ASV,
          tax_table(ps_filtered)[,"Strain"] %>%  as.character() ) %>%  length() 
```

Let's take the Class info here.

## intrinsic:

```{r}
"data/processed/16S/16S_based_intrinsinc_AMR.RDS" %>% 
  here::here() %>% 
  readRDS() -> intrinsic_AMR
```

Wich level to chose? the one with less NA:

```{r}
intrinsic_AMR %>% 
  select(starts_with("intrinsic")) %>% 
  summary()
```

Let's take the Strain info here.

```{r}
ps_rare_gram_int <- ps_rare

# We want sample, reactor..., day_0f_treatment 
ps_rare_gram_int %>% 
  tax_table() %>% 
  data.frame() %>% 
  rownames_to_column("ASVid") %>% 
  left_join(gram %>% dplyr::rename("Strain" = "ASV") %>%  
              select(Strain , gram_neg_class)
  ) %>%  #,
  # by = c("Strain" = "ASV"),
  # suffix = c("", ".y")) %>% 
  select(ASVid:Strain, gram_neg_class) %>% 
  dplyr::rename('gram_stain' = 'gram_neg_class') %>% 
  mutate(gram_stain = replace_na(gram_stain, "unknown-gram")) %>% 
  left_join(., intrinsic_AMR,
            suffix = c("", ".y"),
            by = c("Strain" = "ASV")) %>%
  mutate(intrinsic = ifelse(intrinsic_strain_van == TRUE & intrinsic_strain_ctx == TRUE, "intrinsic-VAN-CTX",  ifelse(intrinsic_strain_van == TRUE, "intrinsic-VAN", ifelse(intrinsic_strain_ctx == TRUE, "intrinsic-CTX", "No-intrinsic")))) %>% 
  select(ASVid:gram_stain,intrinsic) %>%
  mutate(gram_intrinsic = paste0(gram_stain, 
                                 ifelse(is.na(intrinsic), "", paste0("_",intrinsic)))) %>% 
  # dplyr::rename('intrinsic_CTX' = 'intrinsic_strain_ctx',
  # 'intrinsic_VAN' = 'intrinsic_strain_van') %>% 
  column_to_rownames('ASVid') %>% 
  as.matrix() -> tax_table(ps_rare_gram_int)
```

### chicken2:

#### gram:

```{r}
ps_rare_gram_int %>% 
  subset_samples(Model2 == "Chicken2" & Treatment != "DONOR" & Day_of_Treatment >= -1) %>% 
  microViz::ps_mutate(Reactor_Treatment_Dose = paste0(Reactor_Treatment_Dose, "_", Fermentation)) %>% 
  transform_sample_counts(function(x) x/sum(x) * 100) %>%
  physeq_sel_tax_table(c("gram_stain")) %>% 
  # subset_taxa(Strain %in% top_taxa_per_group_h2) %>% # extract only the taxa to display - after percentage normalisation
  microViz::phyloseq_validate() %>%
  # microViz::tax_fix() %>%
  microViz::ps_arrange(Day_of_Treatment) %>% 
  # microViz::tax_mutate(gram = left_join(tax_table(.) %>% data.frame(), gram,
  # by = c("Strain" = "ASV") %>% select(gram_neg_class))) %>% 
  microViz::comp_barplot(
    tax_transform_for_plot = "identity",
    tax_level = "gram_stain",
    label = "Day_of_Treatment", # name an alternative variable to label axes
    palette = microViz::distinct_palette(3, pal = "greenArmytage"),
    n_taxa = 10, # give more taxa unique colours
    merge_other = FALSE, # split the "other" category to display alpha diversity
    bar_width = 0.9, # reduce the bar width to 70% of one row
    bar_outline_colour = "grey5",
    sample_order = "default") +
  ylab("Proportion - %") + xlab( " Day (Treatment) ") -> p_gram_h2

p_gram_h2 + theme_light() + facet_grid(. ~ Reactor_Treatment_Dose, scales = "free", space = "free_x", drop = TRUE) -> p_gram_h2

p_gram_h2 + theme(legend.position = "none") -> p_gram_h2_no_leg

p_gram_h2_no_leg

p_gram_h2 %>%  
  ggpubr::get_legend() %>% 
  ggpubr::as_ggplot() -> p_gram_h2_leg

p_gram_h2_leg
```

```{r}
p_gram_h2$data %>%
  # filter(Model2 == "Human2") %>%
  mutate(gram_stain = factor(gram_stain, levels = c("Gram-positive", "Gram-negative", "unknown-gram"))) %>%
  # mutate(combo = paste0(Model2, "_" , Reactor_F1_F2)) %>%
  # mutate(combo = as.factor(combo)) %>%
  select(SAMPLE, gram_stain,OTU, Abundance, Day_of_Treatment, Reactor_Treatment_Dose) %>% 
  ggplot(.,
         aes(x = Day_of_Treatment, 
             y = Abundance,
             stratum = gram_stain, 
             alluvium = OTU,
             fill = gram_stain,
             label = gram_stain))+
  facet_grid(Reactor_Treatment_Dose  ~ .)+
  # scale_fill_manual(values = c("red", "yellow", "green3"))+
  geom_flow(stat = "alluvium", lode.guidance = "frontback", color = "darkgray")+
  geom_stratum() +
  theme_light() +
  # geom_text(stat = "alluvium", aes(label = top), lode.guidance = "frontback") +
  # scale_fill_manual(values = col_ASVs) +
  scale_color_manual(values  =   microViz::distinct_palette(3, pal = "greenArmytage")) +
  scale_fill_manual(values  =   microViz::distinct_palette(3, pal = "greenArmytage")) +
  # scale_color_manual(values  = microViz::distinct_palette(10, pal = "kelly")) +
  # scale_fill_manual(values  = microViz::distinct_palette(10, pal = "kelly")) +
  ylab("Proportion - %") -> p_allu_human2_gram

p_allu_human2_gram + theme(legend.position = "none") -> p_allu_human2_gram_noleg 

p_allu_human2_gram_noleg 

p_allu_human2_gram %>% 
  ggpubr::get_legend() %>% 
  ggpubr::as_ggplot() ->  p_allu_human2_gram_leg

p_allu_human2_gram_leg
```

#### intrinsic:

```{r}
# ps_rare_gram_int %>% 
#   subset_samples(Model2 == "Chicken1" & Treatment != "DONOR" & Day_of_Treatment >= -1) %>% 
#   microViz::ps_mutate(Reactor_Treatment_Dose = paste0(Reactor_Treatment_Dose, "_", Fermentation)) %>% 
#   transform_sample_counts(function(x) x/sum(x) * 100) %>%
#   physeq_sel_tax_table(c("intrinsic")) %>% 
#   # microbiome::transform("log") %>% 
#   # speedyseq::tax_glom("intrinsic") %>% 
#   # physeq_sel_tax_table(c("gram_stain","intrinsic")) %>% 
#   # physeq_glom_rename(phyloseq = ., speedyseq = TRUE, rename_ASV = FALSE, taxrank = "gram_stain") %>% 
#   # subset_taxa(Strain %in% top_taxa_per_group_h2) %>% # extract only the taxa to display - after percentage normalisation
#   microViz::phyloseq_validate() %>%
#   microViz::tax_fix() %>%
#   microViz::ps_arrange(Day_of_Treatment) %>% 
#   # microViz::tax_mutate(gram = left_join(tax_table(.) %>% data.frame(), gram,
#   # by = c("Strain" = "ASV") %>% select(gram_neg_class))) %>% 
#   microViz::comp_barplot(
#     tax_transform_for_plot = "identity",
#     tax_level = "intrinsic",
#     label = "Day_of_Treatment", # name an alternative variable to label axes
#     # palette = microViz::distinct_palette(3, pal = "brewerPlus"),
#     n_taxa = 10, # give more taxa unique colours
#     merge_other = FALSE, # split the "other" category to display alpha diversity
#     bar_width = 0.9, # reduce the bar width to 70% of one row
#     bar_outline_colour = "grey5",
#     sample_order = "default") +
#   ylab("Proportion - %") + xlab( " Day (Treatment) ") -> p_int_h2
# 
# p_int_h2 + theme_light() +facet_grid(. ~ Reactor_Treatment_Dose, scales = "free", space = "free_x", drop = TRUE) -> p_int_h2
# 
# p_int_h2 %>% 
#   ggpubr::set_palette(p = ., palette = "jco") -> p_int_h2
# 
# p_int_h2 + theme(legend.position = "none") -> p_int_h2_no_leg
# 
# p_int_h2_no_leg
# 
# p_int_h2 %>%  
#   ggpubr::get_legend() %>% 
#   ggpubr::as_ggplot() -> p_int_h2_leg
# 
# p_int_h2_leg
```

```{r}
# p_int_h2$data %>%
#   # filter(Model2 == "Human2") %>%
#   # microViz::ps_mutate(Reactor_Treatment_Dose = paste0(Reactor_Treatment_Dose, "_", Fermentation)) %>% 
#   # mutate(combo = paste0(Model2, "_" , Reactor_F1_F2)) %>%
#   # mutate(combo = as.factor(combo)) %>%
#   select(SAMPLE, top,OTU, Abundance, Day_of_Treatment, Reactor_Treatment_Dose) %>% 
#   ggplot(.,
#          aes(x = Day_of_Treatment, 
#              y = Abundance,
#              stratum = top, 
#              alluvium = OTU,
#              fill = top,
#              label = top))+
#   facet_grid(Reactor_Treatment_Dose  ~ .)+
#   # scale_fill_manual(values = c("red", "yellow", "green3"))+
#   geom_flow(stat = "alluvium", lode.guidance = "frontback", color = "darkgray")+
#   geom_stratum() +
#   theme_light() +
#   # geom_text(stat = "alluvium", aes(label = top), lode.guidance = "frontback") +
#   # scale_fill_manual(values = col_ASVs) +
#   scale_color_manual(values  =  microViz::distinct_palette(3, pal = "greenArmytage")) +
#   # scale_color_manual(values  = microViz::distinct_palette(10, pal = "kelly")) +
#   # scale_fill_manual(values  = microViz::distinct_palette(10, pal = "kelly")) +
#   ylab("Proportion - %") -> p_allu_human2_int
# 
# p_allu_human2_int %>% 
#   ggpubr::set_palette(p = ., palette = "jco") -> p_allu_human2_int
# 
# p_allu_human2_int + theme(legend.position = "none") -> p_allu_human2_int_noleg 
# 
# p_allu_human2_int_noleg 
# 
# p_allu_human2_int %>% 
#   ggpubr::get_legend() %>% 
#   ggpubr::as_ggplot() ->  p_allu_human2_int_leg
# ```
# ```{r}
# ggpubr::ggarrange(p_allu_human2_noleg + xlab(NULL),
#                   p_allu_human2_ASV_noleg + ylab(NULL) + xlab(NULL),
#                   p_allu_human2_int_noleg +  coord_cartesian(ylim = c(50,100)),
#                   p_allu_human2_int_noleg+ ylab(NULL),
#                   align = "v", 
#                   ncol = 2, nrow = 2,
#                   heights = c(1, 0.6), widths = c(1,1),
#                   common.legend = FALSE) -> p_allu_human2
# 
# p_allu_human2
# 
# p_allu_human2 %>%
#   export::graph2ppt(append = TRUE,
#                     width = 317.48031496 * 1,
#                     height = 0.618 * 317.48031496 * 1.75 , paper = "A4",  scaling = 2,
#                     file = out_pptx)
```


```{r}
# ggpubr::ggarrange(p_gram_h2_no_leg + ylab(NULL) + 
#                     theme(axis.title.x=element_blank(),
#                           axis.text.x=element_blank(),
#                           axis.ticks.x=element_blank()),
#                   p_hist_human2_fam_no_leg + 
#                     theme(
#                       strip.background = element_blank(),
#                       strip.text.x = element_blank()
#                     ) +
#                     theme(axis.title.x=element_blank(),
#                           axis.text.x=element_blank(),
#                           axis.ticks.x=element_blank()),
#                   p_hist_human2_strain_no_leg + 
#                     theme(
#                       strip.background = element_blank(),
#                       strip.text.x = element_blank()
#                     ) + ylab(NULL) + 
#                     theme(axis.title.x=element_blank(),
#                           axis.text.x=element_blank(),
#                           axis.ticks.x=element_blank()),
#                   p_int_h2_no_leg + ylab(NULL) + coord_cartesian(ylim = c(50,100)) +
#                     theme(
#                       strip.background = element_blank(),
#                       strip.text.x = element_blank()
#                     ), 
#                   align = "v",
#                   ncol = 1, 
#                   heights = c(0.15, 0.4, 0.8,0.15),
#                   common.legend = FALSE) -> all_plot_h2
# 
# all_plot_h2
# 
# all_plot_h2 %>% 
#   export::graph2ppt(append = TRUE, 
#                     width = 317.48031496 * 3.5, 
#                     height = 0.618 * 317.48031496 * 5 , paper = "A4",  scaling = 2,
#                     file = out_pptx)
# 
# 
# p_gram_h2_leg  %>% 
#   export::graph2ppt(append = TRUE, 
#                     width = 317.48031496 * 1, 
#                     height = 0.618 * 317.48031496 * 1 , paper = "A4",  scaling = 2,
#                     file = out_pptx)

p_hist_human2_fam_leg %>% 
  export::graph2ppt(append = TRUE, 
                    width = 317.48031496 * 1, 
                    height = 0.618 * 317.48031496 * 1 , paper = "A4",  scaling = 2,
                    file = out_pptx)

p_hist_human2_strain_leg  %>% 
  export::graph2ppt(append = TRUE, 
                    width = 317.48031496 * 4, 
                    height = 0.618 * 317.48031496 * 2 , paper = "A3",  scaling = 2,
                    file = out_pptx)
# 
# p_int_h2_leg  %>% 
#   export::graph2ppt(append = TRUE, 
#                     width = 317.48031496 * 1, 
#                     height = 0.618 * 317.48031496 * 1 , paper = "A4",  scaling = 2,
#                     file = out_pptx)
# 
# 
# p_gram_h2_no_leg$data %>% 
#   select(OTU, Sample, Abundance, Treatment, Day_of_Treatment) %>% 
#   xlsx::write.xlsx(file = out_xlsx,
#                    append = TRUE,
#                    sheetName = "p_gram_h2",
#                    showNA = TRUE)
# p_hist_human2_fam$data %>% 
#   select(OTU, Sample, Abundance, Treatment, Day_of_Treatment) %>% 
#   xlsx::write.xlsx(file = out_xlsx,
#                    append = TRUE,
#                    sheetName = "p_hist_human2_fam",
#                    showNA = TRUE)
# p_hist_human2_strain$data %>% 
#   select(Sample, top ,Abundance, Treatment, Day_of_Treatment) %>% 
#   pivot_wider(names_from = top, values_from = Abundance) %>% 
#   xlsx::write.xlsx(file = out_xlsx,
#                    append = TRUE,
#                    sheetName = "p_hist_human2_strain",
#                    showNA = TRUE)
# p_int_h2$data %>% 
#   select(OTU, Sample, Abundance, Treatment, Day_of_Treatment) %>% 
#   xlsx::write.xlsx(file = out_xlsx,
#                    append = TRUE,
#                    sheetName = "p_int_h2",
#                    showNA = TRUE)

```


```{r}
# ggpubr::ggarrange(p_gram_h2_no_leg + ylab(NULL) + 
#                     theme(axis.title.x=element_blank(),
#                           axis.text.x=element_blank(),
#                           axis.ticks.x=element_blank()),
#                   p_hist_human2_fam_no_leg + 
#                     theme(
#                       strip.background = element_blank(),
#                       strip.text.x = element_blank()
#                     ) +
#                     theme(axis.title.x=element_blank(),
#                           axis.text.x=element_blank(),
#                           axis.ticks.x=element_blank()),
#                   treat_speces + 
#                     theme(
#                       strip.background = element_blank(),
#                       strip.text.x = element_blank()
#                     ) + ylab(NULL) + 
#                     theme(axis.title.x=element_blank(),
#                           axis.text.x=element_blank(),
#                           axis.ticks.x=element_blank()),
#                   p_int_h2_no_leg + ylab(NULL) + coord_cartesian(ylim = c(50,100)) +
#                     theme(
#                       strip.background = element_blank(),
#                       strip.text.x = element_blank()
#                     ), 
#                   align = "hv",
#                   ncol = 1, 
#                   heights = c(0.15, 0.4, 0.8,0.15),
#                   common.legend = FALSE) -> all_plot_h2
# 
# all_plot_h2
# 
# all_plot_h2 %>% 
#   export::graph2ppt(append = TRUE, 
#                     width = 317.48031496 * 3.5, 
#                     height = 0.618 * 317.48031496 * 5 , paper = "A4",  scaling = 2,
#                     file = out_pptx)

```



### chicken1 :

#### gram:

```{r}
ps_rare_gram_int %>% 
  subset_samples(Model2 == "Chicken1" & Treatment != "DONOR" & Day_of_Treatment >= -1) %>% 
  # microViz::ps_mutate(Reactor_Treatment_Dose = paste0(Reactor_Treatment_Dose, "_", Fermentation)) %>% 
  transform_sample_counts(function(x) x/sum(x) * 100) %>%
  physeq_sel_tax_table(c("gram_stain")) %>% 
  # subset_taxa(Strain %in% top_taxa_per_group_h2) %>% # extract only the taxa to display - after percentage normalisation
  microViz::phyloseq_validate() %>%
  microViz::tax_fix() %>%
  microViz::ps_arrange(Day_of_Treatment) %>% 
  # microViz::tax_mutate(gram = left_join(tax_table(.) %>% data.frame(), gram,
  # by = c("Strain" = "ASV") %>% select(gram_neg_class))) %>% 
  microViz::comp_barplot(
    tax_transform_for_plot = "identity",
    tax_level = "gram_stain",
    label = "Day_of_Treatment", # name an alternative variable to label axes
    palette = microViz::distinct_palette(2, pal = "greenArmytage"),
    n_taxa = 10, # give more taxa unique colours
    merge_other = FALSE, # split the "other" category to display alpha diversity
    bar_width = 0.9, # reduce the bar width to 70% of one row
    bar_outline_colour = "grey5",
    sample_order = "default") +
  ylab("Proportion - %") + xlab( " Day (Treatment) ") -> p_gram_h1

p_gram_h1 + facet_grid(. ~ Reactor_Treatment_Dose, scales = "free", space = "free_x", drop = TRUE) -> p_gram_h1

p_gram_h1 + theme_light() +theme(legend.position = "none") -> p_gram_h1_no_leg

p_gram_h1_no_leg

p_gram_h1 %>%  
  ggpubr::get_legend() %>% 
  ggpubr::as_ggplot() -> p_gram_h1_leg

```


```{r}
p_gram_h1_f1$data %>%
  # filter(Model2 == "Human2") %>%
  mutate(gram_stain = factor(gram_stain, levels = c("Gram-positive", "Gram-negative", "unknown-gram"))) %>%
  # mutate(combo = paste0(Model2, "_" , Reactor_F1_F2)) %>%
  # mutate(combo = as.factor(combo)) %>%
  select(SAMPLE, gram_stain,OTU, Abundance, Day_of_Treatment, Reactor_Treatment_Dose) %>% 
  ggplot(.,
         aes(x = Day_of_Treatment, 
             y = Abundance,
             stratum = gram_stain, 
             alluvium = OTU,
             fill = gram_stain,
             label = gram_stain))+
  facet_grid(Reactor_Treatment_Dose  ~ .)+
  # scale_fill_manual(values = c("red", "yellow", "green3"))+
  geom_flow(stat = "alluvium", lode.guidance = "frontback", color = "darkgray")+
  geom_stratum() +
  theme_light() +
  # geom_text(stat = "alluvium", aes(label = top), lode.guidance = "frontback") +
  # scale_fill_manual(values = col_ASVs) +
  scale_color_manual(values  =   microViz::distinct_palette(3, pal = "greenArmytage")) +
  scale_fill_manual(values  =   microViz::distinct_palette(3, pal = "greenArmytage")) +
  # scale_color_manual(values  = microViz::distinct_palette(10, pal = "kelly")) +
  # scale_fill_manual(values  = microViz::distinct_palette(10, pal = "kelly")) +
  ylab("Proportion - %") -> p_allu_h1_gram 

p_allu_h1_gram + theme(legend.position = "none") -> p_allu_h1_gram_noleg 

p_allu_h1_gram_noleg 

p_allu_h1_gram %>% 
  ggpubr::get_legend() %>% 
  ggpubr::as_ggplot() ->  p_allu_h1_gram_leg

p_allu_h1_gram_leg
```



#### intrinsic:

```{r}
# ps_rare_gram_int %>% 
#   subset_samples(Model2 == "Chicken1" & Treatment != "DONOR" & Day_of_Treatment >= -1) %>% 
#   microViz::ps_mutate(Reactor_Treatment_Dose = paste0(Reactor_Treatment_Dose, "_", Fermentation)) %>% 
#   transform_sample_counts(function(x) x/sum(x) * 100) %>%
#   physeq_sel_tax_table(c( "intrinsic")) %>% 
#   # microbiome::transform("log") %>% 
#   # speedyseq::tax_glom("intrinsic") %>% 
#   # physeq_sel_tax_table(c("gram_stain","intrinsic")) %>% 
#   # physeq_glom_rename(phyloseq = ., speedyseq = TRUE, rename_ASV = FALSE, taxrank = "gram_stain") %>% 
#   # subset_taxa(Strain %in% top_taxa_per_group_h2) %>% # extract only the taxa to display - after percentage normalisation
#   microViz::phyloseq_validate() %>%
#   microViz::tax_fix() %>%
#   microViz::ps_arrange(Day_of_Treatment) %>% 
#   # microViz::tax_mutate(gram = left_join(tax_table(.) %>% data.frame(), gram,
#   # by = c("Strain" = "ASV") %>% select(gram_neg_class))) %>% 
#   microViz::comp_barplot(
#     tax_transform_for_plot = "identity",
#     tax_level = "intrinsic",
#     label = "Day_of_Treatment", # name an alternative variable to label axes
#     # palette = microViz::distinct_palette(3, pal = "brewerPlus"),
#     n_taxa = 10, # give more taxa unique colours
#     merge_other = FALSE, # split the "other" category to display alpha diversity
#     bar_width = 0.9, # reduce the bar width to 70% of one row
#     bar_outline_colour = "grey5",
#     sample_order = "default") +
#   ylab("Proportion - %") + xlab( " Day (Treatment) ") -> p_int_h1_f1
# 
# p_int_h1_f1 + facet_grid(Kingdom ~ Reactor_Treatment_Dose, scales = "free", space = "free_x", drop = TRUE) -> p_int_h1_f1
# 
# p_int_h1_f1 %>% 
#   ggpubr::set_palette(p = ., palette = "jco") -> p_int_h1_f1
# 
# p_int_h1_f1 + theme_light() +theme(legend.position = "none") -> p_int_h1_f1_no_leg
# 
# p_int_h1_f1_no_leg
# 
# p_int_h1_f1 %>%  
#   ggpubr::get_legend() %>% 
#   ggpubr::as_ggplot() -> p_int_h1_f1_leg
# 
# p_int_h1_f1_leg
```


```{r}
# p_int_h1_f1$data %>%
#   # filter(Model2 == "Human2") %>%
#   # microViz::ps_mutate(Reactor_Treatment_Dose = paste0(Reactor_Treatment_Dose, "_", Fermentation)) %>% 
#   # mutate(combo = paste0(Model2, "_" , Reactor_F1_F2)) %>%
#   # mutate(combo = as.factor(combo)) %>%
#   select(SAMPLE, top,OTU, Abundance, Day_of_Treatment, Reactor_Treatment_Dose) %>% 
#   ggplot(.,
#          aes(x = Day_of_Treatment, 
#              y = Abundance,
#              stratum = top, 
#              alluvium = OTU,
#              fill = top,
#              label = top))+
#   facet_grid(Reactor_Treatment_Dose  ~ .)+
#   # scale_fill_manual(values = c("red", "yellow", "green3"))+
#   geom_flow(stat = "alluvium", lode.guidance = "frontback", color = "darkgray")+
#   geom_stratum() +
#   theme_light() +
#   # geom_text(stat = "alluvium", aes(label = top), lode.guidance = "frontback") +
#   # scale_fill_manual(values = col_ASVs) +
#   scale_color_manual(values  =  microViz::distinct_palette(3, pal = "greenArmytage")) +
#   # scale_color_manual(values  = microViz::distinct_palette(10, pal = "kelly")) +
#   # scale_fill_manual(values  = microViz::distinct_palette(10, pal = "kelly")) +
#   ylab("Proportion - %") -> p_allu_h1_f1_int
# 
# p_allu_h1_f1_int %>% 
#   ggpubr::set_palette(p = ., palette = "jco") -> p_allu_h1_f1_int
# 
# p_allu_h1_f1_int + theme(legend.position = "none") -> p_allu_h1_f1_int_noleg 
# 
# p_allu_h1_f1_int_noleg 
# 
# p_allu_h1_f1_int %>% 
#   ggpubr::get_legend() %>% 
#   ggpubr::as_ggplot() -> p_allu_h1_f1_int_leg
```

```{r}
ggpubr::ggarrange(p_allu_human1_f1_noleg + xlab(NULL),
                  p_allu_human1_ASV_f1_noleg + ylab(NULL) + xlab(NULL),
                  p_allu_h1_f1_int_noleg +  coord_cartesian(ylim = c(50,100)),
                  p_allu_h1_f1_gram_noleg + ylab(NULL),
                  align = "v", 
                  ncol = 2, nrow = 2,
                  heights = c(1, 0.6), widths = c(1,1),
                  common.legend = FALSE) -> p_allu_human1_f1

p_allu_human1_f1

p_allu_human1_f1 %>%
  export::graph2ppt(append = TRUE,
                    width = 317.48031496 * 1,
                    height = 0.618 * 317.48031496 * 1.75 , paper = "A4",  scaling = 2,
                    file = out_pptx)
```

```{r}
ggpubr::ggarrange(p_gram_h1_f1_no_leg + ylab(NULL) + 
                    theme(axis.title.x=element_blank(),
                          axis.text.x=element_blank(),
                          axis.ticks.x=element_blank()),
                  p_hist_human1_f1_fam_no_leg + 
                    theme(
                      strip.background = element_blank(),
                      strip.text.x = element_blank()
                    ) +
                    theme(axis.title.x=element_blank(),
                          axis.text.x=element_blank(),
                          axis.ticks.x=element_blank()),
                  p_hist_human1_f1_strain_no_leg + 
                    theme(
                      strip.background = element_blank(),
                      strip.text.x = element_blank()
                    ) + ylab(NULL) + 
                    theme(axis.title.x=element_blank(),
                          axis.text.x=element_blank(),
                          axis.ticks.x=element_blank()),
                  p_int_h1_f1_no_leg + ylab(NULL) + coord_cartesian(ylim = c(50,100)) +
                    theme(
                      strip.background = element_blank(),
                      strip.text.x = element_blank()
                    ), 
                  align = "v",
                  ncol = 1, 
                  heights = c(0.15, 0.4, 0.8,0.15),
                  common.legend = FALSE) -> all_plot_h1_f1

all_plot_h1_f1

all_plot_h1_f1 %>% 
  export::graph2ppt(append = TRUE, 
                    width = 317.48031496 * 3.5, 
                    height = 0.618 * 317.48031496 * 5 , paper = "A4",  scaling = 2,
                    file = out_pptx)


p_gram_h1_f1_leg  %>% 
  export::graph2ppt(append = TRUE, 
                    width = 317.48031496 * 1, 
                    height = 0.618 * 317.48031496 * 1 , paper = "A4",  scaling = 2,
                    file = out_pptx)

p_hist_human1_f1_fam_leg  %>% 
  export::graph2ppt(append = TRUE, 
                    width = 317.48031496 * 1, 
                    height = 0.618 * 317.48031496 * 1 , paper = "A4",  scaling = 2,
                    file = out_pptx)

p_hist_human1_f1_strain_leg  %>% 
  export::graph2ppt(append = TRUE, 
                    width = 317.48031496 * 4, 
                    height = 0.618 * 317.48031496 * 2 , paper = "A3",  scaling = 2,
                    file = out_pptx)

p_int_h1_f1_leg  %>% 
  export::graph2ppt(append = TRUE, 
                    width = 317.48031496 * 1, 
                    height = 0.618 * 317.48031496 * 1 , paper = "A4",  scaling = 2,
                    file = out_pptx)



p_gram_h1_f1$data %>% 
  select(OTU, Sample, Abundance, Treatment, Day_of_Treatment) %>% 
  xlsx::write.xlsx(file = out_xlsx,
                   append = TRUE,
                   sheetName = "p_gram_h1_f1",
                   showNA = TRUE)
p_hist_human1_f1_fam$data %>% 
  select(OTU, Sample, Abundance, Treatment, Day_of_Treatment) %>% 
  pivot_wider(names_from = OTU, values_from = Abundance) %>% 
  xlsx::write.xlsx(file = out_xlsx,
                   append = TRUE,
                   sheetName = "p_hist_human1_f1_fam",
                   showNA = TRUE)
p_hist_human1_f1_strain$data %>% 
  select(Sample, top ,Abundance, Treatment, Day_of_Treatment) %>% 
  pivot_wider(names_from = top, values_from = Abundance) %>% 
  xlsx::write.xlsx(file = out_xlsx,
                   append = TRUE,
                   sheetName = "p_hist_human1_f1_strain",
                   showNA = TRUE)
p_int_h1_f1$data %>% 
  select(OTU, Sample, Abundance, Treatment, Day_of_Treatment) %>% 
  pivot_wider(names_from = OTU, values_from = Abundance) %>% 
  xlsx::write.xlsx(file = out_xlsx,
                   append = TRUE,
                   sheetName = "p_int_h1_f1",
                   showNA = TRUE)

```


-   explore stability metrics: <https://microbiome.github.io/tutorials/Stability.html> to detect taxa impacted. or filter by coefficient variation. or Hannah's classification

Bimodal population distribution across individuals is often associated with instable intermediate abundances within individuals:

```{r,  eval = FALSE}
# Use relative abundances
pseq <- microbiome::transform(ps_rare %>%  subset_samples(Model2 == "Human2"), "compositional")

# Merge rare taxa to speed up examples
pseq <- microbiome::aggregate_rare(ps_rare, level = "Strain", detection = .1/100, prevalence = 10/100)

# For cross-sectional analysis, include
# only the treated points.

pseq0 <- subset_samples(pseq, Day_of_Treatment > 1)

# intermediate_stability: within subjects:
pseq %>% 
  microViz::ps_mutate(subject = Reactor,
                      time = Day_of_Treatment) %>% 
  intermediate_stability(., #output = "scores", 
                         output = "scores") -> intermediate.stability

intermediate.stability %>% 
  head()
```

```{r, message = FALSE, warning = FALSE, eval = FALSE}
# Bimodality is better estimated from abundances at log scale (such as CLR)
pseq0.clr <- microbiome::transform(pseq0, "clr")

set.seed(4433)

# bimodality: between subjects
bimodality.score <- bimodality(pseq0.clr, method = "potential_analysis",
                               bs.iter = 999, peak.threshold = 5,
                               min.density = 5)

bimodality.score %>% 
  head()
```

```{r, eval=FALSE}
taxa <- taxa(pseq0)
df <- data.frame(group = taxa,
                 intermediate.stability = intermediate.stability[taxa],
                 bimodality = bimodality.score[taxa])
theme_set(theme_bw(20))

library(ggrepel)

p <- ggplot(df,
            aes(x = intermediate.stability, y = bimodality, label = '')) + #Doesn't add labels to points
  geom_text_repel() +
  geom_point()

#Labels only those that have bimodality > 0.4 and intermediate stability < 0, adjusts the placement of labels
p <- p + geom_text(aes(label = ifelse(bimodality > 0.75 | intermediate.stability < -0.25, group, '')), vjust = "inward", hjust = "inward")

print(p)
```

```{r, eval=FALSE}
p$data
```

```{r, eval=FALSE}
# Log10 abundance for a selected taxonomic group
# Pick the most bimodal taxa as an example
tax  <- names(which.max(bimodality.score))

# Detect tipping points at log10 abundances
x <- abundances(microbiome::transform(pseq, "clr"))[tax,]

# Bootstrapped potential analysis to identify potential minima
# in practice, use more bootstrap iterations
potential.minima <- potential_analysis(x, bs.iter = 10)$minima

# Same with earlywarnings package (without bootstrap ie. less robust)
# library(earlywarnings)
# res <- livpotential_ews(x)$min.points

# Identify the potential minimum locations as tipping point candidates
tipping.samples <- sapply(potential.minima, function (m) {names(which.min(abs(sort(x) - m)))})
tipping.point <- abundances(pseq)[tax, tipping.samples]
print(tipping.point)
```


```{r, eval=FALSE}
# Illustrate the detected tipping point
plot(density(x), main = tax)
abline(v = x[tipping.samples])
```


```{r}
# # Bimodality hotplot:
# # Consider a unique sample from each subject: the baseline time point 
# p <- microbiome::hotplot(pseq0, tax, tipping.point = 0.005)
# print(p)
# 
# # Visualize bimodality
# pv <- plot_tipping(pseq, tax, tipping.point = 0.005)
# print(pv)
```

```{r}
sessionInfo()
```
